---
title: "Combine BGC SUs into single SU"
author: "Will MacKenzie"
date: "13/09/2021"
output: html_document
---

```{r setup, include=FALSE}
require(data.table)
require(tidyverse)
require(dplyr)
require(data.tree)
source("./_functions/_TreetoTable.R")
source("./_functions/_TabletoTree.R")
```

#### Import becdb
```{r set folders, include=FALSE}
require(DBI) #loads odbc as well
 becdb <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver13_2026.accdb;")

BGCs <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive")%>% filter(is.na(VersionRetired)) %>% select(BGC_Label, BGC_NoSpace, BGC_Zone) 
SiteSeries<- dbReadTable(becdb, "tblBEC_SiteSeriesWorkingandArchive") %>% filter(!Status == "Retire") %>% select(BGC_NoSpace, SS_NoSpace, SS_Label)
dbDisconnect(becdb)
BGCs_noSS <- anti_join(BGCs , SiteSeries) %>% drop_na

# becdb <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECdb_Working/BECdb_ver12_2021_reduced.accdb;")
# dbWriteTable(becdb, "BGCs_NoSS",  BGCs_noSS, overwrite = TRUE, batch_rows = 1)
# dbDisconnect(becdb)
```

#### Import all BECMaster tables in /raw_data folder
```{r set folders, include=FALSE}
require(DBI) #loads odbc as well
 becmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster/BECMaster_fixing.accdb;")

plot.env <- dbReadTable(becmaster, "BECMaster_fixing_Env") %>% mutate(Longitude = ifelse(Longitude<0, Longitude, 0-Longitude))
plot.admin <- dbReadTable(becmaster, "BECMaster_fixing_Admin")
# VEG <- dbReadTable(becmaster, "BECMaster_fixing_Veg")
env.fields <- colnames(plot.env)
odbc::odbcListColumns(becmaster, table = "BECMaster_fixing_ENV")
dbDisconnect(becmaster)

site.units<- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver13_2026.accdb;")
siteseries <- dbReadTable(site.units, "tblBEC_SiteSeriesWorkingandArchive")
dbDisconnect(site.units)
ssv13 <- siteseries %>% filter(!Status %in% "Retire") %>% select(SS_Label) %>% distinct

```

#### add BGC to association codes in SU
```{r}
# source("../BEC_R_working/_functions/_combined_su.R")
# db <- "D:/BC_Correlation2_Vpro_2023/SUs_to_fix.accdb"
# #db <- "D:/BC_Correlation2_Vpro_2023/Coast_Association_Exercise.accdb"
# su <- combined_su(db)
# library(dplyr)
# library(stringr)
# 
# su <- left_join(su, plot.admin, by = c("PlotNumber" = "Plot")) %>%
#   mutate(
#     GIS_BGC_pad = str_pad(GIS_BGC, width = 9, side = "right"),
#     SiteUnit2   = paste0(GIS_BGC_pad, " /", SiteUnit)
#   ) %>%
#   select(PlotNumber, SiteUnit2)
# 
# fwrite(su, "./clean_tabs/SU_with_BGC_added.csv")

```


### Combine all published SUs
```{r function to create provincial SU table}
library(odbc)
library(dplyr)
library(stringr)

# Function to extract and process _SU tables from a single database
buildSU <- function(db) {
  correlation <- dbConnect(odbc::odbc(), .connection_string = paste0("Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=", db, ";"))
  
  su_tables <- dbListTables(correlation) %>% str_subset("_SU$")  # Get tables ending in "_SU"
  
  all_su <- lapply(setNames(nm = su_tables), dbReadTable, conn = correlation)  # Read tables
  
  dbDisconnect(correlation)  # Close connection
  
  if (length(all_su) > 0) {
    SU <- do.call(rbind.data.frame, all_su) %>%
      mutate(bgc = substr(SiteUnit, 1, 9)) %>%
      drop_na() %>%
      distinct(PlotNumber, .keep_all = TRUE) %>%
      filter(!grepl('poor|low|[$]|add|nudum|_[[:alpha:]]|X|omit|unplaced|moved|support', SiteUnit)) %>%
      arrange(desc(PlotNumber)) %>%
      select(PlotNumber, SiteUnit)  # Keep relevant columns

    return(SU)
  } else {
    return(NULL)  # Return NULL if no _SU tables exist
  }
}

# Process all .accdb files in a folder
process_all_SU_databases <- function(folder_path) {
  db_files <- list.files(folder_path, pattern = "\\.accdb$", full.names = TRUE)  # Get all Access files
  
  SU_combined <- bind_rows(lapply(db_files, buildSU), .id = "source_db")  # Merge results
  
  return(SU_combined)
}

# Example usage (modify folder path as needed)
folder_path <- "D:/BC_Correlation2_Vpro_2023/Sorts_published_LMH/"  # Path to folder containing .accdb files
SU_all <- process_all_SU_databases(folder_path)
SU_all <- SU_all %>% distinct(PlotNumber, SiteUnit) #%>% arrange(desc(PlotNumber))
duplicates <- SU_all %>%
  group_by(PlotNumber) %>%
  filter(n() > 1) %>%  # Keep only duplicate PlotNumbers
  arrange(PlotNumber)  # Sort for clarity

master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb;")
dbWriteTable(master_su, "All_BECv13_published_SU",  SU_all, overwrite = TRUE, batch_rows = 1, row.names = FALSE)
dbDisconnect(master_su)
fwrite(SU_all, "./clean_tabs/All_BECv13_published_SU.csv")
# update the BEC
# View output
head(SU_all)

missing_ss <- anti_join(ssv13, SU_all, by = c("SS_Label" = "SiteUnit"))
```

### Combine all current SU tables into a single AllBGC SU table
CaribooNew
CassiarNew
Coastnew
ColumbiaNew
KB
OK
Skeena
Boreal
Omineca



```{r combine all current and working SU in allBGC}
  require(DBI) #loads odbc as well
buildSU <- function(db){
  correlation <- dbConnect(odbc::odbc(), .connection_string = paste0("Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=",db , ";"))
  all_su <- lapply(setNames(nm = (dbListTables(correlation)%>% str_subset("_SU"))), dbReadTable, conn = correlation)
  dbDisconnect(correlation)
  SU <- do.call(rbind.data.frame, all_su)
  SU <- SU %>% mutate(bgc = substr(SiteUnit,1,9)) %>% drop_na() %>% distinct(PlotNumber, .keep_all = TRUE) %>%
  filter(!grepl('poor|low|[$]|add|nudum|_[[:alpha:]]|X|omit|unplaced', SiteUnit)) %>% 
  arrange(desc(PlotNumber))
  SU2 <- SU %>% dplyr::select(PlotNumber, SiteUnit)
  return(SU)
}


CariboonewSU <- buildSU(db = "D:/BC_Correlation2_Vpro_2023/Cariboo_SUs.accdb")
CassiarnewSU <- buildSU(db = "D:/BC_Correlation2_Vpro_2023/Cassiar_SUs.accdb")
CoastSU <- buildSU(db = "D:/BC_Correlation2_Vpro_2023/published_sorts/CoastArea_Sorts.accdb")
#ColumbianewSU <- buildSU(db = "D:/BC_Correlation2_Vpro_2023/Columbia_FG_draft.accdb")
KBSU <- buildSU(db = "D:/BC_Correlation2_Vpro_2023/published_sorts/KootenayBoundary_Sorts.accdb")
OKSU <- buildSU(db = "D:/BC_Correlation2_Vpro_2023/published_sorts/ThompsonOK_Sorts.accdb")
SkeenaSU <- buildSU(db = "D:/BC_Correlation2_Vpro_2023/published_sorts/Skeena_LMH26_Sorts.accdb")
BorealSU <- buildSU(db = "D:/BC_Correlation2_Vpro_2023/published_sorts/LMH65_Boreal_Sorts.accdb")
OminecaSU <- buildSU(db = "D:/BC_Correlation2_Vpro_2023/published_sorts/OminecaPeace_SU_Tables.accdb")
AllBGC_SU <- rbind(CariboonewSU, CassiarnewSU, CoastSU,  KBSU, OKSU, SkeenaSU, BorealSU, OminecaSU)
##need to build a check for duplicates in here
AllBGC_SU <- AllBGC_SU %>% distinct(PlotNumber, .keep_all = TRUE) %>% arrange(desc(PlotNumber))

master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb;")
dbWriteTable(master_su, "All_BGCv13_SU",  AllBGC_SU, overwrite = TRUE, batch_rows = 1, row.names = FALSE)
dbDisconnect(master_su)

LMH_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/published_sorts/LMH_SU_Tables.accdb;")
dbWriteTable(LMH_su, "LMH77_Published_SU",  CoastSU, overwrite = TRUE, batch_rows = 1, row.names = FALSE)
dbDisconnect(LMH_su)
```
Compare to BECdb list of units
```{r}
site.units<- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver13_2026.accdb;")
siteseries <- dbReadTable(site.units, "tblBEC_SiteSeriesWorkingandArchive")
dbDisconnect(site.units)
ssv13 <- siteseries %>% filter(!Status %in% "Retire") %>% select(SS_Label) %>% distinct
missing_ss <- anti_join(ssv13, AllBGC_SU, by = c("SS_Label" = "SiteUnit"))
```


### New Columbia Guide
```{r}
correlation <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/Columbia_FG_draft.accdb;")
all_su <- lapply(setNames(nm = (dbListTables(correlation)%>% str_subset("_SU"))), dbReadTable, conn = correlation)
dbDisconnect(correlation)
SU <- do.call(rbind.data.frame, all_su)
SU <- SU %>% mutate(bgc = substr(SiteUnit,1,9)) %>% drop_na() %>% distinct(PlotNumber, .keep_all = TRUE) %>%
  arrange(desc(PlotNumber))
ss.unique <- SU %>% select(SiteUnit) %>% distinct
SU2 <- SU %>% dplyr::select(PlotNumber, SiteUnit)

master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/Columbia_FG_draft.accdb;")
dbWriteTable(master_su, "All_Columbia_FG_SU",  SU2, overwrite = TRUE, batch_rows = 1, row.names = FALSE)
dbDisconnect(master_su)

```

## Import SU table and BECdb for current BGC units and site series. Look for missing units. Create complete SU 

###COAST

```{r import SU tables and look for errors}
require(DBI) #loads odbc as well
correlation <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; 
DBQ=D:/BC_Correlation2_Vpro_2023/CoastGuide_Forested.accdb;")
all_su <- lapply(setNames(nm = (dbListTables(correlation)%>% str_subset("_SU"))), dbReadTable, conn = correlation)
dbDisconnect(correlation)
SU <- do.call(rbind.data.frame, all_su)
SU <- SU %>% mutate(bgc = substr(SiteUnit,1,9)) %>% drop_na() %>% distinct(PlotNumber, .keep_all = TRUE) %>% 
  arrange(desc(PlotNumber))
ss.unique <- SU %>% select(SiteUnit) %>% distinct
SU2 <- SU %>% dplyr::select(PlotNumber, SiteUnit)
## *** The field length of plotnumber is changed to 255 in the exported database table. This causes a fail in veg summary reporting from the hierarchy. need to change the Plot Number field length back to 7 **************
master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/CoastGuide_Hierarchy.accdb;")
dbWriteTable(master_su, "All_Coast_Forest_2024v8_SU",  SU2, overwrite = TRUE, batch_rows = 1, row.names = FALSE)
dbDisconnect(master_su)

#%>% mutate(PlotNumber = paste0("'",PlotNumber))
#fwrite(SU, "./clean_tabs/CoastGuide_All_BGC_Forest_SU.csv")
###remove phases and seral units
SU_siteseries <- SU %>% 
  filter(!str_detect(SiteUnit, '[$]'))
##phases = c("a", "b", "c")
SU_siteseries$SiteUnit <- str_replace(SU_siteseries$SiteUnit, "[abc]$", "")
fwrite(SU_siteseries, "./clean_tabs/CoastGuide_All_BGC_Forest_SS.csv")
SU_coast <- SU_siteseries

###Check that there is plot data for all
missingplots <- anti_join(SU_coast, plot.env)
plots <- right_join(SU_coast, plot.env)
### Check for duplicate plot usage
usemorethanone <- SU_coast %>% dplyr::group_by(PlotNumber) %>% dplyr::mutate(dups = n()) %>% filter(dups >1) %>% ungroup() %>% arrange(PlotNumber)
fwrite(usemorethanone, "./review_outputs/Coast_Plots_Used_Morethan_Once.csv")
###Summaries by BGC and by SS
plots_bgc <- SU_coast %>% group_by(bgc) %>% dplyr::summarise(plots = n())
plots_SS <- SU_coast %>% group_by(SiteUnit) %>% dplyr::summarise(plots = n())
SS_count <- length(unique(SU_coast$SiteUnit))
siteunits_toofew <- plots_SS %>% filter(plots <5) #%>% dplyr::select(- PlotNumber) %>% distinct

fwrite(siteunits_toofew, "./review_outputs/Coast_SiteUnits_w_lessthan5plots.csv")

all_zonals <- SU2 %>% filter(str_detect(SiteUnit, "101|/01"))
zonal_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; 
DBQ=D:/BC_Correlation2_Vpro_2023/CoastGuide_Hierarchy.accdb;")
dbWriteTable(zonal_su, "Coast_Zonals_SU", all_zonals, overwrite = TRUE, batch_rows = 1, row.names = FALSE)
dbDisconnect(zonal_su)

```
###SOUTHERN INTERIOR Forested

```{r import SU tables and look for errors}
require(DBI) #loads odbc as well
correlation <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/SIFR_LMH_SU.accdb;")
all_su <- lapply(setNames(nm = (dbListTables(correlation)%>% str_subset("_SU"))), dbReadTable, conn = correlation)
dbDisconnect(correlation)
SU <- do.call(rbind.data.frame, all_su)
SU <- SU %>% mutate(bgc = substr(SiteUnit,1,9)) %>% drop_na() %>% distinct(PlotNumber, .keep_all = TRUE) %>%
  arrange(desc(PlotNumber))
SU2 <- SU %>% dplyr::select(PlotNumber, SiteUnit)

require(DBI) #loads odbc as well
correlation <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/Columbia_FG_draft.accdb;")
all_su <- lapply(setNames(nm = (dbListTables(correlation)%>% str_subset("_SU"))), dbReadTable, conn = correlation)
dbDisconnect(correlation)
SU_draft <- do.call(rbind.data.frame, all_su)
SU_draft <- SU_draft %>% mutate(bgc = substr(SiteUnit,1,9)) %>% drop_na() %>% distinct(PlotNumber, .keep_all = TRUE) %>%
  arrange(desc(PlotNumber))
SU_draft2 <- SU_draft %>% dplyr::select(PlotNumber, SiteUnit)


SU <- rbind(SU,SU_draft) %>% distinct(PlotNumber, .keep_all = TRUE) %>% arrange(desc(PlotNumber))

#%>% mutate(PlotNumber = paste0("'",PlotNumber))
#fwrite(SU, "./clean_tabs/CoastGuide_All_BGC_Forest_SU.csv")
###remove phases and seral units
SU_siteseries <- SU %>%
  filter(!str_detect(SiteUnit, '[$]'))%>% filter(!str_detect(SiteUnit, "/W|/R|/G|/F|/A|/V")) %>% filter(!str_detect(SiteUnit, fixed("(G")))
phases = c("a", "b", "c")
SU_siteseries$SiteUnit <- str_replace(SU_siteseries$SiteUnit, "a$|b$", "")
SU_siteseries2 <- SU_siteseries %>% dplyr::select(-bgc)
fwrite(SU_siteseries, "./clean_tabs/SIFR_Guide_All_BGC_Forest_SS.csv")
SU_zonals <- SU_siteseries2 %>% filter(str_detect(SiteUnit, "101|/01"))
###write combined SU to database
master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/SIFR_Hierarchy.accdb;")
dbWriteTable(master_su, "All_SIFR_Forest_SU",  SU_siteseries2, overwrite = TRUE, batch_rows = 1)
dbWriteTable(master_su, "All_SIFR_Zonals_SU",  SU_zonals, overwrite = TRUE, batch_rows = 1)
dbDisconnect(master_su)

SU_SIFR <- SU_siteseries
###Check that there is plot data for all
missingplots <- anti_join(SU_SIFR, plot.env)
plots <- right_join(SU_SIFR, plot.env)
### Check for duplicate plot usage
usemorethanone <- SU_SIFR %>% dplyr::group_by(PlotNumber) %>% dplyr::mutate(dups = n()) %>% filter(dups >1) %>% ungroup() %>% arrange(PlotNumber)
fwrite(usemorethanone, "./outputs/SIFR_Plots_Used_Morethan_Once.csv")
###Summaries by BGC and by SS
plots_bgc <- SU_SIFR %>% group_by(bgc) %>% dplyr::summarise(plots = n())
plots_SS <- SU_SIFR %>% group_by(SiteUnit) %>% dplyr::summarise(plots = n())
SS_count <- length(unique(SU_SIFR$SiteUnit))
siteunits_toofew <- plots_SS %>% filter(plots <5) #%>% dplyr::select(- PlotNumber) %>% distinct

fwrite(siteunits_toofew, "./outputs/SIFR_SiteUnits_w_lessthan5plots.csv")
```
###Northern INTERIOR Forested

```{r import SU tables and look for errors}
require(DBI) #loads odbc as well
correlation <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/NIFR_SU_Tables.accdb;")
all_su <- lapply(setNames(nm = (dbListTables(correlation)%>% str_subset("_SU"))), dbReadTable, conn = correlation)
dbDisconnect(correlation)
SU <- do.call(rbind.data.frame, all_su)
SU <- SU %>% mutate(bgc = substr(SiteUnit,1,9)) %>% drop_na() %>% distinct(PlotNumber, .keep_all = TRUE) %>%
  arrange(desc(PlotNumber))
SU2 <- SU %>% dplyr::select(PlotNumber, SiteUnit)

####Draft units
# require(DBI) #loads odbc as well
# correlation <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
# DBQ=D:/BC_Correlation2_Vpro_2023/Zonal_Cassiar_2022.accdb;")
# all_su <- lapply(setNames(nm = (dbListTables(correlation)%>% str_subset("_SU"))), dbReadTable, conn = correlation)
# dbDisconnect(correlation)
# SU_draft <- do.call(rbind.data.frame, all_su)
# SU_draft <- SU_draft %>% mutate(bgc = substr(SiteUnit,1,9)) %>% drop_na() %>% distinct(PlotNumber, .keep_all = TRUE) %>%
#   arrange(desc(PlotNumber))
# SU_draft2 <- SU_draft %>% dplyr::select(PlotNumber, SiteUnit)
# 
# 
# SU <- rbind(SU,SU_draft)
#%>% mutate(PlotNumber = paste0("'",PlotNumber))
#fwrite(SU, "./clean_tabs/CoastGuide_All_BGC_Forest_SU.csv")
###remove phases and seral units
SU_siteseries <- SU %>%
  filter(!str_detect(SiteUnit, '[$]'))%>% filter(!str_detect(SiteUnit, "/W|/R|/G|/F|/A|/V")) %>% filter(!str_detect(SiteUnit, fixed("(G")))
phases = c("a", "b", "c")
SU_siteseries$SiteUnit <- str_replace(SU_siteseries$SiteUnit, "a$|b$", "")
SU_siteseries2 <- SU_siteseries %>% dplyr::select(-bgc)
fwrite(SU_siteseries, "./clean_tabs/NIFR_Guide_All_BGC_Forest_SS.csv")
SU_zonals <- SU_siteseries2 %>% filter(str_detect(SiteUnit, "101|/01"))
###write combined SU to database
master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/NIFR_Hierarchy.accdb;")
dbWriteTable(master_su, "All_NIFR_Forest_SU",  SU_siteseries2, overwrite = TRUE, batch_rows = 1)
dbWriteTable(master_su, "All_NIFR_Zonals_SU",  SU_zonals, overwrite = TRUE, batch_rows = 1)
dbDisconnect(master_su)

SU_NIFR <- SU_siteseries
###Check that there is plot data for all
missingplots <- anti_join(SU_NIFR, plot.env)
plots <- right_join(SU_NIFR, plot.env)
### Check for duplicate plot usage
usemorethanone <- SU_NIFR %>% dplyr::group_by(PlotNumber) %>% dplyr::mutate(dups = n()) %>% filter(dups >1) %>% ungroup() %>% arrange(PlotNumber)
fwrite(usemorethanone, "./outputs/NIFR_Plots_Used_Morethan_Once.csv")
###Summaries by BGC and by SS
plots_bgc <- SU_NIFR %>% group_by(bgc) %>% dplyr::summarise(plots = n())
plots_SS <- SU_NIFR %>% group_by(SiteUnit) %>% dplyr::summarise(plots = n())
SS_count <- length(unique(SU_NIFR$SiteUnit))
siteunits_toofew <- plots_SS %>% filter(plots <5) #%>% dplyr::select(- PlotNumber) %>% distinct

fwrite(siteunits_toofew, "./outputs/NIFR_SiteUnits_w_lessthan5plots.csv")
```
```{r write combined SU for province}
all_SS <- rbind(SU_coast, SU_SIFR, SU_NIFR) 
all_SS$bgc <-  str_replace_all(all_SS$bgc, " ", "")
all_SS <- left_join(all_SS, BGCs, by = c("bgc" = "BGC_NoSpace")) %>% dplyr::select(PlotNumber, SiteUnit, bgc)
SS_notBECdb <- anti_join(all_SS, BGCs, by = c("bgc" = "BGC_NoSpace"))
BECdb_noSS <- anti_join(BGCs, all_SS, by = c("BGC_NoSpace" = "bgc"))

all_SS2 <- all_SS %>% select(PlotNumber, SiteUnit)
all_SS_dups <- all_SS2 %>% filter(duplicated(.[["PlotNumber"]])) %>% select(PlotNumber) %>% distinct
Plots_dups <- all_SS2 %>% filter(PlotNumber %in% all_SS_dups$PlotNumber)
fwrite(Plots_dups, "./outputs/PlotsUsedTwice.csv")

all_zonals <- all_SS2 %>% filter(str_detect(SiteUnit, "101|/01"))
all_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; 
DBQ=D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb;")
dbWriteTable(all_su, "All_BC_BGCs_SU", all_SS, overwrite = TRUE, batch_rows = 1, row.names = FALSE)
dbWriteTable(all_su, "All_BC_zonals_SU", all_zonals, overwrite = TRUE, batch_rows = 1, row.names = FALSE)
dbDisconnect(all_su)

```

###Drafts

```{r import SU tables and look for errors}
# require(DBI) #loads odbc as well
# correlation <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; 
# DBQ=D:/BC_Correlation2_Vpro_2023/InteriorGuides_Forested.accdb;")
# all_su <- lapply(setNames(nm = (dbListTables(correlation)%>% str_subset("_SU"))), dbReadTable, conn = correlation)
# dbDisconnect(correlation)
# SU <- do.call(rbind.data.frame, all_su)
# SU <- SU %>% mutate(bgc = substr(SiteUnit,1,9)) %>% drop_na() %>% distinct(PlotNumber, .keep_all = TRUE) %>% 
#   arrange(desc(PlotNumber))
# SU2 <- SU %>% dplyr::select(PlotNumber, SiteUnit)
# 
# 
# #%>% mutate(PlotNumber = paste0("'",PlotNumber))
# #fwrite(SU, "./clean_tabs/CoastGuide_All_BGC_Forest_SU.csv")
# ###remove phases and seral units
# SU_siteseries <- SU %>% 
#   filter(!str_detect(SiteUnit, '[$]'))%>% filter(!str_detect(SiteUnit, "/W|/R|/G|/F|/A|/V")) %>% filter(!str_detect(SiteUnit, fixed("(G")))
# phases = c("a", "b", "c")
# SU_siteseries$SiteUnit <- str_replace(SU_siteseries$SiteUnit, "a$|b$", "")
# SU_siteseries2 <- SU_siteseries %>% dplyr::select(-bgc)
# fwrite(SU_siteseries, "./clean_tabs/SIFR_Guide_All_BGC_Forest_SS.csv")
# ###write combined SU to database
# master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; 
# DBQ=D:/BC_Correlation2_Vpro_2023/SIFR_Hierarchy.accdb;")
# dbWriteTable(master_su, "All_SIFR_Forest_SU",  SU_siteseries2, overwrite = TRUE, batch_rows = 1)
# dbDisconnect(master_su)
# 
# SU <- SU_siteseries
# ###Check that there is plot data for all
# missingplots <- anti_join(SU, plot.env)
# plots <- right_join(SU, plot.env)
# ### Check for duplicate plot usage
# usemorethanone <- SU %>% dplyr::group_by(PlotNumber) %>% dplyr::mutate(dups = n()) %>% filter(dups >1) %>% ungroup() %>% arrange(PlotNumber)
# fwrite(usemorethanone, "./outputs/SIFR_Plots_Used_Morethan_Once.csv")
# ###Summaries by BGC and by SS
# plots_bgc <- SU %>% group_by(bgc) %>% dplyr::summarise(plots = n())
# plots_SS <- SU %>% group_by(SiteUnit) %>% dplyr::summarise(plots = n())
# SS_count <- length(unique(SU$SiteUnit))
# siteunits_toofew <- plots_SS %>% filter(plots <5) #%>% dplyr::select(- PlotNumber) %>% distinct
# 
# fwrite(siteunits_toofew, "./outputs/SIFR_SiteUnits_w_lessthan5plots.csv")
```






Site unit statistics
```{r}
###some evaluative stats for  units
# temp2 <- temp[,c(1,5)]
# VegUnits <- unique(temp2)# number of plots per unit
# numspp <- ddply(temp,~Group,summarise,sppcount=length(unique(Species))) # number of species per unit
# VegUnits2 <-merge(VegUnits, numspp, by= "Group" )
# numspp2<-ddply(temp[temp$Constancy > 20,],~Group,summarise,sppcount=length(unique(Species))) # number of non-accidental species per unit
# VegUnits2 <-merge(VegUnits2, numspp2, by= "Group" )
# constspp <-ddply(temp[temp$Constancy > 59,],~Group,summarise,sppcount=length(unique(Species))) # number of constant species per unit
# VegUnits2 <-merge(VegUnits2, constspp, by= "Group" )
# colnames(VegUnits2)[3:5] <- c("TotalSpp","NonrareSpp","ConstSpp")
# VegUnits2$Poor <- ifelse(VegUnits2$NoPlots < 5 | VegUnits2$ConstSpp<5, "Poor", "OK")
# write.csv(VegUnits2, "GrasslandsUnitSpeciesTotals.csv")
```

### Compare BGCv12 and Hierarchy SU
```{r check for inclusion of classification units}
# Hierarchy <- fread("./inputs/AllForestHier.csv", stringsAsFactors = FALSE)
# colnames(Hierarchy )[1:12]=c("PlotNumber", "Region", "Class", "Order", "SubOrder", "Alliance", "SubAlliance", "Association", "SubAssociation", "Facies", "Working", "SiteUnit")
# #Create lowest working hierarchical units
# Hierarchy$SubAssociation <- ifelse(Hierarchy$SubAssociation == "",Hierarchy$Association,Hierarchy$SubAssociation) ##if SubAssoc blank, fill with Association
# Hierarchy$SubOrder <- ifelse(Hierarchy$SubOrder == "",Hierarchy$Order,Hierarchy$SubOrder)
# Hierarchy$SubAlliance <- ifelse(Hierarchy$SubAlliance == "",Hierarchy$Alliance,Hierarchy$SubAlliance)
# 
# write.csv(Hierarchy, "AllForestHier_filled.csv")
# 
# BGCv12inHier <- left_join(BGCv12_nplots, Hierarchy)

```
`



