---
title: "MachineLearningHigherOrders"
author: "Will MacKenzie & Kiri Daust"
date: "05/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

require(tidyverse)
require(tidymodels)
require(data.table)
require(data.tree)
require(DataExplorer)
require(C50)
require(indicspecies)
require(doParallel)
require(vip)
source("./_functions/_TabletoTree.R")
source("./_functions/_TreetoTable.R")
cloud_dir <- "F:/OneDrive - Personal/OneDrive/BEC_Classification_Paper/"
#cloud_dir <- "F:/OneDrive - Government of BC/CCISSv12/"
```

Steps to build classification hierarchy

1. A machine learning model of high-level hierarchical BEC Forest units
a. Create a ML model of Forested SubOrders using BECv11 site series that meet minimum plot requirements. (going to Orders only causes issues where Pl is a dominant secondary species - many provincial units get placed into a Pinucon Order when applied only to the order level even though they have a subdominant tree species which places them in a different order)
b. Use only tree species for first round of model build for the forested Orders.
c. Predict membership of new BECv13 units and assign to orders then rebuild ML model.
d. Review any mis predicted site series for reassignment.
e. Compare similarity of tree composition within units (noise clustering) to find outliers that may represent new Orders (test by leaving out known units)
f. Create vegetation summaries of Orders and compare vegetation (all species now). Identify climatic indicator species groups.
g. Create climatic and site attributes summary of Orders 
h. Create maps of plot locations for each Order
i. Hierarchy tree graphic

2. Build Alliances. Focus should be on identifying species groups that reflect different site conditions.

a. Use some of the techniques of indicspecies package to create Alliances based on indicator group creation and analysis.
b. Try to build a machine learning model of Alliances
c. Vegetation and Environment summary of Alliances

3. Do pair-wise analysis of site series within Orders/Alliances using high constancy species to analyze site series and create Associations/SubAssociations.
a. Check for similarity between Associations of All orders
b. Build relationship graphic


4. Document hierarchy

5. Build model to assign new units and unassigned plots
Challenge here is to make a constant list of species between model and new units.
a. Predict membership of new site series (draft BECv13). Noise Clustering to test for novel units that do not fit any existing. Use machine learning to hierarchy
b. Predict Order membership of BECMaster plots that are unassigned to existing site series and add to an Order_unplaced site unit under each Order.

6. Build for non-forested units. Classes based on major site level differences rather than climate and using major species groups (e.g. Hydrophytic Carex spp )
May wish to assign a temporary notree pseudo species to all units that have no trees to help separate forested hierarchy from non-forested hierarchy (Hurdle model)

```{r import data}
### Vegetation From BECMaster cleaning script Long form Plot|Species|Cover (across all layers)
### or From Vpro export
#BGCZone <- fread(paste0(cloud_dir,"All_BGCs_Info_v12_2.csv"), data.table=FALSE)
vegDat2 <- fread("../BECMaster_Cleaning/clean_data/BECMaster_VegR_clean.csv", data.table = FALSE)
taxon.all <- fread("D:/CommonTables/SpeciesMaster/SpeciesMaster01Dec2020.csv", header = T, stringsAsFactors = F, strip.white = T) %>% filter(Codetype == "U")
###SU table
#SUTab <- fread("./inputs/ALLBECSU_2021_SU.csv")
SUTab <- fread("./clean_tabs/BECv13Analysis_Forest_17Sept2021_SU.csv")
SUTab$SiteUnit <-  trimws(SUTab$SiteUnit)
SS.names <- unique(SUTab$SiteUnit)

```


##roll up into site series summary data
```{r summarize site series, echo=FALSE}
source("./_functions/_create_su_vegdata.R")

vegsum <- create_su_vegdata(vegDat2, SUTab)

##roll up into site series summary data

source("./_functions/_spp_importance.R")
vegsum <- spp_importance(vegsum) 

### Filter out poor units and low importance species#
vegDat_good <- vegsum %>% dplyr::filter(nPlots > 4) %>% dplyr::filter(spp_importance > 0.05) ## remove rare or very low cover species in SU


Hier.clean <- SUhier$table
 #%>% rename(SiteUnit = SiteUnit))## Select only site series will enough plots
Hier.units <- Hier.clean %>% dplyr::select(SiteUnit, Class, Order, Suborder) %>% distinct()
Hier.data <- left_join(Hier.units, vegDat_good) %>% filter(!is.na(nPlots)) %>% arrange(Species) %>% distinct()
#fwrite(vegSum, './clean_tabs/BECv12SiteUnitSummaryVegData.csv')
```

##Import hierarchy matrix from 0_0_Plot_SU_Hierarchy script with original placements plus all unassigned units added to an unplaced Formation category
```{r import  hierarchy data}
Vpro.hier <- fread("./clean_tabs/BECv13_ForestHierarchy.csv")

###move SU with <4 plots to the UnitsTooFew formation (=2)
#Vpro.hier[Name %in% Units.toofew, Parent:= 2]
#fwrite(Vpro.hier, "./outputs/UpdatedVPROHierarchyv13_updated.csv")

SUhier <- treeToTable(Vpro.hier)
Hier.clean <- SUhier$table

```

```{r reduce summary for analysis, echo=FALSE}
##limit life forms to tree species
trees <- c(1,2)
constcut <- 33 ##remove species less than cutoff
covercut <- 1
treespp <- taxon.all[Lifeform %in% trees, ] %>% dplyr::select(Code)
treespp <- as.vector(treespp$Code)
vegDat_test <- vegsum[Species %in% treespp,]### include only trees for hierarchy build
vegDat_test <- vegDat_test %>% dplyr::filter(MeanCov > covercut) %>% dplyr::filter(Constancy > constcut)


tree.sum <- as.data.table(vegDat_test)#vegDat <- as_tibble(vegDat)
tree.sum$SiteUnit <- as.factor(tree.sum$SiteUnit)


#fwrite(tree.sum, './inputs/SiteUnitConiferSummary.csv')


```


```{r filter and prepare for analysis}

vegSum <- tree.sum
SS_good <- vegSum %>% filter(nPlots >3) %>% filter(Constancy >= 33)  %>% distinct()  #%>% rename(SiteUnit = SiteUnit))## Select only site series will enough plots
Hier.units <- Hier.clean %>% dplyr::select(SiteUnit, Class, Order, Suborder) %>% distinct()
Hier.data <- left_join(Hier.units, SS_good) %>% filter(!is.na(nPlots)) %>% arrange(Species) %>% distinct()
#fwrite(Hier.data, './inputs/SiteUnitForested_w_HierarchyUnits.csv')
```

```{r some hierarchy stats}
classes <- unique(Hier.data$Class)
orders <- unique(Hier.data$Order)
suborders <- unique(Hier.data$Suborder)
suborders
### Choose hierarchical level for analysis

class.dat <-
  Hier.data %>% dplyr::select(SiteUnit, Suborder, Species, MeanCov, Constancy) %>% 
  pivot_wider(id_cols = c(SiteUnit, Suborder),
    names_from = Species,
    values_from = c(MeanCov, Constancy)) %>% 
  mutate(Suborder = ifelse(is.na(Suborder) | Suborder == "", "unplaced", Suborder)) %>% filter(!SiteUnit == "") %>%  mutate_if(is.character, as.factor) %>% 
   replace(is.na(.),0) %>% distinct() %>% droplevels()

#DataExplorer::create_report(class.dat)
```
Data pre-processing includes the following steps:  

```{r prep data, include = TRUE, echo = FALSE}

classID <- class.dat %>% dplyr::select(SiteUnit, Suborder)
#class.dat2 <- class.dat %>% select(-SiteUnit)
BEC_good <- class.dat %>% filter(!Suborder == "unplaced") %>% arrange(SiteUnit)
BEC_good$SiteUnit <- BEC_good$SiteUnit %>% as.factor
BEC_new <- class.dat %>% filter(Suborder == "unplaced") %>% arrange(SiteUnit)

set.seed(123)
BEC_split <- initial_split(BEC_good,strata = Suborder)
BEC_train <- training(BEC_split)
BEC_test  <- testing(BEC_split)

SU_names <- as.data.frame(BEC_good$SiteUnit) %>%  distinct() %>% rowid_to_column('.row') %>% dplyr::rename("SiteUnit" = 2)

BEC_recipe <-
    recipe(Suborder ~ ., data = BEC_train) %>%
     update_role(SiteUnit, new_role = "id variable") %>% 
  step_novel(SiteUnit) %>% 
    prep() 
    summary(BEC_recipe)
    
BEC_fmodel <- rand_forest(mtry = 5, min_n = 2, trees = 501) %>%
  set_mode("classification") %>%
  set_engine("ranger", importance = "impurity") #or "permutations

# note in RF as a tree based model it is not required to scale and normalize covariates and may have negative influence on the model performance  
```


```{r tune cv model of suborders, include = FALSE}
# define model with set parameters from tuning
#set.seed(123)
BEC_split <- initial_split(BEC_good,prop = .9, strata = Suborder)
BEC_train <- training(BEC_split)
BEC_test  <- testing(BEC_split)
BEC_val_set <- validation_split(BEC_train, 
                            strata = Suborder, 
                            prop = 0.70)

BEC_fmodel <- rand_forest(mtry = tune(), min_n = tune(), trees = 101) %>%
  set_mode("classification") %>%
  set_engine("ranger", importance = "impurity") #or "permutations

BEC_cv_workflow <- workflow() %>% 
  add_model(BEC_fmodel) %>% 
  add_recipe(BEC_recipe)

set.seed(345)
rf_res <- 
  BEC_cv_workflow %>% 
  tune_grid(BEC_val_set,
            grid = 25,
            control = control_grid(save_pred = TRUE),
            metrics = metric_set(kap))

autoplot(rf_res)
rf_res %>% 
  show_best(metric = "kap")
rf_best <- rf_res %>% select_best(metric = "kap")
rf_best
mtry1 = as.integer(rf_best$mtry)
min_n1 = as.integer(rf_best$min_n)
```


```{r build  cv model of suborders, include = FALSE}
BEC_fmodel <- rand_forest(mtry = 8, min_n = 3, trees = 1001) %>%
  set_mode("classification") %>%
  set_engine("ranger", importance = "impurity")

BEC_recipe <-
    recipe(Suborder ~ ., data = BEC_train) %>%
     update_role(SiteUnit, new_role = "id variable") %>% 
  step_novel(SiteUnit) %>% 
    prep() 
    summary(BEC_recipe)
    
BEC_cv_workflow <- workflow() %>% 
  add_model(BEC_fmodel) %>% 
  add_recipe(BEC_recipe)
v = 10
reps = 5
  
set.seed(345)
BEC_cvfold <- vfold_cv(BEC_train, 
                         v = v, 
                         repeats = reps,  strata = Suborder)
cv_metrics <- metric_set(accuracy, kap) #roc_auc,, j_index, sens, spec

set.seed(130)
#doParallel::registerDoParallel()
tic()
BEC_cv_fitted <- 
  BEC_cv_workflow %>% 
  fit_resamples(resamples = BEC_cvfold, metrics = cv_metrics, control = control_resamples(save_pred = TRUE))
toc()
gc()

cv_results <- BEC_cv_fitted  %>% collect_metrics(summarize = FALSE)
cv_results_sum <- BEC_cv_fitted  %>% collect_metrics()

```

```{r final cv model and metrics}
# the last model
last_BECmodel <- 
  rand_forest(mtry = 8, min_n = 3, trees = 1001) %>% 
  set_engine("ranger",  importance = "impurity") %>% 
  set_mode("classification")
# last_BEC_recipe <-
#     recipe(Suborder ~ ., data = BEC_split) %>%
#      update_role(SiteUnit, new_role = "id variable") %>% 
#   step_novel(SiteUnit) %>% 
#     prep() 
#     summary(last_BEC_recipe)

last_BEC_workflow <-  BEC_cv_workflow %>% 
  update_model(last_BECmodel) 

last_BEC_fit <-   last_BEC_workflow %>% 
  last_fit(BEC_split, metrics = cv_metrics)

last_BEC_fit  %>% collect_metrics()

final.acc <- last_BEC_fit  %>% 
  collect_metrics()

last_BEC_fit %>% 
  pluck(".workflow", 1) %>%   
  extract_fit_parsnip() %>% 
  vip(num_features = 20)


test.pred <-  collect_predictions(last_BEC_fit) %>%
  mutate(correct = case_when(
    Suborder == .pred_class ~ "Correct",
    TRUE ~ "Incorrect"
  )) %>%
  bind_cols(BEC_test)
test.misID <- test.pred %>% filter(correct == "Incorrect") %>% dplyr::select(SiteUnit, .pred_class, Suborder...4)
test.misID.list <- test.misID$SiteUnit %>% as.character
SS_misID_SU <- SUTab[SUTab$SiteUnit %in% test.misID.list,]
fwrite(SS_misID_SU, "./outputs/WrongSubOrderPredicted_SU.csv")

Hier.orders <- Hier.clean %>% dplyr::select(Order, Suborder) %>% distinct() 
test.order <- test.pred %>% dplyr::select(SiteUnit, .pred_class, Suborder...4) %>% rename(Suborder = Suborder...4) %>% mutate_if(is.factor, as.character)
test.order <- left_join(test.order, Hier.orders, by = "Suborder") %>% rename (Order.hier = Order)
 test.order2 <- left_join(test.order, Hier.orders, by= c(".pred_class" = "Suborder")) %>% rename (Order.pred = Order) %>% mutate(Order.compare = ifelse(Order.hier ==  Order.pred, "Same", "Diff"))%>%
  mutate(Compare = ifelse(.pred_class == Suborder, "Same", "Diff")) %>% filter(Compare == "Diff")

```


```{r final ranger model}
BEC_recipe <-
    recipe(Suborder ~ ., data = BEC_good) %>%
     update_role(SiteUnit, new_role = "id variable") %>% 
  step_novel(SiteUnit) %>% 
    prep() 
    summary(BEC_recipe)

BEC_fmodel <- rand_forest(mtry = 8, min_n = 3, trees = 501) %>%
  set_mode("classification") %>%
  set_engine("ranger", importance = "impurity")

BEC_workflow <- workflow() %>%
  add_model(BEC_fmodel) %>% 
    add_recipe(BEC_recipe,blueprint = hardhat::default_recipe_blueprint(allow_novel_levels = TRUE))
# the last fit
set.seed(345)
BEC_ranger_model <- fit(BEC_workflow, BEC_good)

BEC_ranger_model

save(BEC_ranger_model, file = "./rFmodels/BECv13_Suborders_rFmodel.Rdata")



BEC.pred <- predict(BEC_ranger_model, BEC_good) %>% bind_cols(BEC_good%>% dplyr::select(SiteUnit, Suborder)) 
BEC.acc <- BEC.pred %>%  metrics(truth = Suborder, estimate = .pred_class)
BEC.acc
BEC.pred.list <- BEC.pred$SiteUnit %>% as.character
BEC.missed <- BEC.pred  %>%   mutate(Compare = ifelse(.pred_class == Suborder, "Same", "Diff")) %>% filter(Compare == "Diff")
fwrite(BEC.missed, "./outputs/MisplacedSuborders.csv")
BEC.mis.list <- BEC.missed$SiteUnit %>% as.character
SS_misID_SU <- SUTab[SUTab$SiteUnit %in% BEC.mis.list,]
fwrite(SS_misID_SU, "./outputs/WrongOrderPredicted_SU.csv")


Hier.orders <- Hier.clean %>% dplyr::select(Order, Suborder) %>% distinct()
BEC.order <- BEC.pred %>% dplyr::select(SiteUnit, .pred_class, Suborder)
BEC.order <- left_join(BEC.order, Hier.orders) %>% rename (Order.hier = Order)
BEC.order <- left_join(BEC.order, Hier.orders, by.x = .pred_class, by.y = Suborder) %>% rename (Order.pred = Order) %>% 
  mutate(Order.compare = ifelse(Order.hier == Order.pred, "Same", "Diff")) %>%
  mutate(Compare = ifelse(.pred_class == Suborder, "Same", "Diff")) %>% filter(Compare == "Diff")

```
## Collect accuracy metrics and predictions and improperly predicted site units
```{r collect metrics and predictions, echo = TRUE }
# collect accuracy metrics  

BEC.pred <- BEC_cv_fitted %>% collect_predictions()# collect predictions
#BEC.predx <- left_join( BEC.pred, SU_names)
### accuracy by suborder
BEC.pred_acc <- BEC.pred %>% 
  group_by(Suborder) %>% 
  accuracy(Suborder, .pred_class)
#Summarize prediction
BEC.pred2 <- BEC_cv_fitted %>% collect_predictions(summarize = TRUE)# collect predictions
BEC.pred3 <- left_join(BEC.pred2, SU_names)

## Identify misplaced site units
MisID <- BEC.pred3 %>% dplyr::select(SiteUnit, Suborder, .pred_class) %>% mutate(compare = if_else(Suborder == .pred_class, "Same", "Diff")) %>% filter(compare == "Diff")# %>% rename("SiteUnit" = SiteUnit)
fwrite(MisID, "./outputs/Misplaced Site Series in SubOrders.csv")

```

```{r build  model of suborders, include = FALSE}

# 2: set up cross validation for parameter tuning data sets # note vc is default to 10 fold
BEC_workflow <- workflow() %>%
  add_model(BEC_fmodel) %>% 
    add_recipe(BEC_recipe,blueprint = hardhat::default_recipe_blueprint(allow_novel_levels = TRUE))

BEC_ranger_model <- fit(BEC_workflow, BEC_train)
BEC_ranger_model$fit

save(BEC_ranger_model, file = "./rFmodels/BECv13_Suborders_rFmodel.Rdata")

BEC.pred <- predict(BEC_ranger_model, BEC_test) %>% bind_cols(BEC_test %>% dplyr::select(SiteUnit, Suborder)) 
BEC.acc <- BEC.pred %>%  metrics(truth = Suborder, estimate = .pred_class)
BEC.acc
BEC.pred.list <- BEC.pred$SiteUnit %>% as.character
BEC.missed <- BEC.pred  %>%   mutate(Compare = ifelse(.pred_class == Suborder, "Same", "Diff")) %>% filter(Compare == "Diff")
fwrite(BEC.missed, "./outputs/MidplacedSuborders.csv")
BEC.mis.list <- BEC.missed$SiteUnit %>% as.character
SS_misID_SU <- SUTab[SUTab$SiteUnit %in% BEC.mis.list,]
fwrite(SS_misID_SU, "./outputs/WrongOrderPredicted_SU.csv")
```