---
title: "NoiseClustering to Add Plots to Classification"
author: "William H MacKenzie"
date: "13/01/2023"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

Based on the vegclust package and procedures described by Caceres and
Wiser the basic method is plotmembership \<- vegclust::vegclass(vegclust
object of defined groups, new plot data)

1\. We'll be using existing Associations as predefined groups and this
will be using species importance (sqrt cover \* constancy)

2\. Plot data will have different species so we'll need the function
vegclust::conformveg to harmonize the species lists. May need to do some
cleanup of species to remove low cover and rare species?

3\. Will need experiment to deal with settings of noise level and
defuzzify

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(vegclust)
require(data.table)
require(tidyverse)
require(DBI)
require(labdsv)
source("./_functions/_lump_species.R")
```

read in veg_anal and SU at the Assoc level convert to vegclust object

```{r import data}
#testunit <- c("CDF mm   /101", "CWH vm 2 /103", "MH  mm 1 /112", "CWH xm 1 /101")
coast.ss <- fread ("./clean_tabs/CoastGuide_All_BGC_Forest_SS.csv") %>% distinct(SiteUnit)

veg.dat <- readRDS("./clean_data/Analysis_BECMaster_Veg.rds")

load("./clean_data/SS_sum_analysis.RData")

veglump <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=C:/BC_Correlation2_Vpro_2023/CoastGuide_Spp_Lump.accdb;")
LUMP <- dbReadTable(veglump, "CoastGuide2023_Lump")
dbDisconnect(veglump) 

```

```{r prep data and units}
##remove units with few plots
plots_SS <- coast.ss %>% group_by(SiteUnit) %>% dplyr::summarise(plots = n())
SS_count <- length(unique(coast.ss$SiteUnit))
siteunits_toofew <- plots_SS %>% filter(plots <5) 

##set this number to the max number of units
testunit <- sample_n(coast.ss, 100, replace = FALSE)
## remove non-vasculars
vegdata <- veg.dat %>% filter(Lifeform <9 | Lifeform == 12)
## lump to species and lump file
vegdata <- lump_species(vegdata, LUMP)

assocs <- fread("./outputs/working_SU.csv") %>% arrange(PlotNumber) %>% filter(SiteUnit %in% testunit$SiteUnit) %>% filter(!SiteUnit %in% siteunits_toofew)

plots <- unique(assocs$PlotNumber)
assocs2 <- assocs %>% dplyr::select(-SiteUnit) %>%  column_to_rownames("PlotNumber") %>% t %>% as.vector
##reduce data to only those in assoc list
vegdata2 <- vegdata %>% filter(PlotNumber %in% plots) %>% arrange(PlotNumber)
### reduce list to constant species or high in indiv plots and convet to matrix
const_spp <- unique(veg_anal$Species)

vegdata2 <- vegdata2 %>% mutate(spp_importance = totalcov^0.5) %>% dplyr::select(PlotNumber, Species, spp_importance) %>% filter(spp_importance >2 | Species %in% const_spp ) %>% matrify()
##produce dissimiarity for all plots
all_dissim <- proxy::dist(vegdata2, method="ejaccard", diag=TRUE, upper = FALSE)

##create vegclust object
# assoc_data <- as.vegclust(vegdata2, assocs2, method="KM", dnoise = 0.8)
# xx <- as.data.frame(assoc_data$memb)
# ## Look at variation within cluster - might not be useful here?
# veg.var <- clustvar(assoc_data) %>% as.data.frame

```

################ ###Test using split of guide classification into

training and testing data

```{r split data set and test classification}

veg.test <- sample_n(vegdata2,50)
test.plots <- row.names(veg.test)
assocs.test <- assocs %>% dplyr::filter(PlotNumber %in% test.plots) %>%  arrange(PlotNumber)
veg.test.diss <-  as.data.frame(as.matrix(all_dissim)) %>% select(-one_of(test.plots)) 
veg.test.diss <- veg.test.diss %>% filter(row.names(veg.test.diss) %in% test.plots)  

veg.model <- vegdata2 %>% dplyr::filter(!rownames(vegdata2) %in% test.plots) %>%  arrange(row.names(.))
assocs.model <-assocs %>% dplyr::filter(!PlotNumber %in% test.plots) %>%  arrange(PlotNumber)
assocs.model <- assocs.model %>% dplyr::select(-SiteUnit) %>%  column_to_rownames("PlotNumber") %>% t %>% as.vector
#assoc.model <- as.vegclust(veg.model, assocs.model, method="HNC", dnoise = 10)
assoc.model.diss <- as.vegclust(proxy::dist(veg.model, method="ejaccard", diag=TRUE, upper = FALSE), assocs.model, method="HNC", dnoise = 0.4)

vegComb <- vegclass(assoc.model.diss, veg.test.diss) ##classify vegNew
vegComb.memb <- vegComb[["memb"]]
newGroup <- dematrify(vegComb.memb) ##extract classification
newGroup <- newGroup[,1:2]
colnames(newGroup) <- c("PlotNumber","Class.pred")
newGroup <- newGroup %>% arrange(PlotNumber)
newGroup <- cbind(newGroup, assocs.test)
####merge actual classification for comparison
#colnames(newGroup)[3] <- "Actual"

```

#####Place new plots in predefined classes

Probably better to do this in smaller chunks of a hierarchy rather than
all at once

```{r classify new plots}
##conform data

cf = conformveg(wetland.31, wetland.10)


vegComb <- vegclass(vegOld.clust, vegNew) ##classify vegNew
vegComb.memb <- vegComb[["memb"]]
newGroup <- dematrify(vegComb.memb) ##extract classification
newGroup <- newGroup[,1:2]
colnames(newGroup) <- c("SiteSeries","Class")
newGroup <- cbind(newGroup, actualClass$Class)####merge actual classification for comparison
colnames(newGroup)[3] <- "Actual"


vegDat.chord <- decostand(SUsumMatrix[-1], "normalize") ##standardise data
vegMat <- vegDat.chord
vegMat <- cbind(SUsumMatrix[,1],vegDat.chord)

####create training and testing data sets
vegNew <- vegMat[rownames(vegMat) %in% sample(rownames(vegMat), 100, replace = FALSE),] 
vegOld <- vegMat[!(rownames(vegMat) %in% rownames(vegNew)),]
#actualClass <- vegNew[,1:2] ###siteseries to grouping lookup
#rownames(vegNew)  <- vegNew[1] ###set rownames to siteseries
k <- ncol(vegOld)
n <- nrow(vegOld)
###grouping of training set
grouping <- vegOld[1]
vegOld <- vegOld[,-(1)]
vegNew <- vegNew[,-(1)]
k <- ncol(vegOld)
n <- nrow(vegOld)

grouping <- as.vector.factor(grouping)
vegOld.clust <- as.vegclust(vegOld, grouping)###create noise clustering with grouping as classes
###Kmeans Clustering
vegOld.kmclst <- vegclust(x = vegOld[,-1], mobileCenters=5, method = "KM", nstart=20)###create noise clustering with grouping as classes
t(vegOld.kmclst$memb)
###NC Clustering
vegOld.kmclst <- vegclust(x = vegOld[,-1], mobileCenters=5, method = "NC", m=1.2, dnoise=0.8, nstart=20)###create noise clustering with grouping as classes
round(t(vegOld.kmclst$memb), dig=2)
groups = defuzzify(vegOld.kmclst, method="cut", alpha=0.8)$cluster
table(groups)


vegComb <- vegclass(vegOld.clust, vegNew) ##classify vegNew
vegComb.memb <- vegComb[["memb"]]
newGroup <- dematrify(vegComb.memb) ##extract classification
newGroup <- newGroup[,1:2]
colnames(newGroup) <- c("SiteSeries","Class")
newGroup <- cbind(newGroup, actualClass$Class)####merge actual classification for comparison
colnames(newGroup)[3] <- "Actual"

###MDS for visualisation
MDS <- metaMDS(vegNew, distance = "bray", k = 2, trymax = 200)
MDS.df <- as.data.frame(scores(MDS, display = "sites"))###extract mds scores
MDS.df$SiteSeries <- rownames(MDS.df)
MDS.df <- merge(MDS.df,newGroup, by = "SiteSeries") ##merge predicted and actual classification
MDS.df <- MDS.df[,-1]

colnames(MDS.df)[3:4] <- c("Actual","Predict")

ggplot(MDS.df)+
  geom_point(mapping = aes(x = NMDS1, y = NMDS2, colour = Predict), size = 2.5, shape = 17)+
  coord_fixed()+
  theme_bw()+
  theme(legend.position = "none")
```
