---
title: "Pairwise_Diagnostic"
author: "WH MacKenzie"
date: "08/01/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(reshape2)
require(plyr)
require(dplyr)
require(tidyr)
require(ggplot2)
require(magrittr)
require(foreach)
require(tcltk)
require(openxlsx)
require(doParallel)
require(doBy)
require(doParallel)
require(DBI)
require(data.table)
require(data.tree)
require(labdsv)
require(tidyverse)
library(cluster)
require(ape)
require(factoextra)
require(tictoc)
require(ggthemes)
source("./_functions/_lump_species.R")
source("./_functions/_create_su_vegdata.R")
source("./_functions/_create_analysis_vegsum.R")
source("./_functions/_TabletoTree.R")
source("./_functions/_TreetoTable.R")
source("./_functions/_add_vars.R")
source("./_functions/_do_pairwise.R")
source("./_functions/_create_diagnostic_veg.R")
source("./_functions/_return_similar_pairs.R")
```

## Read in data and sorts and spp lumping

```{r load data}
veg.dat <- readRDS("./clean_data/Analysis_BECMaster_Veg.rds") ###named veg.dat


# master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
# DBQ=D:/GitHub/BEC_Classification_Analysis/Vpro/Coast_Skunkcabbage.accdb;")
# su <- dbReadTable(master_su, "AllBGC_Skunkcabbage_SU")
# dbDisconnect(master_su)

# master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
# DBQ=D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb;")
# su <- dbReadTable(master_su, "All_BC_BGCs_SU")
# dbDisconnect(master_su)
# 
master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
DBQ=D:/BC_Correlation2_Vpro_2023/CoastGuide_Hierarchy.accdb;")
#su <- dbReadTable(master_su, "All_Coast_Forest_2024v4_SU")
hier <- dbReadTable(master_su, "CoastForest_v2024_2_Hierarchy")
dbDisconnect(master_su)

require(DBI) #loads odbc as well
correlation <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; 
DBQ=D:/BC_Correlation2_Vpro_2023/CoastGuide_Forested.accdb;")
all_su <- lapply(setNames(nm = (dbListTables(correlation)%>% str_subset("_SU"))), dbReadTable, conn = correlation)
dbDisconnect(correlation)
SU <- do.call(rbind.data.frame, all_su)
SU <- SU %>% mutate(bgc = substr(SiteUnit,1,9)) %>% drop_na() %>% distinct(PlotNumber, .keep_all = TRUE) %>% 
  arrange(desc(PlotNumber))
ss.unique <- SU %>% select(SiteUnit) %>% distinct
su <- SU %>% dplyr::select(PlotNumber, SiteUnit)
# master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
# DBQ=D:/BC_Correlation2_Vpro_2023/CoastalGuide_NonForested.accdb;")
# su <- dbReadTable(master_su, "Coast_Skunkcabbage_final_SU")
# dbDisconnect(master_su)

su$SiteUnit.orig <- su$SiteUnit
su <- su %>% mutate(SiteUnit = gsub("/", "_", su$SiteUnit)) 
su <- su %>% mutate(SiteUnit = gsub(" ", "", su$SiteUnit, fixed = TRUE)) 
#su.lookup <- su
#su <- su %>% select(-SiteUnit.orig)
su <- su %>% 
  filter(!str_detect(SiteUnit, '[$]')) %>% filter(!str_detect(SiteUnit, "^x"))

#  Fh <- fread("D:/BC_Correlation2_Vpro_2023/Highbench_siteseries.csv")
# Fh <- Fh %>% mutate(highbench_ss = str_replace_all(string=highbench_ss, pattern=" ", repl="")) %>% mutate(highbench_ss = str_replace_all(string=highbench_ss, pattern="/", repl="_"))
# su <- su %>% mutate(SiteUnit = str_replace_all(string=SiteUnit, pattern=" ", repl="")) %>% mutate(SiteUnit = str_replace_all(string=SiteUnit, pattern="/", repl="_"))  %>% filter(SiteUnit %in% Fh$highbench_ss)
# 
# #SU_Fh <- fread("./clean_tabs/CoastGuide_Highbench_SS.csv")
# su <- su %>% filter(SiteUnit %in% SU_Fh$SiteUnit)
sppmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=F:/OneDrive - Personal/OneDrive/BCSpeciesList/SpeciesTaxonomyMaster.accdb;")
# sppmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=C:/Users/whmacken/OneDrive/BCSpeciesList/SpeciesTaxonomyMaster.accdb;")
taxon.all  <- dbReadTable(sppmaster, "USysAllSpecs")
dbDisconnect(sppmaster)
taxon.lifeform <- taxon.all %>% filter(Codetype == "U" | Codetype == "X" | Codetype == "D") %>% dplyr::select(Code, ScientificName, EnglishName, Lifeform) %>% distinct
veg.dat2 <- veg.dat

 trees = c(1, 2)
 #veg.dat2 <- veg.dat2 %>% filter(Lifeform %in% trees)

#  veglump <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BC_Correlation2_Vpro_2023/Correlation2_Spp_Lump.accdb;")
# lump <- dbReadTable(veglump, "CorrelationII_Lump")
# dbDisconnect(veglump)

veglump <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BC_Correlation2_Vpro_2023/CoastGuide_Spp_lump.accdb;")
lump <- dbReadTable(veglump, "CoastGuide2023_Lump")
dbDisconnect(veglump)

veg.dat2 <- lump_species(vegdata = veg.dat2, lump, use.subtaxa = FALSE)
yy <- c("CDFmm_102.2")
  xx <- veg.dat2 %>% left_join(su) %>% filter(SiteUnit %in% yy) #%>% select(Species, Lifeform) %>% distinct


```

## Evaluate units for units for low diagnostic potential

```{r settings, echo=FALSE}
# 
# key.site.indicators <- c("LYSIAME", "OPLOHOR", "ATHYFIL", "TSUGMER", "THUJPLI", "SPHAGNUM", "CLADONIA", "CLADINA", "RACOMITR", "MNIUM")
# ksi.value = 1
# reduced.lifeform = c(9, 10, 11, 12)
# vegsum <- create_diagnostic_veg(veg.dat2, su, minimportance = 0.5, covadj = .33, 
#                                 minconstancy = 55, noiseconstancy = 20,
#                                 minplots = 5,  use.ksi = TRUE, ksi = key.site.indicators, ksi.value = 1,
#                                 reduce.lifeform = TRUE, reduced.lifeform = reduced.lifeform, reduction = 0.1)
# yy <- c("CWHws1_103")
# zz <- vegsum %>% filter(SiteUnit %in% yy)
# units.low.diagnostics <- vegsum %>% select(SiteUnit, unit.diag.sum) %>% distinct %>%  filter(unit.diag.sum < 40) 
```

# If working with higher units apply this chunk to roll up into hierarchy units 
```{r}
# source("./_functions/_TabletoTree.R")
# source("./_functions/_TreetoTable.R")
# hier2 <- treeToTable(hier)
# hier2 <- hier2[[1]]
# unit.filter = "Class conifer"
# hier3 <- hier2 %>% filter(Class %in% unit.filter) 
# hier3 <- hier3 %>% mutate(SiteUnit = gsub("/", "_", hier3$SiteUnit)) 
# hier3 <- hier3 %>% mutate(SiteUnit = gsub(" ", "", hier3$SiteUnit, fixed = TRUE)) 
# su <- su %>% mutate(SiteUnit = gsub(" ", "", su$SiteUnit, fixed = TRUE)) 
# su2 <- left_join(hier3, su, by = "SiteUnit") %>% select(PlotNumber, Assoc) %>% dplyr::rename("SiteUnit" = 2)
```
Run pairwise diagnostic analysis on SU units or loaded hierarchical units
Returns a table of units with high similarity

```{r pair.wise}

#ksi <- c("TSUGMER", "LYSIAME", "OPLOHOR", "ATHYFIL")
#"TSUGMER", "PICEENE", "TSUGHET", "THUJPLI", 
# ksi <- c("SPHAGNUM", "CLADONIA", "CLADINA")
#key.site.indicators <- c("SPHAGNUM", "CLADONIA", "CLADINA", "RACOMITR", "MNIUM")
 key.site.indicators <- c("LYSIAME", "OPLOHOR", "ATHYFIL", "RUBUSPE")
reduced.exceptions <- c("SPHAGNUM", "CLADONIA", "CLADINA", "RACOMITR")
#reduced.lifeforms = c(1,2)
reduced.lifeforms = c(9,10,11)
tic()
###select units to run
su2 <- su
Fh <- fread("D:/BC_Correlation2_Vpro_2023/Highbench_siteseries.csv")
Fh <- Fh %>% mutate(highbench_ss = str_replace_all(string=highbench_ss, pattern=" ", repl="")) %>% mutate(highbench_ss = str_replace_all(string=highbench_ss, pattern="/", repl="_")) %>% 
  filter(!upland %in% 'x')
su2 <- su2 %>% filter(SiteUnit %in% Fh$highbench_ss)
#fwrite(SU_Fh, "./clean_tabs/CoastGuide_Highbench_SS.csv")
# su2 <- su2 %>% filter(!grepl('CWHvh3|CWHwh1|CWHwh2|CWHvh3|ICHvc|MHwh', SiteUnit))###BGC specific|CWHms
# su2 <- su2 %>%filter(!grepl('101.2|101a.2|101b.2|101b|low|-S|add', SiteUnit)) %>%  filter(grepl('01', SiteUnit)) ###zonal specific
# 
# su2 <- su2 %>%filter(!grepl('EBM|TEM|T38', PlotNumber)) ##remove low quality projects by plotnumber series

vegsum.pairs <- do_pairwise(veg.dat2, su = su2, minimportance = 0.1, minconstancy = 55, 
                            noiseconstancy = 10,
                            minplots = 5, 
                            covadj = .33, domcov = 10,
                            use.ksi = TRUE, ksi = key.site.indicators, ksi.value = 1, 
                            reduce.lifeform = TRUE, reduced.lifeforms = reduced.lifeforms, reduction = .1,
                            reduced.exceptions = reduced.exceptions)
toc()
vegsum <- create_su_vegdata(veg.dat2, su2) %>% left_join(taxon.lifeform, by = c("Species" = "Code"))

fwrite(vegsum, "./outputs/example_vegsum2.csv")
fwrite(su2, "./outputs/example_siteunits.csv")
vegsum.pairs <- vegsum.pairs %>%  filter(!is.na(Unit1), !is.na(Unit2)) 

units.low.diagnostics <- vegsum.pairs %>% select(Unit1, unit.diag.sum.x) %>% distinct %>%
  filter(unit.diag.sum.x <40 & unit.diag.sum.x>0)
yy <- c("CWHvh1_111","CWHvh2_112" )
zz <- vegsum.pairs %>% filter(Unit1 %in% yy, Unit2 %in% yy) %>% select(Species, Unit1, Unit2, Constancy.x, MeanCov.x, Constancy.y, MeanCov.y, diagnostic.potential.x, diagnostic.potential.y, d.type.x, dd.type.x, diag.points.x, diag.points.y, d.type.y, dd.type.y, diag.points.x, diag.points.y,sum.shared.diag, diag.tot, diag.ratio)
## Sometimes units get placed in odd places in the cluster analsysis because very similar similarity
## Check here. Probably indicates site unit membership to be fixed.
#su.similar <- return_similar_pairs(vegsum.pairs, similarity = .8, neighbours = 3) 
su.phases <- vegsum.pairs %>% select(Unit1, Unit2, diag.ratio) %>% filter(endsWith(Unit1, c("a","b")), endsWith(Unit2, c( "b"))) %>% distinct()
su.phases$bgc1 <-   stringr::word(su.phases$Unit1,1,sep = "\\_") 
su.phases$bgc2 <-   stringr::word(su.phases$Unit2,1,sep = "\\_") 
su.phases <- su.phases %>%  filter(bgc1 == bgc2)

su.similar <- vegsum.pairs %>% select(Unit1, Unit2, diag.ratio) %>% dplyr::filter(diag.ratio > .1) %>% filter(!Unit1 == Unit2)  %>%  distinct
ss.similar <- su.similar 
ss.similar$bgc1 <-   stringr::word(ss.similar$Unit1,1,sep = "\\_") 
ss.similar$bgc2 <-   stringr::word(ss.similar$Unit2,1,sep = "\\_") 
hg.units <- c("CWHwh1", "CWHwh2", "CWHvh3")
ss.similar <- ss.similar %>%  filter(bgc1 == bgc2) %>% filter(!bgc1 %in% hg.units)
```


```{r cluster analysis}
set.seed(1279)
dis.matrix <- as.data.frame(vegsum.pairs) %>% mutate(diss = 1-diag.ratio) %>% mutate(Unit1 = as.character(Unit1)) %>%  select(Unit1,Unit2, diss) %>% distinct %>%
  filter(!is.na(Unit1), !is.na(Unit2)) %>% arrange(Unit1)
## go from one way to two way matrix for clustering
dis.matrix2 <- dis.matrix %>% dplyr::rename(Unit1 = Unit2, Unit2 = Unit1) %>% select(Unit1, Unit2, diss)%>% arrange(Unit1)
dis.matrix <- rbind(dis.matrix, dis.matrix2) %>% arrange(Unit1, Unit2) %>% pivot_wider(id_cols = Unit1, names_from = Unit2, values_from = diss, names_sort = TRUE) %>%  column_to_rownames("Unit1") %>% mutate_all(~replace(., is.na(.), 1))%>% as.matrix

ss_clst = agnes(dis.matrix , diss = TRUE, stand = TRUE,
      method = "weighted")

dendro_hc <- as.hclust(ss_clst)
dend.dend <- as.dendrogram(dendro_hc)
### returns to cophonetic value - considered good if over 0.75
dend.dis <- as.dist(dis.matrix) 
dend.co <- cophenetic(dendro_hc)
cophonentic <- cor(dend.dis, dend.co)
cophonentic ## 0.891 this value shows how well the clusters align with the data structure. >0.7 is considered good

dendro <- fviz_dend(dendro_hc, cex = .5, lwd = .5, h = .25,
        rect = TRUE,
         type = "rectangle",
          k_colors = "jco",
      rect_border = "black",
          rect_fill = TRUE,
          lower_rect = 0,
          horiz = TRUE,
       ggtheme = theme_calc(),labels=T)
dendro
  ggsave("./graphics/coastal_trees_cluster.pdf", dendro, width=12, height=6)
```


```{r dend with colour bars upper units}
group.level <- "Order"

dend.dend <- as.dendrogram(dendro_hc) %>% set("labels_cex",.4)
unit.lab <- as.data.frame(dendro_hc$labels) %>% dplyr::rename(Assoc = 1) %>% left_join(hier2, by = "Assoc") %>% distinct(Assoc, .keep_all = TRUE) %>% dplyr::select(group.level) 
unit.cats <- factor(unit.lab[,1])
n_units <- length(unique(unit.cats))
#cols_4 <- colorspace::rainbow_hcl(n_units, c = 70, l  = 50)
cols_4 <- colorspace::divergingx_hcl("RdYlBu", n=n_units)
unit.col <- cols_4[unit.cats]
par(mar = c(1,1,1,14)); plot(dend.dend, horiz = TRUE); colored_bars(unit.col, dend.dend, rowLabels = group.level, horiz = TRUE, y_shift = .1);legend("topleft", legend = levels(unit.cats), fill = cols_4, cex = .4)
```


```{r examine pairs}
xx <- vegsum.pairs  %>% select(Unit1, Unit2, Species, diagnostic.potential.x, diagnostic.potential.y, diag.potential.tot, shared.diag, diag.points.x, diag.points.y,  sum.shared.diag, diag.points.x, diag.points.y, diag.tot, diag.ratio) %>% filter(diag.ratio>.7)# %>% filter(Unit1 %in% c("MHmm2_101"), Unit2 %in% c("MHmm2_110"))

xx <- vegsum.pairs %>% filter(Unit1 == "IDFww1_101.1" , Unit2 == "CWHds2_101a") %>% select(Unit1, Unit2, Species, diagnostic.potential.x, diagnostic.potential.y, diag.potential.tot, shared.diag, diag.points.x, diag.points.y,  sum.shared.diag, diag.points.x, diag.points.y, diag.tot, diag.ratio)
```


```{r cluster analysis}
working <- cutree(h = .05, dendro_hc) %>% data.frame %>% rownames_to_column("SiteUnit") %>% rename(Working = ".") %>% mutate(Working = paste0("working-", Working))
facies <- cutree(h = .1, dendro_hc) %>% data.frame %>% rownames_to_column("SiteUnit") %>% rename(Facies = ".") %>% mutate(Facies = paste0("facies-", Facies))
subass <- cutree(h = .12, dendro_hc) %>% data.frame %>% rownames_to_column("SiteUnit") %>% rename(Subass = ".") %>% mutate(Subass = paste0("subass-", Subass))
assoc <- cutree(h = .2, dendro_hc) %>% data.frame %>% rownames_to_column("SiteUnit") %>% rename(Assoc = ".") %>% mutate(Assoc = paste0("assoc-", Assoc))
suball <- cutree(h = .3, dendro_hc) %>% data.frame %>% rownames_to_column("SiteUnit") %>% rename(Suball = ".") %>% mutate(Suball = paste0("suball-", Suball))
alliance <- cutree(h = .40, dendro_hc) %>% data.frame %>% rownames_to_column("SiteUnit") %>% rename(Alliance = ".")%>% mutate(Alliance = paste0("all-", Alliance))
suborder <- cutree(h = .66, dendro_hc) %>% data.frame %>% rownames_to_column("SiteUnit") %>% rename(Suborder = ".")%>% mutate(Suborder = paste0("subord-", Suborder))
order <- cutree(h = .9, dendro_hc) %>% data.frame %>% rownames_to_column("SiteUnit") %>% rename(Order = ".")%>% mutate(Order = paste0("ord-", Order))
class <- cutree(h = .95, dendro_hc) %>% data.frame %>% rownames_to_column("SiteUnit") %>% rename(Class = ".")%>% mutate(Class = paste0("class-", Class))
clst.units <- left_join(class, order, by = "SiteUnit") %>%
                    left_join(suborder, by = "SiteUnit") %>%
                    left_join(alliance, by = "SiteUnit") %>%
                    left_join(suball, by = "SiteUnit") %>%                  
                    left_join(assoc, by = "SiteUnit") %>%
                    left_join(subass, by = "SiteUnit") %>%
                    left_join(facies, by = "SiteUnit") %>%  
                    left_join(working, by = "SiteUnit") %>%
                    arrange(SiteUnit) %>% rowid_to_column("ID") %>%
                    mutate(Formation = "") %>% #, Class = "forest", Suball = "", Subass = "", Faces = "")%>%
                    #mutate(Suborder = "", Alliance = "forest", Assoc = "", Working = "") %>% 
                    left_join(su.lookup) %>% select(-SiteUnit) %>%  rename(SiteUnit = SiteUnit.orig)  %>%
                    select(ID, Formation, Class, Order, Suborder, Alliance, Suball, Assoc, Subass,Facies, Working, SiteUnit) %>% as.data.table
assoc.count <- assoc %>% count(Assoc)
alliance.count <- alliance %>% count(Alliance)
order.count <- order %>% count(Order)
new_su <- left_join(su2, su.lookup)%>% select(PlotNumber, SiteUnit.orig) %>% rename(SiteUnit = SiteUnit.orig)
clst.units2 <- clst.units %>% distinct %>% left_join(new_su) 
fwrite(clst.units2, "./outputs/coastal_orders_cutoffs_to_test_v2.csv")
```


### Write cluster membership into Vpro hierarchy table.
This table can be copied and pasted into a Hierarchy_template table in Vpro. Then two make table queries must be run to align the hierarchy tags with a the autonumbered ID field generated by ACCESS. The table name for the new _Hierarchy table can be named to match the intent of the hierachy
```{r convert into a hierarchy table}
levelNames <- c("Formation", "Class", "Order", "Suborder", "Alliance", "Suball", "Assoc", "Subass", "Facies", "Working", "SiteUnit")
new.hier <- tableToTree(clst.units, levelNames = levelNames) %>% as.data.frame

fwrite(new.hier, "./hierarchy_to_vpro/floodplainfromcluster_v2.csv")
```


Correlation with environmental variables
```{r add BGC and climate data}
# require(sf)
# require(climr)
# coords <- fread("./climate_data/CoastBGCv13_1kmgrid.csv") %>% rename(elev = elev1, long = xcoord, lat=ycoord) %>% dplyr::select(-fid)# long, lat, elev, id, -BGC)
# coords <- coords[!is.na(elev),]
# coords_sf <- st_as_sf(coords, coords = c("lat","long"), crs = 4326)
# coords_sf <- st_transform(coords_sf, 3005)
# coords <- as.data.frame(coords)# %>% dplyr::rename(long = 1, lat = 2)
# setcolorder(coords, c("long","lat","elev","id", "BGC"))
# tic()
# clim_vars <- climr_downscale(coords, which_normal = "auto", vars = list_variables(), return_normal = TRUE, cache = TRUE)
# toc()
# saveRDS(clim_vars, "./climate_data/coastzonal.rds")
```


```{r climate dendro}
clim_vars <- readRDS("./climate_data/coastzonal.rds")
su.bgc <- su2 %>% mutate(bgc = sub("\\_.*", "", su2$SiteUnit)) %>% select(bgc) %>%  distinct %>% mutate(bgc = ifelse(bgc =="IDFww1", "IDFww", 
                                                                                                                     ifelse(bgc == "CWHxm3", "CWHdm", bgc)))
clim_vars2 <- clim_vars %>% filter(BGC %in% su.bgc$bgc)

bgc.climate <- clim_vars2 %>% group_by(BGC) %>%  summarise(across(where(is.numeric), mean, na.rm = TRUE), n = n()) 
bgc.climate2 <- bgc.climate %>% mutate(across(where(is.numeric), scale, center = FALSE))  %>% select_if(~ ! any(is.na(.)))#%>% select(where(is.numeric))
```


```{r climate dendro}
bgc.clim.choose <- bgc.climate2 %>% select(BGC, CMD_sm, PAS, TD, DD18)%>% column_to_rownames("BGC") %>% as.matrix


require(ecodist)
clim.dist <- ecodist::bcdist(bgc.clim.choose) %>%  as.matrix %>% as.data.frame

ss_clst = agnes(clim.dist , diss = TRUE, method = "average")

dendro_clim <- as.hclust(ss_clst)
### returns to cophonetic value - considered good if over 0.75
dend.dis <- as.dist(clim.dist) 
dend.co <- cophenetic(dendro_clim)
cophonentic <- cor(dend.dis, dend.co)
cophonentic ##
```


```{r climate dendro}
dendro.clim <- 
  fviz_dend(dendro_clim, cex = .6, lwd = 1, h=.25,
        rect = TRUE,
         type = "rectangle",
          k_colors = "jco",
      rect_border = "black",
          rect_fill = TRUE,
          lower_rect = -.1,
          horiz = TRUE,
          ggtheme = theme_calc(),labels=F)
dendro.clim
#ggsave("./graphics/coastal_zonals_cluster.pdf", dendro, width=12, height=6)
```


```{r double dendrogram}
require(dendextend)
dend.veg <- as.dendrogram(dendro_hc)
dend.clim <- as.dendrogram(dendro_clim)
dl <- dendlist(dend.veg, dend.clim)
tanglegram(dl, sort = TRUE)#, common_subtrees_color_lines = FALSE, highlight_distinct_edges  = FALSE, highlight_branches_lwd = FALSE)
```




