---
title: "Pairwise_Diagnostic"
author: "WH MacKenzie"
date: "08/01/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(reshape2)
require(plyr)
require(dplyr)
require(tidyr)
require(ggplot2)
require(magrittr)
require(foreach)
require(tcltk)
require(openxlsx)
require(doParallel)
require(doBy)
require(doParallel)

```

## Read in data and sorts and spp lumping

```{r load data}
load(file = "./clean_data/SS_sum_analysis_data.RData")
```

## Settings for pairwise
```{r}
## for convert data in comparisons
sig.classes = c(0,1,2,3,4,5,6,7,8,9)
sig.class.cuts <- c(.1, .3, 2.2, 5, 10, 20, 33, 50, 75)
## function to convert summary data into classes
pres.class <- c(.2, .4, .6, .8)
## function to convert constancy into classes

## for categorization in pair-wise comparison
pair.wise
cons = .6 
  d1 = .8
  d2 = .6
  d3 = .4
  cd = 6 ##sig. class
  ## c will be >cm and <cd
  cm = 1
  
  pair.wise.scores
  d1 = 4
  d2 = 3
  d3 = 1
  no_d = 0
  dd1 - 4
  dd2 = 3
  dd3 = 2
  dd4 = 1
  c = 0
  cd = 0
  cm = -2
  
d_dd = 1.25
moss = .1
round = 1
exclude
  
```

## Calculate constancy and mean cover
Adjust data set to only include species of high constancy in one of the compared untis

```{r settings, echo=FALSE}

```

## Pairwise comparison
unit A vs B and B vs A
compare covers and convert difference into cover class
compare constancy and convert distance into constancy class
assign points based on constancy and cover rules




