---
title: "Generate Associations by Analysis Group"
author: "William H MacKenzie"
date: today
format:
  typst:
    toc: false
    toc-depth: 1
    toc-title: Contents
    section-numbering: 1.1.1
    columns: 1
editor: source
execute:
  echo: false
  error: false
  warning: false
  message: false
  fig.width: 6
  fig.height: 4
  fig.align: 'center'
  fig.cap: true
  fig.pos: H
  out.width: '100%'
  dev: pdf
  fig.ext: pdf
  cache: false
  fig.retina: 2
  dpi: 600
  fig.asp: 1.5
  fig.path: "./figures/"
output-file: "FdPy_Generated_Hierarchy.pdf"
---

```{r}
## Choose analysis set and change name of output-file above
#anal.set <- c("CWH-HwBa-rich")# "CWH-102" "CWH-HwBa" "CWH-HwBa-rich"  "CWH-HwFd" "CWH-HwFd-rich" "coast-wetfor" "MH" "Act"
anal.set <- c("FdPy")# "CWH-102" "CWH-HwBa" "CWH-HwBa-rich"  "CWH-HwFd" "CWH-HwFd-rich" "coast-wetfor" "MH" "Act"
#anal.set <- c("AllFor")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(tidyverse)
require(DBI)
require(data.table)
require(cluster)
require(dendextend)
require(dynamicTreeCut)
require(gtable)
require(gtsummary)
require(colorspace)
require(openxlsx)
require(tictoc)
require(labdsv)
require(indicspecies)
require(Polychrome)
require(ggdendro)
require(purrr)
require(ggplotify)
require(grid)
require(data.tree)
library(stats)
library(overlapping)
require(gt)
require(vegan)
library(tidytext)
library(stringr)

set.seed(1279)
source("../BEC_R_working/_functions/_bec_dist.R")
source("../BEC_R_working/_functions/_bec_dist_matrix.R")
source("../BEC_R_working/_functions/_combine_taxa.R")
source("../BEC_R_working/_functions/_combine_taxa_strata.R")
source("../BEC_R_working/_functions/_create_su_vegdata.R")
source("../BEC_R_working/_functions/_create_analysis_vegsum.R")
source("../BEC_R_working/_functions/_TabletoTree.R")
source("../BEC_R_working/_functions/_TreetoTable.R")
source("../BEC_R_working/_functions/_add_vars.R")
source("../BEC_R_working/_functions/_do_pairwise.R")
source("../BEC_R_working/_functions/_create_diagnostic_veg.R")
source("../BEC_R_working/_functions/_return_similar_pairs.R")
source("../BEC_R_working/_functions/_read_sppmaster.R")
source("../BEC_R_working/_functions/_combined_su.R")
source("../BEC_R_working/_functions/_create_veg_sum.R")
source('../BEC_R_working/_functions/_create_veg_sum_all.R')
source('../BEC_R_working/_functions/_build_species_ordering.R')
source('../BEC_R_working/_functions/_format_veg_table.R')
source('../BEC_R_working/_functions/_format_veg_table_strata.R')
source('../BEC_R_working/_functions/_format_veg_table_nostrata.R')
source('../BEC_R_working/_functions/_encode_veg_sum.R')
source('../BEC_R_working/_functions/_create_dendro.R')
source('../BEC_R_working/_functions/_create_dendro_all.R')
source('../BEC_R_working/_functions/_create_dendro_bybgc.R')
source('../BEC_R_working/_functions/_draw_dendro_split.R')
source('../BEC_R_working/_functions/_cluster_groups.R')
source('../BEC_R_working/_functions/_TabletoTree.R')
source('../BEC_R_working/_functions/_build_species_ordering_hier.R')
source("../BEC_R_working/_functions/_cluster_by_spp_group.R")
source("../BEC_R_working/_functions/_baselga_diagnostics.R")

```

# Introduction

This script compares site unit summary vegetation data to identify associations and subassociations. The analysis is based on a agglomerative clustering approach, which groups site units based on their vegetation community similarity. A BEC similarity metric is constructed from site unit summary data with species diagnostic value assigned to both constancy and mean cover. Following Pojar and Klinka only species with high constancy in a focal unit contribute to diagnostic potential of a unit and only species with high mean cover are assigned dominant differential value.

The process in brief

1.  Analysis group to select related ecosystems and reduce the size of the analysis for review

2.  Site Unit summary values are generated from published sorts and a pair-wise comparison of all units to generate a BEC similarity metric for the paired comparison

3.  The dissimilarity matrix from the pair-wise function is submitted to an AGNES agglomerative clustering analysis and a dendrogram tree produced

4.  Generated hierarchy units are generated but cutting the cluster tree at different similarity thresholds. These initial thresholds were chosen based on tests and iterations to match historical concepts of the levels by Meidinger, Pojar, and Klinka
- 80% similar are Associations
- 60% similar are Alliances
- 20% similar are Orders

5.  Review materials from the quantitative analysis are produced with a focus on the Association and Alliance. 

  - A vegetation summary workbook by site unit, on separate alliance sheets and colour themed by Association. The themeing is completed by macro in Excel.

  - A table of possible alternate association membership for site series based on similarity.
  
  - A summary vegetation workbook with each site unit as a focal unit with all other units with high similarity and those units generated Association and Alliance indicated.
  
  - A PCA of edaphic and climate characteristics of siteunits within generated associations to look for outliers and potentially misplaced site units. 

```{r load data, echo=FALSE}
#------------------------------------------------------------
# 1. Load vegetation data + taxon lifeform classifications
#------------------------------------------------------------
taxon.all <- read_sppmaster()
taxon.exotic <- taxon.all %>%
  filter(Native == "E") %>%
  dplyr::select(Code) %>%
  pull()

taxon.lifeform <- taxon.all %>%
  filter(Codetype %in% c("U", "X", "D")) %>%
  dplyr::select(Code, ScientificName, EnglishName, Lifeform) %>%
  distinct()

trees <- taxon.lifeform %>% 
  filter(Lifeform %in% c("1", "2")) %>% 
  pull(Code)

tree_seedlings <- taxon.lifeform %>% 
  filter(Lifeform %in% c("1", "2")) %>% 
  mutate(Code = paste0(Code, "D")) %>% 
  pull(Code)
veg.dat <- readRDS("./clean_data/Analysis_BECMaster_Veg.rds") %>% filter(!Species %in% taxon.exotic, !startsWith(Species, "POA_PRA"))
#------------------------------------------------------------
# 2. Read lump table and apply combine_taxa()
#------------------------------------------------------------

veglump <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BC_Correlation2_Vpro_2023/Correlation2_Spp_lump.accdb;"
)

lump <- dbReadTable(veglump, "CorrelationII_Lump")
dbDisconnect(veglump)

veg.dat2 <- combine_taxa(vegdata = veg.dat, lump, use.subtaxa = FALSE)%>% 
   filter(!Species %in% taxon.exotic)

#------------------------------------------------------------
# 3. Load SiteUnit data from two LMH databases
#------------------------------------------------------------

#db <- "D:/BC_Correlation2_Vpro_2023/Sorts_published_LMH/LMH77_Coast_Sorts.accdb"
#db2 <- "D:/BC_Correlation2_Vpro_2023/Sorts_published_LMH/LMH68_HG_Sorts.accdb"
db <- "D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb"
su <- rbind(combined_su(db))#, combined_su(db2))

su <- su %>%
  filter(!grepl("X|unplaced|omit|support|nudum|Ww|!", SiteUnit)) %>%
  arrange(SiteUnit)
su2 <- su %>%
  mutate(SiteUnit = str_replace(SiteUnit, "[abc]$", ""))|> #,# ignore phases
         # SiteUnit = if_else(
         #   str_detect(SiteUnit, "101"),
         #   str_replace(SiteUnit, "\\.[123]$", ""), # remove zonal variations
         #   SiteUnit
         # )) %>%
    # # mutate( SU_Label = if_else(
    #        str_detect(SiteUnit.orig, "101"),
    #        str_replace(SiteUnit.orig, "\\.[123]$", ""), # remove zonal variations
    #        SiteUnit.orig))
          mutate(SU_Label = str_replace(SiteUnit.orig, "[abc]$", ""))# ignore phases
  

# Plot counts
n_plots <- su2 %>%
  group_by(SiteUnit) %>%
  summarise(n_plots = n_distinct(PlotNumber), .groups = "drop") 

su2 <- su2 # ignore phases
         # SiteSeries = if_else(
         #   str_detect(SiteUnit.orig, "101"),
         #   str_replace(SiteUnit.orig, "\\.[123]$", ""), # remove zonal variations
         #   SiteUnit.orig
         # ))

#------------------------------------------------------------
# 5. Load BECMaster environmental + edatopic tables
#------------------------------------------------------------

becmaster <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster/BECMaster_fixing.accdb;"
)

plot.env <- dbReadTable(becmaster, "BECMaster_fixing_Env")
plot.edatope <- dbReadTable(becmaster, "BECMaster_fixing_edatopic")
plot.info <- dbReadTable(becmaster, "BECMaster_fixing_Admin")

dbDisconnect(becmaster)


#------------------------------------------------------------
# 6. Write n_plots to BECdb and load hierarchy + edatopic
#------------------------------------------------------------

becdb <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v1.accdb;"
)

dbWriteTable(becdb, "n_plots", n_plots, append = FALSE, overwrite = TRUE,
  batch_rows = 1)

BGC_all <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(FSRegion, BGC_NoSpace)

BEC.hier <- dbReadTable(becdb, "tblBEC_Hierarchy_Class") %>%
  filter(SU_Label %in% su2$SU_Label) %>%
  mutate(SiteUnit = str_replace_all(SU_Label, c("/" = "_", " " = "")))

bgcv13 <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(BGC_NoSpace, SiteClassificationStatus)

edatopic <- dbReadTable(becdb, "Edatopic_v13_3") %>%
  mutate(
    rSMR = paste0("rSMR", substr(Edatopic, 2, 2)),
    SNR = substr(Edatopic, 1, 1)
  ) %>%
  filter(startsWith(Source, "BEC")) %>%
  dplyr::select(BGC, rSMR, SNR, SS_NoSpace, Edatopic) %>%
  distinct() %>%
  mutate(SiteUnit = gsub("/", "_", SS_NoSpace))

dbDisconnect(becdb)


##Use working spreadsheet instead of BECdb
#BEC.hier <- openxlsx::read.xlsx("D:/OneDrive/WNA_BGC/Association_Analysis/Coast_Hierarchy_Analysis/BECdb_Coast_Hierarchy_Working_ver2.xlsx") #|> 
  #filter(Analysis_Group %in% "Forest-Coast_MH")

# Keep only site units present in hierarchy table
su2 <- su2 %>%
  filter(SU_Label %in% BEC.hier$SU_Label)
su.class <- su2 %>%
  left_join(BEC.hier, by = c("SiteUnit" = "SiteUnit", "SU_Label" = "SU_Label"))

#------------------------------------------------------------
# 7. Identify SiteUnits missing from hierarchy
#------------------------------------------------------------

missing_in_hierarchy <- su.class  %>%
  distinct(SiteUnit) %>%
  anti_join(BEC.hier %>% distinct(SiteUnit), by = "SiteUnit") %>%
  mutate(flag = "in_su2_not_hierarchy")
missing_in_hierarchy
# fwrite(missing_in_hierarchy, "./missing_in_hierarchy.csv")


```

## Use previously set hierarchy units for analysis SU where assigned.

This code links Site Units to their BEC hierarchy classifications and determines the most appropriate level for analysis. After joining hierarchy metadata to the Site Unit table, the script walks down the predefined hierarchy—from broad classes to fine Site Units—to identify the first available non‑missing classification for each record. This resolved level becomes the analytic grouping (anal.test). Finally, the script assigns a working hierarchy label (SiteUnit.hier) based on a user‑selected grouping level (e.g., Association), falling back to the root Site Unit name when no Association designation has been provided. The result is a clean, consistent hierarchy assignment ready for downstream ecological analysis.

```{r choose level, echo=FALSE}
## Limit the analysis based on Analysis_Group column in the hierarchy table. 
#, "Deciduous-Coast", 

#------------------------------------------------------------
# 1. Define hierarchy (cleaned)
#------------------------------------------------------------

hierarchy <- c(
  "Class", "Order", "Suborder", "Alliance", "Suballiance","Association", "Subassociation", 
  "generated_subass", "generated_association", "generated_alliance", 
  "working_subass", "working_association", "working_alliance", 
  "SiteUnit"
)

group.to.test <- "Association"## uses data for this field to initially group siteunits where applicablle


#------------------------------------------------------------
# 2. Prepare hierarchy table
#------------------------------------------------------------

#BEC.hier2 <- BEC.hier #%>%
 # dplyr::select(-SiteUnit)   # keep SS_Label + hierarchy columns

#------------------------------------------------------------
# 3. Join hierarchy onto su2 and reduce to anal.set
#------------------------------------------------------------
#anal.set = c("MH", "Bl", "CwHw", "Fd", "CWH-Fd", "CWH")  ## use this to find possible extra Order similar members
su.anal <- su.class %>%
  #left_join(BEC.hier, by = c("SiteSeries" = "SU_Label")) %>%
  filter(Analysis_Group %in% anal.set) #%>%   filter(Order %in% Order.set)
  

#------------------------------------------------------------
# 4. Compute anal.test by walking down the hierarchy
#------------------------------------------------------------
#group.to.test <- "working_Assoc"
# vectorized helper: find first non-NA across hierarchy levels
first_non_na <- function(df, cols) {
  apply(df[cols], 1, function(x) {
    x_no_na <- x[!is.na(x)]
    if (length(x_no_na) == 0) NA_character_ else x_no_na[1]
  })
}

# compute anal.test
su.anal <- su.anal %>%
  mutate(
    anal.test = ifelse(
      !is.na(Analysis_Group),
      Analysis_Group,
      first_non_na(., hierarchy)
    )
  )

#------------------------------------------------------------
# 5. Compute SiteUnit.hier based on group.to.test
#------------------------------------------------------------

# su.anal1 <- su.anal %>%
#   mutate(
#     SiteUnit.hier = coalesce(.data[[group.to.test]], SiteUnit)
#   ) %>%
#   dplyr::select(
#     PlotNumber, SiteUnit, SU_Label, generated_alliance, generated_association, generated_subass,
#     !!sym(group.to.test), anal.test, SiteUnit.hier
#   ) %>%
#   group_by(SiteUnit.hier) %>%# 6. remove any hierarchy units with fewer than 6 plots
#   filter(n() >= 1 | !is.na(!!sym(group.to.test))) %>%
#   ungroup()

```

## Option to add summary data from adjacent juridiction site units for comparison

Other constancy tables for comparison (e.g. USA) table can be added to the analysis for comparison in do.pairwise function.

```{r add extra summary units}
usa_units <- fread("./usa.summary.units/USA_prepped_units3.csv") %>% dplyr::select(-ScientificName) %>% 
   filter(!Species %in% taxon.exotic,!startsWith(Species, "POA_PRA")) %>% filter(startsWith(SiteUnit, "Q"))
usa_units <- usa_units %>%
  mutate(Species = recode(Species,
                          POLYGLY = "POLYPODI",
                          SELAWAL = "SELAGINE"))

#veg.dat2 <- combine_taxa(vegdata = usa_units, lump, use.subtaxa = FALSE)
#usa_units <- readRDS("./usa.summary.units/usa_assocs_vegdat.rds")
# dups <- veg.comb %>%
#   group_by(SiteUnit, Species) %>%
#   filter(n() > 1) %>%
#   arrange(SiteUnit, Species)
```

## Run pairwise vegetation similarity analysis


```{r build pair.wise similarity matrix, echo=FALSE, message=FALSE, warning=FALSE}

#------------------------------------------------------------
# 1. Define indicator species and reduction settings
#------------------------------------------------------------

key.site.indicators <- c("NEPHCRI", "ABIELAS","LYSIAME", "CASSMER", "PHYLLODO", "SPHAGNUM")
#   "POPUTRI","GAULSHA","OPLOHOR","ELLIPYR","ATHYFIL",
#   "RUBUSPE","EQUIARV","GYMNDRY","VALESIT","LUETPEC",
#   "POLYMUN","STRUSPI","RHODGRO","EMPENIG","PICESIT"
# )

reduced.exceptions <- c("SPHAGNUM","CLADONIA","CLADINA","RACOMITR","MNIUM")
reduced.lifeforms  <- c(9, 10, 11)

#------------------------------------------------------------
# 2. Prepare SiteUnit table for analysis
#------------------------------------------------------------
# SiteUnit.hier is replacing site unit with a previously determined Association group
su_working <- su.anal#%>%   rename(SiteUnit.base = SiteUnit, SiteUnit = SiteUnit.hier)

#------------------------------------------------------------
# 3. Run pairwise vegetation similarity analysis
#------------------------------------------------------------
pairwise.params <- list(
  su                = su_working,
  su.field = "SiteUnit",
  minconstancy          = 50,
  noiseconstancy        = 10,
  minplots              = 1,
  minor                 = 5,
  use.ksi               = FALSE,
  ksi   = key.site.indicators,
  ksi.value             = 1.5, ## change this to apply to creation of diagnostic potential as a factor ^1/2 by default
  reduce.lifeform       = TRUE,
  reduced.lifeforms     = reduced.lifeforms,
  reduction             = 0.1, ## keep this as a reduction of actual final diagnostic points
  reduced.exceptions    = reduced.exceptions,
  add_vegsum_table      = NULL,
  mindiff = .2, 
  mindd = 0.35
)

vegsum.pairs <- do.call(
  do_pairwise,
  c(list(veg.dat2), pairwise.params)
)

# Start with everything except `su`
print.params <- pairwise.params[setdiff(names(pairwise.params), "su")]

# Remove ksi if use.ksi == FALSE
if (!pairwise.params$use.ksi) {
  print.params <- print.params[setdiff(names(print.params), c("ksi", "ksi.value"))]
}

# Remove reduction-related fields if reduce.lifeform == FALSE
if (!pairwise.params$reduce.lifeform) {
  print.params <- print.params[setdiff(names(print.params),
                                       c("reduced.lifeforms", "reduction", "reduced.exceptions"))]
}
library(tibble)
library(knitr)

param.table <- tibble(
  parameter = names(print.params),
  value = sapply(print.params, function(x) paste(x, collapse = ", "))
)

param.table %>%
  gt() %>%
  tab_header(title = "Parameter Settings for do_pairwise()") %>%
  tab_options(
    table.font.size = px(8),
    table.border.top.width = px(1),
    table.border.bottom.width = px(1)
  )

#------------------------------------------------------------
# 4. Extract comparison table
#------------------------------------------------------------
#fwrite(vegsum.pairs, "./outputs/vegsum_pairs.csv")

### return unit species information
vegsum.pairs.spp <- vegsum.pairs |> dplyr::select(Unit1, Unit2, Species, ScientificName, MeanCov.unit1, Constancy.unit1, MeanCov.unit2, Constancy.unit2, diff.pts.unit1, diff.pts.unit2)

### return distinct unit comparisons

unit.compare.temp <- vegsum.pairs %>%
  dplyr::select(
    Unit1, Unit2, diff.ratio.unit1, diff.ratio.unit2, nplots.unit1, nplots.unit2, unit.diag.sum.unit1, unit.diag.sum.unit2) %>%
  distinct(Unit1, Unit2, .keep_all = TRUE)


focal.unit.compare <- unit.compare.temp %>% 
    dplyr::select(
      Unit1 = Unit2,
      Unit2 = Unit1,
      diff.ratio.unit1 = diff.ratio.unit2,
      diff.ratio.unit2 = diff.ratio.unit1,
      nplots.unit1 = nplots.unit2,
      nplots.unit2 = nplots.unit1,
      unit.diag.sum.unit1 = unit.diag.sum.unit2,
      unit.diag.sum.unit2 = unit.diag.sum.unit1
      ) |>  rbind(unit.compare.temp) |> 
  distinct() |> mutate(BEC.sim.min = pmin(diff.ratio.unit1, diff.ratio.unit2),
                      BEC.sim.max = pmax(diff.ratio.unit1, diff.ratio.unit2))


#fwrite(focal.unit.compare , "./outputs/focal_unit_compare.csv")


source("../BEC_R_working/_functions/_build_bec_distance_matrix.R")

result <- build_bec_distance_matrix(
  vegsum_pairs = vegsum.pairs,
  su_anal = su.anal,
  preclass_field = "Association",
  use_preclass = TRUE
)

unit.compare2 <- result$unit_compare
bec.dist <- result$bec_dist


# openxlsx::write.xlsx(
#   x = bec.dist,
#   file = "./clean_tabs/Coast_Forest_Distance_matrix.xlsx",
#   rowNames = TRUE
# )
```

## Cluster analysis and dendrogram of all good site series and merged site units

The agglomerative clustering algorithm AGNES successively groups site units based on dissimilarity from the bottom up (most similar units get grouped first). Because units or initial clusters may have near equal dissimilarity to neighbours in two directions, the algorithm may not be able to maintain an exact ordering of dissimilarity. We evaluate the cluster analysis using two metrics. 1) The cophenetic correlation coefficient (CCC) evaluate how well a hierarchical clustering dendrogram preserves the original pairwise distances between observations. The closer CCC is to 1, the better it represents the original distance matrix. Lower numbers indicate that some branchlets of the dendrogram can be associated with more than one branch. Lower CCC values suggest that more review of higher level position of units is warranted 2) The agglomerative coefficient (AC) is a metric that quantifies the strength of the clustering structure. The closer AC is to 1 indicates strong clustering structure and the observations are tightly grouped by low dissimilarity. This will be affected by the breadth/size of the analysis where a larger number of site units will tend to have lower AC values.

We chose to use the clustering method that maximizes the CCC. We tested a number of methods: complete, average, single, ward, weighted, and flexible (at a wide range of settings). The "average" method consistently produced the highest cophenetic values so we applied this as the standard method for clustering.

In the dendrograms, the red line represents the minimum threshold for separating site units into separate site series. This is set at 10% dissimilarity, but in some BGCs we separate some site series to as low as 7% dissimilarity. The green line represents the preliminary dissimilarity threshold for an association. The initial threshold for an association is set at 20% (80% similar) based on review.

```{r }
#| label: fig-first-cluster
#| fig-cap: Cluster of Site Units (including initial predefined groups
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12
source("../BEC_R_working/_functions/_bec_dist_matrix.R")
dendro_hc <- create_dendro_all(unit.compare2, threshold.low = .4)
```

### Cluster diagram subsets from cut.level

Below are branches of the complete dendrogram cut at a selected similar threshold. In this report, a dissimilarity of 40% is used representing prospective Alliance level units. Units that are singles at this threshold presented in table form.

```{r split first cluster}
source("../BEC_R_working/_functions/_bec_dist_matrix.R")
singles <- draw_dendro_split(unit.compare.tbl = unit.compare2, cut.level=.4, show.singles = FALSE)
singles |> 
dplyr::arrange(SiteUnit) 
gt(singles) |> 
  tab_header(
    title = "Single Units"
  ) |>
  tab_options(
    column_labels.hidden = TRUE,
    table.font.size = px(9),
    data_row.padding = px(1),
    row.striping.include_table_body = FALSE
  ) |>
  cols_width(
    everything() ~ px(80)
  ) |>
  opt_vertical_padding(scale = 0.3)

```

## Site Units where quantitative analysis is unreliable

Units with few plots (=<6) or low diagnostic potential (low number of high constancy species) are often not properly assigned to associations in the quantitative analysis. These units will generally require more scrutiny and expert reassignment. 
```{r}
#| label: tab-too-few-plots
#| tab-cap: "Site Units with Fewer than 6 Plots"
#| tbl-cap-location: top
#| warning: false
#| tab-align: "center"

# Identify Site Units with fewer than 8 plots
ss_too.few <- unit.compare.temp %>%
  dplyr::select(Unit1, nplots.unit1) %>%
  filter(nplots.unit1 <= 6 & nplots.unit1>0) %>%
  distinct() %>%
  rename(
    "SiteUnit"      = Unit1,
    "Number of Plots" = nplots.unit1)

# Count unique Site Units for dynamic column label
num_units <- ss_too.few %>%
  pull(SiteUnit) %>%
  n_distinct()

col_label <- paste0("Site Unit (", num_units, ")")

# Build GT table
gt::gt(ss_too.few) %>%
  gt::fmt_number(columns = "Number of Plots", decimals = 0) %>%
  gt::cols_label(SiteUnit = col_label) %>%
  gt::tab_options(table.font.size = 8)

ss_low.diag <- unit.compare.temp %>%
  dplyr::select(Unit1, unit.diag.sum.unit1)  %>% 
  filter(unit.diag.sum.unit1 < 24 & unit.diag.sum.unit1 >0) %>% 
  rename("Diagnostic Potential" = unit.diag.sum.unit1, "SiteUnit" = Unit1) %>%
  distinct()

num_units <- ss_low.diag %>% 
  pull(SiteUnit) %>% 
  n_distinct()
col_label <- paste0("Site Unit (", num_units, ")")

  gt::gt(ss_low.diag) %>%
  gt::fmt_number(columns = "Diagnostic Potential", decimals = 0) %>%
  gt::tab_options(table.font.size = 8) %>%
  gt::cols_label(SiteUnit = col_label) %>% 
  gt::tab_options(table.font.size = 8)

```

```{r,  echo=FALSE}
#| label: tab-low-diagnostic
#| tbl-cap-location: top
#| tbl-cap: "Site Units with Low Diagnostic Potential"
#| warning: false
#| tbl-align: "center"

# ss_low.diag <- unit.compare.temp %>%
#   dplyr::select(Unit1, unit.diag.sum.unit1)  %>% 
#   filter(unit.diag.sum.unit1 < 24 & unit.diag.sum.unit1 >0) %>% 
#   rename("Diagnostic Potential" = unit.diag.sum.unit1, "SiteUnit" = Unit1) %>%
#   distinct()
# 
# num_units <- ss_low.diag %>% 
#   pull(SiteUnit) %>% 
#   n_distinct()
# col_label <- paste0("Site Unit (", num_units, ")")
# 
#   gt::gt(ss_low.diag) %>%
#   gt::fmt_number(columns = "Diagnostic Potential", decimals = 0) %>%
#   gt::tab_options(table.font.size = 8) %>%
#   gt::cols_label(SiteUnit = col_label) %>% 
#   gt::tab_options(table.font.size = 8)
```

## Create generated hierarchy units

From set dissimilarity level cuts in the dendrogram, initial hierarchy unit membership is assigned and the labels exported to spreadsheet for review and are identified by a generated_ suffix. In this table, expert review should then assign a unit to the working_ unit fields and the reason for the assignment given.
if the same as the generated then "analysis"
if the unit is a singleton and get merged with other units then "best fit"
if the unit gets moved from one association to a different association then "alternate"
This table is designed for import into the BECdb tblBECHierarchy_Class table to manage the classification

```{r, echo=FALSE, results= 'hide'}
#| label: fig-second-cluster
#| fig-cap: ("Cluster of Site Units (including initial predefined groups)")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12
source("../BEC_R_working/_functions/_bec_dist_matrix.R")

fig.show='hide'
#--------- do for associations

dendro_hc <- create_dendro_all(unit.compare2, output = FALSE)

clusters <- cutree(dendro_hc, h = 0.10)
cluster_subass<- data.table(
  SiteUnit = names(clusters),
  generated_subass = paste0(anal.set, "_generated_subass_", clusters)
)
clusters <- cutree(dendro_hc, h = 0.20)
cluster_assoc <- data.table(
  SiteUnit = names(clusters),
  generated_association = paste0(anal.set,"_generated_association_", clusters)
)
  clusters <- cutree(dendro_hc, h = 0.30)
cluster_suball <- data.table(
  SiteUnit = names(clusters),
  generated_suball = paste0(anal.set,"_generated_suball_", clusters)
)
clusters <- cutree(dendro_hc, h = 0.40)
cluster_alli <- data.table(
  SiteUnit = names(clusters),
  generated_alliance = paste0(anal.set,"_generated_alliance_", clusters)
)

clusters <- cutree(dendro_hc, h = 0.60)
cluster_subord <- data.table(
  SiteUnit = names(clusters),
  generated_subord= paste0(anal.set,"_generated_subord_", clusters)
)
clusters <- cutree(dendro_hc, h = 0.80)
cluster_ord <- data.table(
  SiteUnit = names(clusters),
  generated_order= paste0(anal.set,"_generated_order_", clusters)
)
# clusters <- cutree(dendro_hc, h = 0.99)
# cluster_class <- data.table(
#   SiteUnit = names(clusters),
#   generated_Class= paste0(anal.set,"_generated_Class_", clusters)
# )

# Combine
cluster_table <- Reduce(
function(x, y) merge(x, y, by = "SiteUnit", all = TRUE),
list(
cluster_subass,
cluster_assoc,
cluster_suball,
cluster_alli,
cluster_subord,
cluster_ord)#,
#cluster_class)
)

#cluster_table
## merge back to retrieve SiteUnit for all and the coalesce.
orig.SU <- su.anal %>% dplyr::select(SU_Label, SiteUnit, working_alliance, working_association, working_subass) %>% distinct
generated.hier <- cluster_table %>%
  left_join(orig.SU, by = c("SiteUnit" = "SiteUnit")) |> group_by(generated_association) |> add_count(generated_association) |> ungroup() |>
  left_join(n_plots) |>  select(n_plots, assoc_n_siteunit= n, everything())

hier.fields <- c(
  "generated_order",
  "generated_subord",
  "generated_alliance",
  "generated_suball",
  "generated_association",
  "generated_subass"
)

su.anal <- su.anal %>%
  select(-any_of(hier.fields)) %>%
  left_join(
    cluster_table, 
    by = "SiteUnit"
  )


fwrite(generated.hier, paste0("./hierarchy/", anal.set, "_generated_hierclasses.csv"))


## write into BECdb a working file to update the tempAssoc field
becdb <- dbConnect(
  odbc::odbc(),
  .connection_string =
    "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
     DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v1.accdb;"
)
dbWriteTable(
  becdb,
  paste0("generated_hier_", anal.set),
  generated.hier, append=FALSE,
  overwrite = TRUE,
  batch_rows = 1
)
dbDisconnect(becdb)

#xx <- cluster_update_assoc %>% filter(SS_Label == "CWHds1_101a")
```

## Review placement of site units with more than one possible pairing at Association level

**See file "*_other_similar_units.xlsx" in review materials**for Focal units with more than one similar unit Highlights other similar units that are not in the same association for review This will look for unit pairs that fall below a similarity threshold and then identify where they have not been put into the same Association. We have set the minimum threshold at 65% similar.

```{r create alternate su matches table}
source("../BEC_R_working/_functions/_export_similar_units.R")
similar_units <- export_similar_units(
  unit.compare = unit.compare2,
  generated.hier = generated.hier,
  anal.set = anal.set,
  same.threshold = 0.65,
  out.dir = "./review_materials"
)
#check that equal number of pair comparisons
#similar_counts <- similar_units %>%
#   count(FocalUnit, name = "n_similar_units") %>%
#   arrange(desc(n_similar_units))
# 
# similar_counts
```

**See file "*_FocalUnit_Comparisons.xlsx" in review materials** where each Focal unit has a worksheet showing its summary vegetation compared to all suitably similar units. This table is particularly useful when looking for either best fit or alternative units.  The summaries include header rows showing n_plots, similarity (to focal unit), and generated hierarchy placement of each unit (Subass, Assoc, Alliance)

```{r summary table of similar units}

source("../BEC_R_working/_functions/_create_focalunit_summary_workbook.R")

create_focalunit_summary_workbook(
  su.anal = su.anal,
  similar.units = similar_units,
  veg.dat = veg.dat2,
  taxon.lifeform = taxon.lifeform,
    anal.set = anal.set
)

```

## Constancy and cover tables

Generate vegetation summary for all site units by generated alliance colour-themed by generated association and subassociation for review. **See file "*_SiteUnit_by_generated_alliance.xlsx" in review materials**

Each sheet has all site units for a generated alliance on a single sheet grouped, labelled, and colour-themed by generated Association for initial review and observation of questionable plot placement. This will be used to review the proposed alliances and make decisions about which site units should be merged into associations, which should be separated into new associations, and which should be left as singletons. The vegetation summary tables will also be used to identify potential diagnostic species for each alliance and to compare the vegetation composition of different alliances.

```{r summary table association within alliances}
source("../BEC_R_working/_functions/_create_veg_summary_workbook.R")

create_veg_summary_workbook(
  su.anal = su.anal,
  veg.dat = veg.dat2,
  hier.tbl = generated.hier,
  taxon.lifeform = taxon.lifeform,
  group_level = "SiteUnit",
  within_level = "generated_alliance",
  classification = "generated",
  anal.set = anal.set,
  out.dir = "./review_materials"
)


```

```{r prep env data}
plot.env2 <- plot.env %>% mutate(lon = ifelse(Longitude<0, Longitude, 0-Longitude)) %>% mutate(Date = as_date(Date)) |> 
  filter(PlotNumber %in% su.anal$PlotNumber) |>
  dplyr::select(PlotNumber, lon, lat = Latitude, elev = Elevation, special = Exposure1, rSMR = MoistureRegime, rSNR = NutrientRegime, slope = SlopeGradient, aspect = Aspect ) 
 plot.info <- plot.info |> dplyr::select(Plot, BECSiteUnit, UserSiteUnit, HumusThickness, BGC=GIS_BGC)  |> filter(Plot %in% su.anal$PlotNumber)
plot.env2 <- left_join(plot.env2, plot.info, by = c("PlotNumber" = "Plot")) |> filter(!is.na (rSMR))


## extract climate data for plots from climr
require(climr)
plot.locs <- plot.env2 %>% dplyr::select(id=PlotNumber, BGC, lon,lat,elev) |> drop_na(lat)
plot.climate <- climr::downscale(setDT(plot.locs),
                              vars = list_vars(set = c("Annual", "Seasonal", "Monthly")),
                              return_refperiod = TRUE,
                              db_option = "local",
                              cache = TRUE)
## calculate aSMR
plot.env3 <- plot.climate |> select(PlotNumber = id, everything()) |> left_join(plot.env2) |> 
  mutate(
    moist = rowSums(
      across(c(PPT_05, PPT_06, PPT_07, PPT_08, PPT_09), ~ .x >= 90),
      na.rm = TRUE
    ),
    moist = ifelse(moist > 4, 1, 0),
    ratio = moist / 5,
    multiplier = 1 + ratio,
    CMD = CMD_an / multiplier
  ) |> mutate(PPT.dorm = PPT_at + PPT_wt) |> 
  mutate(Def = ifelse(CMD >=350, 350 - PPT.dorm, CMD)) |> 
  mutate(CMD.total = CMD + Def) |> 
  mutate(CMD.site = ifelse(slope > 35 & (aspect >180 & aspect <270), CMD.total *1.25,
                           ifelse(slope> 35 & (aspect>315 | aspect <45), CMD.total *0.85, CMD.total))) |> arrange(PlotNumber) |> 
  mutate(
    rSMR.clean = sub("^.*?(\\d+).*$", "\\1", rSMR)) |> 
  mutate(rSMR.clean = as.numeric(rSMR.clean),
  SNR.clean = recode(
    substr(trimws(rSNR), 1, 1),
    "A" = 1L,
    "B" = 2L,
    "C" = 3L,
    "D" = 4L,
    "E" = 5L,
    "F" = 6L,
    .default = NA_integer_
  ))
  ####To adjust in zones with prolonged snowpack remove CMD from early growing season months
## BGCs with or below heavy snow packs
# BGC_Special <- fread("./inputs/BGC_Special.csv", data.table = FALSE)
# allDat2 <- left_join(allDat2,BGC_Special, by = c("BGC" = "BGC"))
# allDat2 <- allDat2 %>% mutate(CMD = ifelse(is.na(Special), CMD, CMD_08)) %>% select(-Special, -CMD_08) %>% drop_na()
#rules <- read.csv("./inputs/aSMR_growingseason_Rules_HalfStep_v13_13Sept2025.csv")
rules <- read.csv("./inputs/aSMR_Rules_HalfStep_v13_23June2025.csv") |>
  mutate(
    rSMR.rule = sub("^rSMR", "", rSMR)) |> dplyr::select(-rSMR)

plot_asmr <- plot.env3 %>% select(PlotNumber, rSMR, CMD.site) |> 
  left_join(rules, by = c("rSMR" = "rSMR.rule")) %>%
  filter(CMD.site <= CMD) %>%
  group_by(PlotNumber) %>%  # ensure row-wise logic if needed
  slice_max(aSMR, with_ties = FALSE) %>%
  ungroup() %>% select(PlotNumber, aSMR)

plot.env3 <- left_join(plot.env3, plot_asmr)

plot.env3 <- plot.env3 |>
  left_join(
    su.anal, 
    by = "PlotNumber")
 
plot.env.vars <- plot.env3 |>   select(-SiteUnit, -working_association, -working_alliance) |> 
   select(PlotNumber, PAS_an,CMD.site, MSP, aSMR , SNR = SNR.clean, Tmin = Tmin_an, AHM,FFP, DD5 = DD5_an, DD18_sm ) |> arrange(PlotNumber)
 keep.rows <- complete.cases(plot.env.vars)

  su.env <- plot.env3 |> rename(assoc = generated_association, alliance = generated_alliance) |> 
  filter(!is.na(alliance)) |> arrange(SiteUnit) |> 
  group_by(SiteUnit, assoc, alliance) |>
    summarise(
  PAS_an = mean(PAS_an, na.rm = TRUE),
  CMD.site = mean(CMD.site, na.rm = TRUE),
  MSP   = mean(MSP, na.rm = TRUE),
  aSMR = mean(aSMR, na.rm = TRUE),
  SNR = mean(SNR.clean, na.rm = TRUE),
  AHM = mean(AHM, na.rm = TRUE),
  #lat = mean(lat, na.rm = TRUE),
  FFP   = mean(FFP, na.rm = TRUE),
  DD5 = mean(DD5_an, na.rm = TRUE),
  DD18_sm = mean(DD18_sm, na.rm = TRUE),
  Tmin = mean(Tmin_an, na.rm = TRUE),
  .groups = "drop"
)

#     summarise(
#   across(
#     where(is.numeric) & !c(lat, lon),
#     ~ mean(.x, na.rm = TRUE)
#   ),
#   .groups = "drop"
# )

# ==============================================================================
# Climate variables for PCA
#==============================================================================
  env.vars <- su.env |>
  select(-assoc, -alliance) 

# Remove rows with missing values
keep.rows <- complete.cases(env.vars)

env.vars <- env.vars[keep.rows, ]

# Remove constant variables if any
env.vars.filt <- env.vars |>
  select(-SiteUnit) |>
  dplyr::select(
    where(~ is.numeric(.x) && sd(.x, na.rm = TRUE) > 0)
  )

keep.rows <- complete.cases(env.vars.filt)

#su.env <- env.vars[keep.rows, ]
su.env <- su.env |>
  dplyr::select(
    SiteUnit, assoc, alliance,
    where(~ is.numeric(.x) && sd(.x, na.rm = TRUE) > 0)
  )

```

```{=typst}
#pagebreak()
```
# Environmental differentiation between units

## PCA of environment variables for analysis group
PCA from a limited set of edaphic (aSMR, SNR) and climatic (DD5, DD18_sm, FFP, Tmin_wt, CMD.site, AHM, MSP, PAS, )

```{r}
#| label: fig-env_pca
#| fig-cap: ("3D environment PCA graphic for SiteUnits within working Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 8


source("../BEC_R_working/_functions/_plot_pca_hierarchy.R")
library(forcats)
require(ggrepel)

su.env2 <- su.env |> select(siteunit = SiteUnit, everything())

pca.env <- plot_pca_hierarchy(su.env2,
                    n_vectors = 10,
    output_file = NULL,
    save_html = FALSE)

pca.env$plot


```

## Review environmental space of site units within alliances

Boxplot of siteunits along PCA axes by Allaince. Look for environmental site units outliers for membership reassessment.
```{r}
#| label: fig-PCA_boxplot
#| fig-cap: ("Environment PCA scores for SiteUnits within working Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 10

source("../BEC_R_working/_functions/_plot_pca_boxplots.R")
pca.boxplots <- plot_pca_boxplots(
  pca = pca.env,
  group_var = "assoc"
)

pca.boxplots

```

```{r}
orig.SU <- su.anal %>% dplyr::select(SU_Label, SiteUnit, working_alliance, working_association, working_subass) %>% distinct
generated.hier <- cluster_table %>%
  left_join(orig.SU, by = c("SiteUnit" = "SiteUnit"))# |> group_by(working_association) |> add_count(working_association) |> select(n_su_in_assoc = n, everything())

# Keep SiteUnit plus columns that are not already in generated.hier
cols.keep <- c("SiteUnit", setdiff(names(BEC.hier), names(generated.hier)))
BEC.hier.join <- BEC.hier %>% select(all_of(cols.keep))
BEC.hier.new <- left_join(
  generated.hier,
  BEC.hier.join,
  by = "SiteUnit"
)%>% select(all_of(names(BEC.hier))) |> group_by(working_association) |> add_count(working_association) |>   mutate(n = as.numeric(n)) |> 
  select(n_siteunits_working = n, everything()) |> ungroup()|>
  group_by(generated_association) |> add_count(generated_association) |> mutate(n = as.numeric(n)) |> select(n_siteunits_gen = n, everything()) |> ungroup()




BEC.hier.new <-BEC.hier.new |> select(Analysis_Group,	SpecialCode,	SU_Label,	n_plots,	SiteUnitLongName,	 working_association,working_subass, working_alliance,	association_decision,	n_siteunits_gen, n_siteunits_working,	generated_subass, generated_association,	generated_suball,	generated_alliance,	generated_subord,	generated_order,	working_suball,	working_subord,	working_order, Comments)


library(openxlsx)
redFill <- createStyle(fgFill = "#F4CCCC")
ylwFill <- createStyle(fgFill = "#FFF2CC")
grnFill <- createStyle(fgFill = "#D9EAD3")
orgFill <- createStyle(fgFill = "#F4B183")

wb <- createWorkbook()

addWorksheet(wb, "BEC_hier")

writeData(
  wb,
  "BEC_hier",
  BEC.hier.new
)

setColWidths(
  wb,
  sheet = "BEC_hier",
  cols = 1:ncol(BEC.hier.new),
  widths = "auto"
)
low_n_plots <- which(BEC.hier.new$n_plots < 7) + 1  # +1 for header row
plot_col <- which(names(BEC.hier.new) == "n_plots")
addStyle(
  wb,
  sheet = "BEC_hier",
  style = redFill,
  rows = low_n_plots,
  cols = plot_col,
  gridExpand = TRUE
)
singletons <- which(BEC.hier.new$n_siteunits_gen < 2) + 1  # +1 for header row
plot_col2 <- which(names(BEC.hier.new) == "n_siteunits_gen")
cols_to_style <- plot_col2:(plot_col2 + 5)

addStyle(
  wb,
  sheet = "BEC_hier",
  style = orgFill,
  rows = singletons,
  cols = cols_to_style,
  gridExpand = TRUE
)

saveWorkbook(
  wb,
  paste0("./review_materials/0_", anal.set, "_BECdb_table_for_edit.xlsx"),
  overwrite = TRUE
)

```
