---
title: "Build Associations for LMH77"
author: "William H MacKenzie"
date: today
format:
  typst:
    toc: false
    toc-depth: 1
    toc-title: Contents
    section-numbering: 1.1.1
    columns: 1
output-file: "CWH-Fd_HierarchyReport.pdf"
editor: source
execute:
  echo: false
  error: false
  warning: false
  message: false
  fig.width: 6
  fig.height: 4
  fig.align: 'center'
  fig.cap: true
  fig.pos: H
  out.width: '100%'
  dev: pdf
  fig.ext: pdf
  cache: false
  fig.retina: 2
  dpi: 600
  fig.asp: 1.5
  fig.path: "./figures/"
---

```{r}
## Choose analysis set and change name of output-file above
anal.set <- c("MH")# "CWH-wetfor" # "CWH"  "MH" "CWH-Fd" "Act" "Beachland" "Estuarine"
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(tidyverse)
require(DBI)
require(data.table)
require(cluster)
require(dendextend)
require(dynamicTreeCut)
require(gtable)
require(gtsummary)
require(colorspace)
require(openxlsx)
require(tictoc)
require(labdsv)
require(indicspecies)
require(Polychrome)
require(ggdendro)
require(purrr)
require(ggplotify)
require(grid)
require(data.tree)
library(stats)
library(overlapping)
require(gt)
require(vegan)
library(tidytext)
library(stringr)

set.seed(1279)
source("../BEC_R_working/_functions/_bec_dist.R")
source("../BEC_R_working/_functions/_bec_dist_matrix.R")
source("../BEC_R_working/_functions/_combine_taxa.R")
source("../BEC_R_working/_functions/_combine_taxa_strata.R")
source("../BEC_R_working/_functions/_create_su_vegdata.R")
source("../BEC_R_working/_functions/_create_analysis_vegsum.R")
source("../BEC_R_working/_functions/_TabletoTree.R")
source("../BEC_R_working/_functions/_TreetoTable.R")
source("../BEC_R_working/_functions/_add_vars.R")
source("../BEC_R_working/_functions/_do_pairwise.R")
source("../BEC_R_working/_functions/_create_diagnostic_veg.R")
source("../BEC_R_working/_functions/_return_similar_pairs.R")
source("../BEC_R_working/_functions/_read_sppmaster.R")
source("../BEC_R_working/_functions/_combined_su.R")
source("../BEC_R_working/_functions/_create_veg_sum.R")
source('../BEC_R_working/_functions/_create_veg_sum_all.R')
source('../BEC_R_working/_functions/_build_species_ordering.R')
source('../BEC_R_working/_functions/_format_veg_table.R')
source('../BEC_R_working/_functions/_format_veg_table_strata.R')
source('../BEC_R_working/_functions/_format_veg_table_nostrata.R')
source('../BEC_R_working/_functions/_encode_veg_sum.R')
source('../BEC_R_working/_functions/_create_dendro.R')
source('../BEC_R_working/_functions/_create_dendro_all.R')
source('../BEC_R_working/_functions/_create_dendro_bybgc.R')
source('../BEC_R_working/_functions/_draw_dendro_split.R')
source('../BEC_R_working/_functions/_cluster_groups.R')
source('../BEC_R_working/_functions/_TabletoTree.R')
source('../BEC_R_working/_functions/_build_species_ordering_hier.R')
source("../BEC_R_working/_functions/_cluster_by_spp_group.R")
source("../BEC_R_working/_functions/_baselga_diagnostics.R")

```

# Introduction

This script compares site unit summary vegetation data to identify associations and subassociations within the CoastArea using published sorts from LMH68 and LMH77. The analysis is based on a agglomerative clustering approach, which groups site units based on their vegetation characteristics. The process of correlation is iterative. Review of the site series for consistency is performed as part of the guide review process so the working root units are considered valid and not modified in this process. As part of correlation the first test the membership of zonal sites for proper designation in BGC mapping. On confirmation of these units, the Association is assigned. All hierarchical membership of site units is stored in the tblBEC_Hierarchy_Class table of the BECdb.

1.  The analysis is done by correlator temporary assigned analysis group which is assigned in the tblBEC_Hierarchy_Class. These analysis groups will largely reflect orders or classes. A complete set of site units relevant to a analysis group/order is a requisite for proper analysis.

2.  There maybe initial assigned associations (usually done through targeted analysis or expert decisions) such as subzone zonal ecosystems, non-forested associations, or 102 conceptual groups. These can be/should be applied a priori in the analysis. Subsequent testing of these should be carried out to verify.

3.  Quantitative analysis of site units with few plots, or low diagnostic potential may not be reliable. For this initial analysis, we include all site units but those with fewer than 6 plots or very low (\<25) diagnostic potential should be subsequently reviewed from proper membership.

4.  The dissimilarity matrix is run through a cluster algorithm and the resulting dendrogram is cut as dissimilarity values to assign initial hierachical groupings for all taxonomic levels. These units are assigned an algorithm generated unit number.

5.  Review materials from the quantitative analysis are produced with a focus on the Association. These materials are reviewed for consistency in vegetation and environment. Potential alternatives are identified from the dissimilarity matrix for units of questionable initial placement. Single units that do not match at the association level should be reviewed for distinctiveness and whether merging with other units should be enforced.

    - A vegetation summary workbook by site unit, on separate alliance sheets and colour themed by Association. The themeing is completed by macro in Excel.

    - A table of alternate association membership for site series based on similarity.

    - Edatopic and climate space comparisons for units

## Read in data

Vegetation data is read in from saved BECMaster.RDS file generated from the BECMaster cleaning scripts. A compiled SU table is build from all BGC_SU tables stored in the coast guides Vpro database. Taxonomy is read in from the species taxonomy database. A species lumping code table is read in from the Correlation2_Spp_lump.accdb database and the vegetation data is lumped using the combine_taxa function.

```{r load data, echo=FALSE}
#------------------------------------------------------------
# 1. Load vegetation data + taxon lifeform classifications
#------------------------------------------------------------
taxon.all <- read_sppmaster()
taxon.exotic <- taxon.all %>%
  filter(Native == "E") %>%
  dplyr::select(Code) %>%
  pull()

taxon.lifeform <- taxon.all %>%
  filter(Codetype %in% c("U", "X", "D")) %>%
  dplyr::select(Code, ScientificName, EnglishName, Lifeform) %>%
  distinct()

trees <- taxon.lifeform %>% 
  filter(Lifeform %in% c("1", "2")) %>% 
  pull(Code)

tree_seedlings <- taxon.lifeform %>% 
  filter(Lifeform %in% c("1", "2")) %>% 
  mutate(Code = paste0(Code, "D")) %>% 
  pull(Code)
veg.dat <- readRDS("./clean_data/Analysis_BECMaster_Veg.rds") %>% filter(!Species %in% taxon.exotic, !startsWith(Species, "POA_PRA"))
#------------------------------------------------------------
# 2. Read lump table and apply combine_taxa()
#------------------------------------------------------------

veglump <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BC_Correlation2_Vpro_2023/Correlation2_Spp_lump.accdb;"
)

lump <- dbReadTable(veglump, "CorrelationII_Lump")
dbDisconnect(veglump)

veg.dat2 <- combine_taxa(vegdata = veg.dat, lump, use.subtaxa = FALSE)%>% 
   filter(!Species %in% taxon.exotic)

#------------------------------------------------------------
# 3. Load SiteUnit data from two LMH databases
#------------------------------------------------------------

db <- "D:/BC_Correlation2_Vpro_2023/Sorts_published_LMH/LMH77_Coast_Sorts.accdb"
#db2 <- "D:/BC_Correlation2_Vpro_2023/Sorts_published_LMH/LMH68_HG_Sorts.accdb"
db <- "D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb"
su <- rbind(combined_su(db))#, combined_su(db2))

su <- su %>%
  filter(!grepl("X|unplaced|omit|support|nudum|Ww|!", SiteUnit)) %>%
  arrange(SiteUnit)
su2 <- su %>%
  mutate(SiteUnit = str_replace(SiteUnit, "[abc]$", ""))|> #,# ignore phases
         # SiteUnit = if_else(
         #   str_detect(SiteUnit, "101"),
         #   str_replace(SiteUnit, "\\.[123]$", ""), # remove zonal variations
         #   SiteUnit
         # )) %>%
    # # mutate( SU_Label = if_else(
    #        str_detect(SiteUnit.orig, "101"),
    #        str_replace(SiteUnit.orig, "\\.[123]$", ""), # remove zonal variations
    #        SiteUnit.orig))
          mutate(SU_Label = str_replace(SiteUnit.orig, "[abc]$", ""))# ignore phases
  

# Plot counts
n_plots <- su2 %>%
  group_by(SiteUnit) %>%
  summarise(n_plots = n_distinct(PlotNumber), .groups = "drop") 

su2 <- su2 # ignore phases
         # SiteSeries = if_else(
         #   str_detect(SiteUnit.orig, "101"),
         #   str_replace(SiteUnit.orig, "\\.[123]$", ""), # remove zonal variations
         #   SiteUnit.orig
         # ))

#------------------------------------------------------------
# 5. Load BECMaster environmental + edatopic tables
#------------------------------------------------------------

becmaster <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster/BECMaster_fixing.accdb;"
)

plot.env <- dbReadTable(becmaster, "BECMaster_fixing_Env")
plot.edatope <- dbReadTable(becmaster, "BECMaster_fixing_edatopic")

dbDisconnect(becmaster)

#------------------------------------------------------------
# 6. Write n_plots to BECdb and load hierarchy + edatopic
#------------------------------------------------------------

becdb <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v0.accdb;"
)

dbWriteTable(becdb, "n_plots", n_plots, append = FALSE, overwrite = TRUE,
  batch_rows = 1)

BGC_all <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(FSRegion, BGC_NoSpace)

BEC.hier <- dbReadTable(becdb, "tblBEC_Hierarchy_Class") %>%
  filter(SU_Label %in% su2$SU_Label) %>%
  mutate(SiteUnit = str_replace_all(SU_Label, c("/" = "_", " " = "")))

bgcv13 <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(BGC_NoSpace, SiteClassificationStatus)

edatopic <- dbReadTable(becdb, "Edatopic_v13_3") %>%
  mutate(
    rSMR = paste0("rSMR", substr(Edatopic, 2, 2)),
    SNR = substr(Edatopic, 1, 1)
  ) %>%
  filter(startsWith(Source, "BEC")) %>%
  dplyr::select(BGC, rSMR, SNR, SS_NoSpace, Edatopic) %>%
  distinct() %>%
  mutate(SiteUnit = gsub("/", "_", SS_NoSpace))

dbDisconnect(becdb)


##Use working spreadsheet instead of BECdb
#BEC.hier <- openxlsx::read.xlsx("D:/OneDrive/WNA_BGC/Association_Analysis/Coast_Hierarchy_Analysis/BECdb_Coast_Hierarchy_Working_ver2.xlsx") #|> 
  #filter(Analysis_Group %in% "Forest-Coast_MH")

# Keep only site units present in hierarchy table
su3 <- su2 %>%
  filter(SU_Label %in% BEC.hier$SU_Label)
su.anal <- su3 %>%
  left_join(BEC.hier, by = c("SiteUnit" = "SiteUnit", "SU_Label" = "SU_Label"))

#------------------------------------------------------------
# 7. Identify SiteUnits missing from hierarchy
#------------------------------------------------------------

missing_in_hierarchy <- su3 %>%
  distinct(SiteUnit) %>%
  anti_join(BEC.hier %>% distinct(SiteUnit), by = "SiteUnit") %>%
  mutate(flag = "in_su2_not_hierarchy")
missing_in_hierarchy
# fwrite(missing_in_hierarchy, "./missing_in_hierarchy.csv")


```

## Use previously set hierarchy units for analysis SU where assigned.

This code links Site Units to their BEC hierarchy classifications and determines the most appropriate level for analysis. After joining hierarchy metadata to the Site Unit table, the script walks down the predefined hierarchy—from broad classes to fine Site Units—to identify the first available non‑missing classification for each record. This resolved level becomes the analytic grouping (anal.test). Finally, the script assigns a working hierarchy label (SiteUnit.hier) based on a user‑selected grouping level (e.g., Association), falling back to the root Site Unit name when no Association designation has been provided. The result is a clean, consistent hierarchy assignment ready for downstream ecological analysis.

```{r choose level, echo=FALSE}
## Limit the analysis based on Analysis_Group column in the hierarchy table. 
#, "Deciduous-Coast", 

#------------------------------------------------------------
# 1. Define hierarchy (cleaned)
#------------------------------------------------------------

hierarchy <- c(
  "Class", "Order", "Suborder", "Alliance", "Suballiance",
  "Association","Working_Assoc",  "Subassociation", "generated_SubAss", "generated_Assoc", "generated_Alliance", "temp_Assoc_Code",
  "tempSubass_Code", "SiteUnit"
)

group.to.test <- "Association"## uses data for this field to initially group siteunits where applicablle


#------------------------------------------------------------
# 2. Prepare hierarchy table
#------------------------------------------------------------

#BEC.hier2 <- BEC.hier #%>%
 # dplyr::select(-SiteUnit)   # keep SS_Label + hierarchy columns

#------------------------------------------------------------
# 3. Join hierarchy onto su2 and reduce to anal.set
#------------------------------------------------------------
#anal.set = c("MH", "Bl", "CwHw", "Fd", "CWH-Fd", "CWH")  ## use this to find possible extra Order similar members
su.anal1 <- su.anal %>%
  #left_join(BEC.hier, by = c("SiteSeries" = "SU_Label")) %>%
  filter(Analysis_Group %in% anal.set) #%>%   filter(Order %in% Order.set)
  

#------------------------------------------------------------
# 4. Compute anal.test by walking down the hierarchy
#------------------------------------------------------------
#group.to.test <- "working_Assoc"
# vectorized helper: find first non-NA across hierarchy levels
first_non_na <- function(df, cols) {
  apply(df[cols], 1, function(x) {
    x_no_na <- x[!is.na(x)]
    if (length(x_no_na) == 0) NA_character_ else x_no_na[1]
  })
}

# compute anal.test
su.anal1 <- su.anal1 %>%
  mutate(
    anal.test = ifelse(
      !is.na(Analysis_Group),
      Analysis_Group,
      first_non_na(., hierarchy)
    )
  )

#------------------------------------------------------------
# 5. Compute SiteUnit.hier based on group.to.test
#------------------------------------------------------------

su.anal2 <- su.anal1 %>%
  mutate(
    SiteUnit.hier = coalesce(.data[[group.to.test]], SiteUnit)
  ) %>%
  dplyr::select(
    PlotNumber, SiteUnit, SU_Label, generated_Alliance, generated_Assoc, generated_SubAss,
    !!sym(group.to.test), anal.test, SiteUnit.hier
  ) %>%
  group_by(SiteUnit.hier) %>%# 6. remove any hierarchy units with fewer than 6 plots
  filter(n() >= 1 | !is.na(!!sym(group.to.test))) %>%
  ungroup()

```

## Add in summary data from adjacent juridictions for comparison

Other summary tables for comparison (e.g. USA) table can be added to the analysis for comparison in do.pairwise function.

```{r add extra summary units}
usa_units <- fread("./usa.summary.units/USA_prepped_units3.csv") %>% dplyr::select(-ScientificName) %>% 
   filter(!Species %in% taxon.exotic,!startsWith(Species, "POA_PRA")) %>% filter(startsWith(SiteUnit, "Q"))
usa_units <- usa_units %>%
  mutate(Species = recode(Species,
                          POLYGLY = "POLYPODI",
                          SELAWAL = "SELAGINE"))

#veg.dat2 <- combine_taxa(vegdata = usa_units, lump, use.subtaxa = FALSE)
#usa_units <- readRDS("./usa.summary.units/usa_assocs_vegdat.rds")
# dups <- veg.comb %>%
#   group_by(SiteUnit, Species) %>%
#   filter(n() > 1) %>%
#   arrange(SiteUnit, Species)
```

## Run pairwise vegetation similarity analysis

This code prepares Site Units for vegetation similarity analysis by assigning each plot to its working hierarchical unit and defining key indicator species and reduction rules. It then runs a pairwise vegetation comparison using do_pairwise(), producing similarity metrics for all Site Unit combinations. From these results, the script extracts a clean comparison table and generates a full BEC similarity distance matrix based on minimum similarity values. The resulting matrix is exported for downstream clustering, ordination, or classification work. One possible option here is to adjust pair-wise similarity downward where environmental differences are significant.

Potentially change the ksi and reduction to exponential factors where by default ksi is sqrt and reduced is \^1/4 rather than an arbitrary points reduction. For Order analysis it might be tree species that get ksi. Mosses would get reduced

```{r build pair.wise similarity matrix, echo=FALSE, message=FALSE, warning=FALSE}

#------------------------------------------------------------
# 1. Define indicator species and reduction settings
#------------------------------------------------------------

key.site.indicators <- c("NEPHCRI", "ABIELAS","LYSIAME", "CASSMER", "PHYLLODO", "SPHAGNUM")
#   "POPUTRI","GAULSHA","OPLOHOR","ELLIPYR","ATHYFIL",
#   "RUBUSPE","EQUIARV","GYMNDRY","VALESIT","LUETPEC",
#   "POLYMUN","STRUSPI","RHODGRO","EMPENIG","PICESIT"
# )

reduced.exceptions <- c("SPHAGNUM","CLADONIA","CLADINA","RACOMITR","MNIUM")
reduced.lifeforms  <- c(9, 10, 11)

#------------------------------------------------------------
# 2. Prepare SiteUnit table for analysis
#------------------------------------------------------------
# SiteUnit.hier is replacing site unit with a previously determined Association group
su_working <- su.anal2 #%>%   rename(SiteUnit.base = SiteUnit, SiteUnit = SiteUnit.hier)

#------------------------------------------------------------
# 3. Run pairwise vegetation similarity analysis
#------------------------------------------------------------
pairwise.params <- list(
  su                = su_working,
  minconstancy          = 60,
  noiseconstancy        = 10,
  minplots              = 1,
  minor                 = 1,
  use.ksi               = FALSE,
  ksi   = key.site.indicators,
  ksi.value             = 1.5, ## change this to apply to creation of diagnostic potential as a factor ^1/2 by default
  reduce.lifeform       = TRUE,
  reduced.lifeforms     = reduced.lifeforms,
  reduction             = 0.1, ## keep this as a reduction of actual final diagnostic points
  reduced.exceptions    = reduced.exceptions,
  add_vegsum_table      = NULL
)

vegsum.pairs <- do.call(
  do_pairwise,
  c(list(veg.dat2), pairwise.params)
)

# Start with everything except `su`
print.params <- pairwise.params[setdiff(names(pairwise.params), "su")]

# Remove ksi if use.ksi == FALSE
if (!pairwise.params$use.ksi) {
  print.params <- print.params[setdiff(names(print.params), c("ksi", "ksi.value"))]
}

# Remove reduction-related fields if reduce.lifeform == FALSE
if (!pairwise.params$reduce.lifeform) {
  print.params <- print.params[setdiff(names(print.params),
                                       c("reduced.lifeforms", "reduction", "reduced.exceptions"))]
}
library(tibble)
library(knitr)

param.table <- tibble(
  parameter = names(print.params),
  value = sapply(print.params, function(x) paste(x, collapse = ", "))
)

param.table %>%
  gt() %>%
  tab_header(title = "Parameter Settings for do_pairwise()") %>%
  tab_options(
    table.font.size = px(8),
    table.border.top.width = px(1),
    table.border.bottom.width = px(1)
  )



#------------------------------------------------------------
# 4. Extract comparison table
#------------------------------------------------------------
fwrite(vegsum.pairs, "./outputs/vegsum_pairs.csv")

### return unit species information
vegsum.pairs.spp <- vegsum.pairs |> dplyr::select(Unit1, Unit2, Species, ScientificName, MeanCov.unit1, Constancy.unit1, MeanCov.unit2, Constancy.unit2, diff.pts.unit1, diff.pts.unit2)

### return distinct unit comparisons

unit.compare.temp <- vegsum.pairs %>%
  dplyr::select(
    Unit1, Unit2, diff.ratio.unit1, diff.ratio.unit2, nplots.unit1, nplots.unit2, unit.diag.sum.unit1, unit.diag.sum.unit2) %>%
  distinct(Unit1, Unit2, .keep_all = TRUE)


focal.unit.compare <- unit.compare.temp %>% 
    dplyr::select(
      Unit1 = Unit2,
      Unit2 = Unit1,
      diff.ratio.unit1 = diff.ratio.unit2,
      diff.ratio.unit2 = diff.ratio.unit1,
      nplots.unit1 = nplots.unit2,
      nplots.unit2 = nplots.unit1,
      unit.diag.sum.unit1 = unit.diag.sum.unit2,
      unit.diag.sum.unit2 = unit.diag.sum.unit1
      ) |>  rbind(unit.compare.temp) |> 
  distinct() |> mutate(BEC.sim.min = pmin(diff.ratio.unit1, diff.ratio.unit2),
                      BEC.sim.max = pmax(diff.ratio.unit1, diff.ratio.unit2))


#fwrite(focal.unit.compare , "./outputs/focal_unit_compare.csv")


source("../BEC_R_working/_functions/_build_bec_distance_matrix.R")

result <- build_bec_distance_matrix(
  vegsum_pairs = vegsum.pairs,
  su_anal = su.anal2,
  preclass_field = "Association",
  use_preclass = FALSE
)

unit.compare2 <- result$unit_compare
bec.dist <- result$bec_dist


# openxlsx::write.xlsx(
#   x = bec.dist,
#   file = "./clean_tabs/Coast_Forest_Distance_matrix.xlsx",
#   rowNames = TRUE
# )
```

## Cluster analysis and dendrogram of all good site series and merged site units

The agglomerative clustering algorithm AGNES successively groups site units based on dissimilarity from the bottom up (most similar units get grouped first). Because units or initial clusters may have near equal dissimilarity to neighbours in two directions, the algorithm may not be able to maintain an exact ordering of dissimilarity. We evaluate the cluster analysis using two metrics. 1) The cophenetic correlation coefficient (CCC) evaluate how well a hierarchical clustering dendrogram preserves the original pairwise distances between observations. The closer CCC is to 1, the better it represents the original distance matrix. Lower numbers indicate that some branchlets of the dendrogram can be associated with more than one branch. 2) The agglomerative coefficient (AC) is a metric that quantifies the strength of the clustering structure. The closer AC is to 1 indicates strong clustering structure and the observations are tightly grouped by low dissimilarity. This will be affected by the breadth/size of the analysis where a larger number of site units will tend to have lower AC values.

We chose to use the clustering method that maximizes the CCC. We tested a number of methods: complete, average, single, ward, weighted, and flexible (at a wide range of settings). The "average" method consistently produced the highest cophenetic values so we applied this as the standard method for clustering.

In the dendrograms, the red line represents the minimum threshold for separating site units into separate site series. This is set at 10% dissimilarity, but in some BGCs we separate some site series to as low as 7% dissimilarity. The green line represents the preliminary dissimilarity threshold for an association. The initial threshold for an association is set at 20% (80% similar) based on review.

```{r }
#| label: fig-first-cluster
#| fig-cap: ("Cluster of Site Units (including initial predefined groups)")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12
source("../BEC_R_working/_functions/_bec_dist_matrix.R")
dendro_hc <- create_dendro_all(unit.compare2, threshold.low = .45)
```

### Draw cluster diagram subsets from tree split at set dissimilarity level

dissimilarity of 17-20% is an approximate threshold for association differentiation.

```{r split first cluster}
source("../BEC_R_working/_functions/_bec_dist_matrix.R")
singles <- draw_dendro_split(unit.compare.tbl = unit.compare2, cut.level=.5, show.singles = TRUE)
```
```{r, echo=FALSE}
#| label: tab-too-few-plots
#| tab-cap: "Site Units with Fewer than 6 Plots"
#| tbl-cap-location: top
#| warning: false
#| tab-align: "left"

# Identify Site Units with fewer than 8 plots
ss_too.few <- unit.compare %>%
  dplyr::select(Unit1, nplots.unit1) %>%
  filter(nplots.unit1 < 6 & nplots.unit1>0) %>%
  distinct() %>%
  rename(
    "SiteUnit"      = Unit1,
    "Number of Plots" = nplots.unit1)

# Count unique Site Units for dynamic column label
num_units <- ss_too.few %>%
  pull(SiteUnit) %>%
  n_distinct()

col_label <- paste0("Site Unit (", num_units, ")")

# Build GT table
gt::gt(ss_too.few) %>%
  gt::fmt_number(columns = "Number of Plots", decimals = 0) %>%
  gt::cols_label(SiteUnit = col_label) %>%
  gt::tab_options(table.font.size = 8)

```

```{r,  echo=FALSE}
#| label: tab-low-diagnostic
#| tbl-cap-location: top
#| tbl-cap: "Site Units with Low Diagnostic Potential"
#| warning: false
#| tbl-align: "left"

ss_low.diag <- unit.compare %>%
  dplyr::select(Unit1, unit.diag.sum.unit1)  %>% 
  filter(unit.diag.sum.unit1 < 24 & unit.diag.sum.unit1 >0) %>% 
  rename("Diagnostic Potential" = unit.diag.sum.unit1, "SiteUnit" = Unit1) %>%
  distinct()

num_units <- ss_low.diag %>% 
  pull(SiteUnit) %>% 
  n_distinct()
col_label <- paste0("Site Unit (", num_units, ")")

  gt::gt(ss_low.diag) %>%
  gt::fmt_number(columns = "Diagnostic Potential", decimals = 0) %>%
  gt::tab_options(table.font.size = 8) %>%
  gt::cols_label(SiteUnit = col_label) %>% 
  gt::tab_options(table.font.size = 8)
```

### Create initial hierarchy from clusters to group and write into BECdb

In the BECdb query to write the initial units in the the tempAssoc_Code field of tblBECHierarchy_Class table for subsequent use in analytics Also return alternate possible site unit matches for review by returning other units with similarity that falls below threshold.

```{r, echo=FALSE, results= 'hide'}
#| label: fig-second-cluster
#| fig-cap: ("Cluster of Site Units (including initial predefined groups)")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12
source("../BEC_R_working/_functions/_bec_dist_matrix.R")

fig.show='hide'
#--------- do for associations

dendro_hc <- create_dendro_all(unit.compare2, output = FALSE)

clusters <- cutree(dendro_hc, h = 0.10)
cluster_subass<- data.table(
  SiteUnit = names(clusters),
  generated_SubAss= paste0(anal.set, "_generated_Subass_", clusters)
)
clusters <- cutree(dendro_hc, h = 0.20)
cluster_assoc <- data.table(
  SiteUnit = names(clusters),
  generated_Assoc = paste0(anal.set,"_generated_Assoc_", clusters)
)
  clusters <- cutree(dendro_hc, h = 0.30)
cluster_suball <- data.table(
  SiteUnit = names(clusters),
  generated_SubAll = paste0(anal.set,"_generated_Suball_", clusters)
)
clusters <- cutree(dendro_hc, h = 0.40)
cluster_alli <- data.table(
  SiteUnit = names(clusters),
  generated_Alliance = paste0(anal.set,"_generated_Alli_", clusters)
)

clusters <- cutree(dendro_hc, h = 0.60)
cluster_subord <- data.table(
  SiteUnit = names(clusters),
  generated_SubOrd= paste0(anal.set,"_generated_SubOrd_", clusters)
)
clusters <- cutree(dendro_hc, h = 0.80)
cluster_ord <- data.table(
  SiteUnit = names(clusters),
  generated_Order= paste0(anal.set,"_generated_Order_", clusters)
)
clusters <- cutree(dendro_hc, h = 0.99)
cluster_class <- data.table(
  SiteUnit = names(clusters),
  generated_Class= paste0(anal.set,"_generated_Class_", clusters)
)

# Combine
cluster_table <- Reduce(
function(x, y) merge(x, y, by = "SiteUnit", all = TRUE),
list(
cluster_subass,
cluster_assoc,
cluster_suball,
cluster_alli,
cluster_subord,
cluster_ord,
cluster_class)
)

cluster_table
## merge back to retrieve SiteUnit for all and the coalesce.
orig.SU <- su.anal1 %>% dplyr::select(SU_Label, SiteUnit, working_Alliance, working_association, working_subass) %>% distinct
generated.hier <- cluster_table %>%
  left_join(orig.SU, by = c("SiteUnit" = "SiteUnit")) |> group_by(working_association) |> add_count(working_association) |> select(n, everything())
fwrite(generated.hier, paste0("./hierarchy/", anal.set, "_generated_hierclasses.csv"))


## write into BECdb a working file to update the tempAssoc field
becdb <- dbConnect(
  odbc::odbc(),
  .connection_string =
    "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
     DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v0.accdb;"
)
dbWriteTable(
  becdb,
  paste0("generated_hier_", anal.set),
  generated.hier, append=FALSE,
  overwrite = TRUE,
  batch_rows = 1
)
dbDisconnect(becdb)

#xx <- cluster_update_assoc %>% filter(SS_Label == "CWHds1_101a")
```

## Review placement of site units with more than one possible pairing at Association level

See excel table for Focal units with more than one similar unit Highlights other similar units that are not in the same association for review This will look for unit pairs that fall below a similarity threshold and then identify where they have not been put into the same Association

```{r create alternate su matches table}
source("../BEC_R_working/_functions/_export_similar_units.R")
similar_units <- export_similar_units(
  unit.compare = unit.compare2,
  generated.hier = generated.hier,
  anal.set = anal.set,
  same.threshold = 0.7
)
#check that equal number of pair comparisons
#similar_counts <- similar_units %>%
#   count(FocalUnit, name = "n_similar_units") %>%
#   arrange(desc(n_similar_units))
# 
# similar_counts
```

See excel vegsummary table for each Focal unit compared to all similar units for review Headers with n_plots, similarity metric, and hierarchy placement (Subass, Assoc, Alliance)

```{r summary table of similar units}

source("../BEC_R_working/_functions/_create_focalunit_summary_workbook.R")

create_focalunit_summary_workbook(
  su.anal = su.anal1,
  similar.units = similar_units,
  veg.dat = veg.dat2,
  taxon.lifeform = taxon.lifeform,
    anal.set = anal.set
)

```



## Constancy and cover tables

Generate vegetation summary for all site units by generated alliance colour-themed by generated association and subassociation for review.

Each sheet has all site units for a proposed alliance on a single sheet grouped, labelled, and colour-themed by generated Association  to help review and editing of membership. This will be used to review the proposed alliances and make decisions about which site units should be merged into associations, which should be separated into new associations, and which should be left as singletons. The vegetation summary tables will also be used to identify potential diagnostic species for each alliance and to compare the vegetation composition of different alliances.

```{r summary table association within alliances}
# compute anal.test
# becdb <- dbConnect(
#   odbc::odbc(),
#   .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v0.accdb;"
# )
# BEC.hier.temp <- dbReadTable(becdb, "tblBEC_Hierarchy_Class") #%>% dplyr::select(-SiteUnit)
# dbDisconnect(becdb)

 group.to.test = "generated_Alliance"
BEC.hier.temp <- generated.hier |>
  filter(!is.na(.data[[group.to.test]]))

su.anal.new <- su.anal1 %>% dplyr::select(PlotNumber, SU_Label) %>% left_join(BEC.hier.temp, by = c("SU_Label" = "SU_Label")) %>% #rename(SiteUnit = SiteSeries) %>%
  mutate(SiteUnit = str_replace(SiteUnit, "[abc]$", ""),# ignore phases
         SiteUnit = if_else(
           str_detect(SiteUnit, "101"),
           str_replace(SiteUnit, "\\.[123]$", ""), # remove zonal variations
           SiteUnit
         ))

source("../BEC_R_working/_functions/_create_veg_summary_workbook.R")

create_veg_summary_workbook(
  su.anal = su.anal.new,
  veg.dat = veg.dat2,
  taxon.lifeform = taxon.lifeform,  
  generated.hier = BEC.hier.temp,
  group.to.test = group.to.test,
  anal.set = anal.set
)

```


## Export table from generated values and BECdb table tbl_BEC_Hierarchy_Class

Edits to membership can be made to the working units in the spreadsheet for incorporation in the assessment scripts.

```{r}
orig.SU <- su.anal1 %>% dplyr::select(SU_Label, SiteUnit, working_Alliance, working_association, working_subass) %>% distinct
generated.hier <- cluster_table %>%
  left_join(orig.SU, by = c("SiteUnit" = "SiteUnit")) |> group_by(working_association) |> add_count(working_association) |> select(n_su_in_assoc = n, everything())

# Keep SiteUnit plus columns that are not already in generated.hier
cols.keep <- c("SiteUnit", setdiff(names(BEC.hier), names(generated.hier)))
BEC.hier.join <- BEC.hier %>% select(all_of(cols.keep))
BEC.hier.new <- left_join(
  generated.hier,
  BEC.hier.join,
  by = "SiteUnit"
)%>% select(all_of(names(BEC.hier)))

openxlsx::write.xlsx(
  x = BEC.hier.new,
  file = paste0("./review_materials/", anal.set, "_BECdb_table_for_edit.xlsx"),
  rowNames = FALSE )



## write into BECdb a working file to update the tempAssoc field
becdb <- dbConnect(
  odbc::odbc(),
  .connection_string =
    "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
     DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v0.accdb;"
)
dbWriteTable(
  becdb,
  paste0(anal.set,"_generated_hier"),
  BEC.hier.new, append=FALSE,
  overwrite = TRUE,
  batch_rows = 1
)
dbDisconnect(becdb)
```

