---
title: "Assess proposed hierarchy for LMH77"
author: "William H MacKenzie"
date: today
format:
  typst:
    toc: false
    toc-depth: 1
    toc-title: Contents
    section-numbering: 1.1.1
    columns: 1
    mainfont: "Calibri"
    fontsize: 10pt
editor: source
execute:
  echo: false
  error: false
  warning: false
  message: false
  fig.width: 6
  fig.height: 4
  fig.align: 'center'
  fig.cap: true
  fig.pos: H
  out.width: '100%'
  dev: pdf
  fig.ext: pdf
  cache: false
  fig.retina: 2
  dpi: 600
  fig.asp: 1.5
  fig.path: "./figures/"
output-file: "FdPy_Generated_Hierarchy.pdf"
---

```{r}
anal.set <- c("FdPy")# "CWH-102" "CWH-HwBa" "CWH-HwBa-rich"  "CWH-HwFd" "CWH-HwFd-rich" "coast-wetfor"  "MH" "Act"

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(tidyverse)
require(DBI)
require(data.table)
require(cluster)
require(dendextend)
require(dynamicTreeCut)
require(gtable)
require(colorspace)
require(gtsummary)
require(openxlsx)
require(tictoc)
require(labdsv)
require(indicspecies)
require(Polychrome)
require(ggdendro)
require(purrr)
require(ggplotify)
require(grid)
library(stringr)
require(vegan)
require(betapart)
library(plotly)
require(gt)
gt <- gt::gt
library(writexl)
set.seed(1279)
source("../BEC_R_working/_functions/_bec_dist.R")
source("../BEC_R_working/_functions/_bec_dist_matrix.R")
source("../BEC_R_working/_functions/_combine_taxa.R")
source("../BEC_R_working/_functions/_combine_taxa_strata.R")
source("../BEC_R_working/_functions/_create_su_vegdata.R")
source("../BEC_R_working/_functions/_create_analysis_vegsum.R")
source("../BEC_R_working/_functions/_TabletoTree.R")
source("../BEC_R_working/_functions/_TreetoTable.R")
source("../BEC_R_working/_functions/_add_vars.R")
source("../BEC_R_working/_functions/_do_pairwise.R")
source("../BEC_R_working/_functions/_create_diagnostic_veg.R")
source("../BEC_R_working/_functions/_return_similar_pairs.R")
source("../BEC_R_working/_functions/_read_sppmaster.R")
source("../BEC_R_working/_functions/_combined_su.R")
source("../BEC_R_working/_functions/_create_veg_sum.R")
source('../BEC_R_working/_functions/_create_veg_sum_all.R')
source('../BEC_R_working/_functions/_build_species_ordering.R')
source('../BEC_R_working/_functions/_format_veg_table.R')
source('../BEC_R_working/_functions/_format_veg_table_strata.R')
source('../BEC_R_working/_functions/_format_veg_table_nostrata.R')
source('../BEC_R_working/_functions/_encode_veg_sum.R')
source('../BEC_R_working/_functions/_create_dendro.R')
source('../BEC_R_working/_functions/_create_dendro_all.R')
source('../BEC_R_working/_functions/_create_dendro_bybgc.R')
source('../BEC_R_working/_functions/_draw_dendro_split.R')
source('../BEC_R_working/_functions/_cluster_groups.R')
source('../BEC_R_working/_functions/_TabletoTree.R')
source('../BEC_R_working/_functions/_build_species_ordering_hier.R')
source("../BEC_R_working/_functions/_cluster_by_spp_group.R")
source("../BEC_R_working/_functions/_baselga_diagnostics.R")
data.path = "D:/OneDrive - Government of BC/GitHub_data"
```


```{r load data, echo=FALSE}
#------------------------------------------------------------
# 1. Load vegetation data + taxon lifeform classifications
#------------------------------------------------------------
taxon.all <- read_sppmaster()
taxon.exotic <- taxon.all %>%
  filter(Native == "E") %>%
  dplyr::select(Code) %>%
  pull()

taxon.lifeform <- taxon.all %>%
  filter(Codetype %in% c("U", "X", "D")) %>%
  dplyr::select(Code, ScientificName, EnglishName, Lifeform) %>%
  distinct()

trees <- taxon.lifeform %>% 
  filter(Lifeform %in% c("1", "2")) %>% 
  pull(Code)

tree_seedlings <- taxon.lifeform %>% 
  filter(Lifeform %in% c("1", "2")) %>% 
  mutate(Code = paste0(Code, "D")) %>% 
  pull(Code)
veg.dat <- readRDS("./clean_data/Analysis_BECMaster_Veg.rds") %>% filter(!Species %in% taxon.exotic, !startsWith(Species, "POA_PRA"))
#------------------------------------------------------------
# 2. Read lump table and apply combine_taxa()
#------------------------------------------------------------

veglump <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BC_Correlation2_Vpro_2023/Correlation2_Spp_lump.accdb;"
)

lump <- dbReadTable(veglump, "CorrelationII_Lump")
dbDisconnect(veglump)

veg.dat2 <- combine_taxa(vegdata = veg.dat, lump, use.subtaxa = FALSE)%>% 
   filter(!Species %in% taxon.exotic)

#------------------------------------------------------------
# 3. Load SiteUnit data from two LMH databases
#------------------------------------------------------------

db <- "D:/BC_Correlation2_Vpro_2023/Sorts_published_LMH/LMH77_Coast_Sorts.accdb"
#db2 <- "D:/BC_Correlation2_Vpro_2023/Sorts_published_LMH/LMH68_HG_Sorts.accdb"
db <- "D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb"
su <- rbind(combined_su(db))#, combined_su(db2))

su <- su %>%
  filter(!grepl("X|unplaced|omit|support|nudum|Ww|!", SiteUnit)) %>%
  arrange(SiteUnit)
su2 <- su %>%
  mutate(SiteUnit = str_replace(SiteUnit, "[abc]$", ""),# ignore phases
         # SiteUnit = if_else(
         #   str_detect(SiteUnit, "101"),
         #   str_replace(SiteUnit, "\\.[123]$", ""), # remove zonal variations
           SiteUnit
         ) %>%
    mutate( SU_Label = str_replace(SiteUnit.orig, "[abc]$", ""))#if_else(
         #   str_detect(SiteUnit.orig, "101"),
         #   str_replace(SiteUnit.orig, "\\.[123]$", ""), # remove zonal variations
         #   SiteUnit.orig
         # )) |> mutate(SU_Label = # ignore phases
  
#------------------------------------------------------------
# 5. Load BECMaster environmental + edatopic tables
#------------------------------------------------------------

becmaster <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster/BECMaster_fixing.accdb;"
)
plot.env <- dbReadTable(becmaster, "BECMaster_fixing_Env")
plot.edatope <- dbReadTable(becmaster, "BECMaster_fixing_edatopic")
plot.info <- dbReadTable(becmaster, "BECMaster_fixing_Admin")
dbDisconnect(becmaster)

#------------------------------------------------------------
# 6. Write n_plots to BECdb and load hierarchy + edatopic
#------------------------------------------------------------

becdb <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v1.accdb;"
)

BGC_all <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(FSRegion, BGC_NoSpace)

BEC.hier <- dbReadTable(becdb, "tblBEC_Hierarchy_Class") %>%
  filter(SU_Label %in% su2$SU_Label) %>%
  mutate(SiteUnit = str_replace_all(SU_Label, c("/" = "_", " " = "")))

bgcv13 <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(BGC_NoSpace, SiteClassificationStatus)

edatopic <- dbReadTable(becdb, "Edatopic_v13_3") %>%
  mutate(
    rSMR = paste0("rSMR", substr(Edatopic, 2, 2)),
    SNR = substr(Edatopic, 1, 1)
  ) %>%
  filter(startsWith(Source, "BEC")) %>%
  dplyr::select(BGC, rSMR, SNR, SS_NoSpace, Edatopic) %>%
  distinct() %>%
  mutate(SiteUnit = gsub("/", "_", SS_NoSpace))

dbDisconnect(becdb)

##Export working BECdb for anal.set
BECdb.anal <- BEC.hier |> filter(Analysis_Group %in% anal.set)

BECdb.anal <- BEC.hier |>
  filter(Analysis_Group %in% anal.set)

write_xlsx(
  BECdb.anal,
  file.path("./review_materials", paste0(anal.set,"_BECdb_for_updates.xlsx"))
)

##Use working spreadsheet instead of BECdb
#BEC.hier <- openxlsx::read.xlsx("D:/OneDrive/WNA_BGC/Association_Analysis/Coast_Hierarchy_Analysis/BECdb_Coast_Hierarchy_Working_ver3.xlsx") #|> 
  #filter(Analysis_Group %in% "Forest-Coast_MH")

# Keep only site units present in hierarchy table
su2 <- su2 %>%
  filter(SU_Label %in% BECdb.anal$SU_Label)
su.missed <- su2 |> filter(!SU_Label %in% BECdb.anal$SU_Label)
su.class <- su2 %>%
  left_join(BECdb.anal, by = c("SiteUnit" = "SiteUnit", "SU_Label" = "SU_Label"))

su.anal <- su.class |> filter(n_plots >1, Analysis_Group %in% anal.set)|> arrange(PlotNumber) 
```

```{r prep env data}
plot.env2 <- plot.env %>% mutate(lon = ifelse(Longitude<0, Longitude, 0-Longitude)) %>% mutate(Date = as_date(Date)) |> 
  filter(PlotNumber %in% su.anal$PlotNumber) |>
  dplyr::select(PlotNumber, lon, lat = Latitude, elev = Elevation, special = Exposure1, rSMR = MoistureRegime, rSNR = NutrientRegime, slope = SlopeGradient, aspect = Aspect ) 
 plot.info <- plot.info |> dplyr::select(Plot, BECSiteUnit, UserSiteUnit, HumusThickness, BGC=GIS_BGC)  |> filter(Plot %in% su.anal$PlotNumber)
plot.env2 <- left_join(plot.env2, plot.info, by = c("PlotNumber" = "Plot")) |> filter(!is.na (rSMR))


## extract climate data for plots from climr
require(climr)
plot.locs <- plot.env2 %>% dplyr::select(id=PlotNumber, BGC, lon,lat,elev) |> drop_na(lat)
plot.climate <- climr::downscale(setDT(plot.locs),
                              vars = list_vars(set = c("Annual", "Seasonal", "Monthly")),
                              return_refperiod = TRUE,
                              db_option = "local",
                              cache = TRUE)
## calculate aSMR
plot.env3 <- plot.climate |> select(PlotNumber = id, everything()) |> left_join(plot.env2) |> 
  mutate(
    moist = rowSums(
      across(c(PPT_05, PPT_06, PPT_07, PPT_08, PPT_09), ~ .x >= 90),
      na.rm = TRUE
    ),
    moist = ifelse(moist > 4, 1, 0),
    ratio = moist / 5,
    multiplier = 1 + ratio,
    CMD = CMD_an / multiplier
  ) |> mutate(PPT.dorm = PPT_at + PPT_wt) |> 
  mutate(Def = ifelse(CMD >=350, 350 - PPT.dorm, CMD)) |> 
  mutate(CMD.total = CMD + Def) |> 
  mutate(CMD.site = ifelse(slope > 35 & (aspect >180 & aspect <270), CMD.total *1.25,
                           ifelse(slope> 35 & (aspect>315 | aspect <45), CMD.total *0.85, CMD.total))) |> arrange(PlotNumber) |> 
  mutate(
    rSMR.clean = sub("^.*?(\\d+).*$", "\\1", rSMR)) |> 
  mutate(rSMR.clean = as.numeric(rSMR.clean),
  SNR.clean = recode(
    substr(trimws(rSNR), 1, 1),
    "A" = 1L,
    "B" = 2L,
    "C" = 3L,
    "D" = 4L,
    "E" = 5L,
    "F" = 6L,
    .default = NA_integer_
  ))
  ####To adjust in zones with prolonged snowpack remove CMD from early growing season months
## BGCs with or below heavy snow packs
# BGC_Special <- fread("./inputs/BGC_Special.csv", data.table = FALSE)
# allDat2 <- left_join(allDat2,BGC_Special, by = c("BGC" = "BGC"))
# allDat2 <- allDat2 %>% mutate(CMD = ifelse(is.na(Special), CMD, CMD_08)) %>% select(-Special, -CMD_08) %>% drop_na()
#rules <- read.csv("./inputs/aSMR_growingseason_Rules_HalfStep_v13_13Sept2025.csv")
rules <- read.csv("./inputs/aSMR_Rules_HalfStep_v13_23June2025.csv") |>
  mutate(
    rSMR.rule = sub("^rSMR", "", rSMR)) |> dplyr::select(-rSMR)

plot_asmr <- plot.env3 %>% select(PlotNumber, rSMR, CMD.site) |> 
  left_join(rules, by = c("rSMR" = "rSMR.rule")) %>%
  filter(CMD.site <= CMD) %>%
  group_by(PlotNumber) %>%  # ensure row-wise logic if needed
  slice_max(aSMR, with_ties = FALSE) %>%
  ungroup() %>% select(PlotNumber, aSMR)

plot.env3 <- left_join(plot.env3, plot_asmr)

plot.env3 <- plot.env3 |>
  left_join(
    su.anal, 
    by = "PlotNumber")
 
plot.env.vars <- plot.env3 |>   select(-SiteUnit, -working_association, -working_alliance) |> 
   select(PlotNumber, PAS_an,CMD.site, MSP, aSMR , SNR = SNR.clean, Tmin = Tmin_an, AHM,FFP, DD5 = DD5_an, DD18_sm ) |> arrange(PlotNumber)
 keep.rows <- complete.cases(plot.env.vars)

  su.env <- plot.env3 |> rename(assoc = working_association, alliance = working_alliance) |> 
  filter(!is.na(alliance)) |> arrange(SiteUnit) |> 
  group_by(SiteUnit, assoc, alliance) |>
    summarise(
  PAS_an = mean(PAS_an, na.rm = TRUE),
  CMD.site = mean(CMD.site, na.rm = TRUE),
  MSP   = mean(MSP, na.rm = TRUE),
  aSMR = mean(aSMR, na.rm = TRUE),
  SNR = mean(SNR.clean, na.rm = TRUE),
  AHM = mean(AHM, na.rm = TRUE),
  #lat = mean(lat, na.rm = TRUE),
  FFP   = mean(FFP, na.rm = TRUE),
  DD5 = mean(DD5_an, na.rm = TRUE),
  DD18_sm = mean(DD18_sm, na.rm = TRUE),
  Tmin = mean(Tmin_an, na.rm = TRUE),
  .groups = "drop"
)

#     summarise(
#   across(
#     where(is.numeric) & !c(lat, lon),
#     ~ mean(.x, na.rm = TRUE)
#   ),
#   .groups = "drop"
# )

# ==============================================================================
# Climate variables for PCA
#==============================================================================
  env.vars <- su.env |>
  select(-assoc, -alliance) 

# Remove rows with missing values
keep.rows <- complete.cases(env.vars)

env.vars <- env.vars[keep.rows, ]

# Remove constant variables if any
env.vars.filt <- env.vars |>
  select(-SiteUnit) |>
  dplyr::select(
    where(~ is.numeric(.x) && sd(.x, na.rm = TRUE) > 0)
  )

keep.rows <- complete.cases(env.vars.filt)

#su.env <- env.vars[keep.rows, ]
su.env <- su.env |>
  dplyr::select(
    SiteUnit, assoc, alliance,
    where(~ is.numeric(.x) && sd(.x, na.rm = TRUE) > 0)
  )

```

# Assess and review the separation associations and alliances

All comparisons use the site unit summary data as the basis for comparison and not plot level data.

An initial generated hierarchical classification from Build_Hierarchy_working can be used but better is to have done initial assessments and created intial working\_ hierarchy levels for application in this script.

This script is to be used iteratively to assess current working unit membership is supported or if site unit membership should be reassigned.

Subsequent analysis here will substantiate the existing working units or suggest where association membership in alliances should be adjusted to maximize indicator species separation and environmental space.

Core analyses are:

1.  NMDS ordination of plant communities showing Alliance, Association and SiteUnits with envfit joint plot for edaphic and select climatic parameters.

2.  Environment separation: PCA of units, boxplot showing position of site units on PCA axes grouped by higher taxons

3.  PERMANOVA (pairwise adonis) with betadispersion metrics to assess vegetation differences between units.

4.  Indicator species analysis identifying species which are statistically associated with units.

5.  Vpro style vegetation summary report by working_alliance organized by working hierarchical taxa

Together these answer:

Are alliances and associations compositionally distinct from one another? How internally consistent are they? How distinct are they? Which species define them? Which alliances overlap? What environmental gradients drive differences?

```{r read in SU data, message=FALSE}
# anal.set = "MH"
# 
# 
# combine.grps <- c("CWH-wetfor")#, "CWH-Fd")
# 
# su.class <- su.class |>
#   dplyr::mutate(
#     Analysis_Group = ifelse(
#       Analysis_Group %in% combine.grps,
#       anal.set,
#       Analysis_Group
#     )
 # ) 





# alliance.remove = c("Ss_shoreline", "CWH_Fraser_Fh", "CDF_rich", "CWHds_102", "CWHws_102", "CWHvmN_102", "CWHvmS_102", "CWH_Fh_south", "CWHms_102", "CWHw_terraces", "CWHvh_102", "CWHv_102")
# su.anal <- su.anal |> filter(!working_alliance %in% alliance.remove)
#   
su.anal2 <- su.anal |> dplyr::select(PlotNumber, SiteUnit, assoc = working_association, alliance = working_alliance) |> 
  mutate(alliance = str_remove(alliance, "^coast_generated_"))  |> distinct()
unique.su <- su.anal2 |> dplyr::select(SiteUnit) |> distinct()|> arrange(SiteUnit)
unique.assocs2 <- su.anal2 |> dplyr::select(assoc) |> distinct()
unique.alliances2 <- su.anal2 |> dplyr::select(alliance) |> distinct()
su.assocs <- su.anal2 |> dplyr::select(SiteUnit, assoc) |> distinct()|> arrange(SiteUnit)
su.alliances <- su.anal2 |> dplyr::select(SiteUnit, alliance) |> distinct()|> arrange(SiteUnit)

veg.anal<- veg.dat2 |> filter(PlotNumber %in% su.class$PlotNumber) |> dplyr::select(PlotNumber, Species, Cover) |>
  arrange(PlotNumber) #|> matrify()

reduced.exceptions <- c("SPHAGNUM","CLADONIA","CLADINA","RACOMITR","MNIUM")
reduced.lifeforms  <- c(9, 10, 11)

pairwise.params <- list(
  su                = su.anal,
  su.field =  "SiteUnit",
  minconstancy          = 50,
  noiseconstancy        = 10,
  minplots              = 1,
  minor                 = 5,
  use.ksi               = FALSE,
  ksi   = NULL,
  ksi.value             = 1.5,
  reduce.lifeform       = TRUE,
  reduced.lifeforms     = reduced.lifeforms,
  reduction             = 0.1,
  reduced.exceptions    = reduced.exceptions,
  add_vegsum_table      = NULL
)

vegsum.pairs <- do.call(
  do_pairwise,
  c(list(veg.anal), pairwise.params)
)



source("../BEC_R_working/_functions/_build_bec_distance_matrix.R")


result <- build_bec_distance_matrix(
  vegsum_pairs = vegsum.pairs,
  su_anal = su.anal,
  preclass_field = "assoc",
  use_preclass = FALSE
)

unit.compare2 <- result$unit_compare
bec.dist <- result$bec_dist

 #bec.dist <- bec_dist_matrix(vegsum.pairs = vegsum.pairs, distance = "BEC.sim.min")

```

```{=typst}
#pagebreak()
```
# Vegetation difference between units

## 3D NMDS ordination plot

NMDS based on BEC dissimilarity using siteunits., working alliances and associations

Visual review: do alliances form distinct clusters? Which alliances overlap? Are some alliances highly variable internally? Includes an envfit joint plot on NMDS from a select set of climate and edaphic variables

[**See interactive 3D NMDS html file in review materials**]{.underline}

```{r}
#| label: fig-nmds
#| fig-cap: "3D NMDS for SiteUnits within working Alliances"
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 8

source("../BEC_R_working/_functions/_plot_nmds_hierarchy.R")
library(dplyr)

# Units present in each object
env_units  <- unique(su.env$SiteUnit)
anal_units <- unique(su.anal$SiteUnit)
dist_units <- rownames(as.matrix(bec.dist))

# Units present in all three
common_units <- Reduce(intersect, list(env_units, anal_units, dist_units))

# Report missing units
missing_from_env  <- setdiff(union(anal_units, dist_units), env_units)
missing_from_anal <- setdiff(union(env_units, dist_units), anal_units)
missing_from_dist <- setdiff(union(env_units, anal_units), dist_units)

cat("Units removed because not present in all datasets:\n")
cat("From su.env :", length(setdiff(env_units, common_units)), "\n")
print(setdiff(env_units, common_units))

cat("\nFrom su.anal:", length(setdiff(anal_units, common_units)), "\n")
print(setdiff(anal_units, common_units))

cat("\nFrom bec.dist:", length(setdiff(dist_units, common_units)), "\n")
print(setdiff(dist_units, common_units))

# Filter datasets
su.env.filt <- su.env %>%
  filter(SiteUnit %in% common_units)

su.anal.filt <- su.anal %>%
  filter(SiteUnit %in% common_units)

bec.dist.filt <- as.matrix(bec.dist)[common_units, common_units]

# Run NMDS
nmds <- plot_nmds_hierarchy(
  bec.dist = bec.dist.filt,
  su.anal = su.anal.filt,
  su.env = su.env.filt,
  k = 3,
  classification = "working",
  output_file = paste0(anal.set, "_NMDS_Hierarchy.html"),
  save_html = TRUE
)

nmds$plot
#nmds$scores
#nmds$stress
```

```{=typst}
#pagebreak()
```

## PERMANOVA Test of Vegetation Differences

Statistical assessment of vegetation composition differences using the **vegan::adonis2** PERMANOVA framework.

See Appendix 1 for details on methods and approach

### Interpretation

- High **R²** values indicate stronger separation among groups.
- Significant **PERMANOVA** results indicate statistically detectable differences in composition.
- Non-significant **betadispersion** results provide greater confidence that differences are due to shifts in composition rather than unequal dispersion among groups.

```{=typst}
#set text(size: 8pt)
#show figure: set block(breakable: true)
```

```{r }
#| label: fig-permanova-alliances
#| fig-cap: "Permanova assessment of Alliances"
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
# fig-width: 8
# fig-height: 12


  #veg.dat = veg.anal;  su.anal = su.anal1; group.field = "working_Alliance"; permutations = 199
source("../BEC_R_working/_functions/_assess_group_support.R")
source("../BEC_R_working/_functions/_pairwise_adonis_dist.R")

alliance.support <- assess_group_support(
  veg.dat = veg.anal,
  su.anal = su.anal,
  group.field = "working_alliance",
    permutations = 999
)
alliance.support$gt_table
# gtsave(
#   res$gt_table,
#   "alliance_support_table.html"
# )
```



```{r}
#| label: fig-permanova-associations
#| fig-cap: ("Permanova assessment of Associations within Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
# fig-width: 8
# fig-height: 12

assoc.support <- assess_group_support(
  veg.dat = veg.anal,
  su.anal = su.anal,
  group.field = "working_association",
  within_level = "working_alliance",
  permutations = 999
)
assoc.support$gt_table
# gtsave(
#   res$gt_table,
#   "alliance_support_table.html"
# )


```

```{=typst}
#pagebreak()
```
## Indicator Species Analysis

Indicator species analysis using plot data to assess what species statistically differentiate units from all other units. Analysis looks at individual hierarchical units but also grouped pairs of units to allow assessment of whether some existing pairs be should combined to maximize Flag hierarchy units with few indicator species and where merged units have a long list of indicator species

```{=typst}
#set text(size: 8pt)
#show figure: set block(breakable: true)
```


```{r}
#| label: fig-alliance-indic
#| fig-cap: ("Indicator species for  Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
# fig-width: 8
# fig-height: 12

source("../BEC_R_working/_functions/_indicator_species_analysis2.R")

ind.alliance <- indicator_species_analysis(
  veg.dat,
  su.anal,
  group_level = "working_alliance",
  within_level = NULL,
  max.order = 2,
  min.constancy = 40,
  min.fidelity = 10,
  p.cutoff = 0.2,
  export.tables = TRUE,
  export.folder = "./review_materials",
  show.tables = TRUE#,
 #top.n = NULL
)
ind.alliance$gt_table
```

```{=typst}
#pagebreak()
```

```{=typst}
#show figure: set block(breakable: true)
```

```{r}
#| label: fig-assoc-indic
#| fig-cap: ("Indicator species for Associations within working Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
# fig-width: 8
# fig-height: 12

ind.assoc.alliance <- indicator_species_analysis(
  veg.dat,
  su.anal,
  group_level = "working_association",
  within_level = "working_alliance",
  max.order = 2,
  min.constancy = 40,
  min.fidelity = 15,
  p.cutoff = 0.2,
  export.tables = TRUE,
  export.folder = "./review_materials",
  show.tables = TRUE#,
  #top.n = NULL

)
ind.assoc.alliance$gt_table

```
```{=typst}
#pagebreak()
```
# Environmental differentiation between units

## PCA of environment variables for analysis group
PCA from a limited set of edaphic (aSMR, SNR) and climatic (DD5, DD18_sm, FFP, Tmin_wt, CMD.site, AHM, MSP, PAS, )

```{r}
#| label: fig-env_pca
#| fig-cap: ("3D environment PCA graphic for SiteUnits within working Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 10


source("../BEC_R_working/_functions/_plot_pca_hierarchy.R")
library(forcats)
require(ggrepel)

su.env2 <- su.env |> select(siteunit = SiteUnit, everything())

pca.env <- plot_pca_hierarchy(su.env2,
                    n_vectors = 10,
    output_file = NULL,
    save_html = FALSE)

pca.env$plot
#pca.env$scores

```
```{=typst}
#pagebreak()
```
## Review environmental space of site units within alliances

Boxplot of siteunits along PCA axes by Allaince. Look for environmental site units outliers for membership reassessment.
```{r}
#| label: fig-PCA_boxplot
#| fig-cap: Environment PCA scores for SiteUnits within working Alliances
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 10

source("../BEC_R_working/_functions/_plot_pca_boxplots.R")


pca.boxplots <- plot_pca_boxplots(
  pca = pca.env,
  group_var = "alliance"
)

pca.boxplots


```

```{r}
#| label: fig-PCA_Assoc_boxplot
#| fig-cap: Environment PCA scores for SiteUnits within working Associations
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 10


pca.boxplots <- plot_pca_boxplots(
  pca = pca.env,
  group_var = "assoc"
)

pca.boxplots


```
```{=typst}
#pagebreak()
```

## Association Vegetation Summary table by Alliance

Each sheet has all site units for a proposed alliance on a single sheet grouped, labelled, and colour-themed by generated Association to help review and editing of membership. This will be used to review the proposed alliances and make decisions about which site units should be merged into associations, which should be separated into new associations, and which should be left as singletons. The vegetation summary tables will also be used to identify potential diagnostic species for each alliance and to compare the vegetation composition of different alliances.

*See exported xlsx table in review materials*

```{r summary table association within alliances}
# compute anal.test
# becdb <- dbConnect(
#   odbc::odbc(),
#   .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v0.accdb;"
# )
# BEC.hier.temp <- dbReadTable(becdb, "tblBEC_Hierarchy_Class") #%>% dplyr::select(-SiteUnit)
# dbDisconnect(becdb)
# 
# group.to.test = "working_alliance"
# BEC.hier.temp <- BEC.hier |>
#   filter(!is.na(.data[[group.to.test]]))
# 
# su.anal.new <- su.anal %>% dplyr::select(PlotNumber, SU_Label) %>% left_join(BEC.hier.temp, by = c("SU_Label" = "SU_Label")) %>% #rename(SiteUnit = SiteSeries) %>%
#   mutate(SiteUnit = str_replace(SiteUnit, "[abc]$", ""),# ignore phases
#          SiteUnit = if_else(
#            str_detect(SiteUnit, "101"),
#            str_replace(SiteUnit, "\\.[123]$", ""), # remove zonal variations
#            SiteUnit
#          ))

source("../BEC_R_working/_functions/_create_veg_summary_workbook.R")

# Site units arranged within generated alliances
create_veg_summary_workbook(
  su.anal,
  veg.dat = veg.dat2,
  taxon.lifeform = taxon.lifeform,
  group_level = "SiteUnit",
  within_level = "working_alliance",
  classification = "working",
  anal.set = anal.set,
  out.dir = "./review_materials"
)

# create_veg_summary_workbook(
#   su.anal = su.anal,
#   veg.dat = veg.dat2,
#   taxon.lifeform = taxon.lifeform,  
#   generated.hier = BEC.hier.temp,
#   group.to.test = group.to.test,
#   anal.set = anal.set
# )

```


## Additional possible work

### Random Forests

If alliances are real compositional groups, you should be able to predict them.

Random Forest rf \<- randomForest( x = comm, y = alliance )

Can hierarchical unit be predicted for new plots or hierarchical units?

### Klinka's Indicator plant analysis

rich-poor, dry-moist, climatic groups (see Klinka 1989)

# Appendix 1

- **PERMANOVA R²**: Proportion of variation explained by the hierarchical grouping.
- **PERMANOVA adjusted p-value**: Statistical significance of compositional differences.
- **Betadispersion adjusted p-value**: Evaluates whether observed differences are driven by true compositional shifts or by differences in within-group dispersion.

## Support Score Calculation

The overall support score is calculated as:

```{=html}
support_score = score_R2 + score_disp + score_perm
```

## 1. PERMANOVA R² Score

- R² ≥ 0.20 → **70 points**
- R² ≥ 0.15 → **60 points**
- R² ≥ 0.10 → **50 points**
- R² ≥ 0.05 → **25 points**
- R² < 0.05 → **0 points**

## 2. PERMANOVA Significance Score

- Adjusted p-value < 0.05 → **10 points**
- Adjusted p-value ≥ 0.05 → **0 points**

## 3. Betadispersion Score

- Adjusted p-value > 0.10 → **20 points**
- Adjusted p-value > 0.05 and ≤ 0.10 → **10 points**
- Adjusted p-value ≤ 0.05 → **0 points**

## Support Classification

| Support Score | Classification |
|---------------|---------------|
| ≥ 80 | Strong |
| ≥ 60 | Moderate |
| ≥ 40 | Weak |
| < 40 | Review |

## Recommendation Rules

Recommendations are assigned in the following order:

1. If **R² < 0.05** → **Merge?**
2. Else if **R² < 0.10** → **Review**
3. Else if **Betadispersion adjusted p-value < 0.05** → **Heterogeneous**
4. Otherwise → **Supported**
