---
title: "Assess proposed hierarchy for LMH77"
author: "William H MacKenzie"
date: today
format:
  typst:
    toc: false
    toc-depth: 1
    toc-title: Contents
    section-numbering: 1.1.1
    columns: 1
output-file: "MH_Hierarchy_Assessment.pdf"
editor: source
execute:
  echo: false
  error: false
  warning: false
  message: false
  fig.width: 6
  fig.height: 4
  fig.align: 'center'
  fig.cap: true
  fig.pos: H
  out.width: '100%'
  dev: pdf
  fig.ext: pdf
  cache: false
  fig.retina: 2
  dpi: 600
  fig.asp: 1.5
  fig.path: "./figures/"
---

```{r}
anal.set = "MH"
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(tidyverse)
require(DBI)
require(data.table)
require(cluster)
require(dendextend)
require(dynamicTreeCut)
require(gtable)
require(colorspace)
require(gtsummary)
require(openxlsx)
require(tictoc)
require(labdsv)
require(indicspecies)
require(Polychrome)
require(ggdendro)
require(purrr)
require(ggplotify)
require(grid)
library(stringr)
require(vegan)
require(betapart)
library(plotly)
require(gt)
gt <- gt::gt
library(writexl)
set.seed(1279)
source("../BEC_R_working/_functions/_bec_dist.R")
source("../BEC_R_working/_functions/_bec_dist_matrix.R")
source("../BEC_R_working/_functions/_combine_taxa.R")
source("../BEC_R_working/_functions/_combine_taxa_strata.R")
source("../BEC_R_working/_functions/_create_su_vegdata.R")
source("../BEC_R_working/_functions/_create_analysis_vegsum.R")
source("../BEC_R_working/_functions/_TabletoTree.R")
source("../BEC_R_working/_functions/_TreetoTable.R")
source("../BEC_R_working/_functions/_add_vars.R")
source("../BEC_R_working/_functions/_do_pairwise.R")
source("../BEC_R_working/_functions/_create_diagnostic_veg.R")
source("../BEC_R_working/_functions/_return_similar_pairs.R")
source("../BEC_R_working/_functions/_read_sppmaster.R")
source("../BEC_R_working/_functions/_combined_su.R")
source("../BEC_R_working/_functions/_create_veg_sum.R")
source('../BEC_R_working/_functions/_create_veg_sum_all.R')
source('../BEC_R_working/_functions/_build_species_ordering.R')
source('../BEC_R_working/_functions/_format_veg_table.R')
source('../BEC_R_working/_functions/_format_veg_table_strata.R')
source('../BEC_R_working/_functions/_format_veg_table_nostrata.R')
source('../BEC_R_working/_functions/_encode_veg_sum.R')
source('../BEC_R_working/_functions/_create_dendro.R')
source('../BEC_R_working/_functions/_create_dendro_all.R')
source('../BEC_R_working/_functions/_create_dendro_bybgc.R')
source('../BEC_R_working/_functions/_draw_dendro_split.R')
source('../BEC_R_working/_functions/_cluster_groups.R')
source('../BEC_R_working/_functions/_TabletoTree.R')
source('../BEC_R_working/_functions/_build_species_ordering_hier.R')
source("../BEC_R_working/_functions/_cluster_by_spp_group.R")
source("../BEC_R_working/_functions/_baselga_diagnostics.R")
data.path = "D:/OneDrive - Government of BC/GitHub_data"
```

```{r load data, echo=FALSE}
#------------------------------------------------------------
# 1. Load vegetation data + taxon lifeform classifications
#------------------------------------------------------------
taxon.all <- read_sppmaster()
taxon.exotic <- taxon.all %>%
  filter(Native == "E") %>%
  dplyr::select(Code) %>%
  pull()

taxon.lifeform <- taxon.all %>%
  filter(Codetype %in% c("U", "X", "D")) %>%
  dplyr::select(Code, ScientificName, EnglishName, Lifeform) %>%
  distinct()

trees <- taxon.lifeform %>% 
  filter(Lifeform %in% c("1", "2")) %>% 
  pull(Code)

tree_seedlings <- taxon.lifeform %>% 
  filter(Lifeform %in% c("1", "2")) %>% 
  mutate(Code = paste0(Code, "D")) %>% 
  pull(Code)
veg.dat <- readRDS("./clean_data/Analysis_BECMaster_Veg.rds") %>% filter(!Species %in% taxon.exotic, !startsWith(Species, "POA_PRA"))
#------------------------------------------------------------
# 2. Read lump table and apply combine_taxa()
#------------------------------------------------------------

veglump <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BC_Correlation2_Vpro_2023/Correlation2_Spp_lump.accdb;"
)

lump <- dbReadTable(veglump, "CorrelationII_Lump")
dbDisconnect(veglump)

veg.dat2 <- combine_taxa(vegdata = veg.dat, lump, use.subtaxa = FALSE)%>% 
   filter(!Species %in% taxon.exotic)

#------------------------------------------------------------
# 3. Load SiteUnit data from two LMH databases
#------------------------------------------------------------

db <- "D:/BC_Correlation2_Vpro_2023/Sorts_published_LMH/LMH77_Coast_Sorts.accdb"
#db2 <- "D:/BC_Correlation2_Vpro_2023/Sorts_published_LMH/LMH68_HG_Sorts.accdb"
db <- "D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb"
su <- rbind(combined_su(db))#, combined_su(db2))

su <- su %>%
  filter(!grepl("X|unplaced|omit|support|nudum|Ww|!", SiteUnit)) %>%
  arrange(SiteUnit)
su2 <- su %>%
  mutate(SiteUnit = str_replace(SiteUnit, "[abc]$", ""),# ignore phases
         # SiteUnit = if_else(
         #   str_detect(SiteUnit, "101"),
         #   str_replace(SiteUnit, "\\.[123]$", ""), # remove zonal variations
           SiteUnit
         ) %>%
    mutate( SU_Label = str_replace(SiteUnit.orig, "[abc]$", ""))#if_else(
         #   str_detect(SiteUnit.orig, "101"),
         #   str_replace(SiteUnit.orig, "\\.[123]$", ""), # remove zonal variations
         #   SiteUnit.orig
         # )) |> mutate(SU_Label = # ignore phases
  
#------------------------------------------------------------
# 5. Load BECMaster environmental + edatopic tables
#------------------------------------------------------------

becmaster <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster/BECMaster_fixing.accdb;"
)

plot.env <- dbReadTable(becmaster, "BECMaster_fixing_Env")
plot.edatope <- dbReadTable(becmaster, "BECMaster_fixing_edatopic")
plot.info <- dbReadTable(becmaster, "BECMaster_fixing_Admin")

dbDisconnect(becmaster)

#------------------------------------------------------------
# 6. Write n_plots to BECdb and load hierarchy + edatopic
#------------------------------------------------------------

becdb <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v0.accdb;"
)

BGC_all <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(FSRegion, BGC_NoSpace)

BEC.hier <- dbReadTable(becdb, "tblBEC_Hierarchy_Class") %>%
  filter(SU_Label %in% su2$SU_Label) %>%
  mutate(SiteUnit = str_replace_all(SU_Label, c("/" = "_", " " = "")))

bgcv13 <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(BGC_NoSpace, SiteClassificationStatus)

edatopic <- dbReadTable(becdb, "Edatopic_v13_3") %>%
  mutate(
    rSMR = paste0("rSMR", substr(Edatopic, 2, 2)),
    SNR = substr(Edatopic, 1, 1)
  ) %>%
  filter(startsWith(Source, "BEC")) %>%
  dplyr::select(BGC, rSMR, SNR, SS_NoSpace, Edatopic) %>%
  distinct() %>%
  mutate(SiteUnit = gsub("/", "_", SS_NoSpace))

dbDisconnect(becdb)

##Export working BECdb for anal.set
BECdb.anal <- BEC.hier |> filter(Analysis_Group %in% anal.set)

BECdb.anal <- BEC.hier |>
  filter(Analysis_Group %in% anal.set)

write_xlsx(
  BECdb.anal,
  file.path("./review_materials", paste0(anal.set,"_BECdb_for_updates.xlsx"))
)

##Use working spreadsheet instead of BECdb
#BEC.hier <- openxlsx::read.xlsx("D:/OneDrive/WNA_BGC/Association_Analysis/Coast_Hierarchy_Analysis/BECdb_Coast_Hierarchy_Working_ver3.xlsx") #|> 
  #filter(Analysis_Group %in% "Forest-Coast_MH")

# Keep only site units present in hierarchy table
su2 <- su2 %>%
  filter(SU_Label %in% BECdb.anal$SU_Label)
su.missed <- su2 |> filter(!SU_Label %in% BECdb.anal$SU_Label)
su.class <- su2 %>%
  left_join(BECdb.anal, by = c("SiteUnit" = "SiteUnit", "SU_Label" = "SU_Label"))

su.anal <- su.class |> filter(n_plots >1, Analysis_Group %in% anal.set)|> arrange(PlotNumber) 
```

```{r prep env data}
plot.env2 <- plot.env %>% mutate(lon = ifelse(Longitude<0, Longitude, 0-Longitude)) %>% mutate(Date = as_date(Date)) |> 
  filter(PlotNumber %in% su.anal$PlotNumber) |>
  dplyr::select(PlotNumber, lon, lat = Latitude, elev = Elevation, special = Exposure1, rSMR = MoistureRegime, rSNR = NutrientRegime, slope = SlopeGradient, aspect = Aspect ) 
 plot.info <- plot.info |> dplyr::select(Plot, BECSiteUnit, UserSiteUnit, HumusThickness, BGC=GIS_BGC)  |> filter(Plot %in% su.anal$PlotNumber)
plot.env2 <- left_join(plot.env2, plot.info, by = c("PlotNumber" = "Plot")) |> filter(!is.na (rSMR))


## extract climate data for plots from climr
require(climr)
plot.locs <- plot.env2 %>% dplyr::select(id=PlotNumber, BGC, lon,lat,elev) |> drop_na(lat)
plot.climate <- climr::downscale(setDT(plot.locs),
                              vars = list_vars(set = c("Annual", "Seasonal", "Monthly")),
                              return_refperiod = TRUE,
                              db_option = "local",
                              cache = TRUE)
## calculate aSMR
plot.env3 <- plot.climate |> select(PlotNumber = id, everything()) |> left_join(plot.env2) |> 
  mutate(
    moist = rowSums(
      across(c(PPT_05, PPT_06, PPT_07, PPT_08, PPT_09), ~ .x >= 90),
      na.rm = TRUE
    ),
    moist = ifelse(moist > 4, 1, 0),
    ratio = moist / 5,
    multiplier = 1 + ratio,
    CMD = CMD_an / multiplier
  ) |> mutate(PPT.dorm = PPT_at + PPT_wt) |> 
  mutate(Def = ifelse(CMD >=350, 350 - PPT.dorm, CMD)) |> 
  mutate(CMD.total = CMD + Def) |> 
  mutate(CMD.site = ifelse(slope > 35 & (aspect >180 & aspect <270), CMD.total *1.25,
                           ifelse(slope> 35 & (aspect>315 | aspect <45), CMD.total *0.85, CMD.total))) |> arrange(PlotNumber) |> 
  mutate(
    rSMR.clean = sub("^.*?(\\d+).*$", "\\1", rSMR)) |> 
  mutate(rSMR.clean = as.numeric(rSMR.clean),
  SNR.clean = recode(
    substr(trimws(rSNR), 1, 1),
    "A" = 1L,
    "B" = 2L,
    "C" = 3L,
    "D" = 4L,
    "E" = 5L,
    "F" = 6L,
    .default = NA_integer_
  ))
  ####To adjust in zones with prolonged snowpack remove CMD from early growing season months
## BGCs with or below heavy snow packs
# BGC_Special <- fread("./inputs/BGC_Special.csv", data.table = FALSE)
# allDat2 <- left_join(allDat2,BGC_Special, by = c("BGC" = "BGC"))
# allDat2 <- allDat2 %>% mutate(CMD = ifelse(is.na(Special), CMD, CMD_08)) %>% select(-Special, -CMD_08) %>% drop_na()
#rules <- read.csv("./inputs/aSMR_growingseason_Rules_HalfStep_v13_13Sept2025.csv")
rules <- read.csv("./inputs/aSMR_Rules_HalfStep_v13_23June2025.csv") |>
  mutate(
    rSMR.rule = sub("^rSMR", "", rSMR)) |> dplyr::select(-rSMR)

plot_asmr <- plot.env3 %>% select(PlotNumber, rSMR, CMD.site) |> 
  left_join(rules, by = c("rSMR" = "rSMR.rule")) %>%
  filter(CMD.site <= CMD) %>%
  group_by(PlotNumber) %>%  # ensure row-wise logic if needed
  slice_max(aSMR, with_ties = FALSE) %>%
  ungroup() %>% select(PlotNumber, aSMR)

plot.env3 <- left_join(plot.env3, plot_asmr)

plot.env3 <- plot.env3 |>
  left_join(
    su.anal, 
    by = "PlotNumber")
 
plot.env.vars <- plot.env3 |>   select(-SiteUnit, -working_association, -working_alliance) |> 
   select(PlotNumber, PAS_an,CMD.site, MSP, aSMR , SNR = SNR.clean, AHM,FFP, DD5 = DD5_an, DD18_sm ) |> arrange(PlotNumber)
 keep.rows <- complete.cases(plot.env.vars)

  su.env <- plot.env3 |> rename(assoc = working_association, alliance = working_alliance) |> 
  filter(!is.na(alliance)) |> arrange(SiteUnit) |> 
  group_by(SiteUnit, assoc, alliance) |>
    summarise(
  PAS_an = mean(PAS_an, na.rm = TRUE),
  CMD.site = mean(CMD.site, na.rm = TRUE),
  MSP   = mean(MSP, na.rm = TRUE),
  aSMR = mean(aSMR, na.rm = TRUE),
  SNR = mean(SNR.clean, na.rm = TRUE),
  AHM = mean(AHM, na.rm = TRUE),
  #lat = mean(lat, na.rm = TRUE),
  FFP   = mean(FFP, na.rm = TRUE),
  DD5 = mean(DD5_an, na.rm = TRUE),
  DD18_sm = mean(DD18_sm, na.rm = TRUE),
  #EXT = mean(EXT, na.rm = TRUE),
  .groups = "drop"
)

#     summarise(
#   across(
#     where(is.numeric) & !c(lat, lon),
#     ~ mean(.x, na.rm = TRUE)
#   ),
#   .groups = "drop"
# )

# ==============================================================================
# Climate variables for PCA
#==============================================================================
  env.vars <- su.env |>
  select(-assoc, -alliance) 

# Remove rows with missing values
keep.rows <- complete.cases(env.vars)

env.vars <- env.vars[keep.rows, ]

# Remove constant variables if any
env.vars.filt <- env.vars |>
  select(-SiteUnit) |>
  dplyr::select(
    where(~ is.numeric(.x) && sd(.x, na.rm = TRUE) > 0)
  )

keep.rows <- complete.cases(env.vars.filt)

#su.env <- env.vars[keep.rows, ]
su.env <- su.env |>
  dplyr::select(
    SiteUnit, assoc, alliance,
    where(~ is.numeric(.x) && sd(.x, na.rm = TRUE) > 0)
  )

```

# Assess and review the separation associations and alliances

All comparisons use the site unit summary data as the basis for comparison and not plot level data.

An initial generated hierarchical classification from Build_Hierarchy_working can be used but better is to have done initial assessments and created intial working\_ hierarchy levels for application in this script.

This script is to be used iteratively to assess current working unit membership is supported or if site unit membership should be reassigned.

Subsequent analysis here will substantiate the existing working units or suggest where association membership in alliances should be adjusted to maximize indicator species separation and environmental space.

### Core analyses are:

1.  NMDS ordination of plant communities showing Alliance, Association and SiteUnits with envfit joint plot for edaphic and select climatic parameters.

2.  Environment separation: PCA of units, boxplot showing position of site units on PCA axes grouped by higher taxons

3.  PERMANOVA (pairwise adonis) with betadispersion metrics to assess vegetation differences between units.

4.  Indicator species analysis identifying species which are statistically associated with units.

5.  Vpro style vegetation summary report by working_alliance organized by working hierarchical taxa

Together these answer:

Are alliances and associations compositionally distinct from one another? How internally consistent are they? How distinct are they? Which species define them? Which alliances overlap? What environmental gradients drive differences?

```{r read in SU data}
# anal.set = "MH"
# 
# 
# combine.grps <- c("CWH-wetfor")#, "CWH-Fd")
# 
# su.class <- su.class |>
#   dplyr::mutate(
#     Analysis_Group = ifelse(
#       Analysis_Group %in% combine.grps,
#       anal.set,
#       Analysis_Group
#     )
 # ) 





# alliance.remove = c("Ss_shoreline", "CWH_Fraser_Fh", "CDF_rich", "CWHds_102", "CWHws_102", "CWHvmN_102", "CWHvmS_102", "CWH_Fh_south", "CWHms_102", "CWHw_terraces", "CWHvh_102", "CWHv_102")
# su.anal <- su.anal |> filter(!working_alliance %in% alliance.remove)
#   
su.anal2 <- su.anal |> dplyr::select(PlotNumber, SiteUnit, assoc = working_association, alliance = working_alliance) |> 
  mutate(alliance = str_remove(alliance, "^coast_generated_"))  |> distinct()
unique.su <- su.anal2 |> dplyr::select(SiteUnit) |> distinct()|> arrange(SiteUnit)
unique.assocs2 <- su.anal2 |> dplyr::select(assoc) |> distinct()
unique.alliances2 <- su.anal2 |> dplyr::select(alliance) |> distinct()
su.assocs <- su.anal2 |> dplyr::select(SiteUnit, assoc) |> distinct()|> arrange(SiteUnit)
su.alliances <- su.anal2 |> dplyr::select(SiteUnit, alliance) |> distinct()|> arrange(SiteUnit)

veg.anal<- veg.dat2 |> filter(PlotNumber %in% su.class$PlotNumber) |> dplyr::select(PlotNumber, Species, Cover) |>
  arrange(PlotNumber) #|> matrify()

reduced.exceptions <- c("SPHAGNUM","CLADONIA","CLADINA","RACOMITR","MNIUM")
reduced.lifeforms  <- c(9, 10, 11)

pairwise.params <- list(
  su                = su.anal,
  su.field =  "SiteUnit",
  minconstancy          = 50,
  noiseconstancy        = 10,
  minplots              = 1,
  minor                 = 5,
  use.ksi               = FALSE,
  ksi   = NULL,
  ksi.value             = 1.5,
  reduce.lifeform       = TRUE,
  reduced.lifeforms     = reduced.lifeforms,
  reduction             = 0.1,
  reduced.exceptions    = reduced.exceptions,
  add_vegsum_table      = NULL
)

vegsum.pairs <- do.call(
  do_pairwise,
  c(list(veg.anal), pairwise.params)
)



source("../BEC_R_working/_functions/_build_bec_distance_matrix.R")


result <- build_bec_distance_matrix(
  vegsum_pairs = vegsum.pairs,
  su_anal = su.anal,
  preclass_field = "assoc",
  use_preclass = FALSE
)

unit.compare2 <- result$unit_compare
bec.dist <- result$bec_dist

 #bec.dist <- bec_dist_matrix(vegsum.pairs = vegsum.pairs, distance = "BEC.sim.min")

```

### 3D NMDS ordination plot

NMDS based on BEC dissimilarity using siteunits., working alliances and associations

Visual review: do alliances form distinct clusters? Which alliances overlap? Are some alliances highly variable internally? Includes an envfit joint plot on NMDS from a select set of climate and edaphic variables

[**See interactive 3D NMDS html file in review materials**]{.underline}

```{r}
#| label: fig-nmds
#| fig-cap: ("3D NMDS for SiteUnits within working Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12

source("../BEC_R_working/_functions/_plot_nmds_hierarchy.R")

nmds <- plot_nmds_hierarchy(
  bec.dist = bec.dist,
  su.anal = su.anal,
  su.env = su.env,
  k=3,
  classification = "working", 
  output_file = paste0(anal.set, "_NMDS_Hierarchy.html"),
  save_html = TRUE
)

nmds$plot
#nmds$scores
#nmds$stress
```

### 2. Permanova test of vegetation differences

Permanova statistical test of composition differences using adonis2 package

Outputs: R² = how much variation is explained by hierarchical group. Betadispersion test within units to adjust the PERMANOVA assessment to assess whether differences are due to real compositional shifts or spread.

Assigning a strength rating based on a sum of points from: r2 of permanova, p.adj from betadisp and permanova

1.  permanova R2 = R2 \>= 0.20= 70, R2 \>= 0.15 = 60, R2 \>= 0.10 = 50, R2 \>= 0.05 = 25, R2 \< 0.05 = 0

2.  permanovap.adj \< 0.05 = 10, p.adj \>= 0.05 = 0

3.  betadisper p.adj. \> 0.10 = 20, p.adj. \> 0.05 = 10, p.adj. \<= 0.05 =0

**support_score** = score_R2 +score_disp + score_perm

**support_class**: score \>= 80 is "Strong", score \>= 60 is "Moderate", score \>= 40 is "Weak", score \<40 is"Review"

**recommendation** = R2 \< 0.05 is "Merge?" R2 \< 0.10 is "Review", betadisper p.adj. \< 0.05 is "Heterogenous", otherwise "Supported"

```{r pairwise permanova}
#| label: fig-permanova-alliances
#| fig-cap: ("Permanova assessment of Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12


  #veg.dat = veg.anal;  su.anal = su.anal1; group.field = "working_Alliance"; permutations = 199
source("../BEC_R_working/_functions/_assess_group_support.R")
source("../BEC_R_working/_functions/_pairwise_adonis_dist.R")

alliance.support <- assess_group_support(
  veg.dat = veg.anal,
  su.anal = su.anal,
  group.field = "working_alliance",
    permutations = 999
)
alliance.support$gt_table
# gtsave(
#   res$gt_table,
#   "alliance_support_table.html"
# )




```

```{r}
#| label: fig-permanova-associations
#| fig-cap: ("Permanova assessment of Associations within Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12

assoc.support <- assess_group_support(
  veg.dat = veg.anal,
  su.anal = su.anal,
  group.field = "working_association",
  within_level = "working_alliance",
  permutations = 999
)
assoc.support$gt_table
# gtsave(
#   res$gt_table,
#   "alliance_support_table.html"
# )


```

## Indicator Species Analysis

Indicator species analysis using plot data to assess what species statistically differentiate units from all other units. Analysis looks at individual hierarchical units but also grouped pairs of units to allow assessment of whether some existing pairs be should combined to maximize Flag hierarchy units with few indicator species and where merged units have a long list of indicator species

```{r}
#| label: fig-alliance-indic
#| fig-cap: ("Indicator species for  Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12

source("../BEC_R_working/_functions/_indicator_species_analysis.R")

ind.alliance <- indicator_species_analysis(
  veg.dat,
  su.anal,
  group_level = "working_alliance",
  max.order = 2,
  min.constancy = 40,
  min.fidelity = 15,
  p.cutoff = 0.05,
  export.tables = TRUE,
  export.folder = "./review_materials",
  show.tables = TRUE,
  gt.top.n = NULL
)
ind.alliance$gt_table
```

```{r}
#| label: fig-assoc-indic
#| fig-cap: ("Indicator species for Associations within working Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12

ind.assoc.alliance <- indicator_species_analysis(
  veg.dat,
  su.anal,
  group_level = "working_association",
  within_level = "working_alliance",
  max.order = 2,
  min.constancy = 50,
  min.fidelity = 15,
  p.cutoff = 0.05,
  export.tables = TRUE,
  export.folder = "./review_materials",
  show.tables = TRUE,
  gt.top.n = NULL
  
)

#ind.assoc.alliance$gt_table


```

11. Environmental differentiation between units

Elevation Slope Aspect MAT MAP Soil moisture Nutrient regime

Use:

RDA rda.mod \<- rda( comm \~ elevation + moisture + nutrients, data = env )

or

## Environment of units

### PCA of environment variables for analysis group

```{r}
#| label: fig-env_pca
#| fig-cap: ("3D environment PCA graphic for SiteUnits within working Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12


source("../BEC_R_working/_functions/_plot_pca_hierarchy.R")
library(forcats)
require(ggrepel)

su.env2 <- su.env |> select(siteunit = SiteUnit, everything())

pca.env <- plot_pca_hierarchy(su.env2,
                    n_vectors = 10,
    output_file = NULL,
    save_html = FALSE)

pca.env$plot

```

```{r}
#| label: fig-PCA_boxplot
#| fig-cap: ("Environment PCA scores for SiteUnits within working Alliances")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12

library(tibble)
library(dplyr)

get_loading_subtitle <- function(loadings, pc, n = 3) {

  x <- loadings |>
    #tibble::rownames_to_column("Variable") |>
    select(variable, all_of(pc))

  top <- x |>
    arrange(desc(abs(.data[[pc]]))) |>
    slice_head(n = n)

  paste(
    paste0(
      top$variable,
      " (",
      sprintf("%.2f", top[[pc]]),
      ")"
    ),
    collapse = "; "
  )
}

pca.var <- pca.env$variance

sub.pc1 <- paste0(
  "Env drivers: ",
  get_loading_subtitle(pca.env$loadings, "PC1"),
  " | Var explained = ",
  round(pca.var["PC1"], 1),
  "%"
)
sub.pc2 <- paste0(
  "Env drivers: ",
  get_loading_subtitle(pca.env$loadings, "PC2"),
  " | Var explained = ",
  round(pca.var["PC2"], 2),
  "%"
)
sub.pc3 <- paste0(
  "Env drivers: ",
  get_loading_subtitle(pca.env$loadings, "PC3"),
  " | Var explained = ",
  round(pca.var["PC3"], 3),
  "%"
)


scores <- pca.env$scores |>
  mutate(
    alliance =
      forcats::fct_reorder(
        alliance,
        PC1,
        median,
        na.rm = TRUE
      )
  )

scores <- scores |>
  group_by(alliance) |>
  mutate(
    outlierPC1 =
      PC1 < quantile(PC1, 0.15, na.rm = TRUE) |
      PC1 > quantile(PC1, 0.85, na.rm = TRUE),
    outlierPC2 =
      PC2 < quantile(PC2, 0.15, na.rm = TRUE) |
      PC2 > quantile(PC2, 0.85, na.rm = TRUE),
    outlierPC3 =
      PC3 < quantile(PC3, 0.15, na.rm = TRUE) |
      PC3 > quantile(PC3, 0.85, na.rm = TRUE)
  ) |>
  ungroup()

library(ggrepel)

p1 <- ggplot(
  scores,
  aes(alliance, PC1, fill = alliance)
) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(
    width = 0.15,
    alpha = 0.6,
    size = 2
  ) +
  geom_text_repel(
    data = filter(scores, outlierPC1),
    aes(label = siteunit),
    size = 2,
    max.overlaps = 50,
    box.padding = 0.3,
    point.padding = 0.2,
    segment.alpha = 0.5
  ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "none") +
  labs(
    title = "PC1 by Alliance",
    subtitle = sub.pc1,
    x = NULL
  )

p2 <- ggplot(
  scores,
  aes(alliance, PC2, fill = alliance)
) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(
    width = 0.15,
    alpha = 0.6,
    size = 2
  ) +
  geom_text_repel(
    data = filter(scores, outlierPC2),
    aes(label = siteunit),
    size = 2,
    max.overlaps = 50,
    box.padding = 0.3,
    point.padding = 0.2,
    segment.alpha = 0.5
  ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "none") +
  labs(
    title = "PC2 by Alliance",
    subtitle = sub.pc2,
    x = NULL
  )

p3 <- ggplot(
  scores,
  aes(alliance, PC3, fill = alliance)
) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(
    width = 0.15,
    alpha = 0.6,
    size = 2
  ) +
  geom_text_repel(
    data = filter(scores, outlierPC3),
    aes(label = siteunit),
    size = 2,
    max.overlaps = 50,
    box.padding = 0.3,
    point.padding = 0.2,
    segment.alpha = 0.5
  ) +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "none") +
  labs(
    title = "PC3 by Alliance",
    subtitle = sub.pc3,
    x = NULL
  )
require(patchwork)
pc.boxplots <- p1 / p2 / p3 #+
  # plot_annotation(
  #   title = "Climate PCA Scores by Alliance"
  # )
pc.boxplots
ggsave(
  "./review_materials/PCA_Axis_Boxplots.png",
  pc.boxplots,
  width = 12,
  height = 16,
  dpi = 300
)

```

## Vegetation Summarize by Alliance

Generate vegetation summary output for all proposed alliances on a single sheet

## Constancy and cover tables

Generate vegetation summary for all site units by generated alliance colour-themed by generated association and subassociation for review.

Each sheet has all site units for a proposed alliance on a single sheet grouped, labelled, and colour-themed by generated Association to help review and editing of membership. This will be used to review the proposed alliances and make decisions about which site units should be merged into associations, which should be separated into new associations, and which should be left as singletons. The vegetation summary tables will also be used to identify potential diagnostic species for each alliance and to compare the vegetation composition of different alliances.

```{r summary table association within alliances}
# compute anal.test
# becdb <- dbConnect(
#   odbc::odbc(),
#   .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v0.accdb;"
# )
# BEC.hier.temp <- dbReadTable(becdb, "tblBEC_Hierarchy_Class") #%>% dplyr::select(-SiteUnit)
# dbDisconnect(becdb)
# 
# group.to.test = "working_alliance"
# BEC.hier.temp <- BEC.hier |>
#   filter(!is.na(.data[[group.to.test]]))
# 
# su.anal.new <- su.anal %>% dplyr::select(PlotNumber, SU_Label) %>% left_join(BEC.hier.temp, by = c("SU_Label" = "SU_Label")) %>% #rename(SiteUnit = SiteSeries) %>%
#   mutate(SiteUnit = str_replace(SiteUnit, "[abc]$", ""),# ignore phases
#          SiteUnit = if_else(
#            str_detect(SiteUnit, "101"),
#            str_replace(SiteUnit, "\\.[123]$", ""), # remove zonal variations
#            SiteUnit
#          ))

source("../BEC_R_working/_functions/_create_veg_summary_workbook.R")

# Site units arranged within generated alliances
create_veg_summary_workbook(
  su.anal,
  veg.dat = veg.dat2,
  taxon.lifeform = taxon.lifeform,
  group_level = "SiteUnit",
  within_level = "working_alliance",
  classification = "working",
  anal.set = "MH",
  out.dir = "./review_materials"
)

# create_veg_summary_workbook(
#   su.anal = su.anal,
#   veg.dat = veg.dat2,
#   taxon.lifeform = taxon.lifeform,  
#   generated.hier = BEC.hier.temp,
#   group.to.test = group.to.test,
#   anal.set = anal.set
# )

```

```{r compare alliances}
reduced.exceptions <- c("SPHAGNUM","CLADONIA","CLADINA","RACOMITR","MNIUM")
reduced.lifeforms  <- c(9, 10, 11)
su.alliance <- su.anal %>% mutate(SiteUnit = working_alliance) %>% select(PlotNumber, SiteUnit)

#------------------------------------------------------------
# 3. Run pairwise vegetation similarity analysis
#------------------------------------------------------------
pairwise.params <- list(
  su                = su.alliance,
  su.field = "SiteUnit",
  minconstancy          = 0,
  noiseconstancy        = 10,
  minplots              = 1,
  minor                 = 5,
  use.ksi               = FALSE,
  ksi   = NULL,
  ksi.value             = 1.5,
  reduce.lifeform       = TRUE,
  reduced.lifeforms     = reduced.lifeforms,
  reduction             = 0.1,
  reduced.exceptions    = reduced.exceptions,
  add_vegsum_table      = NULL
)

vegsum.pairs <- do.call(
  do_pairwise,
  c(list(veg.dat2), pairwise.params)
)

# Start with everything except `su`
print.params <- pairwise.params[setdiff(names(pairwise.params), "su")]

# Remove ksi if use.ksi == FALSE
if (!pairwise.params$use.ksi) {
  print.params <- print.params[setdiff(names(print.params), c("ksi", "ksi.value"))]
}

# Remove reduction-related fields if reduce.lifeform == FALSE
if (!pairwise.params$reduce.lifeform) {
  print.params <- print.params[setdiff(names(print.params),
                                       c("reduced.lifeforms", "reduction", "reduced.exceptions"))]
}
library(tibble)
library(knitr)

param.table <- tibble(
  parameter = names(print.params),
  value = sapply(print.params, function(x) paste(x, collapse = ", "))
)

param.table %>%
  gt::gt() %>%
  gt::tab_header(title = "Parameter Settings for do_pairwise()") %>%
  gt::tab_options(
    table.font.size = gt::px(8),
    table.border.top.width = gt::px(1),
    table.border.bottom.width = gt::px(1)
  )


#------------------------------------------------------------
# 4. Extract comparison table
#------------------------------------------------------------
fwrite(vegsum.pairs, "./outputs/vegsum_pairs.csv")

### return unit species information
vegsum.pairs.spp <- vegsum.pairs |> dplyr::select(Unit1, Unit2, Species, ScientificName, MeanCov.unit1, Constancy.unit1, MeanCov.unit2, Constancy.unit2, diff.pts.unit1, diff.pts.unit2)

### return distinct unit comparisons

unit.compare1 <- vegsum.pairs %>%
  dplyr::select(
    Unit1, Unit2, diff.ratio.unit1, diff.ratio.unit2, nplots.unit1, nplots.unit2, unit.diag.sum.unit1, unit.diag.sum.unit2
  ) %>%
  distinct(Unit1, Unit1, .keep_all = TRUE)


unit.compare <- unit.compare1 %>% 
    dplyr::select(
      Unit1 = Unit2,
      Unit2 = Unit1,
      diff.ratio.unit1 = diff.ratio.unit2,
      diff.ratio.unit2 = diff.ratio.unit1,
      nplots.unit1 = nplots.unit2,
      nplots.unit2 = nplots.unit1,
      unit.diag.sum.unit1 = unit.diag.sum.unit2,
      unit.diag.sum.unit2 = unit.diag.sum.unit1
      ) |>  rbind(unit.compare1) |> 
  distinct() |> mutate(BEC.sim.min = pmin(diff.ratio.unit1, diff.ratio.unit2),
                      BEC.sim.max = pmax(diff.ratio.unit1, diff.ratio.unit2))


fwrite(unit.compare , "./outputs/alliance_compare.csv")
#------------------------------------------------------------
# 5. Build BEC distance matrix and export
#------------------------------------------------------------

source("../BEC_R_working/_functions/_bec_dist_matrix.R")

bec_dist_matrix <- bec_dist_matrix(
  vegsum.pairs,
  distance = "BEC.sim.min"
) %>% 
  data.frame()

openxlsx::write.xlsx(
  x = bec_dist_matrix,
  file = "./clean_tabs/Coast_Forest_Alliance_Distance_matrix.xlsx",
  rowNames = TRUE
)
```

```{r summary report by alliances}
# author = "MH_Forest_Alliance"
# 
# vegsum.wbk <- createWorkbook()
#   # Build vegetation summary for this group
#   vegSum_sub <- create_veg_sum_all(
#     vdat = veg.dat2,
#     siteUnits = su.alliance,
#     minconstancy = 50,
#     noiseconstancy = 25,
#     strata.by = "Lifeform"
#   )
# 
#   # Format table
#   veg.sum.table_sub <- format_veg_table_nostrata(
#     vsum = vegSum_sub,
#     spp = taxon.lifeform
#   )
# sheet_name <- author
#   # Name the sheet
# # sheet_name <- paste0("Group_", g)
# 
# # Trim to 25 chars, then add a 5‑char hash suffix
# # if (nchar(sheet_name) > 31) {
# #   suffix <- substr(digest::digest(sheet_name), 1, 5)
# #   sheet_name <- paste0(substr(sheet_name, 1, 25), "_", suffix)
# # }
# 
# addWorksheet(vegsum.wbk, sheetName = sheet_name)
# 
# 
#   # Write data
#   writeData(vegsum.wbk, sheet = sheet_name, veg.sum.table_sub)
# #}
# 
# saveWorkbook(
#   vegsum.wbk,
#   paste0("./vegsum.tables/", author, ".xlsx"),
#   overwrite = TRUE
# )

```

## Review environmental space of site units within associations

Now look for environmental outliers of site units that have been merged into each association.

### ID Edaphic outliers from central concept of associations

Compare the median aSMR and median SNR value for each site unit to the median value for the entire association and flag those that are more than 1 edatopic step difference from the group median. This is a simple way to identify site units that may be outliers in edaphic space and may need to be separated into a different association or matched to a different association. aSMR is generated using the modelled relationship between rSMR and aSMR for the coastal BGCs.

```{r test for edaphic outliers1}
# source("../BEC_R_working/_functions/_test_edaphic_outliers.R")
# source("../BEC_R_working/_functions/_find_edaphic_similarity.R")
# group = "Association"
# 
# SU_edatope <- plot.edatope %>%
#   left_join(su.anal2, by = "PlotNumber") %>%
#   #rename(SiteUnit_orig = SiteUnit) %>%   # keep original
#   left_join(BEC.hier, by = c("SiteUnit" = "SiteUnit")) %>%
#   mutate(SiteUnit = coalesce(Association, SiteUnit)) %>%  # fallback
#   filter(Analysis_Group == anal.set) %>%
#   dplyr::select(SiteUnit, aSMR, SNR) %>%
#   filter(!is.na(SiteUnit)) %>%
#   mutate(
#     SNR = factor(SNR, levels = c("A", "B", "C", "D", "E"), ordered = TRUE),
#     SNR_num = as.numeric(SNR)
#   ) %>%
#   dplyr::select(-SNR) %>%
#   filter(!is.na(SNR_num))
# 
# 
# 
# median_by_site <- SU_edatope %>%
#   group_by(SiteUnit) %>%
#   summarise(across(where(is.numeric),
#                    \(x) median(x, na.rm = TRUE),
#                    .names = "median_{.col}"))
# 
# 
# su_all_sim2 <- su_all_sim %>% left_join(median_by_site, by = c("Unit1" = "SiteUnit")) %>% left_join(median_by_site, by = c("Unit2" = "SiteUnit")) %>% distinct %>% 
#   dplyr::select(Unit1, Unit2, BEC.sim.min, tempAssoc1, tempAssoc2, smr1 = median_aSMR.x, smr2 = median_aSMR.y, snr1 = median_SNR_num.x, snr2 = median_SNR_num.y, tempAssoc_match )
# 
# 
# su_all_sim2_rev <- su_all_sim2 %>% 
#   dplyr::select(Unit1 = Unit2, Unit2 = Unit1, BEC.sim.min, tempAssoc1 = tempAssoc2, tempAssoc2 = tempAssoc1, smr1 = smr2, smr2 = smr1, snr1 = snr2, snr2 = snr1, tempAssoc_match )
# 
# # bind and sort
# su_all_sim2_full <- bind_rows(su_all_sim2, su_all_sim2_rev) %>%
#   arrange(Unit1, Unit2)
# 
# write.xlsx(
#   su_all_sim2_full,
#   file = "./review_materials/other_similar_su_cluster_review_edatope.xlsx",
#   overwrite = TRUE
# )

#fwrite(su_all_sim2_full, "./review_materials/other_similar_sus_cluster_review_edatope.")
```

```{r test for edaphic outliers2}
# group = "generated_Alliance"
# 
# 
# 
# assoc_edatope <- plot.edatope %>%
#   left_join(su2, by = "PlotNumber") %>%
#   left_join(BEC.hier, by = c("SiteUnit" = "SiteUnit")) %>%
#   filter(Analysis_Group == anal.set) %>%
#   dplyr::select(SiteUnit, aSMR, SNR, groups = group) %>%
#   filter(!is.na(SiteUnit)) %>%
# #  filter(groups %in% select.groups) %>%
#   mutate(
#     SNR = factor(SNR, levels = c("A", "B", "C", "D", "E"), ordered = TRUE),
#     SNR_num = as.numeric(SNR)
#   ) %>% filter(!is.na(SNR))
# 
# all_results <- lapply(sort(unique(assoc_edatope$groups)), function(g) {
#   
#   res <- test_edaphic_outliers(
#     df = assoc_edatope,
#     group_name = g
#   )
#   
#   list(
#     group = g,
#     range_overlap = res$range_overlap %>% mutate(group = g),
#     median_flags  = res$median_flags  %>% mutate(group = g)
#   )
# })
# 
# names(all_results) <- sort(unique(assoc_edatope$groups))
# 
# 
# range_all <- bind_rows(
#   lapply(all_results, function(x) x$range_overlap)
# )
# range_all  <- bind_rows(lapply(all_results, function(x) x$range_overlap))
# edatopic_flag_all  <- bind_rows(lapply(all_results, function(x) x$median_flags)) %>% dplyr::select(group, SiteUnit, flag_aSMR, aSMR_median, group_aSMR_median, flag_SNR,SNR_median, group_SNR_median, flag_any, everything())
# edatopic_outlier <- edatopic_flag_all %>% filter(flag_any == TRUE)
# 
# edatopic_outlier_gt <- edatopic_outlier %>%
#   dplyr::select(
#     Group = group,
#     SiteUnit,
#     aSMR_Median = aSMR_median,
#     Group_aSMR_Median = group_aSMR_median,
#     aSMR_Flag = flag_aSMR,
#     SNR_Median = SNR_median,
#     Group_SNR_Median = group_SNR_median,
#     SNR_Flag = flag_SNR
#   ) 
# 
# gt(edatopic_outlier_gt) %>%
#   gt::tab_header(
#     title = "Edatopic Outlier Summary",
#     subtitle = "SiteUnits flagged as edaphic outliers based on median and range diagnostics"
#   ) %>%
#   gt::cols_label(
#     Group = "Association Group",
#     SiteUnit = "Site Unit",
#     aSMR_Median = "Plot aSMR Median",
#     Group_aSMR_Median = "Group aSMR Median",
#     aSMR_Flag = "aSMR Outlier",
#     SNR_Median = "Plot SNR Median",
#     Group_SNR_Median = "Group SNR Median",
#     SNR_Flag = "SNR Outlier"
#   ) %>%
#   gt::tab_style(
#     style = cell_fill(color = "#FFF3CD"),
#     locations = cells_body(
#       columns = c(aSMR_Flag, SNR_Flag),
#       rows = aSMR_Flag == TRUE | SNR_Flag == TRUE
#     )
#   ) %>%
#   cols_width(everything() ~ px(1)) %>%   # <— auto-fit trick
#   tab_options(
#     table.font.size = px(7),
#     data_row.padding = px(0),
#     table.layout = "auto"
#   )
```

### ID Climatic outliers from central concept of associations

Several approaches are applied. This first looks at univariate range of 4 climate variables (MCMT, DD5, MSP, PAS) for each site unit compared to the range of the rest of the association. This is a simple way to identify site units that may be outliers in climate space and may need to be separated into a different association or matched to a different association. However, this approach does not take into account the covariance between variables and may miss outliers that are extreme in combination but not in any one variable. The second approach is to use mahalanobis distance in multivariate climate space to identify site units that are outliers in climate space compared to the rest of the association. These units may be candidates for merging into a different association with more similar climate or for being separated into their own association if they are sufficiently different from all other associations.

```{r test for climatic outliers}
# climate_vars <- c("MCMT", "DD5_an",  "PAS_an","MSP" ,"CMD_an")
# source("../BEC_R_working/_functions/_climate_outliers.R")
# source("../BEC_R_working/_functions/_climate_comparison.R")
# climate_compared <- climate_outliers(
#   plot_climate = plot.climate,
#   su_anal = su.anal.new,
#   climate_vars = climate_vars,
#   group_col = "generated_Assoc",
#   unit_col = "SiteUnit",
#   low_overlap_threshold = 0.10,
#     alpha_global = .90,
#     alpha_local = .90)
# 
# mahal_outliers <- climate_compared$within_group_outliers %>% dplyr::select(SiteUnit, groups, distance_global, cutoff_global, is_outlier_global) %>% filter(is_outlier_global == TRUE) %>% arrange(groups, desc(distance_global)) %>% drop_na(groups)%>%
#   dplyr::select(
#     Group = groups,
#     SiteUnit,
#     Distance = distance_global,
#     Cutoff = cutoff_global
#   )
# 
#   gt(mahal_outliers) %>%
#   tab_header(
#     title = "Climate Outliers: Global Mahalanobis Distance",
#     subtitle = "SiteUnits exceeding the global multivariate cutoff"
#   ) %>%
#   fmt_number(
#     columns = c(Distance, Cutoff),
#     decimals = 2
#   ) %>%
#   tab_options(
#     table.font.size = px(8),
#     data_row.padding = px(1)
#   )
# 
#   
# univariate_outliers <- climate_compared$climate_outliers %>% dplyr::select(SiteUnit, groups, variable, mean_su, mean_grp) %>% arrange(groups, SiteUnit) %>% drop_na(groups) %>%
#   dplyr::select(
#     Group = groups,
#     SiteUnit,
#     Variable = variable,
#     SU_Mean = mean_su,
#     Group_Mean = mean_grp
#   )
# gt(univariate_outliers) %>%
#   tab_header(
#     title = "Climate Outliers: Univariate Deviations",
#     subtitle = "SiteUnits with mean climate values outside group envelopes"
#   ) %>%
#   fmt_number(
#     columns = c(SU_Mean, Group_Mean),
#     decimals = 2
#   ) %>%
#   cols_width(
#     everything() ~ px(1)   # auto-fit trick
#   ) %>%
#   tab_options(
#     table.font.size = px(7),
#     data_row.padding = px(0),
#     table.layout = "auto"
#   )

#mahalabobis_outliers <- climate_compared$mahal_all  %>% arrange(group, SiteUnit) %>% drop_na(group)
#mahal_dist<- mahal_results$within_group_mahal %>% drop_na(groups)
```

## Look for possible placements for single site units

Find nearest neighbours for single units (singletons) and compare vegetation, edaphic, and climatic similarity to identify potential matches for merging into existing groups. First find the nearest vegetation neighbours for site units and return the min and max BEC similarity Then for each pair which are \> 70% similar compare edatopic medians and climate similarity If high bec.sim.max and close environmental then consider merging the singleton into the neighbour unit Expect that site units in geographically peripheral BGCs may have no BC matches but may match to extrajurisdictional types (for Coast region Alaska or Washington State are most likely sources.)

```{r}
# source("../BEC_R_working/_functions/_find_nearest_neighbours.R")
# group_edaphic_medians <- edatopic_flag_all %>%
#   group_by(group) %>%
#   summarise(
#     group_SNR_median  = median(SNR_median, na.rm = TRUE),
#     group_aSMR_median = median(aSMR_median, na.rm = TRUE),
#     .groups = "drop"
#   )
# ungrouped_SUs <- assoc_edatope %>%
#   filter(is.na(groups)) %>%
#   distinct(SiteUnit)
# 
# edaphic_similarity <- find_edaphic_similarity (
#   edatopic_table = edatopic_flag_all,
#   df_raw = assoc_edatope
# ) %>% 
# arrange(SiteUnit )
# climate_similarity <-  climate_compared$mahal_all %>% arrange(SiteUnit)
# 
# #unit.compare = unit.compare; single.units = single.units
# nearest_units <- find_nearest_neighbours(unit.compare.new, single.units) %>% rename(SiteUnit = Singleton, group = Neighbor) 
# 
# nearest <- nearest_units  %>%
#   left_join(climate_similarity, by = c("SiteUnit", "group")) %>% 
#   left_join(edaphic_similarity, by = c("SiteUnit", "group")) %>%
#   arrange(SiteUnit, distance) %>% drop_na
# 
# fwrite(nearest, "./outputs/veg_climate_edaphic_similarity.csv", row.names = FALSE)
# check.match <- nearest %>% filter(distance < 7.78, BEC.sim.max >.65) %>%
#   dplyr::select(
#     SiteUnit,
#     Group = group,
#     Distance = distance,
#     BEC_Similarity = BEC.sim.max
#   ) %>%
#   arrange(SiteUnit, Distance)
# 
#   gt(check.match) %>%
#   tab_header(
#     title = "Nearest Neighbour Matches",
#     subtitle = "Filtered by climate distance < 7.78 and BEC similarity > 0.65"
#   ) %>%
#   fmt_number(
#     columns = c(Distance, BEC_Similarity),
#     decimals = 2
#   ) %>%
#   tab_style(
#     style = cell_fill(color = "#E8F5E9"),
#     locations = cells_body(
#       rows = Distance < 7.78 & BEC_Similarity > 0.65
#     )
#   ) %>%
#   tab_options(
#     table.font.size = px(8),
#     data_row.padding = px(1)
#   )


```

### Environment summaries

Horizonatal box plots of site unit medians by association and alliance

This links community differences to environmental gradients. function to pull climate data and other environment data

```{r}
# su.anal <- su.anal.new
# ## read in plot data and extract lat/long and environment info
# becmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster/BECMaster_fixing.accdb;")
# plot.env <- dbReadTable(becmaster, "BECMaster_fixing_Env") %>% mutate(lon = ifelse(Longitude<0, Longitude, 0-Longitude)) %>% mutate(Date = as_date(Date)) |> 
#   filter(PlotNumber %in% su.anal$PlotNumber) |>
#   dplyr::select(PlotNumber, lon, lat = Latitude, elev = Elevation, special = Exposure1, rSMR = MoistureRegime, rSNR = NutrientRegime, slope = SlopeGradient, aspect = Aspect ) 
# plot.info <- dbReadTable(becmaster, "BECMaster_fixing_Admin") |> dplyr::select(Plot, BECSiteUnit, UserSiteUnit, HumusThickness, BGC=GIS_BGC)  |> filter(Plot %in% su.anal$PlotNumber)
# plot.env <- left_join(plot.env, plot.info, by = c("PlotNumber" = "Plot")) |> filter(!is.na (rSMR))
# #odbc::odbcListColumns(becmaster, table = "BECMaster_fixing_plot.env")
# dbDisconnect(becmaster)
# 
# ## extract climate data for plots from climr
# require(climr)
# plot.locs <- plot.env %>% dplyr::select(id=PlotNumber, BGC, lon,lat,elev) |> drop_na(lat)
# plot.climate <- climr::downscale(setDT(plot.locs),
#                               vars = list_vars(set = c("Annual", "Seasonal", "Monthly")),
#                               return_refperiod = TRUE,
#                              # obs_periods = c("2001_2020"),
#                               #obs_years = 2001:2024,
#                               # gcms = c("ACCESS-ESM1-5","CNRM-ESM2-1", "EC-Earth3", "GFDL-ESM4", "GISS-E2-1-G", "MIROC6",  "MPI-ESM1-2-HR", "MRI-ESM2-0",  "UKESM1-0-LL" ),
#                               # ssps = c("ssp126", "ssp245", "ssp370"),
#                               # gcm_ssp_years = c(2025:2045),
#                               #ensemble_mean = TRUE,
#                               #obs_ts_dataset = "cru.gpcc",
#                               #obs_ts_dataset = "mswx.blend",
#                               db_option = "local",
#                               cache = TRUE)
# ## calculate aSMR
# plot.env2 <- plot.climate |> select(PlotNumber = id, everything()) |> left_join(plot.env) |> 
#   mutate(
#     rSMR.clean = sub("^.*?(\\d+).*$", "\\1", rSMR)) |> 
#   mutate(rSMR = as.numeric(rSMR.clean),
#   SNR.clean = recode(
#     substr(trimws(rSNR), 1, 1),
#     "A" = 1L,
#     "B" = 2L,
#     "C" = 3L,
#     "D" = 4L,
#     "E" = 5L,
#     "F" = 6L,
#     .default = NA_integer_
#   )) |> 
#   mutate(
#     moist = rowSums(
#       across(c(PPT_05, PPT_06, PPT_07, PPT_08, PPT_09), ~ .x >= 90),
#       na.rm = TRUE
#     ),
#     moist = ifelse(moist > 4, 1, 0),
#     ratio = moist / 5,
#     multiplier = 1 + ratio,
#     CMD = CMD_an / multiplier
#   ) |> mutate(PPT.dorm = PPT_at + PPT_wt) |> 
#   mutate(CMD.total = ifelse(CMD >=350, 350 - PPT.dorm, CMD)) |> 
#     mutate(CMD.site = ifelse(rSMR == 0, CMD.total+500, 
#                            ifelse(rSMR == 1, CMD.total+375, 
#                                   ifelse(rSMR == 2, CMD.total +250,
#                                          ifelse(rSMR == 3, CMD.total +125,
#                                                 ifelse(rSMR == 5, CMD.total/2,
#                                                        ifelse(rSMR == 6, CMD.total/4,
#                                                               ifelse(rSMR == 7, CMD.total/8,
#                                                                      ifelse(rSMR == 8, CMD.total/16, CMD.total))))))))) |> 
#   mutate(CMD.site = ifelse(slope > 35 & (aspect >180 & aspect <270), CMD.site *1.25,
#                            ifelse(slope> 35 & (aspect>315 | aspect <45), CMD.site *0.85, CMD.site))) |> arrange(PlotNumber)
#   
#   ####To adjust in zones with prolonged snowpack remove CMD from early growing season months
# ## BGCs with or below heavy snow packs
# # BGC_Special <- fread("./inputs/BGC_Special.csv", data.table = FALSE)
# # allDat2 <- left_join(allDat2,BGC_Special, by = c("BGC" = "BGC"))
# # allDat2 <- allDat2 %>% mutate(CMD = ifelse(is.na(Special), CMD, CMD_08)) %>% select(-Special, -CMD_08) %>% drop_na()
# #rules <- read.csv("./inputs/aSMR_growingseason_Rules_HalfStep_v13_13Sept2025.csv")
# rules <- read.csv("./inputs/aSMR_Rules_HalfStep_v13_23June2025.csv") |>
#   mutate(rSMR.rule = as.numeric(sub("^rSMR", "", rSMR)), CMD = as.numeric(CMD)) |> dplyr::select(-rSMR)
# 
# plot_asmr <- plot.env2 %>%
#   select(PlotNumber, rSMR, CMD.site) %>%
#   left_join(rules, by = c("rSMR" = "rSMR.rule")) %>%
#   filter(CMD.site <= CMD) %>%
#   group_by(PlotNumber) %>%
#   slice_min(CMD, n = 1) %>%
#   ungroup() %>% select(PlotNumber, aSMR)
# 
# plot.env2 <- left_join(plot.env2, plot_asmr)
# 
# env.alliance <- plot.env2 |>
#   left_join(
#     su.anal, 
#     by = "PlotNumber") |> rename(alliance = generated_Alliance, assoc = generated_Assoc)
# 
# xx <- env.alliance |> filter(SiteUnit == "CWHms5_112")

```

## Environment of units

```{r}
#| label: fig-Assoc-env
#| fig-cap: ("Plot environment by Association")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12
# env.long <- env.alliance |>
#   select(siteunit = SiteUnit, assoc,  aSMR, SNR.clean, MSP, MAT) |>
#   mutate(
#     #alliance = str_remove(alliance, "^.*generated_"),
#     assoc    = str_remove(assoc, "^.*generated_")
#   ) |>
#   pivot_longer(
#     cols = c(aSMR, SNR.clean, MSP, MAT),
#     names_to = "variable",
#     values_to = "value",
#     values_drop_na = TRUE
#   ) |>
#   mutate(
#     assoc = reorder_within(assoc, value, variable, median)
#   )
# 
# library(Polychrome)
# 
# unit_levels <- sort(unique(env.long$siteunit))
# 
# cols <- createPalette(
#   length(unit_levels),
#   seedcolors = c(
#     "#E41A1C",
#     "#377EB8",
#     "#4DAF4A",
#     "#984EA3",
#     "#FF7F00"
#   )
# )
# 
# names(cols) <- unit_levels
# 
# ggplot(env.long, aes(assoc, value)) +
#   geom_boxplot(fill = "grey90", outlier.shape = NA) +
#   geom_jitter(
#     aes(colour = siteunit),
#     width = 0.1,
#     alpha = 0.6,
#     size = 2
#   ) +
#   facet_wrap(~variable, scales = "free") +
#   scale_x_reordered() +
#   coord_flip() +
#   scale_colour_manual(values = cols) +
#   theme(legend.position = "none")

  
```

```{r}
#| label: fig-Alliance-plotenv
#| fig-cap: ("Plot environment by Alliance")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12
# env.long <- env.alliance |>
#   select(siteunit = SiteUnit, assoc, alliance, aSMR, SNR.clean, MSP, MAT) |>
#   mutate(
#     alliance = str_remove(alliance, "^.*generated_"),
#     assoc    = str_remove(assoc, "^.*generated_")
#   ) |>
#   pivot_longer(
#     cols = c(aSMR, SNR.clean, MSP, MAT),
#     names_to = "variable",
#     values_to = "value",
#     values_drop_na = TRUE
#   ) |>
#   mutate(
#     alliance = reorder_within(alliance, value, variable, median)
#   )
# 
# library(Polychrome)
# 
# unit_levels <- sort(unique(env.long$siteunit))
# 
# cols <- createPalette(
#   length(unit_levels),
#   seedcolors = c(
#     "#E41A1C",
#     "#377EB8",
#     "#4DAF4A",
#     "#984EA3",
#     "#FF7F00"
#   )
# )
# 
# names(cols) <- unit_levels
# 
# ggplot(env.long, aes(alliance, value)) +
#   geom_boxplot(fill = "grey90", outlier.shape = NA) +
#   geom_jitter(
#     aes(colour = siteunit),
#     width = 0.2,
#     alpha = 0.6,
#     size = 2
#   ) +
#   facet_wrap(~variable, scales = "free") +
#   scale_x_reordered() +
#   coord_flip() +
#   scale_colour_manual(values = cols) +
#   theme(legend.position = "none")
# 
#   
```

```{r}
#| label: fig-medianSU-env
#| fig-cap: ("Median SiteUnit environment by alliance")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12
# env.sum <- env.alliance |>
#   select(siteunit = SiteUnit, assoc, alliance,aSMR, SNR.clean, MSP, MAT) |>#
#   mutate(
#     alliance = str_remove(alliance, "^.*generated_"),
#     assoc    = str_remove(assoc, "^.*generated_")
#   ) |>
#   pivot_longer(
#     cols = c(aSMR, SNR.clean, MSP, MAT),#
#     names_to = "variable",
#     values_to = "value",
#     values_drop_na = TRUE
#   ) |>
#   group_by(variable, siteunit, alliance) |>
#   summarise(
#     median = median(value, na.rm = TRUE),
#     #mean   = mean(value, na.rm = TRUE),
#     #sd     = sd(value, na.rm = TRUE),
#     .groups = "drop"
#   )
# alliance.order <- env.sum |>
#   group_by(variable, alliance) |>
#   summarise(ord = median(median), .groups = "drop")
# 
# library(tidytext)
# 
# env.sum <- env.sum |>
#   left_join(alliance.order, by = c("variable", "alliance")) |>
#   mutate(
#     alliance_plot = reorder_within(alliance, ord, variable)
#   )
# 
# library(ggrepel)
# 
# ggplot(
#   env.sum,
#   aes(
#     x = alliance_plot,
#     y = median,
#     colour = siteunit
#   )
# ) +
#   geom_errorbar(
#     aes(ymin = median,# - sd,
#         ymax = median),# + sd),
#     width = 0,
#     position = position_dodge(width = 0.4)
#   ) +
#   geom_point(
#     size = 2,
#     position = position_dodge(width = 0.4)
#   ) +
#   geom_text_repel(
#     aes(label = siteunit),
#     direction = "y",
#     size = 2,
#     show.legend = FALSE
#   ) +
#   facet_wrap(~variable, scales = "free") +
#   scale_x_reordered() +
#   coord_flip() +
#   scale_colour_manual(values = cols) +
#   theme_bw() +
#   theme(legend.position = "none")

```

## Additional possible work

### Random Forests

If alliances are real compositional groups, you should be able to predict them.

Random Forest rf \<- randomForest( x = comm, y = alliance )

Can hierarchical unit be predicted for new plots or hierarchical units?

### Klinka's Indicator plant analysis

rich-poor, dry-moist, climatic groups (see Klinka 1989)
