---
title: "Hierarchy Unit Review"
author: "William H MacKenzie"
date: "08/11/2024"
format: 
  pdf: 
    documentclass: scrartcl
    papersize: letter
editor: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(tidyverse)
require(DBI)
require(data.table)
require(cluster)
require(dendextend)
require(dynamicTreeCut)
require(gtable)
require(colorspace)
require(gtsummary)
require(openxlsx)
require(tictoc)
require(labdsv)
require(indicspecies)
require(Polychrome)
require(ggdendro)
require(purrr)
require(ggplotify)
require(grid)
library(stringr)
require(vegan)
require(betapart)
library(plotly)
set.seed(1279)
source("../BEC_R_working/_functions/_bec_dist.R")
source("../BEC_R_working/_functions/_bec_dist_matrix.R")
source("../BEC_R_working/_functions/_combine_taxa.R")
source("../BEC_R_working/_functions/_combine_taxa_strata.R")
source("../BEC_R_working/_functions/_create_su_vegdata.R")
source("../BEC_R_working/_functions/_create_analysis_vegsum.R")
source("../BEC_R_working/_functions/_TabletoTree.R")
source("../BEC_R_working/_functions/_TreetoTable.R")
source("../BEC_R_working/_functions/_add_vars.R")
source("../BEC_R_working/_functions/_do_pairwise.R")
source("../BEC_R_working/_functions/_create_diagnostic_veg.R")
source("../BEC_R_working/_functions/_return_similar_pairs.R")
source("../BEC_R_working/_functions/_read_sppmaster.R")
source("../BEC_R_working/_functions/_combined_su.R")
source("../BEC_R_working/_functions/_create_veg_sum.R")
source('../BEC_R_working/_functions/_create_veg_sum_all.R')
source('../BEC_R_working/_functions/_build_species_ordering.R')
source('../BEC_R_working/_functions/_format_veg_table.R')
source('../BEC_R_working/_functions/_format_veg_table_strata.R')
source('../BEC_R_working/_functions/_format_veg_table_nostrata.R')
source('../BEC_R_working/_functions/_encode_veg_sum.R')
source('../BEC_R_working/_functions/_create_dendro.R')
source('../BEC_R_working/_functions/_create_dendro_all.R')
source('../BEC_R_working/_functions/_create_dendro_bybgc.R')
source('../BEC_R_working/_functions/_draw_dendro_split.R')
source('../BEC_R_working/_functions/_cluster_groups.R')
source('../BEC_R_working/_functions/_TabletoTree.R')
source('../BEC_R_working/_functions/_build_species_ordering_hier.R')
source("../BEC_R_working/_functions/_cluster_by_spp_group.R")
source("../BEC_R_working/_functions/_baselga_diagnostics.R")
data.path = "D:/OneDrive - Government of BC/GitHub_data"
```

# General flow for review

A report to review proposed associations and hierarchical unit. Comparative summary table Ordination Comparative environmental reporting

## Read in data

Read in vegetation data and join to taxonomic read in becdb table with units and hierarchical levels read in working xlsx becdb table with proposed units read in su tables and limit to units in the working xlsx

Vegetation data is read in from saved .RDS file generated from the BECMaster cleaning scripts. Taxonomy is read-in from the species taxonomy database. A species lumping table is read-in from a Vpro database and applied to the vegetation data. The species taxonomy is joined to the vegetation data.

## Read in data

Vegetation data is read in from saved BECMaster.RDS file generated from the BECMaster cleaning scripts. A compiled SU table is build from all BGC_SU tables stored in the coast guides Vpro database. Taxonomy is read in from the species taxonomy database. A species lumping code table is read in from the Correlation2_Spp_lump.accdb database and the vegetation data is lumped using the combine_taxa function.

```{r load data, echo=FALSE}
#------------------------------------------------------------
# 1. Load vegetation data + taxon lifeform classifications
#------------------------------------------------------------
taxon.all <- read_sppmaster()
taxon.exotic <- taxon.all %>%
  filter(Native == "E") %>%
  dplyr::select(Code) %>%
  pull()

taxon.lifeform <- taxon.all %>%
  filter(Codetype %in% c("U", "X", "D")) %>%
  dplyr::select(Code, ScientificName, EnglishName, Lifeform) %>%
  distinct()

trees <- taxon.lifeform %>% 
  filter(Lifeform %in% c("1", "2")) %>% 
  pull(Code)

tree_seedlings <- taxon.lifeform %>% 
  filter(Lifeform %in% c("1", "2")) %>% 
  mutate(Code = paste0(Code, "D")) %>% 
  pull(Code)
veg.dat <- readRDS("./clean_data/Analysis_BECMaster_Veg.rds") %>% filter(!Species %in% taxon.exotic, !startsWith(Species, "POA_PRA"))
#------------------------------------------------------------
# 2. Read lump table and apply combine_taxa()
#------------------------------------------------------------

veglump <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BC_Correlation2_Vpro_2023/Correlation2_Spp_lump.accdb;"
)

lump <- dbReadTable(veglump, "CorrelationII_Lump")
dbDisconnect(veglump)

veg.dat2 <- combine_taxa(vegdata = veg.dat, lump, use.subtaxa = FALSE)%>% 
   filter(!Species %in% taxon.exotic)

#------------------------------------------------------------
# 3. Load SiteUnit data from two LMH databases
#------------------------------------------------------------

db <- "D:/BC_Correlation2_Vpro_2023/Sorts_published_LMH/LMH77_Coast_Sorts.accdb"
#db2 <- "D:/BC_Correlation2_Vpro_2023/Sorts_published_LMH/LMH68_HG_Sorts.accdb"
db <- "D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb"
su <- rbind(combined_su(db))#, combined_su(db2))

su <- su %>%
  filter(!grepl("X|unplaced|omit|support|nudum|Ww|!", SiteUnit)) %>%
  arrange(SiteUnit)
su2 <- su %>%
  mutate(SiteUnit = str_replace(SiteUnit, "[abc]$", ""),# ignore phases
         # SiteUnit = if_else(
         #   str_detect(SiteUnit, "101"),
         #   str_replace(SiteUnit, "\\.[123]$", ""), # remove zonal variations
           SiteUnit
         ) %>%
    mutate( SU_Label = str_replace(SiteUnit.orig, "[abc]$", ""))#if_else(
         #   str_detect(SiteUnit.orig, "101"),
         #   str_replace(SiteUnit.orig, "\\.[123]$", ""), # remove zonal variations
         #   SiteUnit.orig
         # )) |> mutate(SU_Label = # ignore phases
  
#------------------------------------------------------------
# 5. Load BECMaster environmental + edatopic tables
#------------------------------------------------------------

becmaster <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster/BECMaster_fixing.accdb;"
)

plot.env <- dbReadTable(becmaster, "BECMaster_fixing_Env")
plot.edatope <- dbReadTable(becmaster, "BECMaster_fixing_edatopic")

dbDisconnect(becmaster)

#------------------------------------------------------------
# 6. Write n_plots to BECdb and load hierarchy + edatopic
#------------------------------------------------------------

becdb <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v0.accdb;"
)

BGC_all <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(FSRegion, BGC_NoSpace)

BEC.hier <- dbReadTable(becdb, "tblBEC_Hierarchy_Class") %>%
  filter(SU_Label %in% su2$SU_Label) %>%
  mutate(SiteUnit = str_replace_all(SU_Label, c("/" = "_", " " = "")))

bgcv13 <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(BGC_NoSpace, SiteClassificationStatus)

edatopic <- dbReadTable(becdb, "Edatopic_v13_3") %>%
  mutate(
    rSMR = paste0("rSMR", substr(Edatopic, 2, 2)),
    SNR = substr(Edatopic, 1, 1)
  ) %>%
  filter(startsWith(Source, "BEC")) %>%
  dplyr::select(BGC, rSMR, SNR, SS_NoSpace, Edatopic) %>%
  distinct() %>%
  mutate(SiteUnit = gsub("/", "_", SS_NoSpace))

dbDisconnect(becdb)


##Use working spreadsheet instead of BECdb
#BEC.hier <- openxlsx::read.xlsx("D:/OneDrive/WNA_BGC/Association_Analysis/Coast_Hierarchy_Analysis/BECdb_Coast_Hierarchy_Working_ver3.xlsx") #|> 
  #filter(Analysis_Group %in% "Forest-Coast_MH")

# Keep only site units present in hierarchy table
su2 <- su2 %>%
  filter(SU_Label %in% BEC.hier$SU_Label)
su.missed <- su2 |> filter(!SU_Label %in% BEC.hier$SU_Label)
su.class <- su2 %>%
  left_join(BEC.hier, by = c("SiteUnit" = "SiteUnit", "SU_Label" = "SU_Label"))


```

### Core analyses would be:

All comparisons use the site unit summary data as the basis for comparison and not plot level data.

1.  NMDS ordination

2.  PERMANOVA to compare sig differences between units. Alliances, Assocs within Alliances, All Assocs

3.  Betadisper: variability within units

4.  Indicator species analysis: significant indicators associated with single or groups of units. Where few indicators in individual units consider combining with groups

5.  Random Forest classification: prediction of membership

6.  Species constancy tables: standard vegetation summary output

7.  Environmental gradient analysis: environmental separation of units

Together these answer:

Are alliances compositionally distinct from one another? How internally consistent are they? How distinct are they? Which species define them? Which alliances overlap? What environmental gradients drive differences?

For a BEC/alliance classification paper, NMDS + PERMANOVA + Indicator Species + Random Forest + diagnostic species tables would likely provide the strongest and most defensible framework.

```{r read in SU data}

anal.set = "MH"
su.anal <- su.class |> filter(n_plots >0, Analysis_Group %in% anal.set)|> arrange(PlotNumber) 
su.anal2 <- su.anal |> dplyr::select(PlotNumber, SiteUnit, assoc = working_association, alliance = working_alliance) |> 
  mutate(alliance = str_remove(alliance, "^coast_generated_"))  |> distinct()
unique.su <- su.anal2 |> dplyr::select(SiteUnit) |> distinct()|> arrange(SiteUnit)
unique.assocs2 <- su.anal2 |> dplyr::select(assoc) |> distinct()
unique.alliances2 <- su.anal2 |> dplyr::select(alliance) |> distinct()
su.assocs <- su.anal2 |> dplyr::select(SiteUnit, assoc) |> distinct()|> arrange(SiteUnit)
su.alliances <- su.anal2 |> dplyr::select(SiteUnit, alliance) |> distinct()|> arrange(SiteUnit)

veg.anal<- veg.dat2 |> filter(PlotNumber %in% su.class$PlotNumber) |> dplyr::select(PlotNumber, Species, Cover) |>
  arrange(PlotNumber) #|> matrify()

reduced.exceptions <- c("SPHAGNUM","CLADONIA","CLADINA","RACOMITR","MNIUM")
reduced.lifeforms  <- c(9, 10, 11)

pairwise.params <- list(
  su                = su.anal,
  su.field =  "SiteUnit",
  minconstancy          = 50,
  noiseconstancy        = 10,
  minplots              = 1,
  minor                 = 1,
  use.ksi               = FALSE,
  ksi   = NULL,
  ksi.value             = 1.5,
  reduce.lifeform       = TRUE,
  reduced.lifeforms     = reduced.lifeforms,
  reduction             = 0.1,
  reduced.exceptions    = reduced.exceptions,
  add_vegsum_table      = NULL
)

vegsum.pairs <- do.call(
  do_pairwise,
  c(list(veg.anal), pairwise.params)
)
source("../BEC_R_working/_functions/_build_bec_distance_matrix.R")


result <- build_bec_distance_matrix(
  vegsum_pairs = vegsum.pairs,
  su_anal = su.anal,
  preclass_field = "assoc",
  use_preclass = FALSE
)

unit.compare2 <- result$unit_compare
bec.dist <- result$bec_dist

 #bec.dist <- bec_dist_matrix(vegsum.pairs = vegsum.pairs, distance = "BEC.sim.min")

```

##Describe differences in species composition

### 3D NMDS ordination plot

links to interactive html 3D ordination plot

Use BEC dissimilarity and working alliances, associations, siteunits.

Questions answered:

Do alliances form distinct clusters? Which alliances overlap? Are some alliances highly variable internally?

Add alliance centroids and confidence ellipses.

ordiplot(nmds) ordiellipse(nmds, alliance)

```{r NMDS}
source("../BEC_R_working/_functions/_plot_nmds_hierarchy.R")

result <- plot_nmds_hierarchy(
  bec.dist = bec.dist,
  su.anal = su.anal,
  classification = "working", 
  output_file = paste0(anal.set, "_NMDS_Hierarchy.html"),
  save_html = TRUE
)

result$stress
result$plot

```

See the interactive version:

\[[3D_NMDS_Graphic.html](nmds.file)\]

## 2. Test whether alliances differ PERMANOVA

A statistical test of composition differences.

adonis2( comm \~ alliance, data = meta, method = "bray", permutations = 999 )

Outputs:

R² = how much variation is explained by alliance P-value

Given 3450 plots, you'll almost certainly get a significant result. Focus on effect size (R²), not just significance.

Test dispersion first

PERMANOVA is sensitive to unequal dispersion.

disp \<- betadisper( vegdist(comm, method = "bray"), alliance )

anova(disp)

This tells you whether differences are due to centroids (real compositional shifts) or spread.

3.  Quantify within-unit heterogeneity

Some alliances are very tight, others broad.

Beta diversity disp \<- betadisper( vegdist(comm, "bray"), alliance )

boxplot(disp)

Useful outputs:

Mean distance to centroid Rank alliances from homogeneous to heterogeneous

This is often ecologically informative.

```{r pairwise permanova}
  #veg.dat = veg.anal;  su.anal = su.anal1; group.field = "working_Alliance"; permutations = 199
source("../BEC_R_working/_functions/_assess_group_support.R")
source("../BEC_R_working/_functions/_pairwise_adonis_dist.R")

alliance.support <- assess_group_support(
  veg.dat = veg.anal,
  su.anal = su.anal,
  group.field = "working_alliance",
    permutations = 999
)
alliance.support$gt_table
# gtsave(
#   res$gt_table,
#   "alliance_support_table.html"
# )

assoc.support <- assess_group_support(
    veg.dat = veg.anal,
  su.anal = su.anal,
  group.field = "working_association",
    permutations = 999
)
assoc.support$gt_table

# ------------------------------------------------------------
  # Export workbook
  # ------------------------------------------------------------
  out.dir = "./review_materials"
  outfile <- file.path(
    out.dir,
    paste0(anal.set, "_other_similar_units.xlsx")
  )

  replace_sheet <- function(wb, sheet) {
  if (sheet %in% names(wb)) {
    removeWorksheet(wb, sheet)
  }
  addWorksheet(wb, sheet)
}


wb <- if (file.exists(outfile)) {
  loadWorkbook(outfile)
} else {
  createWorkbook()
}

replace_sheet(wb, "alliance_test")

  setColWidths(
    wb,
    sheet = "alliance_test",
    cols = 1:ncol(alliance.support$support),
    widths = "auto"
  )

writeData(wb, "alliance_test", alliance.support$support)

replace_sheet(wb, "assoc_test")

  setColWidths(
    wb,
    sheet = "assoc_test",
    cols = 1:ncol(assoc.support$support),
    widths = "auto"
  )
  writeData(wb, "assoc_test", assoc.support$support)

  saveWorkbook(
    wb,
    outfile,
    overwrite = TRUE
  )

```

What species separate units

indicator species analysis - should some  associations or alliances be combined. Flag site units with few of the hierarchy unit indicator species?


```{r}
source("../BEC_R_working/_functions/_indicator_species_analysis.R")

ind.alliance <- indicator_species_analysis(
  veg.dat,
  su.anal,
  group_level = "working_alliance",
  max.order = 2,
  min.constancy = 40,
  min.fidelity = 15,
  p.cutoff = 0.05
)

alliance.indic <- ind.alliance$indicators

ind.assoc.alliance <- indicator_species_analysis(
  veg.dat,
  su.anal,
  group_level = "working_association",
  within_level = "working_alliance",
  max.order = 1,
  min.constancy = 50
)
assoc.indic <- ind.assoc.alliance$indicators
```



11. Environmental differentiation

Elevation Slope Aspect MAT MAP Soil moisture Nutrient regime

Use:

RDA rda.mod \<- rda( comm \~ elevation + moisture + nutrients, data = env )

or

envfit on NMDS envfit(nmds, env)

This links community differences to environmental gradients. function to pull climate data and other environment data

```{r}
su.anal <- su.anal
## read in plot data and extract lat/long and environment info
becmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster/BECMaster_fixing.accdb;")
plot.env <- dbReadTable(becmaster, "BECMaster_fixing_Env") %>% mutate(lon = ifelse(Longitude<0, Longitude, 0-Longitude)) %>% mutate(Date = as_date(Date)) |> 
  filter(PlotNumber %in% su.anal$PlotNumber) |>
  dplyr::select(PlotNumber, lon, lat = Latitude, elev = Elevation, special = Exposure1, rSMR = MoistureRegime, rSNR = NutrientRegime, slope = SlopeGradient, aspect = Aspect ) 
plot.info <- dbReadTable(becmaster, "BECMaster_fixing_Admin") |> dplyr::select(Plot, BECSiteUnit, UserSiteUnit, HumusThickness, BGC=GIS_BGC)  |> filter(Plot %in% su.anal$PlotNumber)
plot.env <- left_join(plot.env, plot.info, by = c("PlotNumber" = "Plot")) |> filter(!is.na (rSMR))
#odbc::odbcListColumns(becmaster, table = "BECMaster_fixing_plot.env")
dbDisconnect(becmaster)

## extract climate data for plots from climr
require(climr)
plot.locs <- plot.env %>% dplyr::select(id=PlotNumber, BGC, lon,lat,elev) |> drop_na(lat)
plot.climate <- climr::downscale(setDT(plot.locs),
                              vars = list_vars(set = c("Annual", "Seasonal", "Monthly")),
                              return_refperiod = TRUE,
                             # obs_periods = c("2001_2020"),
                              #obs_years = 2001:2024,
                              # gcms = c("ACCESS-ESM1-5","CNRM-ESM2-1", "EC-Earth3", "GFDL-ESM4", "GISS-E2-1-G", "MIROC6",  "MPI-ESM1-2-HR", "MRI-ESM2-0",  "UKESM1-0-LL" ),
                              # ssps = c("ssp126", "ssp245", "ssp370"),
                              # gcm_ssp_years = c(2025:2045),
                              #ensemble_mean = TRUE,
                              #obs_ts_dataset = "cru.gpcc",
                              #obs_ts_dataset = "mswx.blend",
                              db_option = "local",
                              cache = TRUE)
## calculate aSMR
plot.env2 <- plot.climate |> select(PlotNumber = id, everything()) |> left_join(plot.env) |> 
  mutate(
    moist = rowSums(
      across(c(PPT_05, PPT_06, PPT_07, PPT_08, PPT_09), ~ .x >= 90),
      na.rm = TRUE
    ),
    moist = ifelse(moist > 4, 1, 0),
    ratio = moist / 5,
    multiplier = 1 + ratio,
    CMD = CMD_an / multiplier
  ) |> mutate(PPT.dorm = PPT_at + PPT_wt) |> 
  mutate(Def = ifelse(CMD >=350, 350 - PPT.dorm, CMD)) |> 
  mutate(CMD.total = CMD + Def) |> 
  mutate(CMD.site = ifelse(slope > 35 & (aspect >180 & aspect <270), CMD.total *1.25,
                           ifelse(slope> 35 & (aspect>315 | aspect <45), CMD.total *0.85, CMD.total))) |> arrange(PlotNumber) |> 
  mutate(
    rSMR.clean = sub("^.*?(\\d+).*$", "\\1", rSMR)) |> 
  mutate(rSMR.clean = as.numeric(rSMR.clean),
  SNR.clean = recode(
    substr(trimws(rSNR), 1, 1),
    "A" = 1L,
    "B" = 2L,
    "C" = 3L,
    "D" = 4L,
    "E" = 5L,
    "F" = 6L,
    .default = NA_integer_
  ))
  ####To adjust in zones with prolonged snowpack remove CMD from early growing season months
## BGCs with or below heavy snow packs
# BGC_Special <- fread("./inputs/BGC_Special.csv", data.table = FALSE)
# allDat2 <- left_join(allDat2,BGC_Special, by = c("BGC" = "BGC"))
# allDat2 <- allDat2 %>% mutate(CMD = ifelse(is.na(Special), CMD, CMD_08)) %>% select(-Special, -CMD_08) %>% drop_na()
#rules <- read.csv("./inputs/aSMR_growingseason_Rules_HalfStep_v13_13Sept2025.csv")
rules <- read.csv("./inputs/aSMR_Rules_HalfStep_v13_23June2025.csv") |>
  mutate(
    rSMR.rule = sub("^rSMR", "", rSMR)) |> dplyr::select(-rSMR)

plot_asmr <- plot.env2 %>% select(PlotNumber, rSMR, CMD.site) |> 
  left_join(rules, by = c("rSMR" = "rSMR.rule")) %>%
  filter(CMD.site <= CMD) %>%
  group_by(PlotNumber) %>%  # ensure row-wise logic if needed
  slice_max(aSMR, with_ties = FALSE) %>%
  ungroup() %>% select(PlotNumber, aSMR)

plot.env2 <- left_join(plot.env2, plot_asmr)
# fwrite(rsmr_asmr_long, file = "./outputs/aSMR_by_rSMR_long.csv")# for use in droughtprone sites
# all_cmd <- rsmr_asmr_long   %>%
#   arrange(BGC) %>%
#   pivot_wider(names_from = rSMR, values_from = aSMR)
# fwrite (all_cmd, "./outputs/CMDcalc_aSMR_by_rSMR_matrix.csv")

```

## Environment of units

### PCA of environment variables for analysis group

```{r}

source("../BEC_R_working/_functions/_plot_pca_hierarchy.R")

#Use site unit summarized values for graphic
library(forcats)

workin
su.env <- plot.env2 |>
  left_join(
    su.anal, 
    by = "PlotNumber"
  ) |>
  filter(!is.na(alliance)) |>
  group_by(siteunit, assoc, alliance) |>
  summarise(
    PAS_an = mean(PAS_an, na.rm = TRUE),
    #CMD.site = mean(CMD.site, na.rm = TRUE),
    #MSP   = mean(MSP, na.rm = TRUE),
    aSMR = mean(aSMR, na.rm = TRUE),
    SNR = mean(SNR.clean, na.rm = TRUE),
    AHM = mean(AHM, na.rm = TRUE),
    lat = mean(lat, na.rm = TRUE),
    FFP   = mean(FFP, na.rm = TRUE),
    DD5 = mean(DD5_an, na.rm = TRUE),
    DD18_sm = mean(DD18_sm, na.rm = TRUE),
    EXT = mean(EXT, na.rm = TRUE),
    .groups = "drop"
  )

# su.env <- env.alliance |>
#   select(-matches("[0-9]$")) |>
#   group_by(siteunit, assoc, alliance) |>
#   summarise(
#     across(
#       where(is.numeric),
#       ~ mean(.x, na.rm = TRUE)
#     ),
#     .groups = "drop"
#   )


#==============================================================================
# Climate variables for PCA
#==============================================================================

env.vars <- su.env |>
  select(-siteunit, -assoc, -alliance) 

# Remove rows with missing values
keep.rows <- complete.cases(env.vars)

env.vars <- env.vars[keep.rows, ]

# Remove constant variables if any
env.vars <- env.vars |>
  dplyr::select(
    where(~ sd(.x, na.rm = TRUE) > 0)
  )

su.env <- su.env[keep.rows, ]

pca.env <- plot_pca_hierarchy( su.env,
                    n_vectors = 10,
    output_file = NULL,
    save_html = FALSE)

pca.env$plot
boxplot(PC1 ~alliance, data = pca.env$scores)
boxplot(PC2 ~alliance, data = pca.env$scores)
boxplot(PC3 ~alliance, data = pca.env$scores)

```


## Vegetation Summarize by Alliance

Generate vegetation summary output for all proposed alliances on a single sheet

## Constancy and cover tables

Generate vegetation summary for all site units by generated alliance colour-themed by generated association and subassociation for review.

Each sheet has all site units for a proposed alliance on a single sheet grouped, labelled, and colour-themed by generated Association  to help review and editing of membership. This will be used to review the proposed alliances and make decisions about which site units should be merged into associations, which should be separated into new associations, and which should be left as singletons. The vegetation summary tables will also be used to identify potential diagnostic species for each alliance and to compare the vegetation composition of different alliances.

```{r summary table association within alliances}
# compute anal.test
# becdb <- dbConnect(
#   odbc::odbc(),
#   .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v0.accdb;"
# )
# BEC.hier.temp <- dbReadTable(becdb, "tblBEC_Hierarchy_Class") #%>% dplyr::select(-SiteUnit)
# dbDisconnect(becdb)

group.to.test = "generated_Alliance"
BEC.hier.temp <- generated.hier |>
  filter(!is.na(.data[[group.to.test]]))

su.anal.new <- su.anal1 %>% dplyr::select(PlotNumber, SU_Label) %>% left_join(BEC.hier.temp, by = c("SU_Label" = "SU_Label")) %>% #rename(SiteUnit = SiteSeries) %>%
  mutate(SiteUnit = str_replace(SiteUnit, "[abc]$", ""),# ignore phases
         SiteUnit = if_else(
           str_detect(SiteUnit, "101"),
           str_replace(SiteUnit, "\\.[123]$", ""), # remove zonal variations
           SiteUnit
         ))

source("../BEC_R_working/_functions/_create_veg_summary_workbook.R")

create_veg_summary_workbook(
  su.anal = su.anal,
  veg.dat = veg.dat2,
  taxon.lifeform = taxon.lifeform,  
  generated.hier = BEC.hier.temp,
  group.to.test = group.to.test,
  anal.set = anal.set
)

```

```{r compare alliances}
reduced.exceptions <- c("SPHAGNUM","CLADONIA","CLADINA","RACOMITR","MNIUM")
reduced.lifeforms  <- c(9, 10, 11)
su.alliance <- su.anal %>% mutate(SiteUnit = working_Alliance) %>% select(PlotNumber, SiteUnit)

#------------------------------------------------------------
# 3. Run pairwise vegetation similarity analysis
#------------------------------------------------------------
pairwise.params <- list(
  su                = su.alliance,
  minconstancy          = 0,
  noiseconstancy        = 10,
  minplots              = 1,
  minor                 = 1,
  use.ksi               = FALSE,
  ksi   = NULL,
  ksi.value             = 1.5,
  reduce.lifeform       = TRUE,
  reduced.lifeforms     = reduced.lifeforms,
  reduction             = 0.1,
  reduced.exceptions    = reduced.exceptions,
  add_vegsum_table      = NULL
)

vegsum.pairs <- do.call(
  do_pairwise,
  c(list(veg.dat2), pairwise.params)
)

# Start with everything except `su`
print.params <- pairwise.params[setdiff(names(pairwise.params), "su")]

# Remove ksi if use.ksi == FALSE
if (!pairwise.params$use.ksi) {
  print.params <- print.params[setdiff(names(print.params), c("ksi", "ksi.value"))]
}

# Remove reduction-related fields if reduce.lifeform == FALSE
if (!pairwise.params$reduce.lifeform) {
  print.params <- print.params[setdiff(names(print.params),
                                       c("reduced.lifeforms", "reduction", "reduced.exceptions"))]
}
library(tibble)
library(knitr)

param.table <- tibble(
  parameter = names(print.params),
  value = sapply(print.params, function(x) paste(x, collapse = ", "))
)

param.table %>%
  gt::gt() %>%
  gt::tab_header(title = "Parameter Settings for do_pairwise()") %>%
  gt::tab_options(
    table.font.size = gt::px(8),
    table.border.top.width = gt::px(1),
    table.border.bottom.width = gt::px(1)
  )


#------------------------------------------------------------
# 4. Extract comparison table
#------------------------------------------------------------
fwrite(vegsum.pairs, "./outputs/vegsum_pairs.csv")

### return unit species information
vegsum.pairs.spp <- vegsum.pairs |> dplyr::select(Unit1, Unit2, Species, ScientificName, MeanCov.unit1, Constancy.unit1, MeanCov.unit2, Constancy.unit2, diff.pts.unit1, diff.pts.unit2)

### return distinct unit comparisons

unit.compare1 <- vegsum.pairs %>%
  dplyr::select(
    Unit1, Unit2, diff.ratio.unit1, diff.ratio.unit2, nplots.unit1, nplots.unit2, unit.diag.sum.unit1, unit.diag.sum.unit2
  ) %>%
  distinct(Unit1, Unit1, .keep_all = TRUE)


unit.compare <- unit.compare1 %>% 
    dplyr::select(
      Unit1 = Unit2,
      Unit2 = Unit1,
      diff.ratio.unit1 = diff.ratio.unit2,
      diff.ratio.unit2 = diff.ratio.unit1,
      nplots.unit1 = nplots.unit2,
      nplots.unit2 = nplots.unit1,
      unit.diag.sum.unit1 = unit.diag.sum.unit2,
      unit.diag.sum.unit2 = unit.diag.sum.unit1
      ) |>  rbind(unit.compare1) |> 
  distinct() |> mutate(BEC.sim.min = pmin(diff.ratio.unit1, diff.ratio.unit2),
                      BEC.sim.max = pmax(diff.ratio.unit1, diff.ratio.unit2))


fwrite(unit.compare , "./outputs/alliance_compare.csv")
#------------------------------------------------------------
# 5. Build BEC distance matrix and export
#------------------------------------------------------------

source("../BEC_R_working/_functions/_bec_dist_matrix.R")

bec_dist_matrix <- bec_dist_matrix(
  vegsum.pairs,
  distance = "BEC.sim.min"
) %>% 
  data.frame()

openxlsx::write.xlsx(
  x = bec_dist_matrix,
  file = "./clean_tabs/Coast_Forest_Alliance_Distance_matrix.xlsx",
  rowNames = TRUE
)
```

```{r summary report by alliances}
author = "MH_Forest_Alliance"

vegsum.wbk <- createWorkbook()
  # Build vegetation summary for this group
  vegSum_sub <- create_veg_sum_all(
    vdat = veg.dat2,
    siteUnits = su.alliance,
    minconstancy = 50,
    noiseconstancy = 25,
    strata.by = "Lifeform"
  )

  # Format table
  veg.sum.table_sub <- format_veg_table_nostrata(
    vsum = vegSum_sub,
    spp = taxon.lifeform
  )
sheet_name <- author
  # Name the sheet
# sheet_name <- paste0("Group_", g)

# Trim to 25 chars, then add a 5‑char hash suffix
# if (nchar(sheet_name) > 31) {
#   suffix <- substr(digest::digest(sheet_name), 1, 5)
#   sheet_name <- paste0(substr(sheet_name, 1, 25), "_", suffix)
# }

addWorksheet(vegsum.wbk, sheetName = sheet_name)


  # Write data
  writeData(vegsum.wbk, sheet = sheet_name, veg.sum.table_sub)
#}

saveWorkbook(
  vegsum.wbk,
  paste0("./vegsum.tables/", author, ".xlsx"),
  overwrite = TRUE
)

```

```{r build veg guide summary and export to excel, echo = FALSE, warning=FALSE, message=FALSE}
# 
# data.path = "D:/OneDrive - Government of BC/GitHub_data"
# author <- "test.assoc1"
# assoc.list <- bgcv13 %>% filter(!is.na(Association_Code)) %>% group_by(Association_Code) %>% mutate(n=n()) %>% filter(n>2) %>% distinct(Association_Code) %>% pull(Association_Code)
# 
# vegsum.wbk <- createWorkbook()
# for(assoc.choose in assoc.list){
# #assoc.choose = "assoc-13"
# assoc <- bgcv13 %>% filter(Association_Code %in% assoc.choose) %>% dplyr::select(SS_Label, Association_Code) %>% mutate(SiteUnit = gsub(" ", "", SS_Label)) %>% mutate(SiteUnit = str_replace(SiteUnit, "/", "_"))
# su.assoc <- left_join(assoc, su, by = "SiteUnit") %>%  mutate(SiteUnit = str_replace(SiteUnit, "b$|a$", "")) %>% distinct() 
# veg.dat.assoc <- veg.dat2 %>% filter(PlotNumber %in% su.assoc$PlotNumber)
# # vegsum.wbk <- createWorkbook()
# # #bgc.choose = "CDFmmtab2"
# # #for(bgc.choose in bgc.list){
# vegSum <-
#   create_veg_sum_all(vdat = veg.dat.assoc, siteUnits = su.assoc, strata.by = "Lifeform")
# 
# indic.order <- build_species_ordering_hier(vdat = veg.dat.assoc, vsum = vegSum, code.lump=lump, siteUnits = su.assoc, hier.level = "Association_Code")
# veg.sum.table <- format_veg_table2(vsum = vegSum, spp=taxon.lifeform)
# openxlsx::addWorksheet(vegsum.wbk, sheetName = assoc.choose)
# openxlsx::writeData(vegsum.wbk, sheet = assoc.choose, veg.sum.table) 
# }
# 
# saveWorkbook(vegsum.wbk, paste0("./vegsum.tables/LMH77_draft_assoc_tables_", author, ".xlsx"), overwrite = TRUE)

```

In your Baselga implementation, overall similarity is captured by the Sørensen analogue:

Ssor=2A2A+B+CS\_{sor}=\frac{2A}{2A+B+C}

while the dissimilarity is partitioned into:

βsor=βsim+βnes\beta*{sor}=\*\beta{sim}+\beta\_{nes}

where:

Turnover dissimilarity (βsim\beta*{sim}) = replacement of diagnostic potential between units. Nestedness dissimilarity (βnes\*\beta{nes}) = differences caused by one unit essentially being a subset of the other.

The two fractions:

1.  turnover_fraction = turnover_dissimilarity / sorensen_dissimilarity

2.  nestedness_fraction = nestedness_dissimilarity / sorensen_dissimilarity

Practical interpretation guide Pattern Ecological meaning

- High similarity + high nestedness ratio One unit largely contains a subset of the other's diagnostic composition

- High similarity + high turnover ratio Communities are almost identical but differ via species replacement

- Low similarity + high nestedness ratio Strong subset relationship; one unit much poorer than the other

- Low similarity + high turnover ratio Strong compositional replacement; distinct communities

For BEC comparisons, I would interpret high nestedness fractions as indicating a diagnostic signal that is largely shared but differs in strength or completeness, whereas high turnover fractions indicate that different species are carrying the diagnostic signal in each unit.

```{r}
# baselga_pairs <- baselga_diagnostic(vegsum.pairs, join_back = TRUE) |> filter(BEC.sim.min >= 0.8)
# # filter(BEC.sim.min <= 0.8 &  BEC.sim.min >= 0.6 & nestedness_fraction >=0.8)
# 
# ggplot(
#   baselga_pairs,
#   aes(
#     sorensen_similarity,
#     turnover_fraction
#   )
# ) +
#   geom_point(alpha = 0.6) +
#   theme_bw()

```

## Evaluate site series within each BGC

1.  Identify site series that have fewer than 5 plots (difficult to quantitatively analyse)

2.  Identify site series that have low diagnostic potential.

3.  Compare site series to identify site series below the minimal threshold to separate.

Some units with few plots or low diagnostics fall out as singles in the first pass cluster analysis. These are reviewed and merging rules applied before re-running the cluster analysis. List of site series with too few plots or low diagnostic potential for reliable analysis


## Review environmental space of site units within associations

Now look for environmental outliers of site units that have been merged into each association.

### ID Edaphic outliers from central concept of associations

Compare the median aSMR and median SNR value for each site unit to the median value for the entire association and flag those that are more than 1 edatopic step difference from the group median. This is a simple way to identify site units that may be outliers in edaphic space and may need to be separated into a different association or matched to a different association. aSMR is generated using the modelled relationship between rSMR and aSMR for the coastal BGCs.

```{r test for edaphic outliers1}
# source("../BEC_R_working/_functions/_test_edaphic_outliers.R")
# source("../BEC_R_working/_functions/_find_edaphic_similarity.R")
# group = "Association"
# 
# SU_edatope <- plot.edatope %>%
#   left_join(su.anal2, by = "PlotNumber") %>%
#   #rename(SiteUnit_orig = SiteUnit) %>%   # keep original
#   left_join(BEC.hier, by = c("SiteUnit" = "SiteUnit")) %>%
#   mutate(SiteUnit = coalesce(Association, SiteUnit)) %>%  # fallback
#   filter(Analysis_Group == anal.set) %>%
#   dplyr::select(SiteUnit, aSMR, SNR) %>%
#   filter(!is.na(SiteUnit)) %>%
#   mutate(
#     SNR = factor(SNR, levels = c("A", "B", "C", "D", "E"), ordered = TRUE),
#     SNR_num = as.numeric(SNR)
#   ) %>%
#   dplyr::select(-SNR) %>%
#   filter(!is.na(SNR_num))
# 
# 
# 
# median_by_site <- SU_edatope %>%
#   group_by(SiteUnit) %>%
#   summarise(across(where(is.numeric),
#                    \(x) median(x, na.rm = TRUE),
#                    .names = "median_{.col}"))
# 
# 
# su_all_sim2 <- su_all_sim %>% left_join(median_by_site, by = c("Unit1" = "SiteUnit")) %>% left_join(median_by_site, by = c("Unit2" = "SiteUnit")) %>% distinct %>% 
#   dplyr::select(Unit1, Unit2, BEC.sim.min, tempAssoc1, tempAssoc2, smr1 = median_aSMR.x, smr2 = median_aSMR.y, snr1 = median_SNR_num.x, snr2 = median_SNR_num.y, tempAssoc_match )
# 
# 
# su_all_sim2_rev <- su_all_sim2 %>% 
#   dplyr::select(Unit1 = Unit2, Unit2 = Unit1, BEC.sim.min, tempAssoc1 = tempAssoc2, tempAssoc2 = tempAssoc1, smr1 = smr2, smr2 = smr1, snr1 = snr2, snr2 = snr1, tempAssoc_match )
# 
# # bind and sort
# su_all_sim2_full <- bind_rows(su_all_sim2, su_all_sim2_rev) %>%
#   arrange(Unit1, Unit2)
# 
# write.xlsx(
#   su_all_sim2_full,
#   file = "./review_materials/other_similar_su_cluster_review_edatope.xlsx",
#   overwrite = TRUE
# )

#fwrite(su_all_sim2_full, "./review_materials/other_similar_sus_cluster_review_edatope.")
```

```{r test for edaphic outliers2}
# group = "generated_Alliance"
# 
# 
# 
# assoc_edatope <- plot.edatope %>%
#   left_join(su2, by = "PlotNumber") %>%
#   left_join(BEC.hier, by = c("SiteUnit" = "SiteUnit")) %>%
#   filter(Analysis_Group == anal.set) %>%
#   dplyr::select(SiteUnit, aSMR, SNR, groups = group) %>%
#   filter(!is.na(SiteUnit)) %>%
# #  filter(groups %in% select.groups) %>%
#   mutate(
#     SNR = factor(SNR, levels = c("A", "B", "C", "D", "E"), ordered = TRUE),
#     SNR_num = as.numeric(SNR)
#   ) %>% filter(!is.na(SNR))
# 
# all_results <- lapply(sort(unique(assoc_edatope$groups)), function(g) {
#   
#   res <- test_edaphic_outliers(
#     df = assoc_edatope,
#     group_name = g
#   )
#   
#   list(
#     group = g,
#     range_overlap = res$range_overlap %>% mutate(group = g),
#     median_flags  = res$median_flags  %>% mutate(group = g)
#   )
# })
# 
# names(all_results) <- sort(unique(assoc_edatope$groups))
# 
# 
# range_all <- bind_rows(
#   lapply(all_results, function(x) x$range_overlap)
# )
# range_all  <- bind_rows(lapply(all_results, function(x) x$range_overlap))
# edatopic_flag_all  <- bind_rows(lapply(all_results, function(x) x$median_flags)) %>% dplyr::select(group, SiteUnit, flag_aSMR, aSMR_median, group_aSMR_median, flag_SNR,SNR_median, group_SNR_median, flag_any, everything())
# edatopic_outlier <- edatopic_flag_all %>% filter(flag_any == TRUE)
# 
# edatopic_outlier_gt <- edatopic_outlier %>%
#   dplyr::select(
#     Group = group,
#     SiteUnit,
#     aSMR_Median = aSMR_median,
#     Group_aSMR_Median = group_aSMR_median,
#     aSMR_Flag = flag_aSMR,
#     SNR_Median = SNR_median,
#     Group_SNR_Median = group_SNR_median,
#     SNR_Flag = flag_SNR
#   ) 
# 
# gt(edatopic_outlier_gt) %>%
#   gt::tab_header(
#     title = "Edatopic Outlier Summary",
#     subtitle = "SiteUnits flagged as edaphic outliers based on median and range diagnostics"
#   ) %>%
#   gt::cols_label(
#     Group = "Association Group",
#     SiteUnit = "Site Unit",
#     aSMR_Median = "Plot aSMR Median",
#     Group_aSMR_Median = "Group aSMR Median",
#     aSMR_Flag = "aSMR Outlier",
#     SNR_Median = "Plot SNR Median",
#     Group_SNR_Median = "Group SNR Median",
#     SNR_Flag = "SNR Outlier"
#   ) %>%
#   gt::tab_style(
#     style = cell_fill(color = "#FFF3CD"),
#     locations = cells_body(
#       columns = c(aSMR_Flag, SNR_Flag),
#       rows = aSMR_Flag == TRUE | SNR_Flag == TRUE
#     )
#   ) %>%
#   cols_width(everything() ~ px(1)) %>%   # <— auto-fit trick
#   tab_options(
#     table.font.size = px(7),
#     data_row.padding = px(0),
#     table.layout = "auto"
#   )
```

### ID Climatic outliers from central concept of associations

Several approaches are applied. This first looks at univariate range of 4 climate variables (MCMT, DD5, MSP, PAS) for each site unit compared to the range of the rest of the association. This is a simple way to identify site units that may be outliers in climate space and may need to be separated into a different association or matched to a different association. However, this approach does not take into account the covariance between variables and may miss outliers that are extreme in combination but not in any one variable. The second approach is to use mahalanobis distance in multivariate climate space to identify site units that are outliers in climate space compared to the rest of the association. These units may be candidates for merging into a different association with more similar climate or for being separated into their own association if they are sufficiently different from all other associations.

```{r test for climatic outliers}
# climate_vars <- c("MCMT", "DD5_an",  "PAS_an","MSP" ,"CMD_an")
# source("../BEC_R_working/_functions/_climate_outliers.R")
# source("../BEC_R_working/_functions/_climate_comparison.R")
# climate_compared <- climate_outliers(
#   plot_climate = plot.climate,
#   su_anal = su.anal.new,
#   climate_vars = climate_vars,
#   group_col = "generated_Assoc",
#   unit_col = "SiteUnit",
#   low_overlap_threshold = 0.10,
#     alpha_global = .90,
#     alpha_local = .90)
# 
# mahal_outliers <- climate_compared$within_group_outliers %>% dplyr::select(SiteUnit, groups, distance_global, cutoff_global, is_outlier_global) %>% filter(is_outlier_global == TRUE) %>% arrange(groups, desc(distance_global)) %>% drop_na(groups)%>%
#   dplyr::select(
#     Group = groups,
#     SiteUnit,
#     Distance = distance_global,
#     Cutoff = cutoff_global
#   )
# 
#   gt(mahal_outliers) %>%
#   tab_header(
#     title = "Climate Outliers: Global Mahalanobis Distance",
#     subtitle = "SiteUnits exceeding the global multivariate cutoff"
#   ) %>%
#   fmt_number(
#     columns = c(Distance, Cutoff),
#     decimals = 2
#   ) %>%
#   tab_options(
#     table.font.size = px(8),
#     data_row.padding = px(1)
#   )
# 
#   
# univariate_outliers <- climate_compared$climate_outliers %>% dplyr::select(SiteUnit, groups, variable, mean_su, mean_grp) %>% arrange(groups, SiteUnit) %>% drop_na(groups) %>%
#   dplyr::select(
#     Group = groups,
#     SiteUnit,
#     Variable = variable,
#     SU_Mean = mean_su,
#     Group_Mean = mean_grp
#   )
# gt(univariate_outliers) %>%
#   tab_header(
#     title = "Climate Outliers: Univariate Deviations",
#     subtitle = "SiteUnits with mean climate values outside group envelopes"
#   ) %>%
#   fmt_number(
#     columns = c(SU_Mean, Group_Mean),
#     decimals = 2
#   ) %>%
#   cols_width(
#     everything() ~ px(1)   # auto-fit trick
#   ) %>%
#   tab_options(
#     table.font.size = px(7),
#     data_row.padding = px(0),
#     table.layout = "auto"
#   )

#mahalabobis_outliers <- climate_compared$mahal_all  %>% arrange(group, SiteUnit) %>% drop_na(group)
#mahal_dist<- mahal_results$within_group_mahal %>% drop_na(groups)
```

## Look for possible placements for single site units

Find nearest neighbours for single units (singletons) and compare vegetation, edaphic, and climatic similarity to identify potential matches for merging into existing groups. First find the nearest vegetation neighbours for site units and return the min and max BEC similarity Then for each pair which are \> 70% similar compare edatopic medians and climate similarity If high bec.sim.max and close environmental then consider merging the singleton into the neighbour unit Expect that site units in geographically peripheral BGCs may have no BC matches but may match to extrajurisdictional types (for Coast region Alaska or Washington State are most likely sources.)

```{r}
# source("../BEC_R_working/_functions/_find_nearest_neighbours.R")
# group_edaphic_medians <- edatopic_flag_all %>%
#   group_by(group) %>%
#   summarise(
#     group_SNR_median  = median(SNR_median, na.rm = TRUE),
#     group_aSMR_median = median(aSMR_median, na.rm = TRUE),
#     .groups = "drop"
#   )
# ungrouped_SUs <- assoc_edatope %>%
#   filter(is.na(groups)) %>%
#   distinct(SiteUnit)
# 
# edaphic_similarity <- find_edaphic_similarity (
#   edatopic_table = edatopic_flag_all,
#   df_raw = assoc_edatope
# ) %>% 
# arrange(SiteUnit )
# climate_similarity <-  climate_compared$mahal_all %>% arrange(SiteUnit)
# 
# #unit.compare = unit.compare; single.units = single.units
# nearest_units <- find_nearest_neighbours(unit.compare.new, single.units) %>% rename(SiteUnit = Singleton, group = Neighbor) 
# 
# nearest <- nearest_units  %>%
#   left_join(climate_similarity, by = c("SiteUnit", "group")) %>% 
#   left_join(edaphic_similarity, by = c("SiteUnit", "group")) %>%
#   arrange(SiteUnit, distance) %>% drop_na
# 
# fwrite(nearest, "./outputs/veg_climate_edaphic_similarity.csv", row.names = FALSE)
# check.match <- nearest %>% filter(distance < 7.78, BEC.sim.max >.65) %>%
#   dplyr::select(
#     SiteUnit,
#     Group = group,
#     Distance = distance,
#     BEC_Similarity = BEC.sim.max
#   ) %>%
#   arrange(SiteUnit, Distance)
# 
#   gt(check.match) %>%
#   tab_header(
#     title = "Nearest Neighbour Matches",
#     subtitle = "Filtered by climate distance < 7.78 and BEC similarity > 0.65"
#   ) %>%
#   fmt_number(
#     columns = c(Distance, BEC_Similarity),
#     decimals = 2
#   ) %>%
#   tab_style(
#     style = cell_fill(color = "#E8F5E9"),
#     locations = cells_body(
#       rows = Distance < 7.78 & BEC_Similarity > 0.65
#     )
#   ) %>%
#   tab_options(
#     table.font.size = px(8),
#     data_row.padding = px(1)
#   )


```

### Environment summaries

Horizonatal box plots of site unit medians by association and alliance

This links community differences to environmental gradients. function to pull climate data and other environment data

```{r}
# su.anal <- su.anal.new
# ## read in plot data and extract lat/long and environment info
# becmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster/BECMaster_fixing.accdb;")
# plot.env <- dbReadTable(becmaster, "BECMaster_fixing_Env") %>% mutate(lon = ifelse(Longitude<0, Longitude, 0-Longitude)) %>% mutate(Date = as_date(Date)) |> 
#   filter(PlotNumber %in% su.anal$PlotNumber) |>
#   dplyr::select(PlotNumber, lon, lat = Latitude, elev = Elevation, special = Exposure1, rSMR = MoistureRegime, rSNR = NutrientRegime, slope = SlopeGradient, aspect = Aspect ) 
# plot.info <- dbReadTable(becmaster, "BECMaster_fixing_Admin") |> dplyr::select(Plot, BECSiteUnit, UserSiteUnit, HumusThickness, BGC=GIS_BGC)  |> filter(Plot %in% su.anal$PlotNumber)
# plot.env <- left_join(plot.env, plot.info, by = c("PlotNumber" = "Plot")) |> filter(!is.na (rSMR))
# #odbc::odbcListColumns(becmaster, table = "BECMaster_fixing_plot.env")
# dbDisconnect(becmaster)
# 
# ## extract climate data for plots from climr
# require(climr)
# plot.locs <- plot.env %>% dplyr::select(id=PlotNumber, BGC, lon,lat,elev) |> drop_na(lat)
# plot.climate <- climr::downscale(setDT(plot.locs),
#                               vars = list_vars(set = c("Annual", "Seasonal", "Monthly")),
#                               return_refperiod = TRUE,
#                              # obs_periods = c("2001_2020"),
#                               #obs_years = 2001:2024,
#                               # gcms = c("ACCESS-ESM1-5","CNRM-ESM2-1", "EC-Earth3", "GFDL-ESM4", "GISS-E2-1-G", "MIROC6",  "MPI-ESM1-2-HR", "MRI-ESM2-0",  "UKESM1-0-LL" ),
#                               # ssps = c("ssp126", "ssp245", "ssp370"),
#                               # gcm_ssp_years = c(2025:2045),
#                               #ensemble_mean = TRUE,
#                               #obs_ts_dataset = "cru.gpcc",
#                               #obs_ts_dataset = "mswx.blend",
#                               db_option = "local",
#                               cache = TRUE)
# ## calculate aSMR
# plot.env2 <- plot.climate |> select(PlotNumber = id, everything()) |> left_join(plot.env) |> 
#   mutate(
#     rSMR.clean = sub("^.*?(\\d+).*$", "\\1", rSMR)) |> 
#   mutate(rSMR = as.numeric(rSMR.clean),
#   SNR.clean = recode(
#     substr(trimws(rSNR), 1, 1),
#     "A" = 1L,
#     "B" = 2L,
#     "C" = 3L,
#     "D" = 4L,
#     "E" = 5L,
#     "F" = 6L,
#     .default = NA_integer_
#   )) |> 
#   mutate(
#     moist = rowSums(
#       across(c(PPT_05, PPT_06, PPT_07, PPT_08, PPT_09), ~ .x >= 90),
#       na.rm = TRUE
#     ),
#     moist = ifelse(moist > 4, 1, 0),
#     ratio = moist / 5,
#     multiplier = 1 + ratio,
#     CMD = CMD_an / multiplier
#   ) |> mutate(PPT.dorm = PPT_at + PPT_wt) |> 
#   mutate(CMD.total = ifelse(CMD >=350, 350 - PPT.dorm, CMD)) |> 
#     mutate(CMD.site = ifelse(rSMR == 0, CMD.total+500, 
#                            ifelse(rSMR == 1, CMD.total+375, 
#                                   ifelse(rSMR == 2, CMD.total +250,
#                                          ifelse(rSMR == 3, CMD.total +125,
#                                                 ifelse(rSMR == 5, CMD.total/2,
#                                                        ifelse(rSMR == 6, CMD.total/4,
#                                                               ifelse(rSMR == 7, CMD.total/8,
#                                                                      ifelse(rSMR == 8, CMD.total/16, CMD.total))))))))) |> 
#   mutate(CMD.site = ifelse(slope > 35 & (aspect >180 & aspect <270), CMD.site *1.25,
#                            ifelse(slope> 35 & (aspect>315 | aspect <45), CMD.site *0.85, CMD.site))) |> arrange(PlotNumber)
#   
#   ####To adjust in zones with prolonged snowpack remove CMD from early growing season months
# ## BGCs with or below heavy snow packs
# # BGC_Special <- fread("./inputs/BGC_Special.csv", data.table = FALSE)
# # allDat2 <- left_join(allDat2,BGC_Special, by = c("BGC" = "BGC"))
# # allDat2 <- allDat2 %>% mutate(CMD = ifelse(is.na(Special), CMD, CMD_08)) %>% select(-Special, -CMD_08) %>% drop_na()
# #rules <- read.csv("./inputs/aSMR_growingseason_Rules_HalfStep_v13_13Sept2025.csv")
# rules <- read.csv("./inputs/aSMR_Rules_HalfStep_v13_23June2025.csv") |>
#   mutate(rSMR.rule = as.numeric(sub("^rSMR", "", rSMR)), CMD = as.numeric(CMD)) |> dplyr::select(-rSMR)
# 
# plot_asmr <- plot.env2 %>%
#   select(PlotNumber, rSMR, CMD.site) %>%
#   left_join(rules, by = c("rSMR" = "rSMR.rule")) %>%
#   filter(CMD.site <= CMD) %>%
#   group_by(PlotNumber) %>%
#   slice_min(CMD, n = 1) %>%
#   ungroup() %>% select(PlotNumber, aSMR)
# 
# plot.env2 <- left_join(plot.env2, plot_asmr)
# 
# env.alliance <- plot.env2 |>
#   left_join(
#     su.anal, 
#     by = "PlotNumber") |> rename(alliance = generated_Alliance, assoc = generated_Assoc)
# 
# xx <- env.alliance |> filter(SiteUnit == "CWHms5_112")

```

## Environment of units

```{r}
#| label: fig-Assoc-env
#| fig-cap: ("Plot environment by Association")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12
# env.long <- env.alliance |>
#   select(siteunit = SiteUnit, assoc,  aSMR, SNR.clean, MSP, MAT) |>
#   mutate(
#     #alliance = str_remove(alliance, "^.*generated_"),
#     assoc    = str_remove(assoc, "^.*generated_")
#   ) |>
#   pivot_longer(
#     cols = c(aSMR, SNR.clean, MSP, MAT),
#     names_to = "variable",
#     values_to = "value",
#     values_drop_na = TRUE
#   ) |>
#   mutate(
#     assoc = reorder_within(assoc, value, variable, median)
#   )
# 
# library(Polychrome)
# 
# unit_levels <- sort(unique(env.long$siteunit))
# 
# cols <- createPalette(
#   length(unit_levels),
#   seedcolors = c(
#     "#E41A1C",
#     "#377EB8",
#     "#4DAF4A",
#     "#984EA3",
#     "#FF7F00"
#   )
# )
# 
# names(cols) <- unit_levels
# 
# ggplot(env.long, aes(assoc, value)) +
#   geom_boxplot(fill = "grey90", outlier.shape = NA) +
#   geom_jitter(
#     aes(colour = siteunit),
#     width = 0.1,
#     alpha = 0.6,
#     size = 2
#   ) +
#   facet_wrap(~variable, scales = "free") +
#   scale_x_reordered() +
#   coord_flip() +
#   scale_colour_manual(values = cols) +
#   theme(legend.position = "none")

  
```

```{r}
#| label: fig-Alliance-plotenv
#| fig-cap: ("Plot environment by Alliance")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12
# env.long <- env.alliance |>
#   select(siteunit = SiteUnit, assoc, alliance, aSMR, SNR.clean, MSP, MAT) |>
#   mutate(
#     alliance = str_remove(alliance, "^.*generated_"),
#     assoc    = str_remove(assoc, "^.*generated_")
#   ) |>
#   pivot_longer(
#     cols = c(aSMR, SNR.clean, MSP, MAT),
#     names_to = "variable",
#     values_to = "value",
#     values_drop_na = TRUE
#   ) |>
#   mutate(
#     alliance = reorder_within(alliance, value, variable, median)
#   )
# 
# library(Polychrome)
# 
# unit_levels <- sort(unique(env.long$siteunit))
# 
# cols <- createPalette(
#   length(unit_levels),
#   seedcolors = c(
#     "#E41A1C",
#     "#377EB8",
#     "#4DAF4A",
#     "#984EA3",
#     "#FF7F00"
#   )
# )
# 
# names(cols) <- unit_levels
# 
# ggplot(env.long, aes(alliance, value)) +
#   geom_boxplot(fill = "grey90", outlier.shape = NA) +
#   geom_jitter(
#     aes(colour = siteunit),
#     width = 0.2,
#     alpha = 0.6,
#     size = 2
#   ) +
#   facet_wrap(~variable, scales = "free") +
#   scale_x_reordered() +
#   coord_flip() +
#   scale_colour_manual(values = cols) +
#   theme(legend.position = "none")
# 
#   
```

```{r}
#| label: fig-medianSU-env
#| fig-cap: ("Median SiteUnit environment by alliance")
#| fig-cap-location: bottom
#| warning: false
#| fig-align: "center"
#| fig-width: 8
#| fig-height: 12
# env.sum <- env.alliance |>
#   select(siteunit = SiteUnit, assoc, alliance,aSMR, SNR.clean, MSP, MAT) |>#
#   mutate(
#     alliance = str_remove(alliance, "^.*generated_"),
#     assoc    = str_remove(assoc, "^.*generated_")
#   ) |>
#   pivot_longer(
#     cols = c(aSMR, SNR.clean, MSP, MAT),#
#     names_to = "variable",
#     values_to = "value",
#     values_drop_na = TRUE
#   ) |>
#   group_by(variable, siteunit, alliance) |>
#   summarise(
#     median = median(value, na.rm = TRUE),
#     #mean   = mean(value, na.rm = TRUE),
#     #sd     = sd(value, na.rm = TRUE),
#     .groups = "drop"
#   )
# alliance.order <- env.sum |>
#   group_by(variable, alliance) |>
#   summarise(ord = median(median), .groups = "drop")
# 
# library(tidytext)
# 
# env.sum <- env.sum |>
#   left_join(alliance.order, by = c("variable", "alliance")) |>
#   mutate(
#     alliance_plot = reorder_within(alliance, ord, variable)
#   )
# 
# library(ggrepel)
# 
# ggplot(
#   env.sum,
#   aes(
#     x = alliance_plot,
#     y = median,
#     colour = siteunit
#   )
# ) +
#   geom_errorbar(
#     aes(ymin = median,# - sd,
#         ymax = median),# + sd),
#     width = 0,
#     position = position_dodge(width = 0.4)
#   ) +
#   geom_point(
#     size = 2,
#     position = position_dodge(width = 0.4)
#   ) +
#   geom_text_repel(
#     aes(label = siteunit),
#     direction = "y",
#     size = 2,
#     show.legend = FALSE
#   ) +
#   facet_wrap(~variable, scales = "free") +
#   scale_x_reordered() +
#   coord_flip() +
#   scale_colour_manual(values = cols) +
#   theme_bw() +
#   theme(legend.position = "none")

```

## Test whether alliances differ PERMANOVA

A statistical test of compositional differences.

adonis2( comm \~ alliance, data = meta, method = "bray", permutations = 999 )

Outputs:

R² = how much variation is explained by alliance P-value

Given 3450 plots, you'll almost certainly get a significant result. Focus on effect size (R²), not just significance.

Test dispersion first

PERMANOVA is sensitive to unequal dispersion.

disp \<- betadisper( vegdist(comm, method = "bray"), alliance )

anova(disp)

This tells you whether differences are due to centroids (real compositional shifts) or spread.

3.  Quantify within-unit heterogeneity

Some alliances are very tight, others broad.

Beta diversity disp \<- betadisper( vegdist(comm, "bray"), alliance )

boxplot(disp)

Useful outputs:

Mean distance to centroid Rank alliances from homogeneous to heterogeneous

This is often ecologically informative.

```{r}

# pairwise_adonis_dist <- function(
#     dist.mat,
#     groups,
#     permutations = 999,
#     p.adjust.method = "BH"
# ) {
# 
#   groups <- droplevels(as.factor(groups))
# 
#   combs <- combn(levels(groups), 2)
# 
#   results <- data.frame()
# 
#   for (i in seq_len(ncol(combs))) {
# 
#     grp <- combs[, i]
# 
#     keep <- groups %in% grp
# 
#     # subset distance matrix
#     d.sub <- as.dist(
#       as.matrix(dist.mat)[keep, keep]
#     )
# 
#     meta.sub <- data.frame(
#       group = droplevels(groups[keep])
#     )
# 
#     mod <- adonis2(
#       d.sub ~ group,
#       data = meta.sub,
#       permutations = permutations
#     )
# 
#     results <- rbind(
#       results,
#       data.frame(
#         group1 = grp[1],
#         group2 = grp[2],
#         F = mod$F[1],
#         R2 = mod$R2[1],
#         p = mod$`Pr(>F)`[1]
#       )
#     )
#   }
# 
#   results$p.adj <- p.adjust(
#     results$p,
#     method = p.adjust.method
#   )
# 
#   results[order(results$R2, decreasing = TRUE), ]
# }
# 
# library(vegan)
# 
# pairwise_dispersion_dist <- function(
#     dist.mat,
#     groups,
#     p.adjust.method = "BH"
# ) {
# 
#   groups <- as.factor(groups)
# 
#   # Remove NAs
#   keep <- !is.na(groups)
# 
#   groups <- droplevels(groups[keep])
# 
#   dmat <- as.matrix(dist.mat)
#   dmat <- dmat[keep, keep]
# 
#   combs <- combn(levels(groups), 2)
# 
#   results <- data.frame()
# 
#   for (i in seq_len(ncol(combs))) {
# 
#     grp <- combs[, i]
# 
#     keep_pair <- groups %in% grp
# 
#     d.sub <- as.dist(
#       dmat[keep_pair, keep_pair]
#     )
# 
#     groups_sub <- droplevels(groups[keep_pair])
# 
#     bd <- betadisper(d.sub, groups_sub)
# 
#     aov_res <- anova(bd)
# 
#     results <- rbind(
#       results,
#       data.frame(
#         group1 = grp[1],
#         group2 = grp[2],
#         n1 = sum(groups_sub == grp[1]),
#         n2 = sum(groups_sub == grp[2]),
#         mean.dist1 = mean(
#           bd$distances[groups_sub == grp[1]]
#         ),
#         mean.dist2 = mean(
#           bd$distances[groups_sub == grp[2]]
#         ),
#         diff = abs(
#           mean(bd$distances[groups_sub == grp[1]]) -
#           mean(bd$distances[groups_sub == grp[2]])
#         ),
#         F = aov_res[1, "F value"],
#         p = aov_res[1, "Pr(>F)"]
#       )
#     )
#   }
# 
#   results$p.adj <- p.adjust(
#     results$p,
#     method = p.adjust.method
#   )
# 
#   results |>
#     arrange(p.adj)
# }
```

### This seems to produce some wonking results with units not separating that seem quite different.

```{r pairwise permanova}

# assess_group_support <- function(
#     veg.dat,
#     su.anal,
#     group.field = "generated_Alliance",
#     permutations = 999
# ) {
#     veg.matrix.mva <- veg.dat |>
#       filter(PlotNumber %in% su.anal2$PlotNumber) |>
#       dplyr::select(PlotNumber, Species, Cover) |>
#       arrange(PlotNumber) |>
#       matrify()
# 
#     ################################################
#     # Transform cover
#     ################################################
# 
#     veg.matrix.mva <- veg.matrix.mva^(1/3)
# 
#     ################################################
#     # Remove rare species
#     ################################################
# 
#     freq <- colSums(veg.matrix.mva > 0)
# 
#     veg.matrix.mva <- veg.matrix.mva[
#     ,
#     freq >= 0.10 * nrow(veg.matrix.mva),
#     drop = FALSE
#     ]
# 
# 
# plot.dist <- veg.matrix.mva |> vegdist(method = "bray")
#   groups <- su.anal2[[group.field]]
# 
#   adonis.res <- pairwise_adonis_dist(
#     dist.mat = plot.dist,
#     groups = groups,
#     permutations = permutations
#   )
# 
#   disp.res <- pairwise_dispersion_dist(
#     dist.mat = plot.dist,
#     groups = groups
#   )
# 
#   support.tbl <- adonis.res %>%
#     rename(
#       F.adonis = F,
#       p.adonis = p,
#       p.adj.adonis = p.adj
#     ) %>%
#     left_join(
#       disp.res %>%
#         rename(
#           F.disp = F,
#           p.disp = p,
#           p.adj.disp = p.adj
#         ),
#       by = c("group1", "group2")
#     ) %>%
#     mutate(
# 
#       score_R2 = case_when(
#         R2 >= 0.20 ~ 70,
#         R2 >= 0.15 ~ 60,
#         R2 >= 0.10 ~ 50,
#         R2 >= 0.05 ~ 25,
#         TRUE ~ 0
#       ),
# 
#       score_disp = case_when(
#         p.adj.disp > 0.10 ~ 20,
#         p.adj.disp > 0.05 ~ 10,
#         TRUE ~ 0
#       ),
# 
#       score_perm = if_else(
#         p.adj.adonis < 0.05,
#         10,
#         0
#       ),
# 
#       support_score =
#         score_R2 +
#         score_disp +
#         score_perm,
# 
#       support_class = case_when(
#         support_score >= 80 ~ "Strong",
#         support_score >= 60 ~ "Moderate",
#         support_score >= 40 ~ "Weak",
#         TRUE ~ "Review"
#       ),
# 
#       recommendation = case_when(
#         R2 < 0.05 ~ "Consider merge",
#         p.adj.disp < 0.05 ~ "Check heterogeneity",
#         R2 < 0.10 ~ "Review",
#         TRUE ~ "Supported"
#       ),
# 
#       taxonomic_level = group.field
#     ) %>%
#     arrange(support_score)
# 
#   summary.tbl <- support.tbl %>%
#     select(
#       group1,
#       group2,
#       R2,
#       p.adj.adonis,
#       p.adj.disp,
#       support_score,
#       support_class,
#       recommendation
#     ) %>%
#     arrange(support_score)
# 
#   list(
#     adonis = adonis.res,
#     dispersion = disp.res,
#     support = support.tbl,
#     summary = summary.tbl,
#     low_separation = support.tbl %>%
#       filter(R2 < 0.05)
#   )
# }
# 
# alliance.support <- assess_group_support(
#   veg.dat = veg.dat2,
#   su.anal = su.anal2,
#   group.field = "generated_Alliance",
#     permutations = 999
# )
# 
# 
# assoc.support <- assess_group_support(
#     veg.dat = veg.dat2,
#   su.anal = su.anal2,
#   group.field = "generated_Assoc",
#     permutations = 999
# )
# 
# 
# # ------------------------------------------------------------
#   # Export workbook
#   # ------------------------------------------------------------
#   out.dir = "./review_materials"
#   outfile <- file.path(
#     out.dir,
#     paste0(anal.set, "_other_similar_units.xlsx")
#   )
#   
#   replace_sheet <- function(wb, sheet) {
#   if (sheet %in% names(wb)) {
#     removeWorksheet(wb, sheet)
#   }
#   addWorksheet(wb, sheet)
# }
# 
#     
# wb <- if (file.exists(outfile)) {
#   loadWorkbook(outfile)
# } else {
#   createWorkbook()
# }
# 
# replace_sheet(wb, "alliance_test")
# 
#   setColWidths(
#     wb,
#     sheet = "alliance_test",
#     cols = 1:ncol(alliance.support$support),
#     widths = "auto"
#   )
#   
# writeData(wb, "alliance_test", alliance.support$support)
# 
# replace_sheet(wb, "assoc_test")
#   
#   setColWidths(
#     wb,
#     sheet = "assoc_test",
#     cols = 1:ncol(assoc.support$support),
#     widths = "auto"
#   )
#   writeData(wb, "assoc_test", assoc.support$support)
# 
#   saveWorkbook(
#     wb,
#     outfile,
#     overwrite = TRUE
#   )
#   
#   # invisible(
#   #   list(
#   #     summary = unit1_multiple_unit2,
#   #     review_table = units_multiple_assocs,
#   #     file = outfile
#   #   )
#   # )

```

## 4. mvabund analysis (composition and abundance differences) rather than Indicator species analysis (constancy only)

uses the package mvabund

For each significant species ask:

Is the p-value significant? Is frequency in the focal alliance high? Is frequency in other alliances low? Is mean cover meaningfully higher? Is the species present in enough plots (\>3 to 5)? Does the pattern make ecological sense?

Species meeting all six criteria are typically your best alliance indicators.

For vegetation classification, I'd place more weight on Frequency + Mean Cover + Consistency across pairwise comparisons than on the p-value alone. A species can be highly significant statistically but still be a poor diagnostic species if it is rare or occurs in only a few plots.

Prep data for pairwise statistical tests

```{r pairwise indicator species}

# Keep only SiteUnits with an alliance assignment
# alliance.good <- su.anal2 |> filter(!is.na(generated_Alliance)) |> rename(alliance = generated_Alliance)
# veg.matrix <- veg.dat2 |> filter(PlotNumber %in% alliance.good$PlotNumber) |> dplyr::select(PlotNumber, Species, Cover) |>
#   arrange(PlotNumber) |> matrify()
# 
# ##remove rare species
# freq <- colSums(veg.matrix > 0)
# veg.matrix2 <- veg.matrix[, freq >= 10]
# alliance2 <- droplevels(as.factor(alliance.good$alliance))
# 
# 
# table(alliance2)
# ## remove alliances with few members
# keep.grp <- alliance2 %in%
#   names(which(table(alliance2) >= 5))
# 
# veg.matrix2 <- veg.matrix2[keep.grp, ]
# 
# alliance2 <- droplevels(alliance2[keep.grp])
# table(alliance2)
```

```{r}
library(dplyr)
library(purrr)
library(tibble)
library(mvabund)

# ################################################
# # All pairwise combinations
# ################################################
# su.class <- su.anal2 |> rename(alliance = generated_Alliance) |> filter(alliance %in% alliance2)
# pairs <- combn(
#   sort(unique(su.class$alliance)),
#   2,
#   simplify = FALSE
# )
# 
# ################################################
# # Run pairwise comparisons
# ################################################
# #pair <- pairs[[1]]
# pairwise.results <- map_dfr(
#   pairs,
#   function(pair){
# 
#     alliance.choose <- su.class |>
#       filter(alliance %in% pair) |>
#       arrange(PlotNumber)
# 
# 
#     ################################################
#     # Model
#     ################################################
# 
#     alliance2 <- droplevels(
#       factor(alliance.choose$alliance)
#     )
# veg.sub <- veg.matrix2[
#     match(alliance.choose$PlotNumber,
#     rownames(veg.matrix2)),
# ,
# drop = FALSE
# ]
# 
#     mod.ab <- manylm(
#       mvabund(veg.sub) ~ alliance2
#     )
# 
#     ################################################
#     # Test
#     ################################################
# 
#     res <- anova(
#       mod.ab,
#       p.uni = "adjusted"
#     )
# 
# #################################################
# #     # Significant species
# #################################################
# # 
#     sig.species <- tibble(
#       Species = colnames(res$uni.test),
#       P_value = res$uni.p[2, ]
#     ) |>
#       filter(P_value <= 0.05)
# ################################################
# # Diagnostic statistics by alliance
# ################################################
# 
# diag.stats <- alliance.choose |>
#   select(PlotNumber, alliance) |>
#   crossing(
#     Species = sig.species$Species
#   ) |>
#   left_join(
#     veg.dat2 |>
#       select(PlotNumber, Species, Cover),
#     by = c("PlotNumber", "Species")
#   ) |>
#   mutate(
#     Cover = replace_na(Cover, 0)
#   ) |>
#   group_by(Species, alliance) |>
#   summarise(
#     N_Plots = n(),
#     N_Present = sum(Cover > 0),
#     Frequency = N_Present / N_Plots,
#     MeanCover_AllPlots = mean(Cover),
#     MeanCover_PresentOnly = ifelse(
#       N_Present > 0,
#       mean(Cover[Cover > 0]),
#       0
#     ),
#     SD_Cover = sd(Cover),
#     .groups = "drop"
#   )
# 
# 
#     if(nrow(sig.species) == 0)
#       return(NULL)
# 
# 
# ################################################
# # Put alliance diagnostics into columns
# ################################################
# 
# diag_A1 <- diag.stats %>%
#   filter(alliance == pair[1]) %>%
#   rename_with(~paste0(.x, "_A1"),
#               c(N_Plots, N_Present, Frequency,
#                 MeanCover_AllPlots,
#                 MeanCover_PresentOnly,
#                 SD_Cover))
# 
# diag_A2 <- diag.stats %>%
#   filter(alliance == pair[2]) %>%
#   rename_with(~paste0(.x, "_A2"),
#               c(N_Plots, N_Present, Frequency,
#                 MeanCover_AllPlots,
#                 MeanCover_PresentOnly,
#                 SD_Cover))
# 
# diag_pair <- diag_A1 %>%
#   select(-alliance) %>%
#   full_join(
#     diag_A2 %>% select(-alliance),
#     by = "Species"
#   )
# ################################################
# # Return results
# ################################################
# 
# sig.species |>
#   left_join(diag_pair, by = "Species") |>
#   mutate(
# 
#     Higher_Alliance = case_when(
#       MeanCover_AllPlots_A1 > MeanCover_AllPlots_A2 ~ pair[1],
#       MeanCover_AllPlots_A2 > MeanCover_AllPlots_A1 ~ pair[2],
#       TRUE ~ "Tie"
#     ),
# 
#     Indicator_Alliance = case_when(
#       Frequency_A1 > Frequency_A2 ~ pair[1],
#       Frequency_A2 > Frequency_A1 ~ pair[2],
#       MeanCover_AllPlots_A1 > MeanCover_AllPlots_A2 ~ pair[1],
#       MeanCover_AllPlots_A2 > MeanCover_AllPlots_A1 ~ pair[2],
#       TRUE ~ "Tie"
#     ),
# 
#     Comparison = paste0(pair[1], "_vs_", pair[2])
# 
#   ) |>
# 
#   select(
#     Species,
#     P_value,
#     Comparison,
#     Higher_Alliance,
#     Indicator_Alliance,
#     Frequency_A1,
#     Frequency_A2,
#     MeanCover_AllPlots_A1,
#     MeanCover_AllPlots_A2,
#     N_Present_A1,
#     N_Present_A2
#   )
# 
#   }
# )
# 
# ################################################
# # Results
# ################################################
# 
# pairwise.results
# 
# 
# write.csv(
#   pairwise.results,
#   paste0("./review_materials/", anal.set, "Pairwise_Differential_Abundance.csv"),
#   row.names = FALSE
# )


```

```{r}
# library(forcats)
# env.alliance <- plot.env2 |>
#   left_join(
#     su.class, 
#     by = "PlotNumber"
#   ) |>
#   filter(!is.na(alliance))
# env.long <- env.alliance |>
#   select(alliance, aSMR, SNR.clean) |>
#   pivot_longer(
#     cols = c(aSMR, SNR.clean),
#     names_to = "variable",
#     values_to = "value",
#     values_drop_na = TRUE
#   )
# 
# ggplot(
#   env.long |>
#     group_by(variable) |>
#     mutate(alliance = fct_reorder(alliance, value, median, na.rm = TRUE)),
#   aes(x = alliance, y = value)
# ) +
#   geom_boxplot(outlier.shape = NA) +
#   geom_jitter(width = 0.2, alpha = 0.6) +
#   facet_wrap(~variable, scales = "free_y") +
#   theme_bw() +
#   theme(
#     axis.text.x = element_text(angle = 45, hjust = 1)
#   )
#   
```

## Additional possible work

5.  Classification accuracy

If alliances are real compositional groups, you should be able to predict them.

Random Forest rf \<- randomForest( x = comm, y = alliance )

Investigate:

Overall accuracy Confusion matrix Important species

Questions:

Which alliances are consistently confused? Which alliances are highly distinct?

7.  Functional composition

Since you already have lifeforms available, compare:

Tree cover Shrub cover Herb cover Moss cover Lichen cover

Calculate for each plot:

community_summary \<- veg.dat %\>% group_by(PlotID, Lifeform) %\>% summarise(total_cover = sum(Cover))

Then compare among alliances.

This often reveals clearer patterns than species-level analyses.

9.  Differential abundance

Rather than just indicators, identify species significantly more abundant in particular alliances.

Possible methods:

ALDEx2 DESeq2 (presence/abundance transformed) mvabund

Though indicator species analysis is often sufficient for vegetation studies.
