---
title: "Hierarchy Reporting"
author: "William H MacKenzie"
date: "08/11/2024"
format: 
  pdf: 
    documentclass: scrartcl
    papersize: letter
editor: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(tidyverse)
require(DBI)
require(data.table)
require(cluster)
require(dendextend)
require(dynamicTreeCut)
require(gtable)
require(colorspace)
require(gtsummary)
require(openxlsx)
require(tictoc)
require(labdsv)
require(indicspecies)
require(Polychrome)
require(ggdendro)
require(purrr)
require(ggplotify)
require(grid)
set.seed(1279)
source("./_functions/_bec_dist.R")
source("./_functions/_bec_dist_matrix.R")
source("./_functions/_lump_species2.R")
source("./_functions/_create_su_vegdata.R")
source("./_functions/_create_analysis_vegsum.R")
source("./_functions/_TabletoTree.R")
source("./_functions/_TreetoTable.R")
source("./_functions/_add_vars.R")
source("./_functions/_do_pairwise.R")
source("./_functions/_create_diagnostic_veg.R")
source("./_functions/_return_similar_pairs.R")
source("./_functions/_read_sppmaster.R")
source("./_functions/_combined_su.R")
source('./_functions/_create_dendro_all.R')
source('./_functions/_create_dendro_bybgc.R')
source('./_functions/_draw_dendro_split.R')
source('./_functions/_create_veg_sum_all.R')
source('./_functions/_build_species_ordering_hier.R')
source("./_functions/_lump_species.R")
source('./_functions/_format_veg_table2.R')
source('./_functions/_encode_veg_sum.R')
data.path = "D:/OneDrive - Government of BC/GitHub_data"
```

# This script will generate the material for and a formatted report of units from a hierarchical analysis.

Read in vegetation data
read in su tables
read in becdb table with units and hierarchical levels

For each hierarchical level chosen:
  - mapped distribution of plots
  - vegetation summary table for member site units
  - actual edatopic grid unit count
  - summary of climate variables
  - table of related concepts
#### Read in data

Vegetation data is read in from saved .RDS file generated from the BECMaster cleaning scripts. Taxonomy is read-in from the species taxonomy database. A species lumping table is read-in from a Vpro database and applied to the vegetation data. The species taxonomy is joined to the vegetation data.

```{r load data}
veg.dat <- readRDS(file.path(data.path, "BEC_veg_data/Analysis_BECMaster_Veg.rds")) ### named veg.dat
veg.dat2 <- veg.dat
taxon.all <- read_sppmaster()
taxon.lifeform <- taxon.all %>%
  filter(Codetype == "U" | Codetype == "X" | Codetype == "D") %>%
  dplyr::select(Code, ScientificName, EnglishName, Lifeform) %>%
  distinct()
 tree_seedlings <- taxon.lifeform %>% filter(Lifeform %in% c("1", "2")) %>% mutate(Code = paste0(Code, "D")) %>% pull(Code)
trees <- taxon.lifeform %>% filter(Lifeform %in% c("1", "2")) %>% pull(Code)
veglump <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BC_Correlation2_Vpro_2023/Correlation2_Spp_lump.accdb;")
lump <- dbReadTable(veglump, "CorrelationZonal_Lump")
dbDisconnect(veglump)
veg.dat2 <- lump_species(vegdata = veg.dat2, lump, use.subtaxa = TRUE)

bgc.units<- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver13_2024.accdb;")
bgcv12 <- dbReadTable(bgc.units, "tblBGC_UnitsWorkingandArchive")
bgcv13 <- dbReadTable(bgc.units, "tblBEC_SiteSeries_LMH77_forest")
dbDisconnect(bgc.units)
bgcv12 <- bgcv12 %>% filter(is.na(VersionRetired)) %>% select(BGC_NoSpace, SiteClassificationStatus)

```

Project SU tables (e.g. BGC_SU for a LMH) are read-in from regional databases and combined into a project specific table. Then the zonal units are selected for the analysis.

```{r read in SU data}
db <- "D:/BC_Correlation2_Vpro_2023/CoastGuide_Forested.accdb"
su <- combined_su(db)
# db <- "D:/BC_Correlation2_Vpro_2023/Columbia_FG_draft.accdb"
# su2 <- combined_su(db)
# db <- "D:/BC_Correlation2_Vpro_2023/published_sorts/Cariboo_LMH39_Sorts.accdb"
# su3 <- combined_su(db)
# db <- "D:/BC_Correlation2_Vpro_2023/published_sorts/KootenayBoundary_Sorts.accdb"
# su4 <- combined_su(db)
# db <- "D:/BC_Correlation2_Vpro_2023/published_sorts/OminecaPeace_SU_Tables.accdb"
# su5 <- combined_su(db)
# db <- "D:/BC_Correlation2_Vpro_2023/published_sorts/Skeena_LMH26_Sorts.accdb"
# su6 <- combined_su(db)
# db <- "D:/BC_Correlation2_Vpro_2023/published_sorts/ThompsonOK_Sorts.accdb"
# su7 <- combined_su(db)
# db <- "D:/BC_Correlation2_Vpro_2023/published_sorts/LMH65_Boreal_Sorts.accdb"
# su8 <- combined_su(db)
# db <- "D:/BC_Correlation2_Vpro_2023/Cassiar_SUs.accdb"
# su9 <- combined_su(db)
# su <- rbind(su, su2, su3, su4, su5, su6, su7, su8, su9)
#xx <- su %>% filter(PlotNumber == "K780448")
su <- su %>%
  filter(!grepl('101a.2|101b.2|low|-S|add|nudum|poor|yc|_W|_F|ys|moved|101.2|X|omit|unplaced|support', SiteUnit)) 
su <- su %>% select(-SiteUnit.orig) %>% mutate(SiteUnit = str_replace(SiteUnit, "b$|a$", "")) %>% distinct() %>% mutate(zone = str_extract(bgc, "([A-Z]+)"))
bgc.unique <- su %>% distinct(bgc) 
no.su <- anti_join(bgcv12, bgc.unique, by = c("BGC_NoSpace" = "bgc")) %>% filter(!SiteClassificationStatus == "none")
###write combined SU to database
# master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
# DBQ=D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb;")
# dbWriteTable(master_su, "All_BC_zonals2_SU",  su, overwrite = TRUE, batch_rows = 1)
# dbDisconnect(master_su)
xx <- su %>% filter(SiteUnit == "IDFww1_101.2")
su <- su %>%
filter(!grepl('CWHvh3|CWHwh1|CWHwh2|CWHvh3|MHwh', SiteUnit))###BGC specific|CWHms

```

```{r build veg guide summary and export to excel, echo = FALSE, warning=FALSE, message=FALSE}
require(openxlsx)
require(tictoc)
library(stringr)
data.path = "D:/OneDrive - Government of BC/GitHub_data"
author <- "test.assoc1"
assoc.list <- bgcv13 %>% filter(!is.na(Association_Code)) %>% group_by(Association_Code) %>% mutate(n=n()) %>% filter(n>2) %>% distinct(Association_Code) %>% pull(Association_Code)

vegsum.wbk <- createWorkbook()
for(assoc.choose in assoc.list){
#assoc.choose = "assoc-13"
assoc <- bgcv13 %>% filter(Association_Code %in% assoc.choose) %>% select(SS_Label, Association_Code) %>% mutate(SiteUnit = gsub(" ", "", SS_Label)) %>% mutate(SiteUnit = str_replace(SiteUnit, "/", "_"))
su.assoc <- left_join(assoc, su, by = "SiteUnit") %>%  mutate(SiteUnit = str_replace(SiteUnit, "b$|a$", "")) %>% distinct() 
veg.dat.assoc <- veg.dat2 %>% filter(PlotNumber %in% su.assoc$PlotNumber)
# vegsum.wbk <- createWorkbook()
# #bgc.choose = "CDFmmtab2"
# #for(bgc.choose in bgc.list){
vegSum <-
  create_veg_sum_all(vdat = veg.dat.assoc, siteUnits = su.assoc, strata.by = "Lifeform")

indic.order <- build_species_ordering_hier(vdat = veg.dat.assoc, vsum = vegSum, code.lump=lump, siteUnits = su.assoc, hier.level = "Association_Code")
veg.sum.table <- format_veg_table2(vsum = vegSum, spp=taxon.lifeform)
openxlsx::addWorksheet(vegsum.wbk, sheetName = assoc.choose)
openxlsx::writeData(vegsum.wbk, sheet = assoc.choose, veg.sum.table) 
}

saveWorkbook(vegsum.wbk, paste0("./vegsum.tables/LMH77_draft_assoc_tables_", author, ".xlsx"), overwrite = TRUE)

```


