---
title: "Hierarchy Unit Review"
author: "William H MacKenzie"
date: "08/11/2024"
format: 
  pdf: 
    documentclass: scrartcl
    papersize: letter
editor: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(tidyverse)
require(DBI)
require(data.table)
require(cluster)
require(dendextend)
require(dynamicTreeCut)
require(gtable)
require(colorspace)
require(gtsummary)
require(openxlsx)
require(tictoc)
require(labdsv)
require(indicspecies)
require(Polychrome)
require(ggdendro)
require(purrr)
require(ggplotify)
require(grid)
library(stringr)
require(vegan)
require(betapart)
library(plotly)
set.seed(1279)
source("../BEC_R_working/_functions/_bec_dist.R")
source("../BEC_R_working/_functions/_bec_dist_matrix.R")
source("../BEC_R_working/_functions/_combine_taxa.R")
source("../BEC_R_working/_functions/_combine_taxa_strata.R")
source("../BEC_R_working/_functions/_create_su_vegdata.R")
source("../BEC_R_working/_functions/_create_analysis_vegsum.R")
source("../BEC_R_working/_functions/_TabletoTree.R")
source("../BEC_R_working/_functions/_TreetoTable.R")
source("../BEC_R_working/_functions/_add_vars.R")
source("../BEC_R_working/_functions/_do_pairwise.R")
source("../BEC_R_working/_functions/_create_diagnostic_veg.R")
source("../BEC_R_working/_functions/_return_similar_pairs.R")
source("../BEC_R_working/_functions/_read_sppmaster.R")
source("../BEC_R_working/_functions/_combined_su.R")
source("../BEC_R_working/_functions/_create_veg_sum.R")
source('../BEC_R_working/_functions/_create_veg_sum_all.R')
source('../BEC_R_working/_functions/_build_species_ordering.R')
source('../BEC_R_working/_functions/_format_veg_table.R')
source('../BEC_R_working/_functions/_format_veg_table_strata.R')
source('../BEC_R_working/_functions/_format_veg_table_nostrata.R')
source('../BEC_R_working/_functions/_encode_veg_sum.R')
source('../BEC_R_working/_functions/_create_dendro.R')
source('../BEC_R_working/_functions/_create_dendro_all.R')
source('../BEC_R_working/_functions/_create_dendro_bybgc.R')
source('../BEC_R_working/_functions/_draw_dendro_split.R')
source('../BEC_R_working/_functions/_cluster_groups.R')
source('../BEC_R_working/_functions/_TabletoTree.R')
source('../BEC_R_working/_functions/_build_species_ordering_hier.R')
source("../BEC_R_working/_functions/_cluster_by_spp_group.R")
source("../BEC_R_working/_functions/_baselga_diagnostics.R")
data.path = "D:/OneDrive - Government of BC/GitHub_data"
```

# General flow for review

A report to review proposed associations and hierarchical unit. Comparative summary table Ordination Comparative environmental reporting

## Read in data

Read in vegetation data and join to taxonomic read in becdb table with units and hierarchical levels read in working xlsx becdb table with proposed units read in su tables and limit to units in the working xlsx

Vegetation data is read in from saved .RDS file generated from the BECMaster cleaning scripts. Taxonomy is read-in from the species taxonomy database. A species lumping table is read-in from a Vpro database and applied to the vegetation data. The species taxonomy is joined to the vegetation data.

```{r load data}
# =============================================================================
# Vegetation Data
# =============================================================================

veg.dat <- readRDS(
  file.path(data.path, "BEC_veg_data/Analysis_BECMaster_Veg.rds")
)

veg.dat2 <- veg.dat

# =============================================================================
# Species Master and Lifeforms
# =============================================================================

taxon.all <- read_sppmaster()

taxon.lifeform <- taxon.all %>%
  filter(Codetype %in% c("U", "X", "D")) %>%
  dplyr::select(Code, ScientificName, EnglishName, Lifeform) %>%
  distinct()

# Tree species codes
trees <- taxon.lifeform %>%
  filter(Lifeform %in% c("1", "2")) %>%
  pull(Code)

# Tree seedling species codes
tree_seedlings <- taxon.lifeform %>%
  filter(Lifeform %in% c("1", "2")) %>%
  mutate(Code = paste0(Code, "D")) %>%
  pull(Code)

# =============================================================================
# Species Lump Table
# =============================================================================

veglump <- dbConnect(
  odbc::odbc(),
  .connection_string = paste0(
    "Driver={Microsoft Access Driver (*.mdb, *.accdb)};",
    "DBQ=D:/BC_Correlation2_Vpro_2023/Correlation2_Spp_lump.accdb;"
  )
)

lump <- dbReadTable(veglump, "CorrelationZonal_Lump")

dbDisconnect(veglump)

veg.dat2 <- combine_taxa(
  vegdata = veg.dat2,
  lump = lump,
  use.subtaxa = TRUE
)

# =============================================================================
# BEC Database
# =============================================================================

bgc.units <- dbConnect(
  odbc::odbc(),
  .connection_string = paste0(
    "Driver={Microsoft Access Driver (*.mdb, *.accdb)};",
    "DBQ=D:/OneDrive - Government of BC/BECdb_Working/",
    "BECdb_ver14_2026_v0.accdb;"
  )
)

bgcv13 <- dbReadTable(
  bgc.units,
  "tblBEC_SiteSeriesWorkingandArchive"
) %>%
  filter(Status != "Retired")

hierarchy <- dbReadTable(
  bgc.units,
  "tblBEC_Hierarchy_Class"
)

dbDisconnect(bgc.units)

# =============================================================================
# Correlation Database
# =============================================================================

db <- "D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb"

su_all <- combined_su(db)

# Remove retired, support, and other non-working site units
su_all <- su_all %>%
  filter(
    !grepl(
      "101a.2|101b.2|low|-S|add|nudum|poor|yc|_W|_F|ys|moved|X|omit|unplaced|support",
      SiteUnit
    )
  )

#===========================================================================
# Site Unit Cleanup
# =============================================================================

su_all <- su_all %>%
  #dplyr::select(-SiteUnit.orig) %>%
  mutate(
    SiteUnit = str_replace(SiteUnit, "b$|a$", ""),
    zone = str_extract(bgc, "[A-Z]+")
  ) %>%
  distinct()

# =============================================================================
# QA / Check for Oddities
# =============================================================================

bgc.unique <- su_all %>%
  distinct(bgc)

```

### Core analyses would be:

With 3,450 plots assigned to 28 alliances, I'd treat this as a multivariate community ecology problem and analyze it at several complementary levels. All comparisons use the site unit summary data as the basis for comparison and not plot level data.

1.  NMDS ordination

2.  PERMANOVA to compare sig differences between units. Alliances, Assocs within Alliances, All Assocs

3.  Betadisper: variability within units 

4.  Indicator species analysis: significant indicators associated with single or groups of units. Where few indicators in individual units consider combining with groups

5.  Random Forest classification: prediction of membership

6.  Species constancy tables: standard vegetation summary output

7.  Environmental gradient analysis: environmental separation of units

Together these answer:

Are alliances compositionally distinct? How distinct are they? Which species define them? Which alliances overlap? What environmental gradients drive differences? How internally consistent are they?

For a BEC/alliance classification paper, NMDS + PERMANOVA + Indicator Species + Random Forest + diagnostic species tables would likely provide the strongest and most defensible framework.


5.  Classification accuracy

If alliances are real compositional groups, you should be able to predict them.

Random Forest rf \<- randomForest( x = comm, y = alliance )

Investigate:

Overall accuracy Confusion matrix Important species

Questions:

Which alliances are consistently confused? Which alliances are highly distinct? 


7.  Functional composition

Since you already have lifeforms available, compare:

Tree cover Shrub cover Herb cover Moss cover Lichen cover

Calculate for each plot:

community_summary \<- veg.dat %\>% group_by(PlotID, Lifeform) %\>% summarise(total_cover = sum(Cover))

Then compare among alliances.

This often reveals clearer patterns than species-level analyses.

9.  Differential abundance

Rather than just indicators, identify species significantly more abundant in particular alliances.

Possible methods:

ALDEx2 DESeq2 (presence/abundance transformed) mvabund

Though indicator species analysis is often sufficient for vegetation studies.


```{r read in SU data}
hier_working <- openxlsx::read.xlsx("D:/OneDrive/WNA_BGC/Association_Analysis/Coast_Hierarchy_Analysis/BECdb_Coast_Hierarchy_Working_ver2.xlsx") |> filter(Analysis_Group %in% "Forest-Coast_MH")

# Keep only site units present in hierarchy table
su <- su_all %>%
  filter(SiteUnit.orig %in% hier_working$SU_Label)
su.anal <- su %>%
  left_join(hier_working, by = c("SiteUnit.orig" = "SU_Label"))

su.class <- su.anal |> filter(n_plots >1) |> dplyr::select(PlotNumber, siteunit = SiteUnit, assoc = working_Assoc_cleaned, alliance = working_Alliance) |> 
  mutate(alliance = str_remove(alliance, "^coast_generated_"))  |> distinct()|> arrange(PlotNumber)
unique.su <- su.class |> dplyr::select(siteunit) |> distinct()|> arrange(siteunit)
unique.assocs <- su.class |> dplyr::select(siteunit, assoc) |> distinct()|> arrange(siteunit)
unique.alliances <- su.class |> dplyr::select(siteunit, alliance) |> distinct()|> arrange(siteunit)


hier.class <- su.anal |> filter(n_plots >1) |> dplyr::select(n_plots,siteunit = SiteUnit, assoc = working_Assoc_cleaned, alliance = working_Alliance) |> 
  mutate(alliance = str_remove(alliance, "^coast_generated_")) |> arrange(siteunit) |> distinct()
unique.su2 <- hier.class |> dplyr::select(siteunit) |> distinct()
unique.assocs2 <- hier.class |> dplyr::select(assoc) |> distinct()
unique.alliances2 <- hier.class |> dplyr::select(alliance) |> distinct()


veg.anal<- veg.dat2 |> filter(PlotNumber %in% su.class$PlotNumber) |> dplyr::select(PlotNumber, Species, Cover) |>
  arrange(PlotNumber) #|> matrify()

reduced.exceptions <- c("SPHAGNUM","CLADONIA","CLADINA","RACOMITR","MNIUM")
reduced.lifeforms  <- c(9, 10, 11)

pairwise.params <- list(
  su                = su.anal,
  minconstancy          = 50,
  noiseconstancy        = 10,
  minplots              = 1,
  minor                 = 1,
  use.ksi               = FALSE,
  ksi   = NULL,
  ksi.value             = 1.5,
  reduce.lifeform       = TRUE,
  reduced.lifeforms     = reduced.lifeforms,
  reduction             = 0.1,
  reduced.exceptions    = reduced.exceptions,
  add_vegsum_table      = NULL
)

vegsum.pairs <- do.call(
  do_pairwise,
  c(list(veg.anal), pairwise.params)
)

 bec.dist <- bec_dist_matrix(vegsum.pairs = vegsum.pairs, distance = "BEC.sim.min")

```

##Describe differences in species composition

### NMDS ordination

Use BEC dissimilarity and plot alliances.

Questions answered:

Do alliances form distinct clusters? Which alliances overlap? Are some alliances highly variable internally?

Add alliance centroids and confidence ellipses.

ordiplot(nmds) ordiellipse(nmds, alliance)
```{r NMDS}
## NMDS
nmds <- metaMDS(
  bec.dist,
  k = 2
)
nmds$stress

nmds3 <- metaMDS(
  bec.dist, 
  k = 3
)
nmds3$stress
```
## Build 3D NMDS plot
```{r}
library(dplyr)
library(tibble)
library(plotly)
library(vegan)

#==============================================================================
# NMDS scores (SiteUnit level)
#==============================================================================

scores3d <- as.data.frame(
  scores(nmds3, display = "sites")
) |>
  rownames_to_column("siteunit") |>
  left_join(
    hier.class |>
      dplyr::select(siteunit, assoc, alliance),
    by = "siteunit"
  ) |>
  filter(
    !is.na(siteunit),
    !is.na(assoc),
    !is.na(alliance)
  )

#==============================================================================
# Association centroids
#==============================================================================

assoc.cent <- scores3d |>
  group_by(assoc, alliance) |>
  summarise(
    NMDS1 = mean(NMDS1),
    NMDS2 = mean(NMDS2),
    NMDS3 = mean(NMDS3),
    n_siteunits = n(),
    .groups = "drop"
  )

#==============================================================================
# Alliance centroids
#==============================================================================

alliance.cent <- assoc.cent |>
  group_by(alliance) |>
  summarise(
    NMDS1 = mean(NMDS1),
    NMDS2 = mean(NMDS2),
    NMDS3 = mean(NMDS3),
    n_assoc = n(),
    .groups = "drop"
  )

#==============================================================================
# SiteUnit -> Association links
#==============================================================================

links.su <- scores3d |>
  dplyr::select(
    siteunit,
    assoc,
    alliance,
    NMDS1.su = NMDS1,
    NMDS2.su = NMDS2,
    NMDS3.su = NMDS3
  ) |>
  left_join(
    assoc.cent,
    by = c("assoc", "alliance")
  ) |>
  rename(
    NMDS1.assoc = NMDS1,
    NMDS2.assoc = NMDS2,
    NMDS3.assoc = NMDS3
  )

#==============================================================================
# Association -> Alliance links
#==============================================================================

links.assoc <- assoc.cent |>
  left_join(
    alliance.cent,
    by = "alliance",
    suffix = c(".assoc", ".alliance")
  )

#==============================================================================
# Plot
#==============================================================================

p <- plot_ly()


#------------------------------------------------------------------------------
# SiteUnit -> Association lines
#------------------------------------------------------------------------------

for(i in seq_len(nrow(links.su))) {

  p <- p |>
    add_trace(
      type = "scatter3d",
      mode = "lines",
      x = c(
        links.su$NMDS1.su[i],
        links.su$NMDS1.assoc[i]
      ),
      y = c(
        links.su$NMDS2.su[i],
        links.su$NMDS2.assoc[i]
      ),
      z = c(
        links.su$NMDS3.su[i],
        links.su$NMDS3.assoc[i]
      ),
      line = list(
        color = "lightgrey",
        width = 1
      ),
      showlegend = FALSE,
      inherit = FALSE
    )
}


#------------------------------------------------------------------------------
# Association -> Alliance lines
#------------------------------------------------------------------------------

for(i in seq_len(nrow(links.assoc))) {

  p <- p |>
    add_trace(
      type = "scatter3d",
      mode = "lines",
      x = c(
        links.assoc$NMDS1.assoc[i],
        links.assoc$NMDS1.alliance[i]
      ),
      y = c(
        links.assoc$NMDS2.assoc[i],
        links.assoc$NMDS2.alliance[i]
      ),
      z = c(
        links.assoc$NMDS3.assoc[i],
        links.assoc$NMDS3.alliance[i]
      ),
      line = list(
        color = "black",
        width = 4
      ),
      showlegend = FALSE,
      inherit = FALSE
    )
}

#------------------------------------------------------------------------------
# SiteUnit points
#------------------------------------------------------------------------------

p <- p |>
  add_trace(
    data = scores3d,
    x = ~NMDS1,
    y = ~NMDS2,
    z = ~NMDS3,
    color = ~alliance,
    type = "scatter3d",
    mode = "markers+text",
    text = ~siteunit,
    textposition = "top center",
    marker = list(
      size = 3,
      opacity = 0.4
    ),
    hoverinfo = "text",
    hovertext = ~paste(
      "SiteUnit:", siteunit,
      "<br>Association:", assoc,
      "<br>Alliance:", alliance
    )
  )
#------------------------------------------------------------------------------
# Association centroids
#------------------------------------------------------------------------------

p <- p |>
  add_trace(
    data = assoc.cent,
    x = ~NMDS1,
    y = ~NMDS2,
    z = ~NMDS3,
    color = ~alliance,
    type = "scatter3d",
    mode = "markers+text",
    text = ~assoc,
    textposition = "top center",
    marker = list(
      size = 8,
      symbol = "circle",
      line = list(
        color = "black",
        width = 1
      )
    )
  )
#------------------------------------------------------------------------------
# Alliance centroids
#------------------------------------------------------------------------------

p <- p |>
  add_trace(
    data = alliance.cent,
    x = ~NMDS1,
    y = ~NMDS2,
    z = ~NMDS3,
    color = ~alliance,
    type = "scatter3d",
    mode = "markers+text",
    text = ~alliance,
    textposition = "middle right",
    marker = list(
      size = 18,
      symbol = "diamond",
      line = list(
        color = "black",
        width = 3
      )
    ),
    name = "Alliance"
  )

# library(geometry)
# 
# min_n <- 6
# 
# for(a in unique(scores3d$alliance)) {
# 
#   pts <- scores3d |>
#     filter(alliance == a)
# 
#   if(nrow(pts) < min_n) next
# 
#   ell <- make_ellipsoid(pts)
# 
#   p <- p |>
#     add_trace(
#       data = ell,
#       x = ~x,
#       y = ~y,
#       z = ~z,
#       type = "mesh3d",
#       opacity = 0.15,
#       showlegend = FALSE
#     )
# }
p
## Write to stand alone product.
# library(htmlwidgets)
# 
# saveWidget(
#   p,
#   file = "./review_materials/BEC_NMDS_Hierarchy.html",
#   selfcontained = TRUE
#)

```


2.  Test whether alliances differ PERMANOVA

A statistical test of compositional differences.

adonis2( comm \~ alliance, data = meta, method = "bray", permutations = 999 )

Outputs:

R² = how much variation is explained by alliance P-value

Given 3450 plots, you'll almost certainly get a significant result. Focus on effect size (R²), not just significance.

Test dispersion first

PERMANOVA is sensitive to unequal dispersion.

disp \<- betadisper( vegdist(comm, method = "bray"), alliance )

anova(disp)

This tells you whether differences are due to centroids (real compositional shifts) or spread.

3.  Quantify within-unit heterogeneity

Some alliances are very tight, others broad.

Beta diversity disp \<- betadisper( vegdist(comm, "bray"), alliance )

boxplot(disp)

Useful outputs:

Mean distance to centroid Rank alliances from homogeneous to heterogeneous

This is often ecologically informative.

```{r}

pairwise_adonis_dist <- function(
    dist.mat,
    groups,
    permutations = 999,
    p.adjust.method = "BH"
) {

  groups <- droplevels(as.factor(groups))

  combs <- combn(levels(groups), 2)

  results <- data.frame()

  for (i in seq_len(ncol(combs))) {

    grp <- combs[, i]

    keep <- groups %in% grp

    # subset distance matrix
    d.sub <- as.dist(
      as.matrix(dist.mat)[keep, keep]
    )

    meta.sub <- data.frame(
      group = droplevels(groups[keep])
    )

    mod <- adonis2(
      d.sub ~ group,
      data = meta.sub,
      permutations = permutations
    )

    results <- rbind(
      results,
      data.frame(
        group1 = grp[1],
        group2 = grp[2],
        F = mod$F[1],
        R2 = mod$R2[1],
        p = mod$`Pr(>F)`[1]
      )
    )
  }

  results$p.adj <- p.adjust(
    results$p,
    method = p.adjust.method
  )

  results[order(results$R2, decreasing = TRUE), ]
}

library(vegan)

pairwise_dispersion_dist <- function(
    dist.mat,
    groups,
    p.adjust.method = "BH"
) {

  groups <- as.factor(groups)

  # Remove NAs
  keep <- !is.na(groups)

  groups <- droplevels(groups[keep])

  dmat <- as.matrix(dist.mat)
  dmat <- dmat[keep, keep]

  combs <- combn(levels(groups), 2)

  results <- data.frame()

  for (i in seq_len(ncol(combs))) {

    grp <- combs[, i]

    keep_pair <- groups %in% grp

    d.sub <- as.dist(
      dmat[keep_pair, keep_pair]
    )

    groups_sub <- droplevels(groups[keep_pair])

    bd <- betadisper(d.sub, groups_sub)

    aov_res <- anova(bd)

    results <- rbind(
      results,
      data.frame(
        group1 = grp[1],
        group2 = grp[2],
        n1 = sum(groups_sub == grp[1]),
        n2 = sum(groups_sub == grp[2]),
        mean.dist1 = mean(
          bd$distances[groups_sub == grp[1]]
        ),
        mean.dist2 = mean(
          bd$distances[groups_sub == grp[2]]
        ),
        diff = abs(
          mean(bd$distances[groups_sub == grp[1]]) -
          mean(bd$distances[groups_sub == grp[2]])
        ),
        F = aov_res[1, "F value"],
        p = aov_res[1, "Pr(>F)"]
      )
    )
  }

  results$p.adj <- p.adjust(
    results$p,
    method = p.adjust.method
  )

  results |>
    arrange(p.adj)
}
```

```{r pairwise permanova}
adonis.assoc <- pairwise_adonis_dist(
  dist.mat = bec.dist,
  groups = unique.assocs$assoc,
  permutations = 199 ## reset to 999 for final runs
)
low.sep <- adonis.assoc |>
  filter(R2 < 0.05)

disp.assoc <- pairwise_dispersion_dist(
  dist.mat = bec.dist,
  groups = unique.assocs$assoc
)



adonis.alliance <- pairwise_adonis_dist(
  dist.mat = bec.dist,
  groups = unique.alliances$alliance,
  permutations = 199 ## reset to 999 for final runs
)
low.sep <- adonis.alliance |>
  filter(R2 < 0.05)

disp.alliance <- pairwise_dispersion_dist(
  dist.mat = bec.dist,
  groups = unique.alliances$alliance
)

```
4.  mvabund analysis (composition and abundance differences) rather than Indicator species analysis (constancy only)

Questions:

Which species define each alliance? Are your alliances characterized by expected indicators?

Produces species such as:

These become excellent ecological descriptions.

Prep data for pairwise statistical tests
```{r pairwise indicator species}
# Keep only SiteUnits with an alliance assignment
alliance.good <- su.class |> filter(!is.na(alliance))
veg.matrix <- veg.dat2 |> filter(PlotNumber %in% alliance.good$PlotNumber) |> dplyr::select(PlotNumber, Species, Cover) |>
  arrange(PlotNumber) |> matrify()

##remove rare species
freq <- colSums(veg.matrix > 0)
veg.matrix2 <- veg.matrix[, freq >= 10]
alliance2 <- droplevels(as.factor(alliance.good$alliance))


table(alliance2)
## remove alliances with few members
keep.grp <- alliance2 %in%
  names(which(table(alliance2) >= 5))

veg.matrix2 <- veg.matrix2[keep.grp, ]

alliance2 <- droplevels(alliance2[keep.grp])
table(alliance2)
```
## do pairwise comparison between hierarchy units
```{r}
library(dplyr)
library(purrr)
library(tibble)
library(mvabund)

################################################
# All pairwise combinations
################################################
su.class <- su.class |> filter(alliance %in% alliance2)
pairs <- combn(
  sort(unique(su.class$alliance)),
  2,
  simplify = FALSE
)

################################################
# Run pairwise comparisons
################################################
pair <- pairs[[2]]
pairwise.results <- map_dfr(
  pairs,
  function(pair){

    alliance.choose <- su.class |>
      filter(alliance %in% pair) |>
      arrange(PlotNumber)

    ################################################
    # Species matrix
    ################################################

    veg.matrix.mva <- veg.dat2 |>
      filter(PlotNumber %in% alliance.choose$PlotNumber) |>
      dplyr::select(PlotNumber, Species, Cover) |>
      arrange(PlotNumber) |>
      matrify()

    ################################################
    # Transform cover
    ################################################

    veg.matrix.mva <- veg.matrix.mva^(1/3)

    ################################################
    # Remove rare species
    ################################################

    freq <- colSums(veg.matrix.mva > 0)

    veg.matrix.mva <- veg.matrix.mva[
    ,
    freq >= 0.10 * nrow(veg.matrix.mva),
    drop = FALSE
    ]

    ################################################
    # Model
    ################################################

    alliance2 <- droplevels(
      factor(alliance.choose$alliance)
    )

    mod.ab <- manylm(
      mvabund(veg.matrix.mva) ~ alliance2
    )

    ################################################
    # Test
    ################################################

    res <- anova(
      mod.ab,
      p.uni = "adjusted"
    )

    ################################################
    # Significant species
    ################################################

    sig.species <- tibble(
      Species = colnames(res$uni.test),
      P_value = res$uni.p[2, ]
    ) |>
      filter(P_value <= 0.05)

    if(nrow(sig.species) == 0)
      return(NULL)


################################################
# Diagnostic statistics by alliance
################################################

diag.stats <- alliance.choose |>
  select(PlotNumber, alliance) |>
  crossing(
    Species = sig.species$Species
  ) |>
  left_join(
    veg.dat2 |>
      select(PlotNumber, Species, Cover),
    by = c("PlotNumber", "Species")
  ) |>
  mutate(
    Cover = replace_na(Cover, 0)
  ) |>
  group_by(Species, alliance) |>
  summarise(
    N_Plots = n(),
    N_Present = sum(Cover > 0),
    Frequency = N_Present / N_Plots,
    MeanCover_AllPlots = mean(Cover),
    MeanCover_PresentOnly = ifelse(
      N_Present > 0,
      mean(Cover[Cover > 0]),
      0
    ),
    SD_Cover = sd(Cover),
    .groups = "drop"
  )

################################################
# Put alliance diagnostics into columns
################################################

diag_pair <- diag.stats |>
  mutate(
    AllianceCol = case_when(
      alliance == pair[1] ~ "A1",
      alliance == pair[2] ~ "A2"
    )
  ) |>
  pivot_wider(
    names_from = AllianceCol,
    values_from = c(
      N_Plots,
      N_Present,
      Frequency,
      MeanCover_AllPlots,
      MeanCover_PresentOnly,
      SD_Cover
    ),
    values_fill = 0
  )
################################################
# Return results
################################################

sig.species |>
  left_join(
    diag_pair,
    by = c("Species")
  ) |>
  mutate(
    Alliance_1 = pair[1],
    Alliance_2 = pair[2],
    compared = paste0(pair[1], "_vs_", pair[2])) |> 
      mutate(
  Rare_A1 = Frequency_A1 < 0.10,
  Rare_A2 = Frequency_A2 < 0.10,
  Low_N_A1 = N_Present_A1 < 3,
  Low_N_A2 = N_Present_A2 < 3,
  Large_Mean_Diff =
   abs(MeanCover_AllPlots_A1 - MeanCover_AllPlots_A2) > 5
)

  }
)

################################################
# Results
################################################

pairwise.results

# ## pairwise summary
# pairwise.summary <- pairwise.results |>
#   count(
#     Contrast,
#     name = "N_Significant_Species"
#   ) |>
#   arrange(desc(N_Significant_Species))
# 
# pairwise.summary

write.csv(
  pairwise.results,
  "./review_materials/Pairwise_Differential_Abundance.csv",
  row.names = FALSE
)

## compare alliance versus all others
focal <- "Alli_01_woring"

alli_01_results <- pairwise.results |>
  filter(
    Alliance_1 == focal |
    Alliance_2 == focal
  )
species_summary <- alli_01_results |>
  count(Species, sort = TRUE)

focal <- "Alli_01_woring"

alli_01_results <- pairwise.results |>
  filter(
    Alliance_1 == focal |
    Alliance_2 == focal
  )

n_comparisons <- alli_01_results |>
  distinct(compared) |>
  nrow()

species_summary <- alli_01_results |>
  count(Species, name = "n_sig_pairs") |>
  mutate(
    pct_sig_pairs = round(
      100 * n_sig_pairs / n_comparisons,
      1
    )
  ) |>
  arrange(
    desc(pct_sig_pairs),
    desc(n_sig_pairs)
  )

alli_01_results <- alli_01_results |>
  mutate(
    focal_cover = if_else(
      Alliance_1 == focal,
      MeanCover_AllPlots_A1,
      MeanCover_AllPlots_A2
    ),
    other_cover = if_else(
      Alliance_1 == focal,
      MeanCover_AllPlots_A2,
      MeanCover_AllPlots_A1
    ),
    focal_higher = focal_cover > other_cover
  )

indicator_summary <- alli_01_results |>
  filter(focal_higher) |>
  count(Species, name = "n_higher_pairs") |>
  mutate(
    pct_higher_pairs =
      round(100 * n_higher_pairs / n_comparisons, 1)
  ) |>
  arrange(desc(pct_higher_pairs))


alli_01_results <- pairwise.results |>
  filter(
    Alliance_1 == focal |
    Alliance_2 == focal
  ) |>
  mutate(
    Focal_MeanCover = case_when(
      Alliance_1 == focal ~ MeanCover_AllPlots_A1,
      Alliance_2 == focal ~ MeanCover_AllPlots_A2
    ),
    Other_MeanCover = case_when(
      Alliance_1 == focal ~ MeanCover_AllPlots_A2,
      Alliance_2 == focal ~ MeanCover_AllPlots_A1
    ),
    Focal_Higher = Focal_MeanCover > Other_MeanCover
  )
# unique_species <- alli_01_results |>
#   filter(Focal_Higher) |>
#   count(Species, name = "n_sig_pairs") |>
#   arrange(desc(n_sig_pairs))

```


indicator species analysis - should some alliances be combined
```{r}
################################################
# Indicator species analysis
################################################

library(indicspecies)
library(vegan)
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(tictoc)

tic()

ind.cover <- multipatt(
  veg.matrix2,
  alliance2,
  func = "IndVal.g",
  max.order = 3,      # increase as desired
  control = how(nperm = 999)
)

toc()

################################################
# Convert multipatt output to table
################################################

ind.tbl <- ind.cover$sign %>%
  as.data.frame() %>%
  rownames_to_column("species")

group.cols <- grep(
  "^s\\.",
  names(ind.tbl),
  value = TRUE
)

################################################
# Create labels for alliance combinations
################################################

group.names <- sub("^s\\.", "", group.cols)

ind.tbl <- ind.tbl %>%
  rowwise() %>%
  mutate(
    pattern = paste(
      group.names[c_across(all_of(group.cols)) == 1],
      collapse = "+"
    )
  ) %>%
  ungroup()

################################################
# Presence / absence matrix
################################################

veg.pa <- (veg.matrix2 > 0) * 1

################################################
# Fidelity calculations
################################################

fidelity.tbl <- map_dfr(
  seq_len(nrow(ind.tbl)),
  function(i) {

    sp <- ind.tbl$species[i]

    pattern.groups <- strsplit(
      ind.tbl$pattern[i],
      "\\+"
    )[[1]]

    inside <- alliance2 %in% pattern.groups
    outside <- !inside

    tibble(
      species = sp,
      pattern = ind.tbl$pattern[i],

      freq_in = mean(
        veg.pa[inside, sp],
        na.rm = TRUE
      ),

      freq_out = mean(
        veg.pa[outside, sp],
        na.rm = TRUE
      )
    ) %>%
      mutate(
        fidelity = freq_in - freq_out
      )
  }
)

################################################
# Final table
################################################

supp.tbl <- ind.tbl %>%
  dplyr::select(
    species,
    pattern,
    stat,
    p.value
  ) %>%
  left_join(
    fidelity.tbl,
    by = c("species", "pattern")
  ) %>%
  transmute(
    Alliance_Pattern = pattern,
    Species = species,
    IndVal = round(stat, 3),
    P_value = signif(p.value, 3),
    Constancy_in_Pattern = round(freq_in * 100, 1),
    Constancy_Outside_Pattern = round(freq_out * 100, 1),
    Fidelity = round(fidelity * 100, 1)
  ) %>%
  mutate(
    Indicator_Strength = case_when(
      IndVal >= 0.70 ~ "Excellent",
      IndVal >= 0.50 ~ "Strong",
      IndVal >= 0.25 ~ "Moderate",
      TRUE ~ "Weak"
    )
  ) |> 
  filter(
    Alliance_Pattern != "",
    P_value <= 0.2
  ) %>%
  arrange(
    Alliance_Pattern,
    desc(Fidelity),
    desc(IndVal)
  )
supp.tbl
################################################
# Count indicators by alliance pattern
################################################

indicator.counts <- supp.tbl %>%
  count(
    Alliance_Pattern,
    name = "N_Indicators"
  ) %>%
  arrange(desc(N_Indicators))

indicator.counts

################################################
# Export
################################################

write.csv(
  supp.tbl,
  "./review_materials/Indicator_Species_Table.csv",
  row.names = FALSE
)
write.csv(
  indicator.counts,
  "./review_materials/Indicator_Counts_By_Pattern.csv",
  row.names = FALSE
)

supp.tbl.cons <- supp.tbl |> filter(Constancy_in_Pattern >=60)

```

single alliance indicator species
```{r pairwise indicator species for single alliances}
# Indicator species analysis for single alliances
library(indicspecies)
tic()
ind.cover <- multipatt(
  veg.matrix2,
  alliance2,
  func = "IndVal.g",
  max.order = 2,
  control = how(nperm = 199) ## reset to 999 for final runs
)
toc()

library(dplyr)
library(tidyr)
library(tibble)

ind.tbl <- ind.cover$sign %>%
  as.data.frame() %>%
  rownames_to_column("species")

################################################
# Alliance assignment
################################################

group.cols <- grep(
  "^s\\.",
  names(ind.tbl),
  value = TRUE
)

ind.tbl <- ind.tbl %>%
  mutate(
    alliance = apply(
      select(., all_of(group.cols)),
      1,
      function(x){

        idx <- which(x == 1)

        if(length(idx) == 1){
          sub("^s\\.", "", group.cols[idx])
        } else {
          NA_character_
        }
      }
    )
  )

################################################
# Presence / absence matrix
################################################

veg.pa <- (veg.matrix2 > 0) * 1

################################################
# Fidelity calculations
################################################

fidelity.tbl <- bind_rows(

  lapply(sort(unique(alliance2)), function(g){

    inside  <- alliance2 == g
    outside <- alliance2 != g

    tibble(
      alliance = g,
      species = colnames(veg.pa),

      freq_in = colMeans(
        veg.pa[inside, , drop = FALSE]
      ),

      freq_out = colMeans(
        veg.pa[outside, , drop = FALSE]
      )

    ) %>%
      mutate(
        fidelity = freq_in - freq_out
      )

  })

)

################################################
# Final table
################################################

supp.tbl <- ind.tbl %>%
  select(
    species,
    alliance,
    stat,
    p.value
  ) %>%
  left_join(
    fidelity.tbl,
    by = c("alliance", "species")
  ) %>%
  transmute(
    Alliance = alliance,
    Species = species,
    IndVal = round(stat, 3),
    P_value = signif(p.value, 3),
    Constancy_in_Alliance = round(freq_in * 100, 1),
    Constancy_Outside_Alliance = round(freq_out * 100, 1),
    Fidelity = round(fidelity * 100, 1)
  ) %>%
  filter(
    !is.na(Alliance),
    P_value <= 0.05
  ) %>%
  arrange(
    Alliance,
    desc(Fidelity),
    desc(IndVal)
  )
```

11. Environmental differentiation

Elevation Slope Aspect MAT MAP Soil moisture Nutrient regime

Use:

RDA rda.mod \<- rda( comm \~ elevation + moisture + nutrients, data = env )

or

envfit on NMDS envfit(nmds, env)

This links community differences to environmental gradients.


```{r}
#------------------------------------------------------------
# 8. Join edatopic with modeled aSMR ranges
#------------------------------------------------------------

rsmr_asmr.long <- fread("../aSMR_X_rSMR/outputs/modelled_aSMR_by_rSMR_long.csv") %>%
  filter(cmd.type == "CMD", PERIOD == "1961_1990")

SS.edatopic <- left_join(edatopic, rsmr_asmr.long, by = c("BGC", "rSMR")) %>%
  dplyr::select(BGC, rSMR, SNR, SiteUnit, Edatopic, aSMR) %>%
  distinct()

SS.edatopic.range <- SS.edatopic %>%
  group_by(SiteUnit) %>%
  summarise(
    SNR.range = paste0(min(SNR), "_", max(SNR)),
    rSMR = paste0(min(rSMR), "_", max(rSMR)),
    aSMR = paste0(min(aSMR), "_", max(aSMR)),
    .groups = "drop"
  ) %>%
  mutate(SiteUnit = gsub("/", "_", SiteUnit)) %>%
  dplyr::select(SiteUnit, SNR.range, rSMR, aSMR)


```
## Climate of units
```{r}
#------------------------------------------------------------
# 9. Plot climate data
#------------------------------------------------------------
plot.climate <- readRDS("./clean_data/BECMaster_climr.rds")
clim.alliance <- plot.climate |>
  left_join(
    su.class, 
    by = "PlotNumber"
  ) |>
  filter(!is.na(alliance))

kruskal.test(MAT ~ alliance, data = clim.alliance)

kruskal.test(MAP ~ alliance, data = clim.alliance)

kruskal.test(CMD_an ~ alliance, data = clim.alliance)


xx <- pairwise.wilcox.test(
  clim.alliance$MAT,
  clim.alliance$alliance,
  p.adjust.method = "BH"
)

sig.table <- xx$p.value |>
  as.data.frame() |>
  rownames_to_column("alliance1") |>
  pivot_longer(
    -alliance1,
    names_to = "alliance2",
    values_to = "p_adj"
  ) |>
  filter(!is.na(p_adj)) |>
  mutate(
    sig = case_when(
      p_adj <= 0.001 ~ "***",
      p_adj <= 0.01  ~ "**",
      p_adj <= 0.05  ~ "*",
      TRUE           ~ ""
    )
  ) |>
  arrange(p_adj)

gt::gt(sig.table)
```

```{r}
library(dplyr)
library(ggplot2)
library(vegan)

#==============================================================================
# Summarize climate by SiteUnit
#==============================================================================

su.clim <- clim.alliance |>
  group_by(siteunit, assoc, alliance) |>
  summarise(
    PAS_an = mean(PAS_an, na.rm = TRUE),
    CMD_an = mean(CMD_an, na.rm = TRUE),
    MWMT   = mean(MWMT, na.rm = TRUE),
    MCMT   = mean(MCMT, na.rm = TRUE),
    .groups = "drop"
  )

#==============================================================================
# Climate variables for PCA
#==============================================================================

clim.vars <- su.clim |>
  dplyr::select(PAS_an, CMD_an, MWMT, MCMT)

# Remove rows with missing values
keep.rows <- complete.cases(clim.vars)

clim.vars <- clim.vars[keep.rows, ]

# Remove constant variables if any
clim.vars <- clim.vars |>
  dplyr::select(
    where(~ sd(.x, na.rm = TRUE) > 0)
  )

su.clim <- su.clim[keep.rows, ]


# clim.rda <- rda(
#   clim.vars ~ alliance,
#   data = su.clim,
#   scale = TRUE
# )
# anova(clim.rda)
# anova(clim.rda, by = "axis")
#==============================================================================
# PCA
#==============================================================================

clim.pca <- rda(clim.vars, scale = TRUE)

scores.pca <- scores(clim.pca, display = "sites") |>
  as.data.frame() |>
  bind_cols(
    su.clim |>
      dplyr::select(siteunit, assoc, alliance)
  )

#==============================================================================
# Variance explained
#==============================================================================

eig <- eigenvals(clim.pca)
pct <- round(100 * eig / sum(eig), 1)

sum(pct[1:2])
sum(pct[1:3])

loadings <- scores(clim.pca, display = "species")


#==============================================================================
# Association centroids
#==============================================================================

assoc.cent <- scores.pca |>
  group_by(assoc, alliance) |>
  summarise(
    PC1 = mean(PC1),
    PC2 = mean(PC2),
    n_siteunits = n(),
    .groups = "drop"
  )

#==============================================================================
# Alliance centroids
#==============================================================================

alliance.cent <- assoc.cent |>
  group_by(alliance) |>
  summarise(
    PC1 = mean(PC1),
    PC2 = mean(PC2),
    n_assoc = n(),
    .groups = "drop"
  )

#==============================================================================
# SiteUnit -> Association links
#==============================================================================

links.su <- scores.pca |>
  dplyr::select(
    siteunit,
    assoc,
    alliance,
    PC1.su = PC1,
    PC2.su = PC2
  ) |>
  left_join(
    assoc.cent,
    by = c("assoc", "alliance")
  ) |>
  rename(
    PC1.assoc = PC1,
    PC2.assoc = PC2
  )

#==============================================================================
# Association -> Alliance links
#==============================================================================

links.assoc <- assoc.cent |>
  left_join(
    alliance.cent,
    by = "alliance",
    suffix = c(".assoc", ".alliance")
  )

library(dplyr)

hulls <- scores.pca |>
  group_by(alliance) |>
  slice(chull(PC1, PC2)) |>
  ungroup()
#==============================================================================
# Plot
#==============================================================================

ggplot() +

  #--------------------------------------------------------------------------
  # SiteUnit -> Association lines
  #--------------------------------------------------------------------------

  geom_segment(
    data = links.su,
    aes(
      x = PC1.su,
      y = PC2.su,
      xend = PC1.assoc,
      yend = PC2.assoc
    ),
    colour = "grey80",
    linewidth = 0.3
  ) +

  #--------------------------------------------------------------------------
  # Association -> Alliance lines
  #--------------------------------------------------------------------------

  geom_segment(
    data = links.assoc,
    aes(
      x = PC1.assoc,
      y = PC2.assoc,
      xend = PC1.alliance,
      yend = PC2.alliance
    ),
    colour = "black",
    linewidth = 1
  ) +

  #--------------------------------------------------------------------------
  # Alliance ellipses
  #--------------------------------------------------------------------------
geom_polygon(
  data = hulls,
  aes(
    PC1,
    PC2,
    fill = alliance,
    group = alliance
  ),
  alpha = 0.15,
  colour = NA
)+
  # stat_ellipse(
  #   data = scores.pca,
  #   aes(
  #     PC1,
  #     PC2,
  #     colour = alliance
  #   ),
  #   level = 0.68
  #   linewidth = 0.8,
  #   alpha = 0.6
  # ) +

#--------------------------------------------------------------------------
# SiteUnit points
#--------------------------------------------------------------------------

geom_point(
  data = scores.pca,
  aes(
    PC1,
    PC2,
    colour = alliance
  ),
  alpha = 0.4,
  size = 2
) +

geom_text(
  data = scores.pca,
  aes(
    PC1,
    PC2,
    label = siteunit,
    colour = alliance
  ),
  size = 2.5,
  alpha = 0.7,
  check_overlap = TRUE,
  show.legend = FALSE
) +
  #--------------------------------------------------------------------------
  # Association centroids
  #--------------------------------------------------------------------------

  geom_point(
    data = assoc.cent,
    aes(
      PC1,
      PC2,
      fill = alliance
    ),
    shape = 21,
    colour = "black",
    size = 5,
    stroke = 1
  ) +

  geom_text(
    data = assoc.cent,
    aes(
      PC1,
      PC2,
      label = assoc
    ),
    size = 3,
    nudge_y = 0.10,
    show.legend = FALSE
  ) +

  #--------------------------------------------------------------------------
  # Alliance centroids
  #--------------------------------------------------------------------------

  geom_point(
    data = alliance.cent,
    aes(
      PC1,
      PC2,
      fill = alliance
    ),
    shape = 23,
    colour = "black",
    size = 8,
    stroke = 1.5
  ) +

  geom_text(
    data = alliance.cent,
    aes(
      PC1,
      PC2,
      label = alliance
    ),
    fontface = "bold",
    size = 4,
    nudge_y = 0.25,
    show.legend = FALSE
  ) +

  theme_bw() +

  labs(
    x = paste0("PC1 (", pct[1], "%)"),
    y = paste0("PC2 (", pct[2], "%)"),
    colour = "Alliance",
    fill = "Alliance"
  )
```

## Summarize by Alliance

Generate vegetation summary output for all proposed alliances on a single sheet

```{r compare alliances}
reduced.exceptions <- c("SPHAGNUM","CLADONIA","CLADINA","RACOMITR","MNIUM")
reduced.lifeforms  <- c(9, 10, 11)
su.alliance <- su.anal %>% mutate(SiteUnit = working_Alliance) %>% select(PlotNumber, SiteUnit)

#------------------------------------------------------------
# 3. Run pairwise vegetation similarity analysis
#------------------------------------------------------------
pairwise.params <- list(
  su                = su.alliance,
  minconstancy          = 0,
  noiseconstancy        = 10,
  minplots              = 1,
  minor                 = 1,
  use.ksi               = FALSE,
  ksi   = NULL,
  ksi.value             = 1.5,
  reduce.lifeform       = TRUE,
  reduced.lifeforms     = reduced.lifeforms,
  reduction             = 0.1,
  reduced.exceptions    = reduced.exceptions,
  add_vegsum_table      = NULL
)

vegsum.pairs <- do.call(
  do_pairwise,
  c(list(veg.dat2), pairwise.params)
)

# Start with everything except `su`
print.params <- pairwise.params[setdiff(names(pairwise.params), "su")]

# Remove ksi if use.ksi == FALSE
if (!pairwise.params$use.ksi) {
  print.params <- print.params[setdiff(names(print.params), c("ksi", "ksi.value"))]
}

# Remove reduction-related fields if reduce.lifeform == FALSE
if (!pairwise.params$reduce.lifeform) {
  print.params <- print.params[setdiff(names(print.params),
                                       c("reduced.lifeforms", "reduction", "reduced.exceptions"))]
}
library(tibble)
library(knitr)

param.table <- tibble(
  parameter = names(print.params),
  value = sapply(print.params, function(x) paste(x, collapse = ", "))
)

param.table %>%
  gt::gt() %>%
  gt::tab_header(title = "Parameter Settings for do_pairwise()") %>%
  gt::tab_options(
    table.font.size = gt::px(8),
    table.border.top.width = gt::px(1),
    table.border.bottom.width = gt::px(1)
  )



#------------------------------------------------------------
# 4. Extract comparison table
#------------------------------------------------------------
fwrite(vegsum.pairs, "./outputs/vegsum_pairs.csv")

### return unit species information
vegsum.pairs.spp <- vegsum.pairs |> dplyr::select(Unit1, Unit2, Species, ScientificName, MeanCov.unit1, Constancy.unit1, MeanCov.unit2, Constancy.unit2, diff.pts.unit1, diff.pts.unit2)

### return distinct unit comparisons

unit.compare1 <- vegsum.pairs %>%
  dplyr::select(
    Unit1, Unit2, diff.ratio.unit1, diff.ratio.unit2, nplots.unit1, nplots.unit2, unit.diag.sum.unit1, unit.diag.sum.unit2
  ) %>%
  distinct(Unit1, Unit1, .keep_all = TRUE)


unit.compare <- unit.compare1 %>% 
    dplyr::select(
      Unit1 = Unit2,
      Unit2 = Unit1,
      diff.ratio.unit1 = diff.ratio.unit2,
      diff.ratio.unit2 = diff.ratio.unit1,
      nplots.unit1 = nplots.unit2,
      nplots.unit2 = nplots.unit1,
      unit.diag.sum.unit1 = unit.diag.sum.unit2,
      unit.diag.sum.unit2 = unit.diag.sum.unit1
      ) |>  rbind(unit.compare1) |> 
  distinct() |> mutate(BEC.sim.min = pmin(diff.ratio.unit1, diff.ratio.unit2),
                      BEC.sim.max = pmax(diff.ratio.unit1, diff.ratio.unit2))


fwrite(unit.compare , "./outputs/alliance_compare.csv")
#------------------------------------------------------------
# 5. Build BEC distance matrix and export
#------------------------------------------------------------

source("../BEC_R_working/_functions/_bec_dist_matrix.R")

bec_dist_matrix <- bec_dist_matrix(
  vegsum.pairs,
  distance = "BEC.sim.min"
) %>% 
  data.frame()

openxlsx::write.xlsx(
  x = bec_dist_matrix,
  file = "./clean_tabs/Coast_Forest_Alliance_Distance_matrix.xlsx",
  rowNames = TRUE
)
```

```{r summary report by alliances}
author = "MH_Forest_Alliance"

vegsum.wbk <- createWorkbook()
  # Build vegetation summary for this group
  vegSum_sub <- create_veg_sum_all(
    vdat = veg.dat2,
    siteUnits = su.alliance,
    minconstancy = 50,
    noiseconstancy = 25,
    strata.by = "Lifeform"
  )

  # Format table
  veg.sum.table_sub <- format_veg_table_nostrata(
    vsum = vegSum_sub,
    spp = taxon.lifeform
  )
sheet_name <- author
  # Name the sheet
# sheet_name <- paste0("Group_", g)

# Trim to 25 chars, then add a 5‑char hash suffix
# if (nchar(sheet_name) > 31) {
#   suffix <- substr(digest::digest(sheet_name), 1, 5)
#   sheet_name <- paste0(substr(sheet_name, 1, 25), "_", suffix)
# }

addWorksheet(vegsum.wbk, sheetName = sheet_name)


  # Write data
  writeData(vegsum.wbk, sheet = sheet_name, veg.sum.table_sub)
#}

saveWorkbook(
  vegsum.wbk,
  paste0("./vegsum.tables/", author, ".xlsx"),
  overwrite = TRUE
)

```

```{r build veg guide summary and export to excel, echo = FALSE, warning=FALSE, message=FALSE}
# 
# data.path = "D:/OneDrive - Government of BC/GitHub_data"
# author <- "test.assoc1"
# assoc.list <- bgcv13 %>% filter(!is.na(Association_Code)) %>% group_by(Association_Code) %>% mutate(n=n()) %>% filter(n>2) %>% distinct(Association_Code) %>% pull(Association_Code)
# 
# vegsum.wbk <- createWorkbook()
# for(assoc.choose in assoc.list){
# #assoc.choose = "assoc-13"
# assoc <- bgcv13 %>% filter(Association_Code %in% assoc.choose) %>% dplyr::select(SS_Label, Association_Code) %>% mutate(SiteUnit = gsub(" ", "", SS_Label)) %>% mutate(SiteUnit = str_replace(SiteUnit, "/", "_"))
# su.assoc <- left_join(assoc, su, by = "SiteUnit") %>%  mutate(SiteUnit = str_replace(SiteUnit, "b$|a$", "")) %>% distinct() 
# veg.dat.assoc <- veg.dat2 %>% filter(PlotNumber %in% su.assoc$PlotNumber)
# # vegsum.wbk <- createWorkbook()
# # #bgc.choose = "CDFmmtab2"
# # #for(bgc.choose in bgc.list){
# vegSum <-
#   create_veg_sum_all(vdat = veg.dat.assoc, siteUnits = su.assoc, strata.by = "Lifeform")
# 
# indic.order <- build_species_ordering_hier(vdat = veg.dat.assoc, vsum = vegSum, code.lump=lump, siteUnits = su.assoc, hier.level = "Association_Code")
# veg.sum.table <- format_veg_table2(vsum = vegSum, spp=taxon.lifeform)
# openxlsx::addWorksheet(vegsum.wbk, sheetName = assoc.choose)
# openxlsx::writeData(vegsum.wbk, sheet = assoc.choose, veg.sum.table) 
# }
# 
# saveWorkbook(vegsum.wbk, paste0("./vegsum.tables/LMH77_draft_assoc_tables_", author, ".xlsx"), overwrite = TRUE)

```


