---
title: "Final Hierarchy Unit Materials"
author: "William H MacKenzie"
date: today
format:
  typst:
    toc: false
    toc-depth: 1
    toc-title: Contents
    section-numbering: 1.1.1
    columns: 1
editor: source
execute:
  echo: false
  error: false
  warning: false
  message: false
  fig.width: 6
  fig.height: 4
  fig.align: 'center'
  fig.cap: true
  fig.pos: H
  out.width: '100%'
  dev: pdf
  fig.ext: pdf
  cache: false
  fig.retina: 2
  dpi: 600
  fig.asp: 1.5
  fig.path: "./figures/"
output-file: "FdPy_Generated_Hierarchy.pdf"
---
```{r}
knitr::opts_chunk$set(echo = FALSE)
require(tidyverse)
require(DBI)
require(data.table)
require(cluster)
require(dendextend)
require(dynamicTreeCut)
require(gtable)
require(gtsummary)
require(colorspace)
require(openxlsx)
require(tictoc)
require(labdsv)
require(indicspecies)
require(Polychrome)
require(ggdendro)
require(purrr)
require(ggplotify)
require(grid)
require(data.tree)
library(stats)
library(overlapping)
require(gt)
require(vegan)
library(tidytext)
library(stringr)

set.seed(1279)
source("../BEC_R_working/_functions/_bec_dist.R")
source("../BEC_R_working/_functions/_bec_dist_matrix.R")
source("../BEC_R_working/_functions/_combine_taxa.R")
source("../BEC_R_working/_functions/_combine_taxa_strata.R")
source("../BEC_R_working/_functions/_create_su_vegdata.R")
source("../BEC_R_working/_functions/_create_analysis_vegsum.R")
source("../BEC_R_working/_functions/_TabletoTree.R")
source("../BEC_R_working/_functions/_TreetoTable.R")
source("../BEC_R_working/_functions/_add_vars.R")
source("../BEC_R_working/_functions/_do_pairwise.R")
source("../BEC_R_working/_functions/_create_diagnostic_veg.R")
source("../BEC_R_working/_functions/_return_similar_pairs.R")
source("../BEC_R_working/_functions/_read_sppmaster.R")
source("../BEC_R_working/_functions/_combined_su.R")
source("../BEC_R_working/_functions/_create_veg_sum.R")
source('../BEC_R_working/_functions/_create_veg_sum_all.R')
source('../BEC_R_working/_functions/_build_species_ordering.R')
source('../BEC_R_working/_functions/_format_veg_table.R')
source('../BEC_R_working/_functions/_format_veg_table_strata.R')
source('../BEC_R_working/_functions/_format_veg_table_nostrata.R')
source('../BEC_R_working/_functions/_encode_veg_sum.R')
source('../BEC_R_working/_functions/_create_dendro.R')
source('../BEC_R_working/_functions/_create_dendro_all.R')
source('../BEC_R_working/_functions/_create_dendro_bybgc.R')
source('../BEC_R_working/_functions/_draw_dendro_split.R')
source('../BEC_R_working/_functions/_cluster_groups.R')
source('../BEC_R_working/_functions/_TabletoTree.R')
source('../BEC_R_working/_functions/_build_species_ordering_hier.R')
source("../BEC_R_working/_functions/_cluster_by_spp_group.R")
source("../BEC_R_working/_functions/_generate_bec_hierarchy.R")
source("../BEC_R_working/_functions/_link_su_to_hierarchy.R")
source("../BEC_R_working/_functions/_load_veg_data.R")
source("../BEC_R_working/_functions/_load_siteunits.R")
source("../BEC_R_working/_functions/_run_pairwise_similarity.R")
source("../BEC_R_working/_functions/_generate_hierarchy_units.R")
source("../BEC_R_working/_functions/_generate_review_workbooks.R")
```

# This script will generate the material for and a formatted report of units from a hierarchical analysis.

For each hierarchical level chosen:
  - mapped distribution of plots and shaded BGCs (OK)
  - vegetation summary table for unit or optionally member site units or subassociations
  - actual edatopic grid unit with plot count and shade cells (extrapolated)
  - summary of environment and climate variables
  - table and figure of related concepts (high similarity)
  - Photo
  - ID type plot
  - Tichy profile

```{r load data}
veg.dat <- load_veg_data(veg_file = "./clean_data/Analysis_BECMaster_Veg.rds",
                   veg_combine_db = "D:/BC_Correlation2_Vpro_2023/Correlation2_Spp_lump.accdb",
                   veg_combine_tbl = "CorrelationII_Lump")
veg.dat2 <- veg.dat$veg_dat_lumped

becdb <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/OneDrive - Government of BC/BECdb_Working/BECdb_ver14_2026_v1.accdb;"
)
BGC_all <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(FSRegion, BGC_NoSpace)

BEC.hier <- dbReadTable(becdb, "tblBEC_Hierarchy_Class") %>%
  #filter(SU_Label %in% su2$SU_Label) %>%
  mutate(SiteUnit = str_replace_all(SU_Label, c("/" = "_", " " = "")))

bgcv14 <- dbReadTable(becdb, "tblBGC_UnitsWorkingandArchive") %>%
  filter(is.na(BGCVersionRetired)) %>%
  dplyr::select(BGC_NoSpace, SiteClassificationStatus)

edatopic <- dbReadTable(becdb, "Edatopic_v13_3") %>%
  mutate(
    rSMR = paste0("rSMR", substr(Edatopic, 2, 2)),
    SNR = substr(Edatopic, 1, 1)
  ) %>%
  filter(startsWith(Source, "BEC")) %>%
  dplyr::select(BGC, rSMR, SNR, SS_NoSpace, Edatopic) %>%
  distinct() %>%
  mutate(SiteUnit = gsub("/", "_", SS_NoSpace))

dbDisconnect(becdb)

becmaster <- dbConnect(
  odbc::odbc(),
  .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster/BECMaster_fixing.accdb;"
)
plot.env <- dbReadTable(becmaster, "BECMaster_fixing_Env")
plot.info <- dbReadTable(becmaster, "BECMaster_fixing_Admin") |> rename( BGC = GIS_BGC) |> mutate(BGC = gsub("\\s+", "", BGC))
dbDisconnect(becmaster)

db <- "D:/BC_Correlation2_Vpro_2023/All_BC_Correlation.accdb"
su <- rbind(combined_su(db))#, combined_su(db2))

su <- su %>%
  filter(!grepl("X|unplaced|omit|support|nudum|Ww|!", SiteUnit)) %>%
  arrange(SiteUnit)
su2 <- su %>%
  mutate(SiteUnit = str_replace(SiteUnit, "[abc]$", ""),# ignore phases
         # SiteUnit = if_else(
         #   str_detect(SiteUnit, "101"),
         #   str_replace(SiteUnit, "\\.[123]$", ""), # remove zonal variations
           SiteUnit
         ) %>%
    mutate( SU_Label = str_replace(SiteUnit.orig, "[abc]$", ""))

su.class <- su2 %>%
  left_join(BEC.hier, by = c("SiteUnit" = "SiteUnit", "SU_Label" = "SU_Label"))

```

```{r prep env data}
plot.env2 <- plot.env %>% mutate(lon = ifelse(Longitude<0, Longitude, 0-Longitude)) %>% mutate(Date = as_date(Date)) |> 
  filter(PlotNumber %in% su.class$PlotNumber) |>
  dplyr::select(PlotNumber, lon, lat = Latitude, elev = Elevation, special = Exposure1, rSMR = MoistureRegime, rSNR = NutrientRegime, slope = SlopeGradient, aspect = Aspect ) 
 plot.info <- plot.info |> dplyr::select(Plot, BECSiteUnit, UserSiteUnit, HumusThickness, BGC)  |> filter(Plot %in% su.class$PlotNumber)
plot.env2 <- left_join(plot.env2, plot.info, by = c("PlotNumber" = "Plot")) |> filter(!is.na (rSMR))


## extract climate data for plots from climr
require(climr)
plot.locs <- plot.env2 %>% dplyr::select(id=PlotNumber, BGC, lon,lat,elev) |> drop_na(lat)
plot.climate <- climr::downscale(setDT(plot.locs),
                              vars = list_vars(set = c("Annual", "Seasonal", "Monthly")),
                              return_refperiod = TRUE,
                              db_option = "local",
                              cache = TRUE)
## calculate aSMR
plot.env3 <- plot.climate |> select(PlotNumber = id, everything()) |> left_join(plot.env2) |> 
  mutate(
    moist = rowSums(
      across(c(PPT_05, PPT_06, PPT_07, PPT_08, PPT_09), ~ .x >= 90),
      na.rm = TRUE
    ),
    moist = ifelse(moist > 4, 1, 0),
    ratio = moist / 5,
    multiplier = 1 + ratio,
    CMD = CMD_an / multiplier
  ) |> mutate(PPT.dorm = PPT_at + PPT_wt) |> 
  mutate(Def = ifelse(CMD >=350, 350 - PPT.dorm, CMD)) |> 
  mutate(CMD.total = CMD + Def) |> 
  mutate(CMD.site = ifelse(slope > 35 & (aspect >180 & aspect <270), CMD.total *1.25,
                           ifelse(slope> 35 & (aspect>315 | aspect <45), CMD.total *0.85, CMD.total))) |> arrange(PlotNumber) |> 
  mutate(
    rSMR.clean = sub("^.*?(\\d+).*$", "\\1", rSMR)) |> 
  mutate(rSMR.clean = as.numeric(rSMR.clean),
  SNR.clean = recode(
    substr(trimws(rSNR), 1, 1),
    "A" = 1L,
    "B" = 2L,
    "C" = 3L,
    "D" = 4L,
    "E" = 5L,
    "F" = 6L,
    .default = NA_integer_
  ))
  ####To adjust in zones with prolonged snowpack remove CMD from early growing season months
## BGCs with or below heavy snow packs
# BGC_Special <- fread("./inputs/BGC_Special.csv", data.table = FALSE)
# allDat2 <- left_join(allDat2,BGC_Special, by = c("BGC" = "BGC"))
# allDat2 <- allDat2 %>% mutate(CMD = ifelse(is.na(Special), CMD, CMD_08)) %>% select(-Special, -CMD_08) %>% drop_na()
#rules <- read.csv("./inputs/aSMR_growingseason_Rules_HalfStep_v13_13Sept2025.csv")
rules <- read.csv("./inputs/aSMR_Rules_HalfStep_v13_23June2025.csv") |>
  mutate(
    rSMR.rule = sub("^rSMR", "", rSMR)) |> dplyr::select(-rSMR)

plot_asmr <- plot.env3 %>% select(PlotNumber, rSMR, CMD.site) |> 
  left_join(rules, by = c("rSMR" = "rSMR.rule")) %>%
  filter(CMD.site <= CMD) %>%
  group_by(PlotNumber) %>%  # ensure row-wise logic if needed
  slice_max(aSMR, with_ties = FALSE) %>%
  ungroup() %>% select(PlotNumber, aSMR)

plot.env3 <- left_join(plot.env3, plot_asmr)

plot.env3 <- plot.env3 |>
  left_join(
    su.class, 
    by = "PlotNumber")
 
plot.env.vars <- plot.env3 |>   select(-SiteUnit, -working_association, -working_alliance) |> 
   select(PlotNumber, PAS_an,CMD.site, MSP, aSMR , SNR = SNR.clean, Tmin = Tmin_an, AHM,FFP, DD5 = DD5_an, DD18_sm ) |> arrange(PlotNumber)
 keep.rows <- complete.cases(plot.env.vars)

  su.env <- plot.env3 |> rename(assoc = working_association, alliance = working_alliance) |> 
  filter(!is.na(alliance)) |> arrange(SiteUnit) |> 
  group_by(SiteUnit, assoc, alliance) |>
    summarise(
  PAS_an = mean(PAS_an, na.rm = TRUE),
  CMD.site = mean(CMD.site, na.rm = TRUE),
  MSP   = mean(MSP, na.rm = TRUE),
  aSMR = mean(aSMR, na.rm = TRUE),
  SNR = mean(SNR.clean, na.rm = TRUE),
  AHM = mean(AHM, na.rm = TRUE),
  #lat = mean(lat, na.rm = TRUE),
  FFP   = mean(FFP, na.rm = TRUE),
  DD5 = mean(DD5_an, na.rm = TRUE),
  DD18_sm = mean(DD18_sm, na.rm = TRUE),
  Tmin = mean(Tmin_an, na.rm = TRUE),
  .groups = "drop"
)

#     summarise(
#   across(
#     where(is.numeric) & !c(lat, lon),
#     ~ mean(.x, na.rm = TRUE)
#   ),
#   .groups = "drop"
# )

# ==============================================================================
# Climate variables for PCA
#==============================================================================
  env.vars <- su.env |>
  select(-assoc, -alliance) 

# Remove rows with missing values
keep.rows <- complete.cases(env.vars)

env.vars <- env.vars[keep.rows, ]

# Remove constant variables if any
env.vars.filt <- env.vars |>
  select(-SiteUnit) |>
  dplyr::select(
    where(~ is.numeric(.x) && sd(.x, na.rm = TRUE) > 0)
  )

keep.rows <- complete.cases(env.vars.filt)

#su.env <- env.vars[keep.rows, ]
su.env <- su.env |>
  dplyr::select(
    SiteUnit, assoc, alliance,
    where(~ is.numeric(.x) && sd(.x, na.rm = TRUE) > 0)
  )

```


## choose alliance
```{r}
alls <- unique(su.class$working_alliance)
alls
anal.set = "CWH_HwBa"
su.anal <- su.class |> filter(working_alliance %in% anal.set)
```


## Distribution Map for each unit
```{r}
#| label: fig-sites
#| fig-cap: "Locations of plots for alliance"
#| message: false
#| warning: false

library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)
library(grid)

plots <- plot.env3 |> select(alliance = working_alliance, assoc = working_association, lon, lat, bgc, PlotNumber)%>%
filter(!is.na(lon),!is.na(lat),!is.na(assoc))
# plots.bgc.good <- plots |> add_count(alliance, BGC) |> filter (n>3)
# plots.bgc.bad <- plots |> group_by(alliance) |> add_count(BGC) |> filter (n<=3)

plots_sf <- st_as_sf(plots, coords=c("lon", "lat"), crs = 4326)

bgc.anal <- st_read("D:/OneDrive - Government of BC/CommonTables/BGC_maps/BGCv13_simplified.gpkg") |> filter(BGC %in% plots$bgc)

#----------------------------------------
# Site locations
#----------------------------------------


plots_sf$assoc <- factor(
  plots_sf$assoc) 
#----------------------------------------
# Basemap layers
#----------------------------------------

bc <- ne_states(
  country = c(
    "Canada",
    returnclass = "sf" #"United States of America"
  )) |> 
  filter(name_en == "British Columbia")

lakes <- ne_download(
  scale = 10,
  type = "lakes",
  category = "physical",
  returnclass = "sf"
)

rivers <- ne_download(
  scale = 10,
  type = "rivers_lake_centerlines",
  category = "physical",
  returnclass = "sf"
)

# Optional towns if needed later
towns <- ne_download(
  scale = "small",
  type = "populated_places",
  category = "cultural",
  returnclass = "sf"
)

towns_bc <- st_crop(
  towns,
  st_bbox(plots_sf) + c(-1, -1, 1, 1)
)

#----------------------------------------
# Map
#----------------------------------------


ggplot() +
  geom_sf(data = bc, fill = "grey95", colour = "black") +
  geom_sf(data = lakes, fill = "lightblue", colour = NA) +
  geom_sf(data = rivers, colour = "lightblue3", linewidth = 0.2) +
  geom_sf(
    data = bgc.anal,
    aes(fill = BGC),
    colour = NA,
    alpha = 1
  ) +
  geom_sf(
    data = plots_sf,
    aes(colour = assoc),
    size = 1.5,
    alpha = 0.8
  ) +
  coord_sf(expand = FALSE) +
  scale_colour_viridis_d(option = "turbo") +
  theme_bw() +
  theme(
    legend.position = "right"
  ) +
  labs(
    colour = "Association",
    title = paste0("Plot Locations ", anal.set, " Alliance")
  )+

  coord_sf(
    xlim = c(
      min(plots$lon) - 3,
      max(plots$lon) + 3
    ),
    ylim = c(
      min(plots$lat) - 1,
      max(plots$lat) + 1
    )
  ) +

  labs(
    x = "Longitude",
    y = "Latitude"
  ) +
      theme_bw(base_size = 10) +

  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),

    panel.border = element_rect(
      colour = "black",
      fill = NA
    ),

    legend.position = "right",
    legend.direction = "vertical",

    legend.title = element_text(
      face = "bold"
    ),

    legend.text = element_text(
      size = 9
    ),

    legend.key = element_blank()
  ) +

  guides(
    shape = guide_legend(
      nrow = 1,
      byrow = TRUE
    )
  )
```



```{r build veg guide summary and export to excel, echo = FALSE, warning=FALSE, message=FALSE}
require(openxlsx)
require(tictoc)
library(stringr)
data.path = "D:/OneDrive - Government of BC/GitHub_data"
author <- "test.assoc1"
assoc.list <- bgcv13 %>% filter(!is.na(Association_Code)) %>% group_by(Association_Code) %>% mutate(n=n()) %>% filter(n>2) %>% distinct(Association_Code) %>% pull(Association_Code)

vegsum.wbk <- createWorkbook()
for(assoc.choose in assoc.list){
#assoc.choose = "assoc-13"
assoc <- bgcv13 %>% filter(Association_Code %in% assoc.choose) %>% select(SS_Label, Association_Code) %>% mutate(SiteUnit = gsub(" ", "", SS_Label)) %>% mutate(SiteUnit = str_replace(SiteUnit, "/", "_"))
su.assoc <- left_join(assoc, su, by = "SiteUnit") %>%  mutate(SiteUnit = str_replace(SiteUnit, "b$|a$", "")) %>% distinct() 
veg.dat.assoc <- veg.dat2 %>% filter(PlotNumber %in% su.assoc$PlotNumber)
# vegsum.wbk <- createWorkbook()
# #bgc.choose = "CDFmmtab2"
# #for(bgc.choose in bgc.list){
vegSum <-
  create_veg_sum_all(vdat = veg.dat.assoc, siteUnits = su.assoc, strata.by = "Lifeform")

indic.order <- build_species_ordering_hier(vdat = veg.dat.assoc, vsum = vegSum, code.lump=lump, siteUnits = su.assoc, hier.level = "Association_Code")
veg.sum.table <- format_veg_table2(vsum = vegSum, spp=taxon.lifeform)
openxlsx::addWorksheet(vegsum.wbk, sheetName = assoc.choose)
openxlsx::writeData(vegsum.wbk, sheet = assoc.choose, veg.sum.table) 
}

saveWorkbook(vegsum.wbk, paste0("./vegsum.tables/LMH77_draft_assoc_tables_", author, ".xlsx"), overwrite = TRUE)

```
# Polished Ecosystem Unit Summary Report scripts from a copilot session
```{r}
# Load required packages
library(dplyr)
library(readr)
library(sf)
library(ggplot2)
library(glue)

# 1. Import species abundance data
# species_abundance <- read_csv("species_abundance.csv")
# 
# # 2. Import distribution map (shapefile or GeoJSON)
# distribution_map <- st_read("distribution_map.shp")
# 
# # 3. Import environmental summary table
# environment_summary <- read_csv("environment_summary.csv")
# 
# # 4. Summarize species abundance
# top_species <- species_abundance %>%
#   arrange(desc(Abundance)) %>%
#   slice_head(n = 5)
# 
# dominant_species <- paste(top_species$Species, collapse = ", ")
# 
# # 5. Spatial summary (example: count occurrences per species)
# spatial_summary <- distribution_map %>%
#   st_drop_geometry() %>%
#   count(Species, name = "Occurrences") %>%
#   arrange(desc(Occurrences))
# 
# # 6. Environmental drivers
# env_vars <- paste(environment_summary$Variable, collapse = ", ")
# 
# # 7. Generate polished ecosystem summary
# ecosystem_summary <- glue(
#   "Ecosystem Unit Summary:\n\n",
#   "The community is dominated by {dominant_species}, ",
#   "with {top_species$Species[1]} showing the highest abundance. ",
#   "Spatial distribution analysis indicates that {spatial_summary$Species[1]} ",
#   "is most widespread, with {spatial_summary$Occurrences[1]} mapped occurrences. ",
#   "Environmental drivers shaping this unit include {env_vars}, ",
#   "suggesting strong ecological gradients influencing species composition.\n\n",
#   "Together, these patterns highlight the interplay between biological dominance, ",
#   "spatial distribution, and environmental context in structuring the ecosystem."
# )
# 
# cat(ecosystem_summary)
# 
# # 8. Optional visualization
# ggplot(distribution_map) +
#   geom_sf(aes(color = Species)) +
#   theme_minimal() +
#   labs(title = "Species Distribution Map")
```


