---
title: "Create Environment Data sets and summaries"
author: "Will MacKenzie"
date: "16/01/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

require(tidyverse)
require(tidymodels)
require(data.table)
require(data.tree)
require(DataExplorer)
require(C50)
require(indicspecies)
require(doParallel)
require(Hmsc)
source("./_functions/_TabletoTree.R")
source("./_functions/_TreetoTable.R")
cloud_dir <- "F:/OneDrive - Personal/OneDrive/BEC_Classification_Paper/"
#cloud_dir <- "F:/OneDrive - Government of BC/CCISSv12/"
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggridges)
require(ggplot2)
```

## Environmental comparison of hierarchy units
##Inputs are plots with environment data.
Need to overlay clean locations in ClimateBC to get climate dataset for use in this script
1. rSMR converts to aSMR, rSMR, exposure limiting factors, successional status, slope, aspect, meso slope, 
###Climate + BGC
###Site - aSMR, nutrients, topographic 
###Map


```{r ridgeline plot by hierarhcy unit}

ggplot(diamonds, aes(x = price, y = cut, fill = cut)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
