---
title: "BuildOrderbyTrainingML"
author: "WH MacKenzie"
date: "2024-10-14"
output: html_document
---

 
 
@@ -1,292 +0,0 @@ 
Unstage chunk	
 
1
 
---
2
 
title: "MachineLearningHigherOrders"
3
 
author: "Will MacKenzie & Kiri Daust"
4
 
date: "05/07/2021"
5
 
output: html_document
6
 
---
7
 
8
 
```{r setup, include=FALSE}
9
 
knitr::opts_chunk$set(echo = FALSE)
10
 
11
 
require(tidyverse)
12
 
require(tidymodels)
13
 
require(data.table)
14
 
require(data.tree)
15
 
require(DataExplorer)
16
 
require(C50)
17
 
require(indicspecies)
18
 
require(doParallel)
19
 
require(philentropy)
20
 
require(DBI)
21
 
source("./_functions/_TabletoTree.R")
22
 
source("./_functions/_TreetoTable.R")
23
 
source("./_functions/_VegdatSUsummary.R")
24
 
source("./_functions/_spp_importance.R")
25
 
source("./_functions/_create_su_vegdata.R")
26
 
source("./_functions/_create_analysis_vegsum.R")
27
 
cloud_dir <- "F:/OneDrive - Personal/OneDrive/BEC_Classification_Paper/"
28
 
#cloud_dir <- "F:/OneDrive - Government of BC/CCISSv12/"
29
 
```
30
 
31
 
Steps to build classification hierarchy
32
 
33
 
1. A machine learning model of high-level hierarchical BEC Forest units using selected site series as training data
34
 
a. Create a ML model of Forested SubOrders using BECv11 site series that meet minimum plot requirements. (going to Orders only causes issues where Pl is a dominant secondary species - many provincial units get placed into a Pinucon Order when applied only to the order level even though they have a subdominant tree species which places them in a different order)
35
 
b. Use only tree species for first round of model build for the forested Orders.
36
 
c. Predict membership of new BECv13 units and assign to orders then rebuild ML model.
37
 
d. Review any mis predicted site series for reassignment.
38
 
e. Compare similarity of tree composition within units (noise clustering) to find outliers that may represent new Orders (test by leaving out known units)
39
 
f. Create vegetation summaries of Orders and compare vegetation (all species now). Identify climatic indicator species groups.
40
 
g. Create climatic and site attributes summary of Orders 
41
 
h. Create maps of plot locations for each Order
42
 
i. Hierarchy tree graphic
43
 
44
 
2. Build Alliances. Focus should be on identifying species groups that reflect different site conditions.
45
 
46
 
a. Use some of the techniques of indicspecies package to create Alliances based on indicator group creation and analysis.
47
 
b. Try to build a machine learning model of Alliances
48
 
c. Vegetation and Environment summary of Alliances
49
 
50
 
3. Do pair-wise analysis of site series within Orders/Alliances using high constancy species to analyze site series and create Associations/SubAssociations.
51
 
a. Check for similarity between Associations of All orders
52
 
b. Build relationship graphic
53
 
54
 
55
 
4. Document hierarchy
56
 
57
 
5. Build model to assign new units and unassigned plots
58
 
Challenge here is to make a constant list of species between model and new units.
59
 
a. Predict membership of new site series (draft BECv13). Noise Clustering to test for novel units that do not fit any existing. Use machine learning to hierarchy
60
 
b. Predict Order membership of BECMaster plots that are unassigned to existing site series and add to an Order_unplaced site unit under each Order.
61
 
62
 
6. Build for non-forested units. Classes based on major site level differences rather than climate and using major species groups (e.g. Hydrophytic Carex spp )
63
 
May wish to assign a temporary notree pseudo species to all units that have no trees to help separate forested hierarchy from non-forested hierarchy (Hurdle model)
64
 
65
 
#### Import all raw vpro tables in /raw_data folder
66
 
67
 
```{r set folders, include=FALSE}
68
 
#load("./clean_data/veg_plot_data.RData")
69
 
veg.dat <- readRDS("./clean_data/Analysis_BECMaster_Veg.rds")
70
 
# becmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster19/BECMaster19.accdb;")
71
 
# ENV <- dbReadTable(becmaster, "BECMaster19_Env") %>% mutate(Longitude = ifelse(Longitude<0, Longitude, 0-Longitude))
72
 
# ADMIN<- dbReadTable(becmaster, "BECMaster19_Admin")
73
 
# env.fields <- colnames(ENV)
74
 
# odbc::odbcListColumns(becmaster, table = "BECMaster19_ENV")
75
 
# dbDisconnect(becmaster)
76
 
###created in 0_0_Plot_SU_Hierarchy checks script
77
 
master_su <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)};
78
 
DBQ=C:/BC_Correlation2_Vpro_2023/CoastGuide_Hierarchy.accdb;")
79
 
su <- dbReadTable(master_su, "Coast_Forest_2024v4_SU")
80
 
hier <- dbReadTable(master_su, "CoastForest_v2024_2_Hierarchy")
81
 
dbDisconnect(master_su)
82
 
83
 
```
84
 
85
 
##roll up into site series summary data and reduce to analysis set
86
 
```{r summarize site series, echo=FALSE}
87
 
### test run without non-vasculars
88
 
#vegdata <- veg.dat %>% filter(Lifeform <9 | Lifeform == 12)
89
 
vegdata <- veg.dat %>% filter(Lifeform <3)
90
 
vegsum <- create_su_vegdata(vegdata, su)
91
 
veg_anal <- create_analysis_vegsum (vegsum, minimportance = 0.5, minconstancy = 60, minplots = 0) 
92
 
SS_reduced_list <- unique(veg_anal$SiteUnit)
93
 
SU_reduced <- su %>% filter(SiteUnit %in% SS_reduced_list)
94
 
```
95
 
96
 
97
 
##Import hierarchy matrix from 0_0_Plot_SU_Hierarchy script with original placements plus all unassigned units added to an unplaced Formation category
98
 
```{r import  hierarchy data}
99
 
# hierarchy <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BC_Correlation2_Vpro_2023/CoastGuide_Hierarchy.accdb;")
100
 
# Vpro.hier <- dbReadTable(hierarchy, "Coast_Classes_Training_Hierarchy") %>% as.data.table
101
 
# dbDisconnect(hierarchy)
102
 
##Import wide matrix as training data
103
 
#SUhier <- fread("./outputs/BECv12_Hierarchy_Matrix.csv")
104
 
#source("./_functions/_TreetoTable.R")
105
 
unit = "Order"
106
 
107
 
###Import Vpro hierarchy and turn to widematrix
108
 
SUhier <- treeToTable(hier)
109
 
Hier.clean <- SUhier$table
110
 
Hier.units <- Hier.clean %>% dplyr::select(SiteUnit, unit) %>% distinct()
111
 
Hier.data <- left_join(Hier.units, veg_anal)  %>% arrange(Species) %>% distinct() %>% mutate(unit = ifelse(is.na(unit), "unplaced", unit))
112
 
```
113
 
114
 
115
 
```{r some hierarchy stats}
116
 
classes <- unique(Hier.clean$Class)
117
 
orders <- unique(Hier.clean$Order)
118
 
suborders <- unique(Hier.clean$Suborder)
119
 
suborders
120
 
### Choose hierarchical level for analysis
121
 
122
 
class.dat <-
123
 
  Hier.data %>% dplyr::select(SiteUnit, 2, Species, spp_importance) %>% mutate(Species = ifelse(is.na(Species), "None", Species)) %>% filter(!Species %in% "None") %>% 
124
 
  pivot_wider(id_cols = c(SiteUnit, 2),
125
 
    names_from = Species,
126
 
    values_from = spp_importance) %>% distinct() %>% dplyr::rename("Unit" = 2) %>% mutate(Unit = ifelse(is.na(Unit) | Unit == "", "unplaced", Unit)) %>% filter(!SiteUnit == "") %>%  mutate_if(is.character, as.factor) %>% droplevels() #%>% mutate(None = ifelse(is.numeric(complete.cases(.)), 1, 0))
127
 
 # mutate(Suborder = ifelse(is.na(Suborder) | Suborder == "", "unplaced", Suborder)) %>% filter(!SiteUnit == "") %>%  mutate_if(is.character, as.factor) %>% 
128
 
 #  replace(is.na(spp_importance),1) %>%  
129
 
# class.dat <- class.dat %>% as.matrix %>% as.data.frame
130
 
# class.dat <- class.dat$SiteUnit
131
 
class.dat[is.na(class.dat)] <- 0
132
 
133
 
#DataExplorer::create_report(class.dat)
134
 
```
135
 
Data pre-processing includes the following steps:  
136
 
137
 
```{r prep data, include = TRUE, echo = FALSE}
138
 
139
 
classID <- class.dat %>% dplyr::select(SiteUnit, Unit)
140
 
#class.dat2 <- class.dat %>% select(-SiteUnit)
141
 
BEC_good <- class.dat %>% filter(!Unit == "unplaced") %>% arrange(SiteUnit)
142
 
BEC_good$SiteUnit <- BEC_good$SiteUnit %>% as.factor
143
 
BEC_good[is.na(BEC_good)] <- 0
144
 
SU_names <- as.data.frame(BEC_good$SiteUnit) %>%  distinct() %>% rowid_to_column('.row') %>% dplyr::rename("SiteUnit" = 2)
145
 
146
 
147
 
BEC_new <- class.dat %>% filter(Unit == "unplaced") %>% arrange(SiteUnit)
148
 
nospp <- BEC_new  %>%  mutate(spp_count = ncol(dplyr::select(., -c(1:2))) - rowSums(is.na(dplyr::select(., -c(1:2))))) %>%  filter(spp_count == 0) %>%  dplyr::select(SiteUnit)
149
 
BEC_new  <- BEC_new  %>%  filter(!SiteUnit %in% nospp$SiteUnit)
150
 
BEC_new[is.na(BEC_new)] <- 0
151
 
152
 
```
153
 
154
 
```{r build initial model of higher orders, include = FALSE}
155
 
156
 
BEC_recipe <-
157
 
    recipe(Unit ~ ., data = BEC_good) %>%
158
 
     update_role(SiteUnit, new_role = "id variable") %>% 
159
 
  step_novel(SiteUnit) %>% 
160
 
    prep() 
161
 
    summary(BEC_recipe)
162
 
    
163
 
BEC_fmodel <- rand_forest(mtry = 5, min_n = 2, trees = 501) %>%
164
 
  set_mode("classification") %>%
165
 
  set_engine("ranger", importance = "impurity") #or "permutations
166
 
167
 
# note in RF as a tree based model it is not required to scale and normalize covariates and may have negative influence on the model performance  
168
 
# 2: set up cross validation for parameter tuning data sets # note vc is default to 10 fold
169
 
BEC_workflow <- workflow() %>%
170
 
  add_model(BEC_fmodel) %>% 
171
 
    add_recipe(BEC_recipe,blueprint = hardhat::default_recipe_blueprint(allow_novel_levels = TRUE))
172
 
173
 
BEC_ranger_model <- fit(BEC_workflow, BEC_good)
174
 
BEC_ranger_model$fit
175
 
176
 
save(BEC_ranger_model, file = "./rFmodels/BECv12_Orders_rFmodel.Rdata")
177
 
178
 
#BEC.pred <- predict(BEC_ranger_model, BEC_good) %>% bind_cols(BEC_good %>% dplyr::select(SiteUnit, Unit)) 
179
 
180
 
181
 
# %>%  metrics(truth = Class), estimate = .pred_class)
182
 
BEC.pred.list <- BEC.pred$SiteUnit %>% as.character
183
 
BEC.missed <- BEC.pred  %>%    mutate(Compare = ifelse(.pred_class == Unit, "Same", "Diff")) %>% filter(Compare == "Diff")
184
 
fwrite(BEC.missed, "./outputs/MidplacedOrders.csv")
185
 
BEC.mis.list <- BEC.missed$SiteUnit %>% as.character
186
 
SS_misID_SU <- SU_reduced[SU_reduced$SiteUnit %in% BEC.mis.list,]
187
 
fwrite(SS_misID_SU, "./outputs/WrongClassPredicted_SU.csv")
188
 
189
 
BEC.new.pred <- predict(BEC_ranger_model, BEC_new) %>% bind_cols(BEC_new %>% dplyr::select(SiteUnit, Unit)) %>% dplyr::rename(Class_updated = .pred_class)
190
 
setDT(Hier.clean)[setDT(BEC.new.pred), "Class" := Class_updated, on = .(SiteUnit)]
191
 
Hier.clean <- Hier.clean %>% mutate(Formation = ifelse(is.na(Class), "unplaced", Formation))
192
 
193
 
```
194
 
195
 
```{r create C50 decision tree}
196
 
library(C50)
197
 
require(partykit)
198
 
#to do this well need to convert the modified cover values back to percent cover or something
199
 
BEC_good <- BEC_good %>% filter(!Unit %in% "unplaced") %>% droplevels()
200
 
mod <- C5.0(Unit ~ ., data = BEC_good[-1])
201
 
summary(mod)
202
 
c5plot <- as.party.C5.0(mod, type = "extended", cex = .1)
203
 
204
 
plot(c5plot, gp = gpar(fontsize = 5))
205
 
```
206
 
207
 
208
 
```{r update historic hierarchy and then resubmit to create new model}
209
 
##convert matrix to tree, add in new units
210
 
#Hier.update <- fread("./clean_tabs/BEC13_HierarchyMatrixModelled_v1.csv")
211
 
Hier.update <- Hier.clean
212
 
levelNames <- c("Formation", "Class", "Order", "Suborder", "Alliance", "Suball", "Assoc", "Subass", "Facies", "Working", "SiteUnit")
213
 
testReverse <- tableToTree(hierWide = copy(Hier.update),levelNames) ## convert matrix to tree
214
 
testReverse <-testReverse %>% mutate(Parent = ifelse(Parent == 0, NA, Parent))
215
 
fwrite(testReverse, "./outputs/UpdatedCoastHierarchyv12_updated.csv")
216
 
# newBEC <- BEC.pred
217
 
# newBEC2 <- newBEC %>% rename(Name = .pred_class)
218
 
# pred.ID <- newBEC2 %>% select(Name) %>% distinct() %>% left_join(testReverse) %>% select(Name, ID) ## Parent Code for new predictions
219
 
# newBEC3 <- left_join(newBEC2, pred.ID, by = "Name") %>% select(-Name, -Suborder) %>% rename(Parentnew = ID, Name = SiteUnit)
220
 
# 
221
 
# testReverse2 <- left_join(testReverse, newBEC3) %>% mutate(Parent, ifelse(!is.na(Parentnew), Parentnew, Parent)) %>% 
222
 
#             select(-Parent, -Parentnew) %>% rename (Parent = 4) %>% select(ID, Name, Parent, Level)
223
 
# #testReverse2$Parent <- ifelse(testReverse2$Parent == 1, "", testReverse2$Parent) %>% as.integer
224
 
# ##remove old units from hierarchy.
225
 
# SS.v13 <- fread( "./clean_tabs/BECv13_Forest_SiteSeries.csv",  header = T, stringsAsFactors = F, sep = "") %>% rename(Name = SiteUnit)
226
 
# SS.v13$goodv13 <- "good"
227
 
# testReverse3 <- left_join(testReverse2, SS.v13) 
228
 
# testReverse3 <- testReverse3 %>%  mutate(Name = ifelse(!goodv13 %in% "good" & Level == 11, NA, Name)) %>% select(-goodv13) %>% na.omit()
229
 
# 
230
 
# testReverse3$Parent <- ifelse(testReverse3$Parent == 1, "", testReverse3$Parent) %>% as.integer
231
 
# ###write out to Vpro format to review
232
 
# fwrite(testReverse, "./outputs/UpdatedCoastHierarchyv12_updated.csv")
233
 
# 
234
 
# SUhier2 <- treeToTable(testReverse2)
235
 
# Hier.clean2 <- SUhier2$table
236
 
237
 
```
238
 
239
 
```{r build hierarchy data}
240
 
Hier.units <- Hier.clean2 %>% dplyr::select(SiteUnit, Class, Order, Suborder) %>% distinct()
241
 
###new SU table
242
 
243
 
Hier.data2 <- left_join(Hier.units, SS_good) %>% filter(!is.na(nPlots)) %>% arrange(Species) %>% distinct()
244
 
245
 
class.dat2 <-
246
 
  Hier.data2  %>% dplyr::select(SiteUnit, Suborder, Species, MeanCov) %>% 
247
 
  pivot_wider(id_cols = c(SiteUnit, Suborder),
248
 
    names_from = Species,
249
 
    values_from = MeanCov) %>% 
250
 
  mutate(Suborder = ifelse(is.na(Suborder) | Suborder == "", "unplaced", Suborder)) %>% filter(!SiteUnit == "") %>%  mutate_if(is.character, as.factor) %>% 
251
 
   replace(is.na(.),0) %>% distinct() %>% droplevels()
252
 
```
253
 
254
 
```{r build new model}
255
 
classID <- class.dat2 %>% dplyr::select(SiteUnit, Suborder)
256
 
#class.dat2 <- class.dat %>% select(-SiteUnit)
257
 
BEC_good2 <- class.dat2 %>% filter(!is.na(Suborder) & !Suborder %in% "unplaced") %>% arrange(SiteUnit)
258
 
BEC_good2$SiteUnit <- BEC_good2$SiteUnit %>% as.factor
259
 
SU_names <- as.data.frame(BEC_good2$SiteUnit) %>%  distinct() %>% rowid_to_column('.row') %>% dplyr::rename("SiteUnit" = 2)
260
 
261
 
BEC_recipe <-
262
 
    recipe(Suborder ~ ., data = BEC_good2) %>%
263
 
     update_role(SiteUnit, new_role = "id variable") %>% 
264
 
  step_novel(SiteUnit) %>% 
265
 
    prep() 
266
 
    summary(BEC_recipe)
267
 
    
268
 
BEC_fmodel <- rand_forest(mtry = 5, min_n = 2, trees = 501) %>%
269
 
  set_mode("classification") %>%
270
 
  set_engine("ranger", importance = "impurity") #or "permutations
271
 
```
272
 
273
 
```{r build  second model of suborders, include = FALSE}
274
 
275
 
# 2: set up cross validation for parameter tuning data sets # note vc is default to 10 fold
276
 
BEC_workflow <- workflow() %>%
277
 
  add_model(BEC_fmodel) %>% 
278
 
    add_recipe(BEC_recipe,blueprint = hardhat::default_recipe_blueprint(allow_novel_levels = TRUE))
279
 
280
 
BEC_ranger_model <- fit(BEC_workflow, BEC_good2)
281
 
BEC_ranger_model$fit
282
 
283
 
save(BEC_ranger_model, file = "./rFmodels/BECv12_Suborders_rFmodel_Updated.Rdata")
284
 
285
 
BEC.missed2 <- predict(BEC_ranger_model, BEC_good2) %>% bind_cols(BEC_good2 %>% dplyr::select(SiteUnit, Suborder)) %>% 
286
 
  mutate(Compare = ifelse(.pred_class == Suborder, "Same", "Diff")) %>% filter(Compare == "Diff")
287
 
BEC.mis.list2 <- BEC.missed2$SiteUnit %>% as.character
288
 
289
 
SS_misID_SU <- SUTab[SUTab$SiteUnit %in% BEC.mis.list2,]
290
 
fwrite(SS_misID_SU, "./outputs/WrongOrderPredicted_SU.csv")
291
 
```
292
 
