---
title: "Noise Clustering Tests"
author: "WH MacKenzie"
date: "2024-10-14"
output: html_document
---

 
 
@@ -1,187 +0,0 @@ 
Unstage chunk	
 
1
 
--- 
Unstage selection	
 
2
 
title: "NoiseClustering to Add Plots to Classification"
3
 
author: "William H MacKenzie"
4
 
date: "13/01/2023"
5
 
output: html_document
6
 
editor_options: 
7
 
  markdown: 
8
 
    wrap: 72
9
 
---
10
 
11
 
Based on the vegclust package and procedures described by Caceres and
12
 
Wiser the basic method is plotmembership \<- vegclust::vegclass(vegclust
13
 
object of defined groups, new plot data)
14
 
15
 
1\. We'll be using existing Associations as predefined groups and this
16
 
will be using species importance (sqrt cover \* constancy)
17
 
18
 
2\. Plot data will have different species so we'll need the function
19
 
vegclust::conformveg to harmonize the species lists. May need to do some
20
 
cleanup of species to remove low cover and rare species?
21
 
22
 
3\. Will need experiment to deal with settings of noise level and
23
 
defuzzify
24
 
25
 
```{r setup, include=FALSE}
26
 
knitr::opts_chunk$set(echo = TRUE)
27
 
require(vegclust)
28
 
require(data.table)
29
 
require(tidyverse)
30
 
require(DBI)
31
 
require(labdsv)
32
 
source("./_functions/_lump_species.R")
33
 
```
34
 
35
 
read in veg_anal and SU at the Assoc level convert to vegclust object
36
 
37
 
```{r import data}
38
 
#testunit <- c("CDF mm   /101", "CWH vm 2 /103", "MH  mm 1 /112", "CWH xm 1 /101")
39
 
coast.ss <- fread ("./clean_tabs/CoastGuide_All_BGC_Forest_SS.csv") %>% distinct(SiteUnit)
40
 
41
 
veg.dat <- readRDS("./clean_data/Analysis_BECMaster_Veg.rds")
42
 
43
 
load("./clean_data/SS_sum_analysis.RData")
44
 
45
 
veglump <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=C:/BC_Correlation2_Vpro_2023/CoastGuide_Spp_Lump.accdb;")
46
 
LUMP <- dbReadTable(veglump, "CoastGuide2023_Lump")
47
 
dbDisconnect(veglump) 
48
 
49
 
```
50
 
51
 
```{r prep data and units}
52
 
##remove units with few plots
53
 
plots_SS <- coast.ss %>% group_by(SiteUnit) %>% dplyr::summarise(plots = n())
54
 
SS_count <- length(unique(coast.ss$SiteUnit))
55
 
siteunits_toofew <- plots_SS %>% filter(plots <5) 
56
 
57
 
##set this number to the max number of units
58
 
testunit <- sample_n(coast.ss, 100, replace = FALSE)
59
 
## remove non-vasculars
60
 
vegdata <- veg.dat %>% filter(Lifeform <9 | Lifeform == 12)
61
 
## lump to species and lump file
62
 
vegdata <- lump_species(vegdata, LUMP)
63
 
64
 
assocs <- fread("./outputs/working_SU.csv") %>% arrange(PlotNumber) %>% filter(SiteUnit %in% testunit$SiteUnit) %>% filter(!SiteUnit %in% siteunits_toofew)
65
 
66
 
plots <- unique(assocs$PlotNumber)
67
 
assocs2 <- assocs %>% dplyr::select(-SiteUnit) %>%  column_to_rownames("PlotNumber") %>% t %>% as.vector
68
 
##reduce data to only those in assoc list
69
 
vegdata2 <- vegdata %>% filter(PlotNumber %in% plots) %>% arrange(PlotNumber)
70
 
### reduce list to constant species or high in indiv plots and convet to matrix
71
 
const_spp <- unique(veg_anal$Species)
72
 
73
 
vegdata2 <- vegdata2 %>% mutate(spp_importance = totalcov^0.5) %>% dplyr::select(PlotNumber, Species, spp_importance) %>% filter(spp_importance >2 | Species %in% const_spp ) %>% matrify()
74
 
##produce dissimiarity for all plots
75
 
all_dissim <- proxy::dist(vegdata2, method="ejaccard", diag=TRUE, upper = FALSE)
76
 
77
 
##create vegclust object
78
 
# assoc_data <- as.vegclust(vegdata2, assocs2, method="KM", dnoise = 0.8)
79
 
# xx <- as.data.frame(assoc_data$memb)
80
 
# ## Look at variation within cluster - might not be useful here?
81
 
# veg.var <- clustvar(assoc_data) %>% as.data.frame
82
 
83
 
```
84
 
85
 
################ ###Test using split of guide classification into
86
 
87
 
training and testing data
88
 
89
 
```{r split data set and test classification}
90
 
91
 
veg.test <- sample_n(vegdata2,50)
92
 
test.plots <- row.names(veg.test)
93
 
assocs.test <- assocs %>% dplyr::filter(PlotNumber %in% test.plots) %>%  arrange(PlotNumber)
94
 
veg.test.diss <-  as.data.frame(as.matrix(all_dissim)) %>% select(-one_of(test.plots)) 
95
 
veg.test.diss <- veg.test.diss %>% filter(row.names(veg.test.diss) %in% test.plots)  
96
 
97
 
veg.model <- vegdata2 %>% dplyr::filter(!rownames(vegdata2) %in% test.plots) %>%  arrange(row.names(.))
98
 
assocs.model <-assocs %>% dplyr::filter(!PlotNumber %in% test.plots) %>%  arrange(PlotNumber)
99
 
assocs.model <- assocs.model %>% dplyr::select(-SiteUnit) %>%  column_to_rownames("PlotNumber") %>% t %>% as.vector
100
 
#assoc.model <- as.vegclust(veg.model, assocs.model, method="HNC", dnoise = 10)
101
 
assoc.model.diss <- as.vegclust(proxy::dist(veg.model, method="ejaccard", diag=TRUE, upper = FALSE), assocs.model, method="HNC", dnoise = 0.4)
102
 
103
 
vegComb <- vegclass(assoc.model.diss, veg.test.diss) ##classify vegNew
104
 
vegComb.memb <- vegComb[["memb"]]
105
 
newGroup <- dematrify(vegComb.memb) ##extract classification
106
 
newGroup <- newGroup[,1:2]
107
 
colnames(newGroup) <- c("PlotNumber","Class.pred")
108
 
newGroup <- newGroup %>% arrange(PlotNumber)
109
 
newGroup <- cbind(newGroup, assocs.test)
110
 
####merge actual classification for comparison
111
 
#colnames(newGroup)[3] <- "Actual"
112
 
113
 
```
114
 
115
 
#####Place new plots in predefined classes
116
 
117
 
Probably better to do this in smaller chunks of a hierarchy rather than
118
 
all at once
119
 
120
 
```{r classify new plots}
121
 
##conform data
122
 
123
 
cf = conformveg(wetland.31, wetland.10)
124
 
125
 
126
 
vegComb <- vegclass(vegOld.clust, vegNew) ##classify vegNew
127
 
vegComb.memb <- vegComb[["memb"]]
128
 
newGroup <- dematrify(vegComb.memb) ##extract classification
129
 
newGroup <- newGroup[,1:2]
130
 
colnames(newGroup) <- c("SiteSeries","Class")
131
 
newGroup <- cbind(newGroup, actualClass$Class)####merge actual classification for comparison
132
 
colnames(newGroup)[3] <- "Actual"
133
 
134
 
135
 
vegDat.chord <- decostand(SUsumMatrix[-1], "normalize") ##standardise data
136
 
vegMat <- vegDat.chord
137
 
vegMat <- cbind(SUsumMatrix[,1],vegDat.chord)
138
 
139
 
####create training and testing data sets
140
 
vegNew <- vegMat[rownames(vegMat) %in% sample(rownames(vegMat), 100, replace = FALSE),] 
141
 
vegOld <- vegMat[!(rownames(vegMat) %in% rownames(vegNew)),]
142
 
#actualClass <- vegNew[,1:2] ###siteseries to grouping lookup
143
 
#rownames(vegNew)  <- vegNew[1] ###set rownames to siteseries
144
 
k <- ncol(vegOld)
145
 
n <- nrow(vegOld)
146
 
###grouping of training set
147
 
grouping <- vegOld[1]
148
 
vegOld <- vegOld[,-(1)]
149
 
vegNew <- vegNew[,-(1)]
150
 
k <- ncol(vegOld)
151
 
n <- nrow(vegOld)
152
 
153
 
grouping <- as.vector.factor(grouping)
154
 
vegOld.clust <- as.vegclust(vegOld, grouping)###create noise clustering with grouping as classes
155
 
###Kmeans Clustering
156
 
vegOld.kmclst <- vegclust(x = vegOld[,-1], mobileCenters=5, method = "KM", nstart=20)###create noise clustering with grouping as classes
157
 
t(vegOld.kmclst$memb)
158
 
###NC Clustering
159
 
vegOld.kmclst <- vegclust(x = vegOld[,-1], mobileCenters=5, method = "NC", m=1.2, dnoise=0.8, nstart=20)###create noise clustering with grouping as classes
160
 
round(t(vegOld.kmclst$memb), dig=2)
161
 
groups = defuzzify(vegOld.kmclst, method="cut", alpha=0.8)$cluster
162
 
table(groups)
163
 
164
 
165
 
vegComb <- vegclass(vegOld.clust, vegNew) ##classify vegNew
166
 
vegComb.memb <- vegComb[["memb"]]
167
 
newGroup <- dematrify(vegComb.memb) ##extract classification
168
 
newGroup <- newGroup[,1:2]
169
 
colnames(newGroup) <- c("SiteSeries","Class")
170
 
newGroup <- cbind(newGroup, actualClass$Class)####merge actual classification for comparison
171
 
colnames(newGroup)[3] <- "Actual"
172
 
173
 
###MDS for visualisation
174
 
MDS <- metaMDS(vegNew, distance = "bray", k = 2, trymax = 200)
175
 
MDS.df <- as.data.frame(scores(MDS, display = "sites"))###extract mds scores
176
 
MDS.df$SiteSeries <- rownames(MDS.df)
177
 
MDS.df <- merge(MDS.df,newGroup, by = "SiteSeries") ##merge predicted and actual classification
178
 
MDS.df <- MDS.df[,-1]
179
 
180
 
colnames(MDS.df)[3:4] <- c("Actual","Predict")
181
 
182
 
ggplot(MDS.df)+
183
 
  geom_point(mapping = aes(x = NMDS1, y = NMDS2, colour = Predict), size = 2.5, shape = 17)+
184
 
  coord_fixed()+
185
 
  theme_bw()+
186
 
  theme(legend.position = "none")
187
 
```