---
title: "NewSiteUnitAnalysis"
author: "Will MacKenzie"
date: "05/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

require(tidyverse)
require(tidymodels)
require(data.table)
require(data.tree)
require(DataExplorer)
require(C50)
require(indicspecies)
require(doParallel)
require(philentropy)
require(harrietr)
require(factoextra)
source("./_functions/_TabletoTree.R")
source("./_functions/_TreetoTable.R")
source("./_functions/_VegdatSUsummary.R")
source("./_functions/_spp_importance.R")
cloud_dir <- "F:/OneDrive - Personal/OneDrive/BEC_Classification_Paper/"
#cloud_dir <- "F:/OneDrive - Government of BC/CCISSv12/"
```

Steps to build classification hierarchy
This script joins site series into Associations using cluster-analysis.
The to success is appropriate transformation of the site series data
1. Power 0.5 transform the data as per Chytry et al.
2. Multiply transformed value by % constancy = species_importance
3. Remove all records with low importance (<0.5)
4. Remove all non-vascular species (these may be used to separate Associations at a SubAssociation level)
** could look at including edatopic/climate similarity in the clustering equation (avalanching)

5. Cluster using a quantitative distance metric (extended Jaccard)
6. Use a complete chain clustering methods with will retain the dissimilarity metric as tree height
7. Cut hierarchy at set dissimilarity height 1st approximation (to be reviewed)
      <25%  = association level similarity
      <50% = alliance level
      <75% = suborder level
      <95% = order level
Add cluster info to a clean wide hierarchy with only site series
Add Class level to hold unplaced Site Series (those with fewer than 5 plots)
Export to Vpro Tree structure.

1. A machine learning model of high-level hierarchical BEC Forest units
a. Create a ML model of Forested SubOrders using BECv11 site series that meet minimum plot requirements. (going to Orders only causes issues where Pl is a dominant secondary species - many provincial units get placed into a Pinucon Order when applied only to the order level even though they have a subdominant tree species which places them in a different order)
b. Use only tree species for first round of model build for the forested Orders.
c. Predict membership of new BECv13 units and assign to orders then rebuild ML model.
d. Review any mis predicted site series for reassignment.
e. Compare similarity of tree composition within units (noise clustering) to find outliers that may represent new Orders (test by leaving out known units)
f. Create vegetation summaries of Orders and compare vegetation (all species now). Identify climatic indicator species groups.
g. Create climatic and site attributes summary of Orders 
h. Create maps of plot locations for each Order
i. Hierarchy tree graphic

2. Build Alliances. Focus should be on identifying species groups that reflect different site conditions.

a. Use some of the techniques of indicspecies package to create Alliances based on indicator group creation and analysis.
b. Try to build a machine learning model of Alliances
c. Vegetation and Environment summary of Alliances

3. Do pair-wise analysis of site series within Orders/Alliances using high constancy species to analyze site series and create Associations/SubAssociations.
a. Check for similarity between Associations of All orders
b. Build relationship graphic


4. Document hierarchy

5. Build model to assign new units and unassigned plots
Challenge here is to make a constant list of species between model and new units.
a. Predict membership of new site series (draft BECv13). Noise Clustering to test for novel units that do not fit any existing. Use machine learning to hierarchy
b. Predict Order membership of BECMaster plots that are unassigned to existing site series and add to an Order_unplaced site unit under each Order.

6. Build for non-forested units. Classes based on major site level differences rather than climate and using major species groups (e.g. Hydrophytic Carex spp )
May wish to assign a temporary notree pseudo species to all units that have no trees to help separate forested hierarchy from non-forested hierarchy (Hurdle model)

```{r import data}
### Vegetation From BECMaster cleaning script Long form Plot|Species|Cover (across all layers)
### or From Vpro export
#BGCZone <- fread(paste0(cloud_dir,"All_BGCs_Info_v12_2.csv"), data.table=FALSE)
vegDat2 <- fread("./clean_tabs/BECMaster_VegR_clean.csv", data.table = FALSE)
taxon.all <- fread("D:/CommonTables/SpeciesMaster/SpeciesMaster01Dec2020.csv", header = T, stringsAsFactors = F, strip.white = T) %>% filter(Codetype == "U")

nonvasc = c(9, 10, 11)
vegDat2 <- vegDat2 %>% filter(!Lifeform %in% nonvasc)
###SU table
#SUTab <- fread("./inputs/ALLBECSU_2021_SU.csv")
SUTab <- fread("./inputs/BECv13Analysis_Forest_17Sept2021_SU.csv")
SUTab$SiteUnit <-  trimws(SUTab$SiteUnit)
SS.names <- unique(SUTab$SiteUnit)
PlotsSS <- SUTab %>% group_by(SiteUnit) %>% summarize(plotcount = n())

Vpro.hier <- fread("./clean_tabs/BECv13_ForestHierarchy.csv")
SUhier <- treeToTable(Vpro.hier)
#vegdat_test <- left_join(SUTab, vegDat2)
# #####list of official current and future correlated units from BECdb
# current <- c('Current', 'Future')
# ecotypes <- c("Forest", "Deciduous")
# #### cleaned list of official site series
# BECdb_SS <- fread("./inputs/BECdb_SiteUnitv12.csv") %>% filter(Status %in% current) %>% filter(ArchivedinVersion == "") %>% filter(`Forest-NonForest` %in% ecotypes) %>% 
#   select(MergedBGC_SS, SS_NoSpace, SiteUnitLongName, SiteUnitScientificName) %>% rename (SiteUnit = 1, SSName = 3, SciName = 4) %>% distinct(SiteUnit, .keep_all = TRUE)
# #### compare SUTab to BECdb.  Whan all BECdb units to have data and SUTab to include no unofficial units
# SS_missing <- full_join(SUTab, BECdb_SS) %>% filter(is.na(PlotNumber))
##Import Vpro hierarchy table widen and reformat
```


##roll up into site series summary data
### create species importance values by sqrt cover * constancy
### filter out site units with few plots and species with low importance values
```{r summarize site series, echo=FALSE}
##roll up into site series summary data

vegSum <- VegdatSUsummary(vegDat2, SUTab)
source("./_functions/_spp_importance.R")
vegSum <- spp_importance(vegSum) 

### Filter out poor units and low importance species#
vegDat_good <- vegSum %>% dplyr::filter(nPlots > 4) %>% dplyr::filter(spp_importance > 0.05) ## remove rare or very low cover species in SU

Hier.clean <- SUhier$table
 #%>% rename(SiteUnit = SiteUnit))## Select only site series will enough plots
Hier.units <- Hier.clean %>% dplyr::select(SiteUnit, Class, Order, Suborder) %>% distinct()
Hier.data <- left_join(Hier.units, vegDat_good) %>% filter(!is.na(nPlots)) %>% arrange(Species) %>% distinct()
```


```{r some hierarchy stats}
classes <- unique(Hier.data$Class)
orders <- unique(Hier.data$Order)
suborders <- unique(Hier.data$Suborder)
orders
### Choose hierarchical level for analysis

class.dat <-
  Hier.data %>% dplyr::select(SiteUnit, Species, spp_importance) %>% matrify() #%>% filter(Order == "ORDER Pyi") 
#%>% filter(Suborder == "SUBORDER PyFd-Pseuspi" |  Suborder == "SUBORDER FdPy-Pseuspi (PP transition)" ) 

```
## Cluster Analysis Species by Site eris

``` {r cluster analysis, echo = F, result = F}
### determine statistical number of clusters to use
# n_clust <- fviz_nbclust(test.data, kmeans, nstart = 3, method = "gap_stat", nboot = 50, k.max = 5)#+
# #geom_vline(xintercept = 6, linetype = 2)
# #  labs(subtitle = "Gap statistic method")
# plot(n_clust) ## "optimal" cluster number
# 
# n_clust<-n_clust$data
# best.clust<-as.numeric(n_clust$clusters[which.max(n_clust$ymax)])
# 
# ## clusterability###numbers below 0.5 have low clusterability
# res <- get_clust_tendency(class.dat, n = nrow(class.dat)-1, graph = F)
# res$hopkins_stat
# res

```

```{r cluster and cut  }
 test.data <- class.dat# %>% filter(str_detect(row.names(class.dat), "ESSFwh"))%>% as.matrix# [7:8,] 
# test_jac.vegan <- vegan::vegdist(test.data, method = "sorensen", binary = F) %>% as.matrix
#test_jac <- philentropy::distance(test.data, method = "jaccard")
test_jac <- proxy::dist(test.data, method="ejaccard", diag=TRUE, upper = TRUE) ## ejaccard and philentropy jaccard are real-data versions extended jaccard opr taimoto

# , use.row.names = TRUE, as.dist.obj = TRUE)
test_jac.tab <- test_jac %>% as.matrix
test_jac.list <- melt_dist(test_jac.tab) %>% as.data.table %>% filter(dist < .25) %>% filter(str_detect(iso1, "ESSFwh"))

ss_clst <- hclust(test_jac, method = "complete")
 assocs <- cutree( ss_clst ,h =.25)
 alliances <- cutree(ss_clst, h=.5)
 suborders <- cutree(ss_clst, h=.75) 
 orders <- cutree(ss_clst, h=.95)
k = length(unique(assocs))
k2 <- length(unique(alliances))
k3 <- length(unique(suborders))
k4 <- length(unique(orders))
ss_dend <- fviz_dend (ss_clst , k=k4, color_labels_by_k = T, lwd = .5,
                    rect = T, cex = .5, horiz = T) %>%    plot()
  # 
assoc_SU <- as.data.frame(assocs) %>% rownames_to_column("SiteUnit") %>% mutate(assocs = paste0("ASSOC-", assocs))
newassocs <- unique(assoc_SU$assocs)
alliance_SU <- as.data.frame(alliances) %>% rownames_to_column("SiteUnit") %>% mutate(alliances = paste0("ALLIANCE-", alliances))
newalliance <- unique(alliance_SU$alliances)
suborder_SU <- as.data.frame(suborders) %>% rownames_to_column("SiteUnit") %>% mutate(suborders = paste0("SUBORDER-", suborders))
newsuborder <- unique(suborder_SU$suborders)
order_SU <- as.data.frame(orders) %>% rownames_to_column("SiteUnit") %>% mutate(orders = paste0("ORDER-", orders))
neworder <- unique(order_SU$orders)
#(#left_join(names)
    #ss_clst <- hcut(test_jac1,isdiss = TRUE, hc_method = "ward.D2")#k = best.clust,, hc_metric = "sorensen"
#test.data[test.data == 0] <- NA
# a=test.data[1,] %>% janitor::remove_empty(which = "cols") %>% colnames() ; b=test.data[2,]%>% janitor::remove_empty(which = "cols")%>% colnames()
# jaccard <- function(a, b) {
#     intersection = length(intersect(a, b))
#     union = length(a) + length(b) - intersection
#     return (intersection/union)
# }
# 
# #find Jaccard Similarity between the two sets 
# jaccard(a, b)
# 
# Stats123 <- dist( test.data, method="euclidean", diag=TRUE, upper = TRUE)
# Proxy123 <- proxy::dist(test.data, method="ejaccard", diag=TRUE, upper = TRUE)
# Philentropy123 <- philentropy::distance(test.data, method="jaccard")
```



```{r update hierarchy and then resubmit to create new model}
##convert matrix to tree, add in new units
#Hier.update <- fread("./clean_tabs/BEC13_HierarchyMatrixModelled_v1.csv")
Vpro.hier <- fread("./clean_tabs/BECv13_ForestHierarchy.csv")
SUhier <- treeToTable(Vpro.hier)
Hier.clean <- SUhier$table
Hier.clean[,c(1:11)] <- NA
#Hier.new <- Hier.clean[0, ]
#test <- left_join(assocs3, Hier.clean, by = c("siteunit" = "SiteUnit")) %>% dplyr::rename(SiteUnit = "siteunit")
Hier2 <- Hier.clean[assoc_SU, Assoc := i.assocs, on = .(SiteUnit)] 
Hier2 <- Hier2[alliance_SU, Alliance := i.alliances, on = .(SiteUnit)] 
Hier2 <- Hier2[suborder_SU, Suborder := i.suborders, on = .(SiteUnit)] 
Hier2 <- Hier2[order_SU, Order := i.orders, on = .(SiteUnit)] 
Hier2 <- Hier2 %>% mutate(Class = ifelse(is.na(Assoc), "Unplaced", "Placed" ))
#Hier3[,c(1:5,7)] <- NA#Assoc := assoc,
#Hier2 <- Hier2 %>% dplyr::select(-ID)
# # test2 <- test2 %>%  filter(Assoc == "Assoc_1")
# Hier.data2 <- left_join(Hier2, vegDat_good) %>% filter(!is.na(nPlots)) %>% arrange(Species) %>% distinct()
# ### prepare to rerun with new associations and unplaced site series
# class.dat2 <-
#   Hier.data2 %>% dplyr::select(Assoc, SiteUnit, Working, Species, spp_importance) %>% mutate(Working = ifelse(!is.na(Assoc), Assoc, SiteUnit)) %>% dplyr::select(Working, Species, spp_importance)
# class.dat2 <- class.dat2[ , mean(spp_importance, na.rm = TRUE), by = c("Working", "Species")] %>% dplyr::rename(spp_importance = 3) %>% dplyr::filter(spp_importance > 0.05) %>% matrify() #%>% filter(Order == "ORDER Pyi") 
# #%>% filter(Suborder == "SUBORDER PyFd-Pseuspi" |  Suborder == "SUBORDER FdPy-Pseuspi (PP transition)" ) 

```



```{r export to Vpro hierarchy tree}
levelNames <- c("Formation", "Class", "Order", "Suborder", "Alliance", "Suball", "Assoc", "Subass", "Facies", "Working", "SiteUnit")
#Hier.new <- Hier.clean[0, ]
Hier.updated <- tableToTree(Hier2,levelNames)## convert matrix to tree

Hier.updated$Parent[Hier.updated$Parent == 0] <- NA
#Hier.updated <- Hier.updated %>% mutate(Parent = ifelse(Name %in% newassocs, NA, Parent))

fwrite(Hier.updated, "./outputs/UpdatedVPROHierarchy_v13_assocs.csv")
```



```{r dendrogram output, fig.cap="Bray-Curtis Cluster of Mountain Flora with 6 Groups", fig.width = 8, fig.height=6 }
# xx <- sNNclust(test_jac1, k = 1, eps = .2, minPts = 1)
# xx$cluster
# yy <- frNN(test_jac1, eps = .5, query = NULL, sort = TRUE, search = "kdtree", bucketSize = 10, splitRule = "suggest", approx = 0)
# yy$dist
# yy$eps
# 
#  clu.h80
# test = c("CWH xm 1 /102", "CWH xm 2 /102")
# #%>% filter(row.names(class.dat) %in% test)
#  #test.data[test.data == 0] = NA
# test_jac <- vegan::vegdist(test.data, method = "bray", binary = F, na.rm = TRUE)# %>% as.matrix
# # %>% as.matrix
#  test_jac2 <- philentropy::distance(test.data, method = "jaccard")
# test_jac2 <- 1 - test_jac2
# test_jac3 <- melt_dist(test_jac2) %>% as.data.table
#  # ggsave("./outputs/Test_ClusterDendrogram.jpeg")
#  test_jac1 <- vegan::vegdist(class.dat, method = "jaccard", binary = FALSE) 
# 
#  test_jac <- philentropy::distance(class.dat, method = "jaccard")
# unitnames <- row.names(class.dat)
# test_jac2 <- 1 - test_jac %>% as.data.frame 
# names(test_jac2) <- unitnames
# #test_jac2 <-  cbind(unitnames, test_jac2)
# #test_jac2[test_jac2<.8]=NA ; test_jac2[test_jac2==1.00]=NA
# #test_jac2 <- join_metadata(test_jac2,unitnames)#cbind(rownames(class.dat), test_jac2) %>% as.data.frame %>% remove_rownames() %>% column_to_rownames("V1")
# 
# 
# test_jac <- philentropy::distance(class.dat, method = "jaccard")
# unitnames <- row.names(class.dat)
# test_jac3 <- melt_dist(test_jac2) %>% as.data.table
#   test_jac3[unitnames, iso1 := V2, on = c(iso1 = 'V1')] 
#   test_jac3[unitnames, iso2 := V2, on = c(iso2 = 'V1')] 
#   
#   assoc <- test_jac3 %>% filter(dist > .8)
#   assoc1 <- test_jac3 %>% filter(dist > .8) %>% dplyr::rename("SiteUnit" = iso1, "Same" = iso2 )
#   assoc2 <- test_jac3 %>% filter(dist > .8) %>% dplyr::rename("SiteUnit" = iso2, "Same" = iso1) %>% dplyr::select(SiteUnit, Same, dist)
#   assocs_all <- rbind(assoc1,assoc2)
# 
#     assoc_num <- unique(assocs_all$SiteUnit) %>% data.frame %>% dplyr::rename("SiteUnit" = 1)%>% arrange(SiteUnit) %>% rownames_to_column("assoc")%>% mutate(assoc = paste0("Assoc-", assoc)) 
#   assocs2 <- left_join(assoc_num, assocs_all)
#   
# assoc1 <- assocs 
# 
# 
#   assocs <- test_jac3 %>% filter(dist > .7)  %>% rownames_to_column("assoc")#%>% data.frame
# 
# assoc1 <- assocs %>% dplyr::select(assoc, iso1, dist) %>% dplyr::rename("SiteUnit" = iso1)
# assoc2 <- assocs %>% dplyr::select(assoc, iso2, dist) %>% dplyr::rename("SiteUnit" = iso2)
# 
# ##joins the closest pairs
# assocs3 <- assocs2[ , .SD[which.max(dist)], by = SiteUnit] 
# #assocs3 <- test_jac3 %>% filter(dist > .80) %>% group_by(iso1) %>% top_n(3, dist)
# #combine units in to associations then rerun jaccard
# 
# assocs_test <- test_jac3 %>% filter(dist > .7)
#  assocs_test2 <-  assocs_test [ , .SD[which.max(dist)], by = iso1]
# assocs_add <- anti_join(assocs_test, assocs_test2)
###build a dendrogram/cluster object and plot using dendextend
# mnt_clst <- siteseries %>% # list data
#         matrify %>% # convert to matrix
#         vegdist (method = "bray", binary = TRUE) %>% # calculate a distance matrix,
#         hclust(method = "ward.D2")%>% # Hierarchical clustering
#         plot()# Turn the object into a dendrogram and plot
# # 
# ss_clst2 <- as.data.frame(cutree (ss_clst, best.clust))
# colnames (ss_clst2) [1] <- "cluster"
# mnt_clst2 <- tibble::rownames_to_column(ss_clst2, var = "mountain")
# clust <- ss_clst2$cluster %>% count()

```